---
title: "Script to validate points in ReSurvey database using RS data (Satellite Embeddings Dataset)"
subtitle: "Validation done with ALL points (all observations)"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

This R script is used to validate the points in the ReSurvey database using only bands from the Google Satellite Embedding dataset: https://developers.google.com/earth-engine/datasets/catalog/GOOGLE_SATELLITE_EMBEDDING_V1_ANNUAL#description

# Load libraries

```{r}
library(tidyverse)
library(here)
library(gridExtra)
library(readxl)
library(scales)
library(sf)
library(rnaturalearth)
library(dtplyr)
library(lme4)
library(lmerTest)
library(car)
library(ggeffects)
library(party)
library(partykit)
library(moreparty)
library(doParallel)
library(strucchange)
library(ggparty)
library(caret)
library(moreparty)
library(randomForest)
library(pROC)
library(corrplot)
library(rlang)
library(stringr)
library(beepr)
library(foreach)
library(permimp)
library(yardstick)
```

# Define printall function

```{r}
printall <- function(tibble) {
  print(tibble, width = Inf)
  }
```

# Load geom_flat_violin plot

```{r}
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

# Load previously created objects

```{r}
# Define the folder path
folder_path <- here("objects", "RF", "Satellite_Embeddings")

# List all .RData or .rda files in the folder
rdata_files <- list.files(folder_path, full.names = TRUE)

# Load each file
lapply(rdata_files, load, envir = .GlobalEnv)
```

# Read data

```{r}
data_validation <- read_tsv(here(
  "data", "clean","final_RS_data_SatEmb_20250922.csv"))
```

No parsing issues!

# Some data managenemt

## TO-DO: Missing data checks

Do when all RS data is ready!

# Distributions all bioregions

```{r}
# Define a function to create histograms
plot_histogram <- function(data, x_var, x_label) {
  ggplot(data %>%
           dplyr::filter(EUNISa_1 %in% c("T", "R", "S", "Q")),
         aes(x = !!sym(x_var))) +
    geom_histogram(color = "black", fill = "white") +
    labs(x = x_label, y = "Frequency") +
    theme_bw()
}
```

```{r}
# Define a function to create plots with violin + boxplot + points
distr_plot <- function(data, y_vars, y_labels) {
  for (i in seq_along(y_vars)) {
    y_var <- y_vars[[i]]
    y_label <- y_labels[[i]]
    
    p <- ggplot(data = data %>%
                  dplyr::filter(EUNISa_1 %in% c("T", "R", "S", "Q")),
                aes(x = EUNISa_1_descr, y = !!sym(y_var), fill = EUNISa_1_descr)) +
      geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
      geom_point(aes(y = !!sym(y_var), color = EUNISa_1_descr),
                 position = position_jitter(width = 0.15), size = 1, alpha = 0.25) +
      geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
      stat_summary(fun.y = mean, geom = "point", shape = 20, size = 1) +
      stat_summary(fun.data = function(x) data.frame(y = max(x) + 0.1,
                                                     label = length(x)),
                   geom = "text", aes(label = ..label..), vjust = 0.5) +
      labs(y = y_label, x = "EUNIS level 1") +
      scale_x_discrete(labels = function(x) str_wrap(x, width = 15)) +
      guides(fill = FALSE, color = FALSE) +
      theme_bw() + coord_flip()
    
    print(p)
  }
}
```


Histograms to check that max and min values are ok:

```{r}
plot_histogram(data_validation, "A00", "A00")
plot_histogram(data_validation, "A01", "A01")
plot_histogram(data_validation, "A23", "A23")
plot_histogram(data_validation, "A32", "A32")
plot_histogram(data_validation, "A50", "A50")
plot_histogram(data_validation, "A61", "A61")
plot_histogram(data_validation, "A63", "A63")
```

Distribution plots:

```{r message=FALSE, warning=FALSE}
distr_plot(data_validation,
           c("A00", "A01", "A10", "A22", "A33",
             "A40", "A53", "A61", "A62", "A63"),
           c("A00", "A01", "A10", "A22", "A33",
             "A40", "A53", "A61", "A62", "A63"))
```

# Distributions per bioregion

```{r}
# Define a function to create plots with violin + boxplot + points
distr_plot_biogeo <- function(data, y_vars, y_labels) {
  plots <- list()
  
  for (i in seq_along(y_vars)) {
    y_var <- y_vars[[i]]
    y_label <- y_labels[[i]]
    
    p <- ggplot(data = data %>%
                  dplyr::filter(EUNISa_1 %in% c("T", "R", "S", "Q")),
                aes(x = EUNISa_1_descr, y = !!sym(y_var), fill = EUNISa_1_descr)) +
      geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
      geom_point(aes(y = !!sym(y_var), color = EUNISa_1_descr),
                 position = position_jitter(width = 0.15), size = 1, alpha = 0.25) +
      geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
      stat_summary(fun.y = mean, geom = "point", shape = 20, size = 1) +
      stat_summary(fun.data = function(x) data.frame(y = max(x) + 0.1,
                                                     label = length(x)),
                   geom = "text", aes(label = ..label..), vjust = 0.5) +
      labs(y = y_label, x = "EUNISa_1_descr") +
      scale_x_discrete(labels = function(x) str_wrap(x, width = 15)) +
      guides(fill = FALSE, color = FALSE) +
      theme_bw() + coord_flip() + facet_wrap(~ biogeo)
    
    plots[[y_var]] <- p
  }
  
  return(plots)
}
```

Distribution plots:

```{r message=FALSE, warning=FALSE}
distr_plot_biogeo(data_validation,
           c("A00", "A01", "A10", "A22", "A33",
             "A40", "A53", "A61", "A62", "A63"),
           c("A00", "A01", "A10", "A22", "A33",
             "A40", "A53", "A61", "A62", "A63"))
```

# Define functions for RF models

## Function for fitting RF models

RF models fitted using the conditional inference version of random forest (first cforest in package party, now fastcforest in package moreparty). Suggested if the data are highly correlated. Cforest is more stable in deriving variable importance values in the presence of highly correlated variables, thus providing better accuracy in calculating variable importance (ref below).

Hothorn, T., Hornik, K. and Zeileis, A. (2006) Unbiased Recursive Portioning: A Conditional Inference Framework. Journal of Computational and Graphical Statistics, 15, 651-
674. http://dx.doi.org/10.1198/106186006X133933

Choose the hyperparameter mtry based on the square root of the number of predictor variables:

Hastie, T., Tibshirani, R., & Friedman, J. (2009). The elements of statistical
learning: Data mining, inference, and prediction. Springer Science &
Business Media.

Maybe TO_DO:
We variated ntree from 50 to 800 in steps of 50, leaving mtry constant at 2. Tis parameter variation showed that ntree=500 was optimal, while higher ntree led to no further model improvement (Supplementary Fig. S10). Subsequently, the hyperparameter mtry was varied from 2 to 8 with constant ntree=500. Here, mtry=3 led to the best results in almost all cases (Supplementary Fig. S11). Consequently, we chose ntree=500 and mtry=3 for our main analysis across all study sites.

Define a function to run fastcforest models:

```{r}
run_rf <- function(vars_RF, train_data, response_var, ntree = 500) 
  {
  
  # Detect and register available cores (leave one free)
  n_cores <- parallel::detectCores() - 1
  cl <- makeCluster(n_cores)
  registerDoParallel(cl)
  
  train_name <- deparse(substitute(train_data))
  
  # Export necessary variables to the cluster
  clusterExport(cl, varlist = c("vars_RF", "train_data", "response_var"),
                envir = environment())

  
  # Set seed for reproducibility
  set.seed(123)
  
  # Measure execution time
  execution_time <- system.time({
    rf_model <- fastcforest(
      formula = reformulate(vars_RF, response = response_var),
      data = train_data,
      controls = party::cforest_control(
        mtry = round(sqrt(length(vars_RF))),
        ntree = ntree
      ),
      parallel = TRUE
    )
  })
  
  # Stop the cluster
  stopCluster(cl)
  
  # Return both the model and execution time
  list(model = rf_model, time = execution_time)
}
```

## Function to compute variable importance

```{r}
# compute_varimp <- function(model, nperm = 100, 
#                                    n_cores = parallel::detectCores() - 1) {
#   # Set up parallel backend
#   cl <- makeCluster(n_cores)
#   registerDoParallel(cl)
#   
#   # Measure execution time
#   execution_time <- system.time({
#     varimp_list <- foreach(i = 1:nperm, .combine = '+', 
#                            .packages = "party") %dopar% {
#       varimp(model, conditional = FALSE, nperm = 1)
#     }
#   })
#   
#   stopCluster(cl)
#   
#   # Average the results
#   varimp_avg <- varimp_list / nperm
#   
#   return(list(varimp = varimp_avg, time = execution_time))
# }
```

Using permimp() en permimp package:
https://cran.r-project.org/web/packages/permimp/vignettes/permimp-package.html#fn1

```{r}
compute_varimp <- function(model, nperm = 100) {

  # Measure execution time
  execution_time <- system.time({
    varimp_result <- permimp(model, conditional = FALSE, progressBar = TRUE)
  })

  return(list(varimp = varimp_result, time = execution_time))
}
```

## Function to compute CONDITIONAL variable importance

```{r}
compute_varimp_cond <- function(model, nperm = 100) {

  # Measure execution time
  execution_time <- system.time({
    varimp_result <- permimp(model, conditional = TRUE, progressBar = TRUE)
  })

  return(list(varimp = varimp_result, time = execution_time))
}
```

## Function to compute ROC (level 1)

```{r}
compute_roc_level1 <- function(model, test_data) {
  # Measure execution time
  execution_time <- system.time({
    # Step 1: Predict probabilities
    probabilities <- predict(model, newdata = test_data, type = "prob")
    
    # Step 2: Convert list of matrices to a proper data frame
    prob_matrix <- t(sapply(probabilities, as.vector))
    prob_df <- as.data.frame(prob_matrix)
    colnames(prob_df) <- c("Q", "R", "S", "T")
    
    # Step 3: Prepare actual class labels
    actual <- factor(test_data$EUNISa_1, levels = c("Q", "R", "S", "T"))
    
    # Step 4: Binarize actual labels
    actual_bin <- model.matrix(~ actual - 1)
    colnames(actual_bin) <- gsub("actual", "", colnames(actual_bin))
    
    # Step 5: Compute ROC data for each class
    roc_data <- lapply(levels(actual), function(class) {
      roc_obj <- roc(actual_bin[, class], prob_df[[class]])
      auc_val <- round(auc(roc_obj), 3)
      data.frame(
        FPR = rev(roc_obj$specificities),
        TPR = rev(roc_obj$sensitivities),
        Class = paste0(class, " (AUC = ", auc_val, ")")
      )
    }) %>% bind_rows()
  })
  
  # Return both ROC data and execution time
  return(list(roc = roc_data, time = execution_time))
}
```

## Function to compute ROC (level 2)

```{r}
compute_roc_level2 <- function(model, test_data) {
  # Measure execution time
  execution_time <- system.time({
    # Step 1: Predict probabilities
    probabilities <- predict(model, newdata = test_data, type = "prob")
    
    # Step 2: Convert list of matrices to a proper data frame
    prob_matrix <- t(sapply(probabilities, as.vector))
    prob_df <- as.data.frame(prob_matrix)
    colnames(prob_df) <- c("Q1", "Q2", "Q4", "Q5", "R1", "R2", "R3", "R4", "R5",
                           "R6", "S3", "S4", "T1", "T3")
    
    # Step 3: Prepare actual class labels
    actual <- factor(test_data$EUNISa_2, 
                     levels = c("Q1", "Q2", "Q4", "Q5", "R1", "R2", "R3", "R4",
                                "R5", "R6", "S3", "S4", "T1", "T3"))
    
    # Step 4: Binarize actual labels
    actual_bin <- model.matrix(~ actual - 1)
    colnames(actual_bin) <- gsub("actual", "", colnames(actual_bin))
    
    # Step 5: Compute ROC data for each class
    roc_data <- lapply(levels(actual), function(class) {
      roc_obj <- roc(actual_bin[, class], prob_df[[class]])
      auc_val <- round(auc(roc_obj), 3)
      data.frame(
        FPR = rev(roc_obj$specificities),
        TPR = rev(roc_obj$sensitivities),
        Class = paste0(class, " (AUC = ", auc_val, ")")
      )
    }) %>% bind_rows()
  })
  
  # Return both ROC data and execution time
  return(list(roc = roc_data, time = execution_time))
}
```

# Define list of predictor vars

```{r}
vars_RF_SatEmb <- colnames(
  data_validation)[startsWith(colnames(data_validation), "A")]
vars_RF_SatEmb_CH <- c(vars_RF_SatEmb, "canopy_height")
```

# Correlation

Correlation of all variables to be included in RF models:

```{r}
corrplot(data_validation %>% 
           dplyr::select(starts_with("A")) %>%
           cor(use = "pairwise.complete.obs"),
         method = "color", type = "upper", tl.col = "black", tl.srt = 45)
corrplot(data_validation %>% 
           dplyr::select(starts_with("A"), canopy_height) %>%
           cor(use = "pairwise.complete.obs"),
         method = "color", type = "upper", tl.col = "black", tl.srt = 45)
```

# Rough validation

Define a set of rules for a first validation of ALL ReSurvey data ("Expert-based" rules). Not very ambitious, only taking out observations that are clearly wrong on the basis of indicator values.

Create column for first validation based on different indicators, where "wrong" is noted when the validation rule is not met. 

Here only using validation based on canopy height, not NDWI.

Define rules:

```{r}
data_validation <-
  data_validation %>%
  mutate(
    valid_1_CH = case_when(
      # R & Q points with high CH
      EUNISa_1 %in% c("R", "Q") & canopy_height > 2 ~ "wrong",
      # T points with low CH
      EUNISa_1 == "T" & canopy_height < 3 ~ "wrong",
      # S points with high CH
      EUNISa_1 == "S" & canopy_height > 3 ~ "wrong",
      TRUE ~ NA_character_),
    # Count how many validation rules are not met
    valid_1_count = rowSums(across(c(valid_1_CH), ~ . == "wrong"),
                            na.rm = TRUE),
    # Points where at least 1 rule not met
    valid_1 = if_else(valid_1_count > 0, "At least 1 rule broken",
                      "No rules broken so far")
    )
```

# Fit RF models (level 1)

## Without refinement

### Get filtered data

```{r}
# No validation
data_validation_novalid <- data_validation %>%
  # Select only GPS points
  dplyr::filter(Lctnmth == "Location with GPS" |
                  Lctnmth == "Location with differential GPS") %>%
  select(-Lctnmth) %>%
  mutate(EUNISa_1 = as.factor(EUNISa_1))

# Rough validation
data_validation_roughvalid <- data_validation %>%
  # Select only GPS points
  dplyr::filter(Lctnmth == "Location with GPS" |
                  Lctnmth == "Location with differential GPS") %>%
  # Filter out points that have not passed the rough validation
  dplyr::filter(valid_1 == "No rules broken so far") %>%
  select(-Lctnmth, -valid_1) %>%
  mutate(EUNISa_1 = as.factor(EUNISa_1))
```

### N points per category

```{r}
bind_rows(
  data_validation_novalid %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(data = "data_validation_novalid") %>%
    select(data, Q, R, S, T),
  data_validation_roughvalid %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(data = "data_validation_roughvalid") %>%
    select(data, Q, R, S, T),
  )
```

### Split into training and test data sets

```{r}
set.seed(123)
```

```{r}
train_indices_novalid <- sample(1:nrow(data_validation_novalid), 
                         0.7 * nrow(data_validation_novalid))
train_indices_roughvalid <- sample(1:nrow(data_validation_roughvalid), 
                         0.7 * nrow(data_validation_roughvalid))
```

```{r}
train_data_novalid <- data_validation_novalid[train_indices_novalid, ]
train_data_roughvalid <- data_validation_roughvalid[train_indices_roughvalid, ]
```

```{r}
test_data_novalid <- data_validation_novalid[-train_indices_novalid, ]
test_data_roughvalid <- data_validation_roughvalid[-train_indices_roughvalid, ]
```

### Fit models

```{r eval=FALSE, include=FALSE}
# 1: No validation, SatEmb bands
rf1 <- run_rf(vars_RF_SatEmb, train_data_novalid, "EUNISa_1")
# 2: No validation, SatEmb bands + CH
rf2 <- run_rf(vars_RF_SatEmb_CH, train_data_novalid, "EUNISa_1")
# 3: Rough validation, SatEmb bands
rf3 <- run_rf(vars_RF_SatEmb, train_data_roughvalid, "EUNISa_1")
# 4: Rough validation, SatEmb bands + CH
rf4 <- run_rf(vars_RF_SatEmb_CH, train_data_roughvalid, "EUNISa_1")
```

```{r}
print(rf1$time)
print(rf2$time)
print(rf3$time)
print(rf4$time)
```

```{r eval=FALSE, include=FALSE}
save(rf1, file = "objects/RF/Satellite_Embeddings/rf1_model_l1_novalid_norefin_SatEmb.Rdata")
save(rf2, file = "objects/RF/Satellite_Embeddings/rf2_model_l1_novalid_norefin_Sat_Emb_CH.Rdata")
save(rf3, file = "objects/RF/Satellite_Embeddings/rf3_model_l1_roughvalid_norefin_Sat_Emb.Rdata")
save(rf4, file = "objects/RF/Satellite_Embeddings/rf4_model_l1_roughvalid_norefin_sat_Emb_CH.Rdata")
```

### Predictions

```{r eval=FALSE, include=FALSE}
predictions_rf1 <- predict(rf1$model, newdata = test_data_novalid,OOB = TRUE,
                           type = "response")
predictions_rf2 <- predict(rf2$model, newdata = test_data_novalid,OOB = TRUE,
                           type = "response")
predictions_rf3 <- predict(rf3$model, newdata = test_data_roughvalid,OOB = TRUE,
                           type = "response")
predictions_rf4 <- predict(rf4$model, newdata = test_data_roughvalid,OOB = TRUE,
                           type = "response")
```

```{r eval=FALSE, include=FALSE}
save(predictions_rf1, file = "objects/RF/Satellite_Embeddings/rf1_pred_l1_novalid_norefin_SatEmb.Rdata")
save(predictions_rf2, file = "objects/RF/Satellite_Embeddings/rf1_pred_l1_novalid_norefin_SatEmb_CH.Rdata")
save(predictions_rf3, file = "objects/RF/Satellite_Embeddings/rf3_pred_l1_roughvalid_norefin_Sat_Emb.Rdata")
save(predictions_rf4, file = "objects/RF/Satellite_Embeddings/rf3_pred_l1_roughvalid_norefin_Sat_Emb_CH.Rdata")
```

### Confusion matrices

```{r}
# 1: No validation, SatEmb bands
confusionMatrix(predictions_rf1, test_data_novalid$EUNISa_1)
# 2: No validation, SatEmb bands + CH
confusionMatrix(predictions_rf2, test_data_novalid$EUNISa_1)
# 3: Rough validation, SatEmb bands
confusionMatrix(predictions_rf3, test_data_roughvalid$EUNISa_1)
# 4: Rough validation, SatEmb bands + CH
confusionMatrix(predictions_rf4, test_data_roughvalid$EUNISa_1)
```

### Variable importance

#### Unconditional

```{r eval=FALSE, include=FALSE}
varimp_rf1 <- compute_varimp(rf1$model, nperm = 100)
varimp_rf2 <- compute_varimp(rf2$model, nperm = 100)
varimp_rf3 <- compute_varimp(rf3$model, nperm = 100)
varimp_rf4 <- compute_varimp(rf4$model, nperm = 100)
```

```{r eval=FALSE, include=FALSE}
print(varimp_rf1$time)
print(varimp_rf2$time)
print(varimp_rf3$time)
print(varimp_rf4$time)
```

```{r eval=FALSE, include=FALSE}
save(varimp_rf1, file = "objects/RF/Satellite_Embeddings/rf1_varimp_l1_novalid_norefin_SatEmb.Rdata")
save(varimp_rf2, file = "objects/RF/Satellite_Embeddings/rf2_varimp_l1_novalid_norefin_SatEmb_CH.Rdata")
save(varimp_rf3, file = "objects/RF/Satellite_Embeddings/rf3_varimp_l1_roughvalid_norefin_SatEmb.Rdata")
save(varimp_rf4, file = "objects/RF/Satellite_Embeddings/rf4_varimp_l1_roughvalid_norefin_SatEmb_CH.Rdata")
```

##### Plots

```{r}
plot(varimp_rf1$varimp, margin = c(10, 6, 2, 2))
plot(varimp_rf2$varimp, margin = c(10, 6, 2, 2))
plot(varimp_rf3$varimp, margin = c(10, 6, 2, 2))
plot(varimp_rf4$varimp, margin = c(10, 6, 2, 2))
```




# HERE



## No validation

### No refinement

#### SatEmb bands

#### SatEmb + CH

### Refinement 10-90th: 1 variable

### Refinement 10-90th: 2 variables

### Refinement 10-90th: 3 variables

### Refinement 10-90th: 4 variables


## Rough validation


## Without refinement

### Get filtered data

```{r}
data_validation_GPS <- data_validation %>%
  # Select only GPS points
  dplyr::filter(Lctnmth == "Location with GPS" |
                  Lctnmth == "Location with differential GPS") %>%
  select(-Lctnmth) %>%
  mutate(EUNISa_1 = as.factor(EUNISa_1))
```

#### N points per category

```{r}
data_validation_GPS %>% count(EUNISa_1)
```

### Split into training and test data sets

```{r}
set.seed(123)
```

```{r}
train_indices1 <- sample(1:nrow(data_validation_GPS), 
                         0.7 * nrow(data_validation_GPS))
```

```{r}
train_data1 <- data_validation_GPS[train_indices1, ]
```

```{r}
test_data1 <- data_validation_GPS[-train_indices1, ]
```

### Fit models

```{r eval=FALSE, include=FALSE}
rf1 <- run_rf(vars_RF_SatEmb, train_data1, "EUNISa_1")
rf1_CH <- run_rf(vars_RF_SatEmb_CH, train_data1, "EUNISa_1")
```

```{r}
print(rf1$time)
print(rf1_CH$time)
```

```{r eval=FALSE, include=FALSE}
save(rf1, file = "objects/RF/Satellite_Embeddings/rf1_model_l1_novalid_GPS_SatEmb.Rdata")
save(rf1_CH, file = "objects/RF/Satellite_Embeddings/rf1_model_l1_novalid_GPS_SatEmb_CH.Rdata")
```

### Predictions

```{r eval=FALSE, include=FALSE}
predictions_rf1 <- predict(rf1$model, newdata = test_data1,OOB = TRUE,
                           type = "response")
predictions_rf1_CH <- predict(rf1_CH$model, newdata = test_data1,OOB = TRUE,
                           type = "response")
```

```{r eval=FALSE, include=FALSE}
save(predictions_rf1, file = "objects/RF/Satellite_Embeddings/rf1_pred_l1_novalid_GPS_SatEmb.Rdata")
save(predictions_rf1_CH, file = "objects/RF/Satellite_Embeddings/rf1_pred_l1_novalid_GPS_SatEmb_CH.Rdata")
```

### Confusion matrix

```{r}
confusionMatrix(predictions_rf1, test_data1$EUNISa_1)
confusionMatrix(predictions_rf1_CH, test_data1$EUNISa_1)
```
### Variable importance

#### Unconditional

```{r eval=FALSE, include=FALSE}
varimp_rf1 <- compute_varimp(rf1$model, nperm = 100)
```

```{r eval=FALSE, include=FALSE}
print(varimp_rf1$time)
```

```{r eval=FALSE, include=FALSE}
save(varimp_rf1, file = "objects/RF/S2/rf1_varimp_l1_slope_novalid_GPS_S2.Rdata")
```

##### Plot

```{r}
plot(varimp_rf1$varimp, margin = c(10, 6, 2, 2))
```

#### Conditional (TBD)

```{r eval=FALSE, include=FALSE}
varimp-cond_rf1 <- compute_varimp_cond(rf1$model, nperm = 100)
```

```{r eval=FALSE, include=FALSE}
print(varimp-cond_rf1$time)
```

```{r eval=FALSE, include=FALSE}
save(varimp-cond_rf1, 
     file = "objects/RF/S2/rf1_varimp-cond_l1_slope_novalid_GPS_S2.Rdata")
```

##### Plot

```{r}
plot(varimp-cond_rf1$varimp, margin = c(9, 3, 2, 2))
```

### ROC curves (TBD)

```{r eval=FALSE, include=FALSE}
roc_data1 <- compute_roc_level1(rf1$model, test_data1)
```

```{r}
print(roc_data1$time)
```

```{r}
save(roc_data1, 
     file = "objects/RF/S2/rf1_rocdata_l1_slope_novalid_GPS_S2.Rdata")
```

#### Plot

```{r}
roc1 <- ggplot(roc_data1$roc, aes(x = FPR, y = TPR, color = Class)) +
  geom_line(size = 1.2) + geom_abline(linetype = "dashed", color = "gray") +
  labs(title = "Multiclass ROC Curves with AUC", x = "False Positive Rate",
       y = "True Positive Rate", color = "Class (AUC)") +
  theme_minimal() + theme(legend.position = "bottom")
```

```{r}
roc1
```

## With refinement

### Refinement

Based on the 1st, 2nd, 3rd, 4th... most important variables: A59, A15, A12, A60.

```{r}
distr_plot_percentiles <- function(data, y_vars, y_labels) {
  for (i in seq_along(y_vars)) {
    y_var <- y_vars[[i]]
    y_label <- y_labels[[i]]
    
    # Calculate percentiles per EUNISa_1 group
    percentiles <- data %>%
      group_by(EUNISa_1) %>%
      summarise(
        p10 = quantile(.data[[y_var]], 0.1, na.rm = TRUE),
        p90 = quantile(.data[[y_var]], 0.9, na.rm = TRUE),
        .groups = "drop"
      )
    
    # Join percentiles back to data
    data_flagged <- data %>%
      left_join(percentiles, by = "EUNISa_1") %>%
      mutate(outlier_flag = case_when(
        .data[[y_var]] < p10 ~ "low",
        .data[[y_var]] > p90 ~ "high",
        TRUE ~ "mid"
      ))
    
    # Filter and plot
    p <- ggplot(data = data_flagged %>%
                  dplyr::filter(EUNISa_1 %in% c("T", "R", "S", "Q")),
                aes(x = EUNISa_1_descr, y = .data[[y_var]])) +
      geom_flat_violin(aes(fill = EUNISa_1_descr),
                       position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
      geom_point(aes(color = ifelse(outlier_flag == "mid",
                                    EUNISa_1_descr, "grey")),
                 position = position_jitter(width = 0.15), size = 1,
                 alpha = 0.6) +
      geom_boxplot(aes(fill = EUNISa_1_descr), width = 0.2, outlier.shape = NA,
                   alpha = 0.5) +
      stat_summary(fun = mean, geom = "point", shape = 20, size = 1) +
      stat_summary(fun.data = function(x) data.frame(y = max(x, na.rm = TRUE) +
                                                       0.1, label = length(x)),
                   geom = "text", aes(label = ..label..), vjust = 0.5) +
      labs(y = y_label, x = "EUNIS level 1") +
      scale_x_discrete(labels = function(x) str_wrap(x, width = 15)) +
      guides(fill = FALSE, color = FALSE) +
      theme_bw() + coord_flip() +
      scale_color_manual(values = c(
        "Forests and other wooded land" = "#F8766D",
        "Grasslands" = "#7CAE00",
        "Heathlands, scrub and tundra" = "#00BFC4",
        "Wetlands" = "#C77CFF",
        "grey" = "grey"))
    
    print(p)
  }
}
```

```{r}
distr_plot_percentiles(
  # GPS points
  data_validation_GPS,
  c("A59", "A15", "A12", "A60"),
  c("A59", "A15", "A12", "A60"))
```

Calculate percentiles:

```{r}
percentiles <- 
  data_validation_GPS %>%
  group_by(EUNISa_1) %>%
  summarize(
    perc_10_A59 = quantile(A59, probs = 0.10, na.rm = T),
    perc_20_A59 = quantile(A59, probs = 0.20, na.rm = T),
    perc_80_A59 = quantile(A59, probs = 0.80, na.rm = T),
    perc_90_A59 = quantile(A59, probs = 0.90, na.rm = T),
    perc_10_A15 = quantile(A15, probs = 0.10, na.rm = T),
    perc_20_A15 = quantile(A15, probs = 0.20, na.rm = T),
    perc_80_A15 = quantile(A15, probs = 0.80, na.rm = T),
    perc_90_A15 = quantile(A15, probs = 0.90, na.rm = T),
    perc_10_A12 = quantile(A12, probs = 0.10, na.rm = T),
    perc_20_A12 = quantile(A12, probs = 0.20, na.rm = T),
    perc_80_A12 = quantile(A12, probs = 0.80, na.rm = T),
    perc_90_A12 = quantile(A12, probs = 0.90, na.rm = T),
    perc_10_A60 = quantile(A60, probs = 0.10, na.rm = T),
    perc_20_A60 = quantile(A60, probs = 0.20, na.rm = T),
    perc_80_A60 = quantile(A60, probs = 0.80, na.rm = T),
    perc_90_A60 = quantile(A60, probs = 0.90, na.rm = T),
    )
```

### Get filtered data

```{r}
data_validation_GPS_refin1 <- 
  data_validation_GPS %>%
  left_join(percentiles, by = "EUNISa_1") %>%
  mutate(EUNISa_1 = as.factor(EUNISa_1)) %>%
  dplyr::filter(
    (A59 >= perc_10_A59 & A59 <= perc_90_A59)
    )
data_validation_GPS_refin2 <- 
  data_validation_GPS %>%
  left_join(percentiles, by = "EUNISa_1") %>%
  mutate(EUNISa_1 = as.factor(EUNISa_1)) %>%
  dplyr::filter(
    (A59 >= perc_10_A59 & A59 <= perc_90_A59) &
      (A15 >= perc_10_A15 & A15 <= perc_90_A15)
    )
data_validation_GPS_refin3 <- 
  data_validation_GPS %>%
  left_join(percentiles, by = "EUNISa_1") %>%
  mutate(EUNISa_1 = as.factor(EUNISa_1)) %>%
  dplyr::filter(
    (A59 >= perc_10_A59 & A59 <= perc_90_A59) &
      (A15 >= perc_10_A15 & A15 <= perc_90_A15) &
      (A12 >= perc_10_A12 & A12 <= perc_90_A12)
    )
data_validation_GPS_refin4 <- 
  data_validation_GPS %>%
  left_join(percentiles, by = "EUNISa_1") %>%
  mutate(EUNISa_1 = as.factor(EUNISa_1)) %>%
  dplyr::filter(
    (A59 >= perc_10_A59 & A59 <= perc_90_A59) &
      (A15 >= perc_10_A15 & A15 <= perc_90_A15) &
      (A12 >= perc_10_A12 & A12 <= perc_90_A12) &
      (A60 >= perc_10_A60 & A60 <= perc_90_A60)
    )
```

#### N plots per category

```{r}
bind_rows(
  data_validation_GPS_refin1 %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(data = "data_validation_GPS_refin1") %>%
    select(data, Q, R, S, T),
  data_validation_GPS_refin2 %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(data = "data_validation_GPS_refin2") %>%
    select(data, Q, R, S, T),
  data_validation_GPS_refin3 %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(data = "data_validation_GPS_refin3") %>%
    select(data, Q, R, S, T),
  data_validation_GPS_refin4 %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(data = "data_validation_GPS_refin4") %>%
    select(data, Q, R, S, T)
  )
```

### Split into training and test data sets

```{r}
set.seed(123)
```

```{r}
train_indices2 <- sample(1:nrow(data_validation_GPS_refin1), 
                         0.7 * nrow(data_validation_GPS_refin1))
train_indices3 <- sample(1:nrow(data_validation_GPS_refin2), 
                         0.7 * nrow(data_validation_GPS_refin3))
train_indices4 <- sample(1:nrow(data_validation_GPS_refin3), 
                         0.7 * nrow(data_validation_GPS_refin3))
train_indices5 <- sample(1:nrow(data_validation_GPS_refin4), 
                         0.7 * nrow(data_validation_GPS_refin4))
```

```{r}
train_data2 <- data_validation_GPS_refin1[train_indices2, ]
train_data3 <- data_validation_GPS_refin2[train_indices3, ]
train_data4 <- data_validation_GPS_refin3[train_indices4, ]
train_data5 <- data_validation_GPS_refin4[train_indices5, ]
```

```{r}
test_data2 <- data_validation_GPS_refin1[-train_indices2, ]
test_data3 <- data_validation_GPS_refin2[-train_indices3, ]
test_data4 <- data_validation_GPS_refin3[-train_indices4, ]
test_data5 <- data_validation_GPS_refin4[-train_indices5, ]
```

### Fit models

```{r eval=FALSE, include=FALSE}
rf2 <- run_rf(vars_RF_SatEmb, train_data2, "EUNISa_1")
rf3 <- run_rf(vars_RF_SatEmb, train_data3, "EUNISa_1")
rf4 <- run_rf(vars_RF_SatEmb, train_data4, "EUNISa_1")
rf5 <- run_rf(vars_RF_SatEmb, train_data5, "EUNISa_1")
```

```{r}
print(rf2$time)
print(rf3$time)
print(rf4$time)
print(rf5$time)
```

```{r eval=FALSE, include=FALSE}
save(rf2, 
     file = "objects/RF/Satellite_Embeddings/rf2_model_l1_novalid-refin1090_1v_GPS_SatEmb.Rdata")
save(rf3, 
     file = "objects/RF/Satellite_Embeddings/rf3_model_l1_novalid-refin1090_2v_GPS_SatEmb.Rdata")
save(rf4, 
     file = "objects/RF/Satellite_Embeddings/rf4_model_l1_novalid-refin1090_3v_GPS_SatEmb.Rdata")
save(rf5, 
     file = "objects/RF/Satellite_Embeddings/rf5_model_l1_novalid-refin1090_4v_GPS_SatEmb.Rdata")
```

### Predictions

```{r eval=FALSE, include=FALSE}
predictions_rf2 <- predict(rf2$model, newdata = test_data2,OOB = TRUE,
                           type = "response")
predictions_rf3 <- predict(rf3$model, newdata = test_data3,OOB = TRUE,
                           type = "response")
predictions_rf4 <- predict(rf4$model, newdata = test_data4,OOB = TRUE,
                           type = "response")
predictions_rf5 <- predict(rf5$model, newdata = test_data5,OOB = TRUE,
                           type = "response")
```

```{r eval=FALSE, include=FALSE}
save(predictions_rf2, 
     file = "objects/RF/Satellite_Embeddings/rf2_pred_l1_novalid-refin1090_1v_GPS_SatEmb.Rdata.Rdata")
save(predictions_rf3, 
     file = "objects/RF/Satellite_Embeddings/rf3_pred_l1_novalid-refin1090_2v_GPS_SatEmb.Rdata.Rdata")
save(predictions_rf4, 
     file = "objects/RF/Satellite_Embeddings/rf4_pred_l1_novalid-refin1090_3v_GPS_SatEmb.Rdata.Rdata")
save(predictions_rf5, 
     file = "objects/RF/Satellite_Embeddings/rf5_pred_l1_novalid-refin1090_4v_GPS_SatEmb.Rdata.Rdata")
```

### Confusion matrices

```{r}
confusionMatrix(predictions_rf2, test_data2$EUNISa_1)
confusionMatrix(predictions_rf3, test_data3$EUNISa_1)
confusionMatrix(predictions_rf4, test_data4$EUNISa_1)
confusionMatrix(predictions_rf5, test_data5$EUNISa_1)
```

### Variable importance

#### Unconditional

```{r eval=FALSE, include=FALSE}
varimp_rf2 <- compute_varimp(rf2$model, nperm = 100)
varimp_rf3 <- compute_varimp(rf3$model, nperm = 100)
varimp_rf4 <- compute_varimp(rf4$model, nperm = 100)
varimp_rf5 <- compute_varimp(rf5$model, nperm = 100)
```

```{r eval=FALSE, include=FALSE}
print(varimp_rf2$time)
print(varimp_rf3$time)
print(varimp_rf4$time)
print(varimp_rf5$time)
```

```{r eval=FALSE, include=FALSE}
save(varimp_rf2,
     file = "objects/RF/Satellite_Embeddings/rf2_varimp_l1_novalid-refin1090_1v_GPS_SatEmb.Rdata.Rdata")
save(varimp_rf3,
     file = "objects/RF/Satellite_Embeddings/rf3_varimp_l1_novalid-refin1090_2v_GPS_SatEmb.Rdata.Rdata")
save(varimp_rf4,
     file = "objects/RF/Satellite_Embeddings/rf4_varimp_l1_novalid-refin1090_3v_GPS_SatEmb.Rdata.Rdata")
save(varimp_rf5,
     file = "objects/RF/Satellite_Embeddings/rf5_varimp_l1_novalid-refin1090_4v_GPS_SatEmb.Rdata.Rdata")
```

##### Plots

```{r}
plot(varimp_rf2$varimp, margin = c(10, 6, 2, 2))
plot(varimp_rf3$varimp, margin = c(10, 6, 2, 2))
plot(varimp_rf4$varimp, margin = c(10, 6, 2, 2))
plot(varimp_rf5$varimp, margin = c(10, 6, 2, 2))
```

# INSERT ABOVE: Fit RF models (level 1) with CH



# Session info

```{r}
sessionInfo()
```

