---
title: "Script to prepare data from ReSurvey database v2025.2"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

This R script is used to prepare data of the ReSurvey database for further analyses, starting from the raw data downloaded from:
https://onedrive.live.com/?redeem=aHR0cHM6Ly8xZHJ2Lm1zL2YvYy8zOWE2ZDZiNzFmMzM1ZTE4L0VoaGVNeC0zMXFZZ2dEbmo0d0lBQUFBQkEwdmo5RjRyYTRnd1VyUGVXQy1UbEE%5FZT1yMUNHVlk&id=39A6D6B71F335E18%21sc5fde9e31e1d49aaab78bfe905b546d1&cid=39A6D6B71F335E18&sb=name&sd=1


# Load libraries

```{r}
library(tidyverse)
library(here)
library(readxl)
```

# Read data

```{r}
data_V2025.2 <- read_tsv(
  "C:/Data/MOTIVATE/MOTIVATE_data/Data/EVA and ReSurvey Data/2025 data - V2025.2/Core data/200_Motivate20250415_notJUICE_header_with_precise_coordinates.csv")
```

No parsing issues!

# Keep only the columns needed

```{r}
data_V2025.2 <- data_V2025.2 %>%
  select(PlotObservationID, PlotID, Country, `Date of recording`, 
         `Expert system`, `ReSurvey plot (Y/N)`, `For EVA (Y/N)`, 
         `ReSurvey project`, `ReSurvey site`, `ReSurvey plot`,
         `ReSurvey observation`, `Manipulate (y/n)`, `Type of manipulation`,
         `Location method`, RS_CODE, RS_DUPL, RS_PROJTYP, Dataset, Longitude, 
         Latitude, Lon_prec, Lat_prec)
```

# Filter only ReSurvey part

```{r}
data_V2025.2 %>% count(`ReSurvey plot (Y/N)`)
```

What about NAs?

```{r}
data_V2025.2 <- data_V2025.2 %>%
  dplyr::filter(`ReSurvey plot (Y/N)` == "Y") %>%
  select(-`ReSurvey plot (Y/N)`)
```

# Data manipulation

```{r}
data_V2025.2 <- data_V2025.2 %>%
  # Update coordinates
  mutate(Lon_updated = ifelse(is.na(Lon_prec), Longitude, Lon_prec),
        Lat_updated = ifelse(is.na(Lat_prec), Latitude, Lat_prec)) %>%
  # Remove old coordinate columns
  select(-Longitude, -Latitude, -Lon_prec, -Lat_prec) %>%
  # Remove 586 observations without coordinates
  dplyr::filter(!is.na(Lon_updated)) %>%
  # Translate codes from Location method to words
  # Update `Location method`if edit_loc_method = TRUE
  mutate(`Location method` = 
           str_replace_all(
             `Location method`,
             c("01" = "Permanently marked plot isolated (i.e. somewhere within the site)",
               "02" = "Marked plot in a grid (i.e. with regularly spaced neighbor plots)",
               "03" = "Location with differential GPS",
               "04" = "Location with GPS",
               "05" = "Location from accurate map",
               "06" = "Location from a description",
               "07" = "Other",
               "08" = "Marked plot in a transect")), `Location method`) %>%
  # Unify codes for ReSurvey project types
  mutate(
    RS_PROJTYP = str_replace(RS_PROJTYP, "^Resampling$", "resampling"), 
    RS_PROJTYP = str_replace(RS_PROJTYP, 
                             "^Permanent \\(man\\)$", "permanent (man)")
    ) %>%
  # Add columns date and year
  mutate(date = dmy(`Date of recording`), year = year(date))
```

### Clean info on Expert system column and separate it

```{r}
data_V2025.2 <- data_V2025.2 %>% 
  mutate(
    # Clean 'Expert System' column by removing "!" and replacing "~" with NA
    `Expert system` = case_when(
      `Expert system` == "~" ~ NA_character_,  # Replace "~" with NA
      TRUE ~ str_replace_all(`Expert system`, "!", "")  # Remove "!"
    )
  ) %>%
  # Separate the values in 'Expert System' into multiple columns
  separate(
    `Expert system`,
    into = c("EUNISa", "EUNISb", "EUNISc"), # At most, there are 3 diff values
    sep = ",",
    extra = "drop",  # Drop extra values if there are more than columns
    fill = "right",   # Fill missing values with NA for cases with fewer values
    remove = FALSE    # Keep the original 'Expert System' column
  ) %>%
  # Add columns for the different EUNIS levels
  mutate(
    # EUNISa levels
    EUNISa_1 = substr(EUNISa, 1, ifelse(str_starts(EUNISa, "MA"), 2, 1)),
    EUNISa_2 = ifelse(
      nchar(EUNISa) >= ifelse(str_starts(EUNISa, "MA"), 3, 2), 
      substr(EUNISa, 1, ifelse(str_starts(EUNISa, "MA"), 3, 2)),
      NA_character_
    ),
    EUNISa_3 = ifelse(
      nchar(EUNISa) >= ifelse(str_starts(EUNISa, "MA"), 4, 3), 
      substr(EUNISa, 1, ifelse(str_starts(EUNISa, "MA"), 4, 3)),
      NA_character_
      ),
    EUNISa_4 = ifelse(
      nchar(EUNISa) >= ifelse(str_starts(EUNISa, "MA"), 5, 4), 
      substr(EUNISa, 1, ifelse(str_starts(EUNISa, "MA"), 5, 4)),
      NA_character_
    ),
    # EUNISb levels
    EUNISb_1 = substr(EUNISb, 1, ifelse(str_starts(EUNISb, "MA"), 2, 1)),
    EUNISb_2 = ifelse(
      nchar(EUNISb) >= ifelse(str_starts(EUNISb, "MA"), 3, 2), 
      substr(EUNISb, 1, ifelse(str_starts(EUNISb, "MA"), 3, 2)),
      NA_character_
    ),
    EUNISb_3 = ifelse(
      nchar(EUNISb) >= ifelse(str_starts(EUNISb, "MA"), 4, 3), 
      substr(EUNISb, 1, ifelse(str_starts(EUNISb, "MA"), 4, 3)),
      NA_character_
    ),
    EUNISb_4 = ifelse(
      nchar(EUNISb) >= ifelse(str_starts(EUNISb, "MA"), 5, 4), 
      substr(EUNISb, 1, ifelse(str_starts(EUNISb, "MA"), 5, 4)),
      NA_character_
    ),
    # EUNISc levels
    EUNISc_1 = substr(EUNISc, 1, ifelse(str_starts(EUNISc, "MA"), 2, 1)),
    EUNISc_2 = ifelse(
      nchar(EUNISc) >= ifelse(str_starts(EUNISc, "MA"), 3, 2), 
      substr(EUNISc, 1, ifelse(str_starts(EUNISc, "MA"), 3, 2)),
      NA_character_
    ),
    EUNISc_3 = ifelse(
      nchar(EUNISc) >= ifelse(str_starts(EUNISc, "MA"), 4, 3), 
      substr(EUNISc, 1, ifelse(str_starts(EUNISc, "MA"), 4, 3)),
      NA_character_
    ),
    EUNISc_4 = ifelse(
      nchar(EUNISc) >= ifelse(str_starts(EUNISc, "MA"), 5, 4), 
      substr(EUNISc, 1, ifelse(str_starts(EUNISc, "MA"), 5, 4)),
      NA_character_
    )
  )
```

### Add descriptions

#### Level 1

```{r}
data_V2025.2 <- data_V2025.2 %>%
  mutate(
    EUNISa_1_descr = case_when(
      EUNISa_1 == "V" ~ "Vegetated man-made habitats",
      EUNISa_1 == "U" ~ "Inland habitats with no or little soil",
      EUNISa_1 == "T" ~ "Forests and other wooded land",
      EUNISa_1 == "S" ~ "Heathlands, scrub and tundra",
      EUNISa_1 == "R" ~ "Grasslands",
      EUNISa_1 == "Q" ~ "Wetlands",
      EUNISa_1 == "P" ~ "Inland waters",
      EUNISa_1 == "N" ~ "Coastal habitats",
      EUNISa_1 == "MA" ~ "Marine habitats",
      TRUE ~ NA_character_
    ),
    EUNISb_1_descr = case_when(
      EUNISb_1 == "V" ~ "Vegetated man-made habitats",
      EUNISb_1 == "U" ~ "Inland habitats with no or little soil",
      EUNISb_1 == "T" ~ "Forests and other wooded land",
      EUNISb_1 == "S" ~ "Heathlands, scrub and tundra",
      EUNISb_1 == "R" ~ "Grasslands",
      EUNISb_1 == "Q" ~ "Wetlands",
      EUNISb_1 == "P" ~ "Inland waters",
      EUNISb_1 == "N" ~ "Coastal habitats",
      EUNISb_1 == "MA" ~ "Marine habitats",
      TRUE ~ NA_character_
    ),
    EUNISc_1_descr = case_when(
      EUNISc_1 == "V" ~ "Vegetated man-made habitats",
      EUNISc_1 == "U" ~ "Inland habitats with no or little soil",
      EUNISc_1 == "T" ~ "Forests and other wooded land",
      EUNISc_1 == "S" ~ "Heathlands, scrub and tundra",
      EUNISc_1 == "R" ~ "Grasslands",
      EUNISc_1 == "Q" ~ "Wetlands",
      EUNISc_1 == "P" ~ "Inland waters",
      EUNISc_1 == "N" ~ "Coastal habitats",
      EUNISc_1 == "MA" ~ "Marine habitats",
      TRUE ~ NA_character_
    )
  )
```

#### Level 2

```{r}
descriptions<-read_excel(here("..", "DB_first_check", "data", "edited",
                                 "EUNIS-Habitats-2021-06-01_modified.xlsx"))
```

```{r}
# Define the columns and corresponding description column names
eunis_cols <- c("EUNISa_2", "EUNISa_3", "EUNISa_4",
                "EUNISb_2", "EUNISb_3", "EUNISb_4", 
                "EUNISc_2", "EUNISc_3", "EUNISc_4")

# Create corresponding description column names
descr_col_names <- paste0(eunis_cols, "_descr")

# Use reduce to loop through the columns and join dynamically based on level
data_V2025.2 <- reduce(seq_along(eunis_cols), function(data, i) {
  # Extract level number from the column name (e.g., EUNISa_2 -> 2)
  level <- as.numeric(gsub("\\D", "", eunis_cols[i]))
  
  # Filter descriptions for the corresponding level
  descriptions_level <- descriptions %>%
    dplyr::filter(level == level) %>%
    select(`EUNIS 2020 code`, `EUNIS-2021 habitat name`)
  
  # Perform the left_join and rename the column dynamically
  data %>%
    left_join(
      descriptions_level,
      by = setNames("EUNIS 2020 code", eunis_cols[i])
    ) %>%
    rename(!!descr_col_names[i] := `EUNIS-2021 habitat name`)
}, .init = data_V2025.2)
```

See where matching did not work:

```{r}
data_V2025.2 %>% 
  dplyr::filter(!is.na(EUNISa_2) & is.na(EUNISa_2_descr)) %>% count(EUNISa_2)
data_V2025.2 %>% 
  dplyr::filter(!is.na(EUNISa_3) & is.na(EUNISa_3_descr)) %>% count(EUNISa_3)
data_V2025.2 %>% 
  dplyr::filter(!is.na(EUNISa_4) & is.na(EUNISa_4_descr)) %>% count(EUNISa_4)
```

The matching did not work sometimes, correct!

```{r}
# Correct EUNISa levels 2-4 descriptions
data_V2025.2 <- data_V2025.2 %>%
  mutate(EUNISa_2_descr = 
           ifelse(!is.na(EUNISa_2_descr), EUNISa_2_descr,
                  case_when(
                    EUNISa_2 == "P2" ~ "Running aquatic habitats",
                    EUNISa_2 == "P3" ~ "Aquatic plant communities",
                    EUNISa_2 == "Pf" ~ "Fresh-water submerged vegetation",
                    EUNISa_2 == "Pj" ~ "Stonewort vegetation",
                    EUNISa_2 == "R4" ~ "Alpine and subalpine grasslands",
                    EUNISa_2 == "Pb" ~ "Calcareous spring and spring brook",
                    EUNISa_2 == "Qb" ~ "Wetlands",
                    EUNISa_2 == "R3" ~ "Seasonally wet and wet grasslands",
                    EUNISa_2 == "Qa" ~ "Mires",
                    EUNISa_2 == "Pa" ~ "Base-poor spring and spring brook",
                    EUNISa_2 == "Ph" ~ "Oligotrophic-water vegetation",
                    EUNISa_2 == "Pg" ~ "Fresh-water nymphaeid vegetation",
                    EUNISa_2 ==
                      "Pd" ~ "Fresh-water small pleustophyte vegetation",
                    EUNISa_2 == "Pc" ~ "Brackish-water vegetation",
                    EUNISa_2 ==
                      "Pe" ~ "Fresh-water large pleustophyte vegetation",
                    EUNISa_2 == "Pi" ~ "Dystrophic-water vegetation",
                    EUNISa_2 == "S1" ~ "Tundra",
                    EUNISa_2 ==
                      "U7" ~ "Unvegetated or sparsely vegetated gravel bars",
                    EUNISa_2 == "Q6" ~ "Periodically exposed shores",
                    EUNISa_2 == "R3" ~ "Seasonally wet and wet grasslands",
                    EUNISa_2 == "R4" ~ "Alpine and subalpine grasslands",
                    EUNISa_2 == "U7" ~ "Unvegetated or sparsely vegetated gravel bars",
                    # Sa & Sb, not sure what these are, left without descr
                    TRUE ~ NA_character_)
                  ),
         EUNISa_3_descr = 
           ifelse(!is.na(EUNISa_3_descr), EUNISa_3_descr,
                  case_when(
                    EUNISa_3 =="U71" ~ "Unvegetated or sparsely vegetated gravel bar in montane and alpine regions",
                    EUNISa_3 =="Q61" ~ "Periodically exposed shore with stable, eutrophic sediments with pioneer or ephemeral vegetation",
                    EUNISa_3 =="Q62" ~ "Periodically exposed shore with stable, mesotrophic sediments with pioneer or ephemeral vegetation",
                    EUNISa_3 == "P2N" ~ "Spring",
                    EUNISa_3 == "P3a" ~ "Brackish-water vegetation",
                    EUNISa_3 == "P3b" ~ "Fresh-water small pleustophyte vegetation",
                    EUNISa_3 == "P3c" ~ "Fresh-water large pleustophyte vegetation",
                    EUNISa_3 == "P3d" ~ "Fresh-water submerged vegetation",
                    EUNISa_3 == "P3e" ~ "Fresh-water nymphaeid vegetation",
                    EUNISa_3 == "P3f" ~ "Oligotrophic-water vegetation",
                    EUNISa_3 == "P3g" ~ "Dystrophic-water vegetation",
                    EUNISa_3 == "P3h" ~ "Stonewort vegetation",
                    TRUE ~ NA_character_
                    ))
         )
```

# Add EUNIS from info on HabitatID from DK 

Based on information got from Jesper.

### Read the data sent by Jesper from DK

```{r}
db_DK_J<-read_tsv(
  "C:/Data/MOTIVATE/DK_Naturdata_Res_habitat_hab_codes_Jesper/DK_Naturdata_Res_habitat_hab_codes.txt")
```

## Add info on HabitatID

```{r}
data_V2025.2 <- data_V2025.2 %>%
  # Keeping all obs in data_V2025.2 but not all in db_DK_J
  left_join(db_DK_J %>% select(PlotObservationID, HabitatID))
```

## Change some Annex I habitat codes that were wrong

```{r}
data_V2025.2 <- data_V2025.2 %>%
  mutate(HabitatID = as.character(HabitatID)) %>%
  mutate(HabitatID = ifelse(HabitatID == "9998", "91D0",
                            ifelse(HabitatID == "9999", "91E0", HabitatID)))
```

## Add info on correspondences HabitatID (DK, Jesper) - EUNIS

Read correspondences file:

```{r}
correspondences<-read_excel(here("..", "DB_first_check", "data", "edited",
                                 "correspondence_HabitatID_DK.xlsx"))
```

Add info to data_V2025.2:

```{r}
data_V2025.2 <- data_V2025.2 %>%
  # Keeping all obs in data_V2025.2 but not all in db_DK_J
  left_join(correspondences %>% select(HabitatID, EUNIS))
```

Correct NA values in EUNIS

```{r}
data_V2025.2 <- data_V2025.2 %>%
  mutate(EUNIS = ifelse(EUNIS == "NA", NA, EUNIS))
```

Add info on EUNIS (DK) to EUNISa:

```{r}
data_V2025.2 <- data_V2025.2 %>%
  mutate(EUNISa =
           # If EUNIS (DK) is available, add as EUNISa
           ifelse(!is.na(EUNIS), EUNIS, 
                  # Otherwise keep EUNISa
                  EUNISa),
         EUNIS_assignation = ifelse(!is.na(EUNIS), "Info from DK",
                                    ifelse(is.na(EUNISa), "Not possible",
                                           "Expert system"))) %>%
  # Remove unneeded columns
  select(-EUNIS, -HabitatID)
```

## Update columns for EUNIS levels and descriptions

Update the columns for the different EUNIS levels:

```{r}
data_V2025.2 <- data_V2025.2 %>%
  mutate(
    # EUNISa levels
    EUNISa_1 = substr(EUNISa, 1, ifelse(str_starts(EUNISa, "MA"), 2, 1)),
    EUNISa_2 = ifelse(
      nchar(EUNISa) >= ifelse(str_starts(EUNISa, "MA"), 3, 2), 
      substr(EUNISa, 1, ifelse(str_starts(EUNISa, "MA"), 3, 2)),
      NA_character_
    ),
    EUNISa_3 = ifelse(
      nchar(EUNISa) >= ifelse(str_starts(EUNISa, "MA"), 4, 3), 
      substr(EUNISa, 1, ifelse(str_starts(EUNISa, "MA"), 4, 3)),
      NA_character_
      ),
    EUNISa_4 = ifelse(
      nchar(EUNISa) >= ifelse(str_starts(EUNISa, "MA"), 5, 4), 
      substr(EUNISa, 1, ifelse(str_starts(EUNISa, "MA"), 5, 4)),
      NA_character_
    )
  ) 
```

Update columns with descriptions for the level 1 codes:

```{r}
data_V2025.2 <- data_V2025.2 %>%
  mutate(
    EUNISa_1_descr = case_when(
      EUNISa_1 == "V" ~ "Vegetated man-made habitats",
      EUNISa_1 == "U" ~ "Inland habitats with no or little soil",
      EUNISa_1 == "T" ~ "Forests and other wooded land",
      EUNISa_1 == "S" ~ "Heathlands, scrub and tundra",
      EUNISa_1 == "R" ~ "Grasslands",
      EUNISa_1 == "Q" ~ "Wetlands",
      EUNISa_1 == "P" ~ "Inland waters",
      EUNISa_1 == "N" ~ "Coastal habitats",
      EUNISa_1 == "MA" ~ "Marine habitats",
      TRUE ~ NA_character_
    )
  )
```

## Update number of different EUNIS codes

Recalculate how many different EUNIS codes have been assigned:

```{r}
data_V2025.2 <- data_V2025.2 %>%
  mutate(
    # Count the number of non-NA values across the EUNIS columns
    n_EUNIS = rowSums(!is.na(select(., EUNISa:EUNISc)))
  )
```

New plot for EUNISa_1 (the first assigned EUNIS in cases of multiple assignations, level 1):

```{r}
ggplot(data_V2025.2, aes(EUNISa_1_descr)) +
         geom_bar(aes(y = (..count..) / sum(..count..) * 100)) +
  labs(y = "Percentage of ReSurvey observations",
       x = "EUNIS level 1") + coord_flip()
ggplot(data_V2025.2 %>% filter(!is.na(EUNISa_1_descr)),
       aes(EUNISa_1_descr)) +
         geom_bar(aes(y = (..count..) / sum(..count..) * 100)) +
  labs(y = "Percentage of ReSurvey observations",
       x = "EUNIS level 1") + coord_flip()
```

# Notes EUNIS codes - to change?

This classification of Q + numbers is now coexisting in the database with Qa & Qb (metadata). How to proceed?

```{r}
data_V2025.2 %>% filter(EUNISa_1 == "Q") %>% distinct(EUNISa_2)
```

What are Sa and Sb in S?

# Save to clean datafiles

## All ReSurvey data

```{r}
write_csv(data_V2025.2,here("data", "clean","ReSurvey_data_V2025.2_20251003.csv"))
```

## Only Q, R, S & T

```{r}
write_csv(data_V2025.2 %>% 
            dplyr::filter(EUNISa_1 %in% c("Q", "R", "S", "T")),
          here("data", "clean","ReSurvey_QRST_data_V2025.2_20251003.csv"))
```

## Only Q, R, S & T & from 2016 onwards

```{r}
write_csv(data_V2025.2 %>% 
            dplyr::filter(EUNISa_1 %in% c("Q", "R", "S", "T")) %>%
            dplyr::filter(year >= 2016),
          here("data", "clean","ReSurvey_QRST_from2016_data_V2025.2_20251003.csv"))
```

# Session info

```{r}
sessionInfo()
```
