---
title: "Script to merge all RS data"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
---

This R script is used to merge RS data downloaded the Google Drive folder shared by Bea with the ReSurvey database.

# Load libraries

```{r}
library(tidyverse)
library(here)
```

# Read files with RS data from Bea

```{r}
# Set the folder path
folder_path <- here("data","raw","S2")

# List only CSV files that contain "max_Filtered" in their filename
csv_files <- list.files(folder_path, pattern = "max_Filtered.*\\.csv$",
                        full.names = TRUE, recursive = TRUE)

# Function to read each file and extract info from the filename
read_and_label <- function(file_path) {
  file_name <- basename(file_path)
  
  # Extract region and subregion from the filename
  # Updated regular expression to handle the case where subregion is missing
  components <- str_match(file_name, "(.*?)_(.*?)_Sentinel.*?_(.*?)\\.csv")
  
  # Extract the biogeo and unit, handling missing subregion (unit)
  biogeo <- components[2]
  
  # If subregion (unit) is missing, set it as NA
  unit <- ifelse(is.na(components[3]), NA, components[3])
  
  # Check if biogeo is missing, and if so,
  # assign the first part of the filename (region name)
  if (is.na(biogeo) && grepl("Sentinel", file_name)) {
    # Capture the first part (biogeo) directly
    biogeo <- str_match(file_name, "^(.*?)_Sentinel")[2]
  }
  
  # If biogeo is still NA, print a warning
  if (is.na(biogeo)) {
    warning(paste("Failed to extract biogeo for file:", file_name))
  }

  # Read CSV and add columns for extracted info
  read_csv(file_path) %>%
    mutate(biogeo = biogeo,
           unit = unit)
}
# Read and merge all CSV files
data_RS <- map_dfr(csv_files, read_and_label)

# View the resulting tibble
print(data_RS)
```

# Data cleaning RS

```{r}
data_RS <- data_RS %>%
  # Keep the columns we need
  select(obs_unique, biogeo, unit, year, source, Lat_update, Lon_update,
         NDVI, NDMI) %>%
  # Rename Lat and Lon, these are only kept in case there is differrence with
  # those in the ReSurvey database due to updates based on Ilona's info
  rename(Lat_RS = Lat_update, Lon_RS = Lon_update) %>%
  # Same for year
  rename(year_RS = year)
```

# Read file db_Europa

In this file, there is the correspondence obs_unique - PlotObservationID.

```{r}
db_Europa <- read_csv(
  here("..", "DB_first_check", "data", "clean","db_Europa_20250107.csv")
  )
```

Get only the columns PlotObservationID (original unique identifier) obs_unique_id (unique identified created by me) and year.

```{r}
db_Europa <- db_Europa %>% select(PlotObservationID, obs_unique_id)
```

# Merge data_RS and db_Europa

```{r}
data_RS_ID <- db_Europa %>%
  right_join(data_RS %>%
              # Rename to be able to join on this column
              rename(obs_unique_id = obs_unique))
```

Now we have PlotObservationID in data_RS_ID.

# Read file db_resurv_updated_clean

This is the ReSurvey database after updates (to be continued).

```{r}
db_resurv <- read_tsv(
  here("..", "DB_first_check","data", "clean","db_resurv_updated_clean.csv"),
  col_types = cols(
    # Dynamically specify EUNIS columns as character
    .default = col_guess(),  # Default guessing for other columns
    EUNISa = col_character(),
    EUNISb = col_character(),
    EUNISc = col_character(),
    EUNISd = col_character(),
    EUNISa_1 = col_character(),
    EUNISa_2 = col_character(),
    EUNISa_3 = col_character(),
    EUNISa_4 = col_character(),
    EUNISb_1 = col_character(),
    EUNISb_2 = col_character(),
    EUNISb_3 = col_character(),
    EUNISb_4 = col_character(),
    EUNISc_1 = col_character(),
    EUNISc_2 = col_character(),
    EUNISc_3 = col_character(),
    EUNISc_4 = col_character(),
    EUNISd_1 = col_character(),
    EUNISd_2 = col_character(),
    EUNISd_3 = col_character(),
    EUNISd_4 = col_character(),
    EUNISa_1_descr = col_character(),
    EUNISb_1_descr = col_character(),
    EUNISc_1_descr = col_character(),
    EUNISd_1_descr = col_character(),
    EUNIS_assignation = col_character(),
    EUNISa_2_descr = col_character(),
    EUNISa_3_descr = col_character(),
    EUNISa_4_descr = col_character(),
    EUNISb_2_descr = col_character(),
    EUNISb_3_descr = col_character(),
    EUNISb_4_descr = col_character(),
    EUNISc_2_descr = col_character(),
    EUNISc_3_descr = col_character(),
    EUNISc_4_descr = col_character(),
    EUNISd_2_descr = col_character(),
    EUNISd_3_descr = col_character(),
    EUNISd_4_descr = col_character()
    )
  )
```

No parsing issues!

# Merge RS data to the ReSurvey database

```{r}
db_resurv_RS <- db_resurv %>%
  left_join(data_RS_ID %>%
              # This column is not needed anymore
              select(-obs_unique_id)) %>%
  mutate(RS_data = !is.na(biogeo))
```

```{r}
db_resurv_RS %>% count(RS_data)
```

# Save to clean data

Save clean file for analyses (to be updated continuosly due to updates in ReSurvey database and updates on RS data).

```{r}
write_tsv(db_resurv_RS,here("data", "clean","db_resurv_RS.csv"))
```

# Session info

```{r}
sessionInfo()
```

