---
title: "Script to get validated points from the ReSurvey database using best RF models"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

This R script is used get validated points from the ReSurvey database using the best RF models.

Later, get validated points using each of the fitted models and see if / how they differ.

# Load libraries

```{r}
library(tidyverse)
library(here)
```

# Load functions

```{r}
knitr::purl("functions.Rmd", output = "functions.R")
source("functions.R")
```

# Load previously created objects

```{r}
# Define the folder path
folder_path <- here("objects", "validated_points")

# List all .RData or .rda files in the folder
rdata_files <- list.files(folder_path, full.names = TRUE)

# Load each file
lapply(rdata_files, load, envir = .GlobalEnv)
```

# Load best models

Model with SatEmb + CH for predicting where there is SatEmb data:

```{r}
load(here("objects", "RF_1500", "rf_SatEmb_CH_model.RData"))
```

```{r}
rf_SatEmb_CH <- rf_SatEmb_CH$model
```

Model with months, Landsat for predicting where there is no SatEmb data:

```{r}
load(here("objects", "RF_1500", "rf_months_Landsat_model.RData"))
```

```{r}
rf_months_Landsat <- rf_months_Landsat$model
```

# Read data

Read data validation SatEmb and Landsat:

```{r}
data_validation_SatEmb <- read_tsv(here(
  "data", "clean","final_RS_data_SatEmb_20251010.csv")) %>%
  rename(EUNISa_1_descr = EUNIS_1_d)
data_validation_Landsat <- read_tsv(here(
  "data", "clean","final_RS_data_bands_Landsat_all_20251021.csv"))
```

No parsing issues!

```{r}
nrow(data_validation_Landsat)
nrow(data_validation_SatEmb)
```

# Filter out manipulated observations

## Read list of PlotObservationIDs to keep

```{r}
PlotObservationID_list <- read_csv(
  here("data", "clean","PlotObservationIDs_filtered_QRST_20251211.csv"))
```

## Filter out data to keep only those PlotObservationIDs

```{r}
data_validation_SatEmb <- data_validation_SatEmb %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
data_validation_Landsat <- data_validation_Landsat %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
```

```{r}
nrow(data_validation_Landsat)
nrow(data_validation_SatEmb)
```

# Merge both datasets

```{r}
data_validation <- data_validation_Landsat %>% 
  full_join(data_validation_SatEmb) %>%
  # Add SatEmb T/F
  mutate(SatEmb = ifelse(is.na(A00), FALSE, TRUE))
```

```{r}
data_validation %>% count(SatEmb)
```

# Predictions from best models

Model with SatEmb + CH for predicting where there is SatEmb data:

```{r eval=FALSE, include=FALSE}
prediction_SatEmb_resp <- predict(rf_SatEmb_CH,
                             newdata = data_validation %>% filter(SatEmb == T), 
                             type = "response")
```

```{r eval=FALSE, include=FALSE}
prediction_SatEmb_prob <- predict(rf_SatEmb_CH,
                             newdata = data_validation %>% filter(SatEmb == T), 
                             type = "prob")
```

Model with months, Landsat for predicting where there is no SatEmb data:

```{r eval=FALSE, include=FALSE}
prediction_Landsat_resp <- predict(rf_months_Landsat,
                             newdata = data_validation, 
                             type = "response")
```

```{r eval=FALSE, include=FALSE}
prediction_Landsat_prob <- predict(rf_months_Landsat,
                             newdata = data_validation, 
                             type = "prob")
```

```{r}
predictions <- c("prediction_SatEmb_resp", "prediction_SatEmb_prob",
                 "prediction_Landsat_resp", "prediction_Landsat_prob")
```

Save:

```{r eval=FALSE, include=FALSE}
for (prediction_name in predictions) {
  file_path <- paste0("objects/validated_points/", prediction_name, ".Rdata")
  save(list = prediction_name, file = file_path)
}
```

Add predictions:

```{r}
data_validated <- data_validation %>%
  left_join(
    # Predictions from model with SatEmb + CH
    cbind(
      data.frame(
        # PlotObservationIDs
        PlotObservationID = data_validation %>% 
          filter(SatEmb == T) %>% pull(PlotObservationID),
        # Prediction response
        pred_SatEmb_resp = prediction_SatEmb_resp
        ),
      # Prediction probabilities
      do.call(rbind, prediction_SatEmb_prob) %>% as.data.frame() %>%
        rename(pred_SatEmb_Q = EUNISa_1.Q, pred_SatEmb_R = EUNISa_1.R,
               pred_SatEmb_S = EUNISa_1.S, pred_SatEmb_T = EUNISa_1.T)
      )
  ) %>%
  # ADD
  left_join(
    # Predictions from model with months, Landsat
    cbind(
      data.frame(
        # PlotObservationIDs - SEE IF SAME NR
        PlotObservationID = data_validation %>% pull(PlotObservationID),
        # Prediction response
        pred_Landsat_resp = prediction_Landsat_resp
        ),
      # Prediction probabilities
      do.call(rbind, prediction_Landsat_prob) %>% as.data.frame() %>%
        rename(pred_Landsat_Q = EUNISa_1.Q, pred_Landsat_R = EUNISa_1.R,
               pred_Landsat_S = EUNISa_1.S, pred_Landsat_T = EUNISa_1.T)
      )
  ) %>%
  # Validated or not (SatEmb)
  mutate(valid_SatEmb = ifelse(pred_SatEmb_resp == EUNISa_1, TRUE, FALSE)) %>%
  # Validated or not (Landsat)
  mutate(valid_Landsat = ifelse(pred_Landsat_resp == EUNISa_1, TRUE, FALSE))
```

# Counts

```{r}
data_validated %>% count(valid_SatEmb)
data_validated %>% count(valid_Landsat)
```

# Data cleaning

```{r}
data_validated_clean <- data_validated %>%
  # Keep only needed cols
  select(
    # These three cols are supposed to uniquely define each resurvey plot
    RS_CODE, RSrvyst, RSrvypl,
    # The three previous cols + this one are supposed to uniquely define
    # each resurvey observation
    RSrvyob,
    # Unique identifier for each ovservation
    PlotObservationID,
    Dtofrcr, year,
    biogeo, unit,
    Lctnmth,
    EUNISa_1, EUNISa_1_descr, EUNISa_2, EUNISa_2_descr,
    SatEmb, valid_SatEmb, valid_Landsat
    ) %>%
  # Rename some cols as in original db
  rename(`ReSurvey site` = RSrvyst, `ReSurvey plot` = RSrvypl, 
         `ReSurvey observation` = RSrvyob, `Date of recording` = Dtofrcr,
         `Location method` = Lctnmth ) %>%
  # Create new valid col
  mutate(valid = case_when(
    # If there is no SatEmb info, use Landsat validation
    SatEmb == F ~ valid_Landsat,
    # If there is SatEmb info, use SatEmb validation
    SatEmb == T ~ valid_SatEmb
  ))
```

How many NAs for each row?

```{r}
data_validated_clean %>% 
  summarise(across(everything(), ~ sum(is.na(.))))
```

# Get info from original db

## Read original db

```{r}
data_V2025.2 <- read_tsv(
  "C:/Data/MOTIVATE/MOTIVATE_data/Data/EVA and ReSurvey Data/2025 data - V2025.2/Core data/200_Motivate20250415_notJUICE_header_with_precise_coordinates.csv")
```

## Get info for cols where there were NAs

```{r}
data_validated_clean <- rows_patch( # Updates only the NAs
  data_validated_clean,
  data_V2025.2 %>%
    # Select cols where there were NAs
    select(RS_CODE, `ReSurvey site`, `ReSurvey plot`, `ReSurvey observation`, 
           PlotObservationID, `Date of recording`, `Location method`),
  by = "PlotObservationID",
  unmatched = "ignore" # Only patches matching rows and ignores the rest
  )
```

## Add coordinates

```{r}
data_validated_clean <- data_validated_clean %>%
  left_join(
    data_V2025.2 %>% 
      # Update coordinates
      mutate(Lon_updated = ifelse(is.na(Lon_prec), Longitude, Lon_prec),
             Lat_updated = ifelse(is.na(Lat_prec), Latitude, Lat_prec)) %>%
      # Select columns needed
      select(PlotObservationID, Lon_updated, Lat_updated)
    )
```

How many NAs for each row?

```{r}
data_validated_clean %>% 
  summarise(across(everything(), ~ sum(is.na(.))))
```

# More data cleaning

```{r}
data_validated_clean <- data_validated_clean %>%
  # Convert to date
  mutate(`Date of recording` = dmy(`Date of recording`))
```

# Obs with more than one row

```{r}
data_validated_clean %>%
  # These 4 cols uniquely identify each obs
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`, `ReSurvey observation`) %>%
  # Filter obs with more than 1 row
  filter(n() > 1) %>%
  arrange(RS_CODE, `ReSurvey site`, `ReSurvey plot`, `ReSurvey observation`)
```

For each unique combination of RS_CODE, ReSurvey site, ReSurvey plot and ReSurvey observation, check whether any of the other columns (except PlotObservationID) differ across the rows in that group:

```{r}
data_validated_clean %>%
  group_by(
    RS_CODE, `ReSurvey site`, `ReSurvey plot`, `ReSurvey observation`
  ) %>%
  filter(n() > 1) %>%
  group_modify(~ {
    
    # Exclude columns that should NOT be checked
    cols_to_check <- setdiff(names(.x), "PlotObservationID")
    
    # Identify columns where at least one value differs
    diffs <- map_lgl(.x[cols_to_check], ~ n_distinct(.x, na.rm = FALSE) > 1)
    
    tibble(
      differing_columns = paste(names(diffs)[diffs], collapse = ", ")
      )
    }) %>%
  ungroup() %>%
  # Keep only groups where at least one difference exists
  filter(differing_columns != "")
```

# Add info on nr resurveys (n different years)

Some have only 1 year, either because the first year(s) were prior to Landsat, or because they were still not resurveyed. 

```{r}
data_validated_clean <- data_validated_clean %>%
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`) %>%
  mutate(
    n_resurveys = n_distinct(year, na.rm = TRUE),   # count unique years
    first_year  = min(year, na.rm = TRUE),          # earliest year
    last_year   = max(year, na.rm = TRUE)           # latest year
  ) %>%
  ungroup()
```

# Get only latest obs for each plot

```{r}
data_validated_clean_latest <- data_validated_clean %>%
  # Keep only rows from the last year
  filter(year == last_year) %>%
  # These 3 cols are supposed to uniquely identify each plot
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`) %>%
  # If more than one row in the same year, keep the latest date
  #slice_max(`Date of recording`, n = 1, with_ties = FALSE) %>%
  ungroup()
```

## Obs with more than one row

```{r}
data_validated_clean_latest %>%
  # These 3 cols are supposed to uniquely identify each plot
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`) %>%
  # Filter obs with more than 1 row
  filter(n() > 1) %>%
  ungroup() %>%
  arrange(RS_CODE, `ReSurvey site`, `ReSurvey plot`)
```

# Write csvs

```{r}
write_csv(data_validated_clean,
          here("data", "clean","data_validated_clean_20260120.csv"))
write_csv(data_validated_clean_latest,
          here("data", "clean","data_validated_clean_latest_20260120.csv"))
```

# Session info

```{r}
sessionInfo()
```

