---
title: "Script to get validated points from the ReSurvey database using best RF models"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

This R script is used get validated points from the ReSurvey database using the best RF models.

Later, get validated points using each of the fitted models and see if / how they differ.

# Load libraries

```{r}
library(tidyverse)
library(here)
library(rnaturalearth)
library(sf)
library(patchwork)
library(scales)
```

# Load functions

```{r}
knitr::purl("functions_20260122.Rmd", output = "functions_20260122.R")
source("functions_20260122.R")
```

# Load previously created objects

```{r}
# Define the folder path
folder_path <- here("objects", "validated_points")

# List all .RData or .rda files in the folder
rdata_files <- list.files(folder_path, full.names = TRUE)

# Load each file
lapply(rdata_files, load, envir = .GlobalEnv)
```

# Load best models

Model with SatEmb + CH for predicting where there is SatEmb data:

```{r eval=FALSE, include=FALSE}
load(here("objects", "RF_1500", "rf_SatEmb_CH_model.RData"))
```

```{r eval=FALSE, include=FALSE}
rf_SatEmb_CH <- rf_SatEmb_CH$model
```

Model with months, Landsat for predicting where there is no SatEmb data:

```{r eval=FALSE, include=FALSE}
load(here("objects", "RF_1500", "rf_months_Landsat_model.RData"))
```

```{r eval=FALSE, include=FALSE}
rf_months_Landsat <- rf_months_Landsat$model
```

# Read data

Read data validation SatEmb and Landsat:

```{r}
data_validation_SatEmb <- read_tsv(here(
  "data", "clean","final_RS_data_SatEmb_20251010.csv")) %>%
  rename(EUNISa_1_descr = EUNIS_1_d)
data_validation_Landsat <- read_tsv(here(
  "data", "clean","final_RS_data_bands_Landsat_all_20251021.csv"))
```

No parsing issues!

```{r}
nrow(data_validation_Landsat)
nrow(data_validation_SatEmb)
```

Read original db:

```{r}
data_V2025.2 <- read_tsv(
  "C:/Data/MOTIVATE/MOTIVATE_data/Data/EVA and ReSurvey Data/2025 data - V2025.2/Core data/200_Motivate20250415_notJUICE_header_with_precise_coordinates.csv")
```

# Filter out manipulated observations

## Read list of PlotObservationIDs to keep

```{r}
PlotObservationID_list <- read_csv(
  here("data", "clean","PlotObservationIDs_filtered_QRST_20251211.csv"))
```

## Filter out data to keep only those PlotObservationIDs

```{r}
data_validation_SatEmb <- data_validation_SatEmb %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
data_validation_Landsat <- data_validation_Landsat %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
```

```{r}
nrow(data_validation_Landsat)
nrow(data_validation_SatEmb)
```

# Merge both datasets

```{r}
data_validation <- data_validation_Landsat %>% 
  full_join(data_validation_SatEmb) %>%
  # Add SatEmb T/F
  mutate(SatEmb = ifelse(is.na(A00), FALSE, TRUE))
```

```{r}
data_validation %>% count(SatEmb)
```

# Predictions from best models

Model with SatEmb + CH for predicting where there is SatEmb data:

```{r eval=FALSE, include=FALSE}
prediction_SatEmb_resp <- predict(rf_SatEmb_CH,
                             newdata = data_validation %>% filter(SatEmb == T), 
                             type = "response")
```

```{r eval=FALSE, include=FALSE}
prediction_SatEmb_prob <- predict(rf_SatEmb_CH,
                             newdata = data_validation %>% filter(SatEmb == T), 
                             type = "prob")
```

Model with months, Landsat for predicting where there is no SatEmb data:

```{r eval=FALSE, include=FALSE}
prediction_Landsat_resp <- predict(rf_months_Landsat,
                             newdata = data_validation, 
                             type = "response")
```

```{r eval=FALSE, include=FALSE}
prediction_Landsat_prob <- predict(rf_months_Landsat,
                             newdata = data_validation, 
                             type = "prob")
```

```{r}
predictions <- c("prediction_SatEmb_resp", "prediction_SatEmb_prob",
                 "prediction_Landsat_resp", "prediction_Landsat_prob")
```

Save:

```{r eval=FALSE, include=FALSE}
for (prediction_name in predictions) {
  file_path <- paste0("objects/validated_points/", prediction_name, ".Rdata")
  save(list = prediction_name, file = file_path)
}
```

Add predictions:

```{r}
data_validated <- data_validation %>%
  left_join(
    # Predictions from model with SatEmb + CH
    cbind(
      data.frame(
        # PlotObservationIDs
        PlotObservationID = data_validation %>% 
          filter(SatEmb == T) %>% pull(PlotObservationID),
        # Prediction response
        pred_SatEmb_resp = prediction_SatEmb_resp
        ),
      # Prediction probabilities
      do.call(rbind, prediction_SatEmb_prob) %>% as.data.frame() %>%
        rename(pred_SatEmb_Q = EUNISa_1.Q, pred_SatEmb_R = EUNISa_1.R,
               pred_SatEmb_S = EUNISa_1.S, pred_SatEmb_T = EUNISa_1.T)
      )
  ) %>%
  # ADD
  left_join(
    # Predictions from model with months, Landsat
    cbind(
      data.frame(
        # PlotObservationIDs - SEE IF SAME NR
        PlotObservationID = data_validation %>% pull(PlotObservationID),
        # Prediction response
        pred_Landsat_resp = prediction_Landsat_resp
        ),
      # Prediction probabilities
      do.call(rbind, prediction_Landsat_prob) %>% as.data.frame() %>%
        rename(pred_Landsat_Q = EUNISa_1.Q, pred_Landsat_R = EUNISa_1.R,
               pred_Landsat_S = EUNISa_1.S, pred_Landsat_T = EUNISa_1.T)
      )
  ) %>%
  # Validated or not (SatEmb)
  mutate(valid_SatEmb = ifelse(pred_SatEmb_resp == EUNISa_1, TRUE, FALSE)) %>%
  # Validated or not (Landsat)
  mutate(valid_Landsat = ifelse(pred_Landsat_resp == EUNISa_1, TRUE, FALSE))
```

# Counts

```{r}
data_validated %>% count(valid_SatEmb)
data_validated %>% count(valid_Landsat)
```

# Data cleaning

```{r}
data_validated_clean <- data_validated %>%
  # Keep only needed cols
  select(
    # These three cols are supposed to uniquely define each resurvey plot
    RS_CODE, RSrvyst, RSrvypl,
    # The three previous cols + this one are supposed to uniquely define
    # each resurvey observation
    RSrvyob,
    # Unique identifier for each ovservation
    PlotObservationID,
    Dtofrcr, year,
    biogeo, unit,
    Lctnmth,
    EUNISa_1, EUNISa_1_descr, EUNISa_2, EUNISa_2_descr,
    SatEmb, 
    pred_SatEmb_resp, pred_SatEmb_Q, pred_SatEmb_R, pred_SatEmb_S, pred_SatEmb_T,
    pred_Landsat_resp, pred_Landsat_Q, pred_Landsat_R, pred_Landsat_S, pred_Landsat_T,
    valid_SatEmb, valid_Landsat, 
    ) %>%
  # Rename some cols as in original db
  rename(`ReSurvey site` = RSrvyst, `ReSurvey plot` = RSrvypl, 
         `ReSurvey observation` = RSrvyob, `Date of recording` = Dtofrcr,
         `Location method` = Lctnmth ) %>%
  # Create new valid col
  mutate(valid = case_when(
    # If there is no SatEmb info, use Landsat validation
    SatEmb == F ~ valid_Landsat,
    # If there is SatEmb info, use SatEmb validation
    SatEmb == T ~ valid_SatEmb
    )) %>%
  # Create new cols with predictions for each habitat type
  mutate(
    pred_Q = case_when(
      # If there is no SatEmb info, use Landsat prediction
      SatEmb == F ~ pred_Landsat_Q,
      # If there is SatEmb info, use SatEmb prediction
      SatEmb == T ~ pred_SatEmb_Q),
    pred_R = case_when(
      # If there is no SatEmb info, use Landsat prediction
      SatEmb == F ~ pred_Landsat_R,
      # If there is SatEmb info, use SatEmb prediction
      SatEmb == T ~ pred_SatEmb_R),
    pred_S = case_when(
      # If there is no SatEmb info, use Landsat prediction
      SatEmb == F ~ pred_Landsat_S,
      # If there is SatEmb info, use SatEmb prediction
      SatEmb == T ~ pred_SatEmb_S),
    pred_T = case_when(
      # If there is no SatEmb info, use Landsat prediction
      SatEmb == F ~ pred_Landsat_T,
      # If there is SatEmb info, use SatEmb prediction
      SatEmb == T ~ pred_SatEmb_T)
    )
```

```{r}
data_validated_clean %>% count(valid)
```


How many NAs for each row?

```{r}
data_validated_clean %>% 
  summarise(across(everything(), ~ sum(is.na(.))))
```

# Get info from original db

## Read original db

```{r}
data_V2025.2 <- read_tsv(
  "C:/Data/MOTIVATE/MOTIVATE_data/Data/EVA and ReSurvey Data/2025 data - V2025.2/Core data/200_Motivate20250415_notJUICE_header_with_precise_coordinates.csv")
```

## Get info for cols where there were NAs

```{r}
data_validated_clean <- rows_patch( # Updates only the NAs
  data_validated_clean,
  data_V2025.2 %>%
    # Select cols where there were NAs
    select(RS_CODE, `ReSurvey site`, `ReSurvey plot`, `ReSurvey observation`, 
           PlotObservationID, `Date of recording`, `Location method`),
  by = "PlotObservationID",
  unmatched = "ignore" # Only patches matching rows and ignores the rest
  )
```

## Add coordinates

```{r}
data_validated_clean <- data_validated_clean %>%
  left_join(
    data_V2025.2 %>% 
      # Update coordinates
      mutate(Lon_updated = ifelse(is.na(Lon_prec), Longitude, Lon_prec),
             Lat_updated = ifelse(is.na(Lat_prec), Latitude, Lat_prec)) %>%
      # Select columns needed
      select(PlotObservationID, Lon_updated, Lat_updated)
    )
```

How many NAs for each row?

```{r}
data_validated_clean %>% 
  summarise(across(everything(), ~ sum(is.na(.))))
```

# More data cleaning

```{r}
data_validated_clean <- data_validated_clean %>%
  # Convert to date
  mutate(`Date of recording` = dmy(`Date of recording`))
```

# Obs with more than one row

```{r}
data_validated_clean %>%
  # These 4 cols uniquely identify each obs
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`, `ReSurvey observation`) %>%
  # Filter obs with more than 1 row
  filter(n() > 1) %>%
  arrange(RS_CODE, `ReSurvey site`, `ReSurvey plot`, `ReSurvey observation`)
```

For each unique combination of RS_CODE, ReSurvey site, ReSurvey plot and ReSurvey observation, check whether any of the other columns (except PlotObservationID) differ across the rows in that group:

```{r}
data_validated_clean %>%
  group_by(
    RS_CODE, `ReSurvey site`, `ReSurvey plot`, `ReSurvey observation`
  ) %>%
  filter(n() > 1) %>%
  group_modify(~ {
    
    # Exclude columns that should NOT be checked
    cols_to_check <- setdiff(names(.x), "PlotObservationID")
    
    # Identify columns where at least one value differs
    diffs <- map_lgl(.x[cols_to_check], ~ n_distinct(.x, na.rm = FALSE) > 1)
    
    tibble(
      differing_columns = paste(names(diffs)[diffs], collapse = ", ")
      )
    }) %>%
  ungroup() %>%
  # Keep only groups where at least one difference exists
  filter(differing_columns != "")
```

# Add info on nr resurveys (n different years)

Some have only 1 year, either because the first year(s) were prior to Landsat, or because they were still not resurveyed. 

```{r}
data_validated_clean <- data_validated_clean %>%
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`) %>%
  mutate(
    n_resurveys = n_distinct(year, na.rm = TRUE),   # count unique years
    first_year  = min(year, na.rm = TRUE),          # earliest year
    last_year   = max(year, na.rm = TRUE)           # latest year
  ) %>%
  ungroup()
```

# Get only latest obs for each plot

```{r}
data_validated_clean_latest <- data_validated_clean %>%
  # Keep only rows from the last year
  filter(year == last_year) %>%
  # These 3 cols are supposed to uniquely identify each plot
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`) %>%
  # If more than one row in the same year, keep the latest date
  #slice_max(`Date of recording`, n = 1, with_ties = FALSE) %>%
  ungroup()
```

## Obs with more than one row

```{r}
data_validated_clean_latest %>%
  # These 3 cols are supposed to uniquely identify each plot
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`) %>%
  # Filter obs with more than 1 row
  filter(n() > 1) %>%
  ungroup() %>%
  arrange(RS_CODE, `ReSurvey site`, `ReSurvey plot`)
```

# Write csvs

```{r}
write_csv(data_validated_clean,
          here("data", "clean","data_validated_clean_20260120.csv"))
write_csv(data_validated_clean_latest,
          here("data", "clean","data_validated_clean_latest_20260120.csv"))
```

# Plots validated / not-validated

## Plot

```{r}
plot1 <- data_validated_clean %>%
  count(EUNISa_1, valid, name = "n") %>% 
  ggplot(aes(x = EUNISa_1, y = n, fill = valid)) +
  geom_col(position = position_stack()) +
  xlab(NULL) + ylab("Number of observations") +
  scale_y_continuous(labels = label_number(scale = 1e-3, suffix = "k")) +
  scale_fill_manual(values = c("FALSE" = "#D55E00", "TRUE" = "#0072B2"),
                     name = "Validated") +
  theme_minimal() + theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot1
```

## Map

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
world_3035 <- st_transform(world, 3035)
points_3035 <- st_as_sf(
  data_validated_clean,
  coords = c("Lon_updated", "Lat_updated"),
  crs = 4326
) |>
  st_transform(3035)

europe_bbox_3035 <- st_bbox(
  c(
    xmin = 2500000,
    ymin = 1400000,
    xmax = 6500000,
    ymax = 5500000
  ),
  crs = st_crs(3035)
)
world_3035_crop <- st_crop(world_3035, europe_bbox_3035)
points_3035_crop <- st_crop(points_3035, europe_bbox_3035)
```

```{r}
map1 <- ggplot() +
  geom_sf(data = world_3035_crop, fill = "lightgrey", color = "gray") +
  geom_sf(
    data = points_3035_crop,
    aes(color = valid),
    size = 0.1
  ) +
  facet_wrap(~ EUNISa_1) +
  scale_color_manual(values = c("FALSE" = "#D55E00", "TRUE"  = "#0072B2"),
                     name = "Validated") +
  theme_minimal() +
  theme(legend.position = "none")
map1
```

## Combined

```{r}
map1 + plot1  +
  plot_layout(widths = c(4, 1)) +
  plot_annotation(tag_levels = "A", tag_suffix = ")") +
  theme(plot.margin = margin(1, 1, 1, 1))
```

# Plots GPS observations used for models

```{r}
# Excluding EVI because many values outside of range -1, 1
vars_RF_months_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, and SAVI between march and oct
  "NDVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI and SAVI at march, pos and oct
  "NDVI_avg_value_03", "NDVI_pos_value", "NDVI_avg_value_10",
  "SAVI_avg_value_03", "SAVI_pos_value", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "SAVI_diff_pos_march_value",
  "NDVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "SAVI_diff_pos_oct_value",
  "NDVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy",
  # Canopy height
  "canopy_height")
```

```{r}
filter_data_Landsat <- function(data, variables) {
  data %>%
    mutate(EUNISa_1 = as.factor(EUNISa_1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    # With Landsat data, many observations removed because of high/low EVI values
    # In Landsat_extraction, values of EVI outside of expected range were set as NA,
    # so better not to use EVI here
    # dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Select only variables needed
    dplyr::select(PlotObservationID, EUNISa_1, all_of(variables), Lctnmth,
           valid_1) %>%
    # Keep only rows with differences > 0 in variables
    filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0))
}
```

```{r}
data_models <- filter_data_Landsat(
  data_validation_Landsat %>%
    rough_validation_Landsat(), # Rough validation based on NDWI
  vars_RF_months_Landsat) %>%
  left_join(data_V2025.2 %>%
              # Update coordinates
              mutate(Lon_updated = ifelse(is.na(Lon_prec), Longitude, Lon_prec),
                     Lat_updated = ifelse(is.na(Lat_prec), Latitude, Lat_prec)) %>%
              select(PlotObservationID, `Location method`, Lon_updated, Lat_updated)) %>%
  mutate(GPS = case_when(`Location method` %in% 
                           c("Location with GPS", 
                             "Location with differential GPS") ~ TRUE,
                         !`Location method` %in% 
                           c("Location with GPS", 
                             "Location with differential GPS") ~ FALSE,
                         is.na(`Location method`) ~ FALSE))
```

```{r}
points_3035 <- st_as_sf(data_models,
                        coords = c("Lon_updated", "Lat_updated"),
                        crs = 4326
                        ) |>
  st_transform(3035)

points_3035_crop <- st_crop(points_3035, europe_bbox_3035)
```

```{r}
map2 <- ggplot() +
  geom_sf(data = world_3035_crop, fill = "lightgrey", color = "gray") +
  geom_sf(
    data = points_3035_crop,
    aes(color = GPS),
    size = 0.1
  ) +
  facet_wrap(~ EUNISa_1) +
  scale_color_manual(values = c("FALSE" = "#D55E00", "TRUE"  = "#0072B2"),
                     name = "Recorded with GPS") +
  theme_minimal() + theme(legend.position = "none")
map2
```

# Session info

```{r}
sessionInfo()
```

