---
title: "Script to validate points in ReSurvey database using RS data (S2 + SatEmb + Landsat)"
subtitle: "Fitting Random Forest models with ntree = 1500, excluding manipulated observations"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

This R script is used to validate the points in the ReSurvey database using either RS indicators from S2 or Landsat (indices + phenology + canopy height), or bands from the Google Satellite Embedding dataset: https://developers.google.com/earth-engine/datasets/catalog/GOOGLE_SATELLITE_EMBEDDING_V1_ANNUAL#description
Here, we filter out manipulated observations to avoid using them in analyses.

# Load libraries

```{r}
library(tidyverse)
library(here)
library(parallel)
library(doParallel)
library(moreparty)
library(caret)
library(randomForest)
library(pROC)
library(rlang)
library(stringr)
library(beepr)
library(permimp)
library(purrr)
```

# Load functions

```{r}
knitr::purl("functions.Rmd", output = "functions.R")
source("functions.R")
```

# Load previously created objects

```{r}
# Define the folder path
folder_path <- here("objects", "RF_1500")

# List all .RData or .rda files in the folder
rdata_files <- list.files(folder_path, full.names = TRUE)

# Load each file
lapply(rdata_files, load, envir = .GlobalEnv)
```

# Read data

```{r}
data_validation_S2 <- read_tsv(here(
  "data", "clean","final_RS_data_bands_S2_all_20251021.csv"))
data_validation_SatEmb <- read_tsv(here(
  "data", "clean","final_RS_data_SatEmb_20251010.csv")) %>%
  rename(EUNISa_1_descr = EUNIS_1_d)
data_validation_Landsat <- read_tsv(here(
  "data", "clean","final_RS_data_bands_Landsat_all_20251021.csv"))
```

No parsing issues!

```{r}
nrow(data_validation_S2)
nrow(data_validation_Landsat)
nrow(data_validation_SatEmb)
```

# Filter out manipulated observations

## Read list of PlotObservationIDs to keep

```{r}
PlotObservationID_list <- read_csv(
  here("data", "clean","PlotObservationIDs_filtered_QRST_20251211.csv"))
```

## Filter out data to keep only those PlotObservationIDs

```{r}
data_validation_S2 <- data_validation_S2 %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
data_validation_SatEmb <- data_validation_SatEmb %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
data_validation_Landsat <- data_validation_Landsat %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
```

```{r}
nrow(data_validation_S2)
nrow(data_validation_Landsat)
nrow(data_validation_SatEmb)
```

# Descriptives

Number of observations with S2 data:

```{r}
data_validation_S2 %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  nrow()
```

Number of observations with Landsat data:

```{r}
data_validation_Landsat %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  nrow()
```

Number of observations with SatEmb data:

```{r}
data_validation_SatEmb %>%
  filter(if_any(c(2:65), ~ !is.na(.))) %>%
  nrow()
```

# Define lists of predictor vars

## S2 and Landsat

### Moments with slope method

Not using threshold method cause it gave worse results earlier.

```{r}
vars_RF_slope <- c(
  # Min values of all indices
  "NDVI_min", "EVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, EVI and SAVI
  "NDVI_auc_slope", "EVI_auc_slope", "SAVI_auc_slope",
  # Values of NDVI, EVI and SAVI at sos, pos and eos
  "NDVI_sos_slope_value", "NDVI_pos_value", "NDVI_eos_slope_value",
  "EVI_sos_slope_value", "EVI_pos_value", "EVI_eos_slope_value", 
  "SAVI_sos_slope_value", "SAVI_pos_value", "SAVI_eos_slope_value",
  # Differences pos-sos in value and doy
  "NDVI_diff_pos_sos_slope_value", "EVI_diff_pos_sos_slope_value",
  "SAVI_diff_pos_sos_slope_value","NDVI_diff_pos_sos_slope_doy", 
  "EVI_diff_pos_sos_slope_doy", "SAVI_diff_pos_sos_slope_doy",
  # Differences pos-eos in value and doy
  "NDVI_diff_pos_eos_slope_value", "EVI_diff_pos_eos_slope_value",
  "SAVI_diff_pos_eos_slope_value", "NDVI_diff_pos_eos_slope_doy", 
  "EVI_diff_pos_eos_slope_doy", "SAVI_diff_pos_eos_slope_doy",
  # Growing season duration
  "NDVI_slope_gsd", "EVI_slope_gsd", "SAVI_slope_gsd",
  # Canopy height
  "canopy_height")
```

```{r}
# Excluding EVI because many values outside of range -1, 1
vars_RF_slope_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI and SAVI
  "NDVI_auc_slope", "SAVI_auc_slope",
  # Values of NDVI and SAVI at sos, pos and eos
  "NDVI_sos_slope_value", "NDVI_pos_value", "NDVI_eos_slope_value",
  "SAVI_sos_slope_value", "SAVI_pos_value", "SAVI_eos_slope_value",
  # Differences pos-sos in value and doy
  "NDVI_diff_pos_sos_slope_value","SAVI_diff_pos_sos_slope_value",
  "NDVI_diff_pos_sos_slope_doy", "SAVI_diff_pos_sos_slope_doy",
  # Differences pos-eos in value and doy
  "NDVI_diff_pos_eos_slope_value", "SAVI_diff_pos_eos_slope_value",
  "NDVI_diff_pos_eos_slope_doy", "SAVI_diff_pos_eos_slope_doy",
  # Growing season duration
  "NDVI_slope_gsd", "SAVI_slope_gsd",
  # Canopy height
  "canopy_height")
```

### Months method

```{r}
vars_RF_months <- c(
  # Min values of all indices
  "NDVI_min", "EVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, EVI and SAVI between march and oct
  "NDVI_auc_mar_oct", "EVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI, EVI and SAVI at march, pos and oct
  "NDVI_avg_value_03", "NDVI_pos_value", "NDVI_avg_value_10",
  "EVI_avg_value_03", "EVI_pos_value", "EVI_avg_value_10", 
  "SAVI_avg_value_03", "SAVI_pos_value", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "EVI_diff_pos_march_value",
  "SAVI_diff_pos_march_value","NDVI_diff_pos_march_doy", 
  "EVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "EVI_diff_pos_oct_value",
  "SAVI_diff_pos_oct_value", "NDVI_diff_pos_oct_doy", 
  "EVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy",
  # Canopy height
  "canopy_height")
```

```{r}
# Excluding EVI because many values outside of range -1, 1
vars_RF_months_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, and SAVI between march and oct
  "NDVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI and SAVI at march, pos and oct
  "NDVI_avg_value_03", "NDVI_pos_value", "NDVI_avg_value_10",
  "SAVI_avg_value_03", "SAVI_pos_value", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "SAVI_diff_pos_march_value",
  "NDVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "SAVI_diff_pos_oct_value",
  "NDVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy",
  # Canopy height
  "canopy_height")
```

## SatEmb

```{r}
vars_RF_SatEmb <- colnames(
  data_validation_SatEmb)[startsWith(colnames(data_validation_SatEmb), "A")]
vars_RF_SatEmb_CH <- c(vars_RF_SatEmb, "canopy_height")
```

# Filter data to use in RF models

## Functions

```{r}
filter_data_S2 <- function(data, variables) {
  data %>%
    mutate(EUNISa_1 = as.factor(EUNISa_1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Select only variables needed
    select(PlotObservationID, EUNISa_1, all_of(variables), Lctnmth,
           valid_1) %>%
    # Keep only rows with differences > 0 in variables
    filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0)) %>%
    # Select only GPS points
    dplyr::filter(Lctnmth == "Location with GPS" | 
                    Lctnmth == "Location with differential GPS")
}
```

```{r}
filter_data_Landsat <- function(data, variables) {
  data %>%
    mutate(EUNISa_1 = as.factor(EUNISa_1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    # With Landsat data, many observations removed because of high/low EVI values
    # In Landsat_extraction, values of EVI outside of expected range were set as NA,
    # so better not to use EVI here
    # dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Select only variables needed
    select(PlotObservationID, EUNISa_1, all_of(variables), Lctnmth,
           valid_1) %>%
    # Keep only rows with differences > 0 in variables
    filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0)) %>%
    # Select only GPS points
    dplyr::filter(Lctnmth == "Location with GPS" | 
                    Lctnmth == "Location with differential GPS")
}
```

```{r}
filter_data_SatEmb <- function(data, variables) {
  data %>%
    mutate(EUNISa_1 = as.factor(EUNISa_1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Select only variables needed
    select(PlotObservationID, EUNISa_1, all_of(variables), 
           Lctnmth, valid_1) %>%
    # Select only GPS points
    dplyr::filter(Lctnmth == "Location with GPS" |
                    Lctnmth == "Location with differential GPS")
  }
```

## Filtered data

Select only GPS points that have passed the rough validation in each case.

For S2 and Landsat, use only slope and months methods, as threshold method has many missing values.

```{r}
# S2
filtered_data_slope_S2 <- filter_data_S2(
  data_validation_S2 %>%
    rough_validation_S2(), # Rough validation based on CH and NDWI
  vars_RF_slope)
filtered_data_months_S2 <- filter_data_S2(
  data_validation_S2 %>%
    rough_validation_S2(), # Rough validation based on CH and NDWI
  vars_RF_months)

# Landsat
filtered_data_slope_Landsat <- filter_data_Landsat(
  data_validation_Landsat %>%
    rough_validation_Landsat(), # Rough validation based on NDWI
  vars_RF_slope_Landsat)
filtered_data_months_Landsat <- filter_data_Landsat(
  data_validation_Landsat %>%
    rough_validation_Landsat(), # Rough validation based on NDWI
  vars_RF_months_Landsat)

# SatEmb
filtered_data_SatEmb <- filter_data_SatEmb(
  data_validation_SatEmb %>%
    rough_validation_SatEmb(), # Rough validation based on CH
  vars_RF_SatEmb)
filtered_data_SatEmb_CH <- filter_data_SatEmb(
  data_validation_SatEmb %>%
    rough_validation_SatEmb(), # Rough validation based on CH
  vars_RF_SatEmb_CH)
```

# Fit RF models (level 1)

## Without refinement

### N points per category

```{r}
bind_rows(
  # S2
  filtered_data_slope_S2 %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "S2", method = "slope") %>% 
    select(sensor, method, Q, R, S, T),
  filtered_data_months_S2 %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "S2", method = "months") %>% 
    select(sensor, method, Q, R, S, T),
  # # Landsat
  filtered_data_slope_Landsat %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>%
    mutate(sensor = "Landsat", method = "slope") %>% 
    select(sensor, method, Q, R, S, T),
  filtered_data_months_Landsat %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>%
    mutate(sensor = "Landsat", method = "months") %>% 
    select(sensor, method, Q, R, S, T),
  # SatEmb
  filtered_data_SatEmb %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "SatEmb", method ="SatEmb only") %>% 
    select(sensor, method, Q, R, S, T),
  filtered_data_SatEmb_CH %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "SatEmb", method ="SatEmb + CH") %>% 
    select(sensor, method, Q, R, S, T)
)
```

```{r}
bind_rows(
  # S2
  filtered_data_slope_S2 %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "S2", method = "slope") %>% 
    select(sensor, method, Q, R, S, T),
  filtered_data_months_S2 %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "S2", method = "months") %>% 
    select(sensor, method, Q, R, S, T),
  # # Landsat
  filtered_data_slope_Landsat %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>%
    mutate(sensor = "Landsat", method = "slope") %>% 
    select(sensor, method, Q, R, S, T),
  filtered_data_months_Landsat %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>%
    mutate(sensor = "Landsat", method = "months") %>% 
    select(sensor, method, Q, R, S, T),
  # SatEmb
  filtered_data_SatEmb %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "SatEmb", method ="SatEmb only") %>% 
    select(sensor, method, Q, R, S, T),
  filtered_data_SatEmb_CH %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "SatEmb", method ="SatEmb + CH") %>% 
    select(sensor, method, Q, R, S, T)
) %>%
  pivot_longer(cols = 3:6, names_to = "EUNIS", values_to = "Count") %>%
  ggplot(aes(x = EUNIS, y = Count, fill = EUNIS, group = EUNIS)) +
  geom_col() +
  facet_wrap(~ sensor + method) + theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  
```

### Split into training and test data sets

Set seed for reproducibility:

```{r}
set.seed(123)
```

Split data:

```{r}
# Split intro training and testing
# S2
split_slope_S2 <- split_train_test(filtered_data_slope_S2)
train_data_slope_S2 <- split_slope_S2$train
test_data_slope_S2 <- split_slope_S2$test
split_months_S2 <- split_train_test(filtered_data_months_S2)
train_data_months_S2 <- split_months_S2$train
test_data_months_S2 <- split_months_S2$test

# Landsat
split_slope_Landsat <- split_train_test(filtered_data_slope_Landsat)
train_data_slope_Landsat <- split_slope_Landsat$train
test_data_slope_Landsat <- split_slope_Landsat$test
split_months_Landsat <- split_train_test(filtered_data_months_Landsat)
train_data_months_Landsat <- split_months_Landsat$train
test_data_months_Landsat <- split_months_Landsat$test

# SatEmb
split_SatEmb <- split_train_test(filtered_data_SatEmb)
train_data_SatEmb <- split_SatEmb$train
test_data_SatEmb <- split_SatEmb$test
split_SatEmb_CH <- split_train_test(filtered_data_SatEmb_CH)
train_data_SatEmb_CH <- split_SatEmb_CH$train
test_data_SatEmb_CH <- split_SatEmb_CH$test
```

### Fit models

```{r eval=FALSE, include=FALSE}
# S2
rf_slope_S2 <- run_rf(vars_RF_slope, train_data_slope_S2, "EUNISa_1",
                      ntree = 1500)
rf_months_S2 <- run_rf(vars_RF_months, train_data_months_S2, "EUNISa_1",
                      ntree = 1500)

# Landsat
rf_slope_Landsat <- run_rf(vars_RF_slope_Landsat, train_data_slope_Landsat,
                           "EUNISa_1", ntree = 1500)
rf_months_Landsat <- run_rf(vars_RF_months_Landsat, train_data_months_Landsat,
                            "EUNISa_1", ntree = 1500)

# SatEmb
rf_SatEmb <- run_rf(vars_RF_SatEmb, train_data_SatEmb, "EUNISa_1", ntree = 1500)
rf_SatEmb_CH <- run_rf(vars_RF_SatEmb_CH, train_data_SatEmb_CH, "EUNISa_1",
                      ntree = 1500)
```

```{r}
models_no_refin <- c("rf_slope_S2", "rf_months_S2", "rf_slope_Landsat", 
                     "rf_months_Landsat", "rf_SatEmb", "rf_SatEmb_CH")
```

Print the time:

```{r}
for (model_name in models_no_refin) {
  obj <- get(model_name)
  cat("Time for", model_name, ":\n")
  print(obj$time)
}
```

Save:

```{r eval=FALSE, include=FALSE}
for (model_name in models_no_refin) {
  file_path <- paste0("objects/RF_1500/", model_name, "_model.Rdata")
  save(list = model_name, file = file_path)
}
```

### Predictions

```{r eval=FALSE, include=FALSE}
for (model_name in models_no_refin) {
  # Derive test data name by replacing "rf_" with "test_data_"
  test_data_name <- sub("^rf_", "test_data_", model_name)
  
  # Create prediction variable name
  prediction_name <- paste0(model_name, "_prediction")
  
  # Get model and test data
  model <- get(model_name)$model
  test_data <- get(test_data_name)
  
  # Run prediction
  prediction <- predict(model, newdata = test_data, 
                        type = "response")
  
  # Assign prediction to a variable
  assign(prediction_name, prediction)
}
```

```{r}
predictions_no_refin <- c("rf_slope_S2_prediction", "rf_months_S2_prediction",
                          "rf_slope_Landsat_prediction", 
                          "rf_months_Landsat_prediction",
                          "rf_SatEmb_prediction", "rf_SatEmb_CH_prediction")
```

Save:

```{r eval=FALSE, include=FALSE}
for (prediction_name in predictions_no_refin) {
  file_path <- paste0("objects/RF_1500/", prediction_name, ".Rdata")
  save(list = prediction_name, file = file_path)
}
```

### Confusion matrices

```{r}
for (prediction_name in predictions_no_refin) {
  # Derive test data name by replacing "_prediction" with "" 
  # and "rf_" with "test_data_"
  test_data_name <- paste0("test_data_", sub("^rf_", "", sub("_prediction$", "",
                                                             prediction_name)))
  
  # Get prediction and test data
  prediction <- get(prediction_name)
  test_data <- get(test_data_name)
  
  # Compute confusion matrix
  cat("Confusion Matrix for", sub("_prediction$", "", prediction_name), ":\n")
  print(confusionMatrix(prediction, test_data$EUNISa_1))
  cat("\n")
}
```

#### Plots

```{r}
for (prediction_name in predictions_no_refin) {
  # Derive name of test data
  test_data_name <- paste0("test_data_", sub("^rf_", "", sub("_prediction$", "",
                                                             prediction_name)))
  
  # Obtain objects
  prediction <- get(prediction_name)
  test_data <- get(test_data_name)

  # # Personalized title
  # title <- paste(#"Confusion Matrix for", 
  #   sub("_prediction$", "",
  #                                             prediction_name))

  # Plot
  print(plot_confusion_matrix(prediction, test_data$EUNISa_1, title))
}
```

### Variable importance

#### Unconditional

```{r eval=FALSE, include=FALSE}
for (model_name in models_no_refin) {

  # Compute variable importance
  varimp <- compute_varimp(get(model_name)$model, nperm = 100)
  
  # Create variable importance variable name
  varimp_name <- paste0(model_name, "_varimp")
  
  # Save result with a dynamic name
  assign(varimp_name, varimp)
}
```

```{r}
varimp_no_refin <- c("rf_slope_S2_varimp", "rf_months_S2_varimp",
                     "rf_slope_Landsat_varimp", "rf_months_Landsat_varimp", 
                     "rf_SatEmb_varimp", "rf_SatEmb_CH_varimp")
```

Save:

```{r eval=FALSE, include=FALSE}
for (varimp_name in varimp_no_refin) {
  file_path <- paste0("objects/RF_1500/", varimp_name, ".Rdata")
  save(list = varimp_name, file = file_path)
}
```

##### Plots

```{r}
for (varimp_name in varimp_no_refin) {
  cat("Plotting variable importance for", varimp_name, "...\n")
  plot(get(varimp_name)$varimp, margin = c(10, 6, 2, 2))
}
```

#### Conditional (TBD)

```{r eval=FALSE, include=FALSE}
for (model_name in models_no_refin) {

  # Compute variable importance (conditional)
  varimp_cond <- compute_varimp_cond(get(model_name)$model, nperm = 100)
  
  # Create variable importance variable name
  varimp_cond_name <- paste0(model_name, "_varimp_cond")
  
  # Save result with a dynamic name
  assign(varimp_cond_name, varimp_cond)
}
```

```{r}
# varimp_cond_no_refin <- c("rf_slope_S2_varimp_cond", "rf_months_S2_varimp_cond",
#                           "rf_slope_Landsat_varimp_cond", 
#                           "rf_months_Landsat_varimp_cond", 
#                           "rf_SatEmb_varimp_cond", "rf_SatEmb_CH_varimp_cond")
```

Save:

```{r eval=FALSE, include=FALSE}
for (varimp_cond_name in varimp_cond_no_refin) {
  file_path <- paste0("objects/RF_1500/", varimp_cond_name, ".Rdata")
  save(list = varimp_cond_name, file = file_path)
}
```

##### Plots

```{r}
# for (varimp_cond_name in varimp_cond_no_refin) {
#   cat("Plotting variable importance for", varimp_cond_name, "...\n")
#   plot(get(varimp_cond_name)$varimp, margin = c(10, 6, 2, 2))
# }
```

### ROC curves

```{r eval=FALSE, include=FALSE}
for (model_name in models_no_refin) {
  
  # Derive test data name by replacing "rf_" with "test_data_"
  test_data_name <- paste0("test_data_", sub("^rf_", "", model_name))

  # Compute ROC data
  roc_data <- compute_roc_level1(get(model_name)$model, get(test_data_name))

  # Create ROC data variable name
  roc_data_name <- paste0(model_name, "_roc_data")
  
  # Save result with a dynamic name
  assign(roc_data_name, roc_data)
}
```

Save:

```{r eval=FALSE, include=FALSE}
for (model_name in models_no_refin) {
  roc_data_name <- paste0(model_name, "_roc_data")
  file_path <- paste0("objects/RF_1500/", roc_data_name, ".Rdata")
  save(list = roc_data_name, file = file_path)
}
```

#### Plots

```{r}
roc_plots_no_refin <- list()

for (model_name in models_no_refin) {
  roc_data_name <- paste0(model_name, "_roc_data")
  roc_df <- get(roc_data_name)$roc

  roc_plot <- ggplot(roc_df, aes(x = FPR, y = TPR, color = Class)) +
    geom_line(linewidth = 1.2) +
    geom_abline(linetype = "dashed", color = "gray") +
    labs(title = paste("Multiclass ROC Curve -", model_name),
         x = "False Positive Rate",
         y = "True Positive Rate",
         color = "Class (AUC)") +
    theme_minimal() +
    theme(legend.position = "bottom")

  roc_plots_no_refin[[model_name]] <- roc_plot
}
```

```{r}
roc_plots_no_refin
```

## With refinement 

Based on PAs "Morralla protocol".

### Compute predictions for all data and combine with original data

```{r eval=FALSE, include=FALSE}
# S2, slope
filtered_data_slope_S2_with_probs <- bind_cols(
  filtered_data_slope_S2 %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_slope_S2$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_slope_S2,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_slope_S2$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_slope_S2,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# S2, months
filtered_data_months_S2_with_probs <- bind_cols(
  filtered_data_months_S2 %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_months_S2$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_months_S2,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_months_S2$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_months_S2,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# Landsat, slope
filtered_data_slope_Landsat_with_probs <- bind_cols(
  filtered_data_slope_Landsat %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_slope_Landsat$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_slope_Landsat,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_slope_Landsat$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_slope_Landsat,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# Landsat, months
filtered_data_months_Landsat_with_probs <- bind_cols(
  filtered_data_months_Landsat %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_months_Landsat$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_months_Landsat,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_months_Landsat$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_months_Landsat,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# SatEmb
filtered_data_SatEmb_with_probs <- bind_cols(
  filtered_data_SatEmb %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_SatEmb$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_SatEmb,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_SatEmb$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_SatEmb,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# SatEmb_CH
filtered_data_SatEmb_CH_with_probs <- bind_cols(
  filtered_data_SatEmb_CH %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_SatEmb_CH$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_SatEmb_CH,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_SatEmb_CH$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_SatEmb_CH,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)
```

```{r}
filtered_data_with_probs <- c("filtered_data_slope_S2_with_probs",
                              "filtered_data_months_S2_with_probs",
                              "filtered_data_slope_Landsat_with_probs", 
                              "filtered_data_months_Landsat_with_probs",
                              "filtered_data_SatEmb_with_probs", 
                              "filtered_data_SatEmb_CH_with_probs")
```

Save:

```{r eval=FALSE, include=FALSE}
for (filtered_data_name in filtered_data_with_probs) {
  file_path <- paste0("objects/RF_1500/", filtered_data_name, ".Rdata")
  save(list = filtered_data_name, file = file_path)
}
```

### Detect probability columns

```{r}
prob_cols_slope_S2 <- names(filtered_data_slope_S2_with_probs)[grepl("^prob_", names(filtered_data_slope_S2_with_probs))]
prob_cols_months_S2 <- names(filtered_data_months_S2_with_probs)[grepl("^prob_", names(filtered_data_months_S2_with_probs))]
prob_cols_slope_Landsat <- names(filtered_data_slope_Landsat_with_probs)[grepl("^prob_", names(filtered_data_slope_Landsat_with_probs))]
prob_cols_months_Landsat <- names(filtered_data_months_Landsat_with_probs)[grepl("^prob_", names(filtered_data_months_Landsat_with_probs))]
prob_cols_SatEmb <- names(filtered_data_SatEmb_with_probs)[grepl("^prob_", names(filtered_data_SatEmb_with_probs))]
prob_cols_SatEmb_CH <- names(filtered_data_SatEmb_CH_with_probs)[grepl("^prob_", names(filtered_data_SatEmb_CH_with_probs))]
```

### Calculate auxiliary metrics

```{r}
df_aux_slope_S2 <- calculate_df_aux(filtered_data_slope_S2_with_probs,
                                     prob_cols_slope_S2)
df_aux_months_S2 <- calculate_df_aux(filtered_data_months_S2_with_probs,
                                     prob_cols_months_S2)
df_aux_slope_Landsat <- calculate_df_aux(filtered_data_slope_Landsat_with_probs,
                                     prob_cols_slope_Landsat)
df_aux_months_Landsat <- calculate_df_aux(filtered_data_months_Landsat_with_probs,
                                     prob_cols_months_Landsat)
df_aux_SatEmb <- calculate_df_aux(filtered_data_SatEmb_with_probs,
                                  prob_cols_SatEmb)
df_aux_SatEmb_CH <- calculate_df_aux(filtered_data_SatEmb_CH_with_probs,
                                     prob_cols_SatEmb_CH)
```

### Filtering ONCE

Applying "Morralla" protocol once, removing negatives until 35% of the class and, if there is room, uncertain positives until 10% of the rest.

#### Apply filtering function

```{r}
df_final_35_10_slope_S2 <- df_aux_slope_S2 %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg35_pos10,
          df = df_aux_slope_S2)
df_final_35_10_months_S2 <- df_aux_months_S2 %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg35_pos10,
          df = df_aux_months_S2)
df_final_35_10_slope_Landsat <- df_aux_slope_Landsat %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg35_pos10,
          df = df_aux_slope_Landsat)
df_final_35_10_months_Landsat <- df_aux_months_Landsat %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg35_pos10,
          df = df_aux_months_Landsat)
df_final_35_10_SatEmb <- df_aux_SatEmb %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg35_pos10,
          df = df_aux_SatEmb)
df_final_35_10_SatEmb_CH <- df_aux_SatEmb_CH %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg35_pos10,
          df = df_aux_SatEmb_CH)
```

```{r}
filtered_data_35_10_slope_S2_probs <- filtered_data_slope_S2 %>%
  dplyr::filter(PlotObservationID %in% df_final_35_10_slope_S2$PlotObservationID)
filtered_data_35_10_months_S2_probs <- filtered_data_months_S2 %>%
  dplyr::filter(PlotObservationID %in% df_final_35_10_months_S2$PlotObservationID)
filtered_data_35_10_slope_Landsat_probs <- filtered_data_slope_Landsat %>%
  dplyr::filter(PlotObservationID %in% df_final_35_10_slope_Landsat$PlotObservationID)
filtered_data_35_10_months_Landsat_probs <- filtered_data_months_Landsat %>%
  dplyr::filter(PlotObservationID %in% df_final_35_10_months_Landsat$PlotObservationID)
filtered_data_35_10_SatEmb_probs <- filtered_data_SatEmb %>%
  dplyr::filter(PlotObservationID %in% df_final_35_10_SatEmb$PlotObservationID)
filtered_data_35_10_SatEmb_CH_probs <- filtered_data_SatEmb_CH %>%
  dplyr::filter(PlotObservationID %in% df_final_35_10_SatEmb_CH$PlotObservationID)
```

#### Compare classes before and after

```{r}
compare_classes(filtered_data_slope_S2, filtered_data_35_10_slope_S2_probs)
compare_classes(filtered_data_months_S2, filtered_data_35_10_months_S2_probs)
compare_classes(filtered_data_slope_Landsat, filtered_data_35_10_slope_Landsat_probs)
compare_classes(filtered_data_months_Landsat, filtered_data_35_10_months_Landsat_probs)
compare_classes(filtered_data_SatEmb, filtered_data_35_10_SatEmb_probs)
compare_classes(filtered_data_SatEmb_CH, filtered_data_35_10_SatEmb_CH_probs)
```

#### Split into training and testing

Set seed for reproducibility:

```{r}
set.seed(123)
```

```{r}
train_data_35_10_slope_S2_probs <-
  split_train_test(filtered_data_35_10_slope_S2_probs)$train
test_data_35_10_slope_S2_probs <-
  split_train_test(filtered_data_35_10_slope_S2_probs)$test

train_data_35_10_months_S2_probs <-
  split_train_test(filtered_data_35_10_months_S2_probs)$train
test_data_35_10_months_S2_probs <-
  split_train_test(filtered_data_35_10_months_S2_probs)$test

train_data_35_10_slope_Landsat_probs <-
  split_train_test(filtered_data_35_10_slope_Landsat_probs)$train
test_data_35_10_slope_Landsat_probs <-
  split_train_test(filtered_data_35_10_slope_Landsat_probs)$test

train_data_35_10_months_Landsat_probs <-
  split_train_test(filtered_data_35_10_months_Landsat_probs)$train
test_data_35_10_months_Landsat_probs <-
  split_train_test(filtered_data_35_10_months_Landsat_probs)$test

train_data_35_10_SatEmb_probs <-
  split_train_test(filtered_data_35_10_SatEmb_probs)$train
test_data_35_10_SatEmb_probs <-
  split_train_test(filtered_data_35_10_SatEmb_probs)$test

train_data_35_10_SatEmb_CH_probs <-
  split_train_test(filtered_data_35_10_SatEmb_CH_probs)$train
test_data_35_10_SatEmb_CH_probs <-
  split_train_test(filtered_data_35_10_SatEmb_CH_probs)$test
```

#### Fit models

```{r eval=FALSE, include=FALSE}
rf_slope_S2_refin_35_10 <- run_rf(vars_RF_slope, 
                                  train_data_35_10_slope_S2_probs, "EUNISa_1",
                                  ntree = 1500)
rf_months_S2_refin_35_10 <- run_rf(vars_RF_months,
                                   train_data_35_10_months_S2_probs, "EUNISa_1",
                                  ntree = 1500)
rf_slope_Landsat_refin_35_10 <- run_rf(vars_RF_slope_Landsat,
                                       train_data_35_10_slope_Landsat_probs, 
                                       "EUNISa_1", ntree = 1500)
rf_months_Landsat_refin_35_10 <- run_rf(vars_RF_months_Landsat,
                                        train_data_35_10_months_Landsat_probs, 
                                        "EUNISa_1", ntree = 1500)
rf_SatEmb_refin_35_10 <- run_rf(vars_RF_SatEmb, train_data_35_10_SatEmb_probs, 
                                "EUNISa_1", ntree = 1500)
rf_SatEmb_CH_refin_35_10 <- run_rf(vars_RF_SatEmb_CH, 
                                   train_data_35_10_SatEmb_CH_probs, "EUNISa_1",
                                  ntree = 1500)
```

```{r}
models_refin_35_10 <- c("rf_slope_S2_refin_35_10", "rf_months_S2_refin_35_10",
                        "rf_slope_Landsat_refin_35_10",
                        "rf_months_Landsat_refin_35_10",
                        "rf_SatEmb_refin_35_10", "rf_SatEmb_CH_refin_35_10")
```

Save:

```{r eval=FALSE, include=FALSE}
for (model_name in models_refin_35_10) {
  file_path <- paste0("objects/RF_1500/", model_name, "_model.Rdata")
  save(list = model_name, file = file_path)
}
```

#### Predictions

```{r eval=FALSE, include=FALSE}
for (model_name in models_refin_35_10) {
  # Derive test data name
  test_data_name <- paste0(gsub("_refin_35_10$", "", 
                                sub("^rf_", "test_data_35_10_", model_name)), 
                           "_probs")

  # Create prediction variable name
  prediction_name <- paste0(model_name, "_prediction")
  
  # Get model and test data
  model <- get(model_name)$model
  test_data <- get(test_data_name)
  
  # Run prediction
  prediction <- predict(model, newdata = test_data, 
                        type = "response")
  
  # Assign prediction to a variable
  assign(prediction_name, prediction)
}
```

```{r}
predictions_refin_35_10 <- c("rf_slope_S2_refin_35_10_prediction",
                        "rf_months_S2_refin_35_10_prediction",
                        "rf_slope_Landsat_refin_35_10_prediction",
                        "rf_months_Landsat_refin_35_10_prediction",
                        "rf_SatEmb_refin_35_10_prediction",
                        "rf_SatEmb_CH_refin_35_10_prediction")
```

Save:

```{r eval=FALSE, include=FALSE}
for (prediction_name in predictions_refin_35_10) {
  file_path <- paste0("objects/RF_1500/", prediction_name, ".Rdata")
  save(list = prediction_name, file = file_path)
  }
```

#### Confusion matrices

```{r}
for (prediction_name in predictions_refin_35_10) {
  # Derive test data name
  test_data_name <- paste0(gsub("_refin_35_10$", "",
                              sub("^rf_", "test_data_35_10_", 
                                  sub("_prediction$", "", prediction_name))), 
                         "_probs")
  
  # Get prediction and test data
  prediction <- get(prediction_name)
  test_data <- get(test_data_name)
  
  # Compute confusion matrix
  cat("Confusion Matrix for", sub("_prediction$", "", prediction_name), ":\n")
  print(confusionMatrix(prediction, test_data$EUNISa_1))
  cat("\n")
}
```

##### Plots

```{r}
for (prediction_name in predictions_refin_35_10) {
  # Derive test data name
  test_data_name <- paste0(gsub("_refin_35_10$", "",
                              sub("^rf_", "test_data_35_10_", 
                                  sub("_prediction$", "", prediction_name))), 
                         "_probs")
  
  # Obtain objects
  prediction <- get(prediction_name)
  test_data <- get(test_data_name)

  # # Personalized title
  # title <- paste("Confusion Matrix for", sub("_prediction$", "",
  #                                            prediction_name))

  # Plot
  print(plot_confusion_matrix(prediction, test_data$EUNISa_1, title))
}
```

#### Variable importance

##### Unconditional

```{r eval=FALSE, include=FALSE}
for (model_name in models_refin_35_10) {

  # Compute variable importance
  varimp <- compute_varimp(get(model_name)$model, nperm = 100)
  
  # Create variable importance variable name
  varimp_name <- paste0(model_name, "_varimp")
  
  # Save result with a dynamic name
  assign(varimp_name, varimp)
}
```

```{r}
varimp_refin_35_10 <- c("rf_slope_S2_refin_35_10_varimp",
                        "rf_months_S2_refin_35_10_varimp",
                        "rf_slope_Landsat_refin_35_10_varimp",
                        "rf_months_Landsat_refin_35_10_varimp",
                        "rf_SatEmb_refin_35_10_varimp",
                        "rf_SatEmb_CH_refin_35_10_varimp")
```

Save:

```{r eval=FALSE, include=FALSE}
for (varimp_name in varimp_refin_35_10) {
  file_path <- paste0("objects/RF_1500/", varimp_name, ".Rdata")
  save(list = varimp_name, file = file_path)
}
```

###### Plots

```{r}
for (varimp_name in varimp_refin_35_10) {
  cat("Plotting variable importance for", varimp_name, "...\n")
  plot(get(varimp_name)$varimp, margin = c(10, 6, 2, 2))
}
```

##### Conditional (TBD)

```{r eval=FALSE, include=FALSE}
for (model_name in models_refin_35_10) {

  # Compute variable importance (conditional)
  varimp_cond <- compute_varimp_cond(get(model_name)$model, nperm = 100)
  
  # Create variable importance variable name
  varimp_cond_name <- paste0(model_name, "_varimp_cond")
  
  # Save result with a dynamic name
  assign(varimp_cond_name, varimp_cond)
}
```

```{r}
# varimp_cond_refin_35_10 <- c("rf_slope_S2_refin_35_10_varimp_cond",
#                         "rf_months_S2_refin_35_10_varimp_cond",
#                         "rf_slope_Landsat_refin_35_10_varimp_cond",
#                         "rf_months_Landsat_refin_35_10_varimp_cond",
#                         "rf_SatEmb_refin_35_10_varimp_cond",
#                         "rf_SatEmb_CH_refin_35_10_varimp_cond")
```

Save:

```{r eval=FALSE, include=FALSE}
for (varimp_cond_name in varimp_cond_refin_35_10) {
  file_path <- paste0("objects/RF_1500/", varimp_cond_name, ".Rdata")
  save(list = varimp_cond_name, file = file_path)
}
```

###### Plots

```{r}
# for (varimp_cond_name in varimp_cond_refin_35_10) {
#   cat("Plotting variable importance for", varimp_cond_name, "...\n")
#   plot(get(varimp_cond_name)$varimp, margin = c(10, 6, 2, 2))
# }
```

#### ROC curves

```{r eval=FALSE, include=FALSE}
for (model_name in models_refin_35_10) {
  
  # Derive test data name
  test_data_name <- paste0(gsub("_refin_35_10$", "", 
                                sub("^rf_", "test_data_35_10_", model_name)), 
                           "_probs")

  # Compute ROC data
  roc_data <- compute_roc_level1(get(model_name)$model, get(test_data_name))

  # Create ROC data variable name
  roc_data_name <- paste0(model_name, "_roc_data")
  
  # Save result with a dynamic name
  assign(roc_data_name, roc_data)
}
```

Save:

```{r eval=FALSE, include=FALSE}
for (model_name in models_refin_35_10) {
  roc_data_name <- paste0(model_name, "_roc_data")
  file_path <- paste0("objects/RF_1500/", roc_data_name, ".Rdata")
  save(list = roc_data_name, file = file_path)
}
```

##### Plots

```{r}
roc_plots_refin_35_10 <- list()

for (model_name in models_refin_35_10) {
  roc_data_name <- paste0(model_name, "_roc_data")
  roc_df <- get(roc_data_name)$roc

  roc_plot <- ggplot(roc_df, aes(x = FPR, y = TPR, color = Class)) +
    geom_line(linewidth = 1.2) +
    geom_abline(linetype = "dashed", color = "gray") +
    labs(title = paste("Multiclass ROC Curve -", model_name),
         x = "False Positive Rate",
         y = "True Positive Rate",
         color = "Class (AUC)") +
    theme_minimal() +
    theme(legend.position = "bottom")

  roc_plots_refin_35_10[[model_name]] <- roc_plot
}
```

```{r}
roc_plots_refin_35_10
```


### HERE: Filtering TWICE

Applying "Morralla" protocol once, removing negatives until 20% of the class and, if there is room, uncertain positives until 5% of the rest.

Then refitting the model and applying "Morralla" protocol again with same %s.




# Session info

```{r}
sessionInfo()
```

