---
title: "Script to get correlations between embeddings and other data"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

This R script is used to get correlations between embeddings and other data.

# Load libraries

```{r}
library(tidyverse)
library(here)
```

# Read data

Read data validation S2, SatEmb and Landsat:

```{r}
data_validation_S2 <- read_tsv(here(
  "data", "clean","final_RS_data_bands_S2_all_20251021.csv"))
data_validation_SatEmb <- read_tsv(here(
  "data", "clean","final_RS_data_SatEmb_20251010.csv")) %>%
  rename(EUNISa_1_descr = EUNIS_1_d)
data_validation_Landsat <- read_tsv(here(
  "data", "clean","final_RS_data_bands_Landsat_all_20251021.csv"))
```

No parsing issues!

```{r}
nrow(data_validation_S2)
nrow(data_validation_Landsat)
nrow(data_validation_SatEmb)
```

# Filter out manipulated observations

## Read list of PlotObservationIDs to keep

```{r}
PlotObservationID_list <- read_csv(
  here("data", "clean","PlotObservationIDs_filtered_QRST_20251211.csv"))
```

## Filter out data to keep only those PlotObservationIDs

```{r}
data_validation_S2 <- data_validation_S2 %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
data_validation_SatEmb <- data_validation_SatEmb %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
data_validation_Landsat <- data_validation_Landsat %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
```

```{r}
nrow(data_validation_S2)
nrow(data_validation_Landsat)
nrow(data_validation_SatEmb)
```

# Merge the three datasets

```{r}
data_validation <- data_validation_Landsat %>% 
  # Add "Landsat_" prefix to RS vars
  rename_with(
    ~ paste0("Landsat_", .x),
    matches("^(NDVI|SAVI|EVI|NDMI|NDWI)")
    ) %>%
  # Join S2 data
  full_join(
    data_validation_S2 %>%
      # Add "S2_ prefix to RS vars
      rename_with(
        ~ paste0("S2_", .x),
        matches("^(NDVI|SAVI|EVI|NDMI|NDWI)")
      )
    ) %>%
  # Add S2 T/F
  mutate(S2 = rowSums(!is.na(across(starts_with("S2_"))))>0) %>%
  # Join SatEmb data
  full_join(data_validation_SatEmb) %>%
  # Add SatEmb T/F
  mutate(SatEmb = ifelse(is.na(A00), FALSE, TRUE))
```

```{r}
data_validation %>% count(S2)
data_validation %>% count(SatEmb)
```

# Define lists of predictor vars

```{r}
vars_RF_both <- c(
  # Min values of all indices
  "NDVI_min", "EVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, EVI and SAVI
  "NDVI_auc_slope", "EVI_auc_slope", "SAVI_auc_slope",
  # AUC of NDVI, EVI and SAVI between march and oct
  "NDVI_auc_mar_oct", "EVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI, EVI and SAVI at sos, pos and eos
  "NDVI_sos_slope_value", "NDVI_pos_value", "NDVI_eos_slope_value",
  "EVI_sos_slope_value", "EVI_pos_value", "EVI_eos_slope_value", 
  "SAVI_sos_slope_value", "SAVI_pos_value", "SAVI_eos_slope_value",
  # Differences pos-sos in value and doy
  "NDVI_diff_pos_sos_slope_value", "EVI_diff_pos_sos_slope_value",
  "SAVI_diff_pos_sos_slope_value","NDVI_diff_pos_sos_slope_doy", 
  "EVI_diff_pos_sos_slope_doy", "SAVI_diff_pos_sos_slope_doy",
  # Differences pos-eos in value and doy
  "NDVI_diff_pos_eos_slope_value", "EVI_diff_pos_eos_slope_value",
  "SAVI_diff_pos_eos_slope_value", "NDVI_diff_pos_eos_slope_doy", 
  "EVI_diff_pos_eos_slope_doy", "SAVI_diff_pos_eos_slope_doy",
  # Growing season duration
  "NDVI_slope_gsd", "EVI_slope_gsd", "SAVI_slope_gsd",
  # Values of NDVI, EVI and SAVI at march and oct
  "NDVI_avg_value_03", "NDVI_avg_value_10",
  "EVI_avg_value_03", "EVI_avg_value_10", 
  "SAVI_avg_value_03", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "EVI_diff_pos_march_value",
  "SAVI_diff_pos_march_value","NDVI_diff_pos_march_doy", 
  "EVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "EVI_diff_pos_oct_value",
  "SAVI_diff_pos_oct_value", "NDVI_diff_pos_oct_doy", 
  "EVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy",
  # Canopy height
  "canopy_height")
```

```{r}
# Excluding EVI because many values outside of range -1, 1
vars_RF_both_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI and SAVI
  "NDVI_auc_slope", "SAVI_auc_slope",
  # AUC of NDVI, and SAVI between march and oct
  "NDVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI and SAVI at sos, pos and eos
  "NDVI_sos_slope_value", "NDVI_pos_value", "NDVI_eos_slope_value",
  "SAVI_sos_slope_value", "SAVI_pos_value", "SAVI_eos_slope_value",
  # Differences pos-sos in value and doy
  "NDVI_diff_pos_sos_slope_value","SAVI_diff_pos_sos_slope_value",
  "NDVI_diff_pos_sos_slope_doy", "SAVI_diff_pos_sos_slope_doy",
  # Differences pos-eos in value and doy
  "NDVI_diff_pos_eos_slope_value", "SAVI_diff_pos_eos_slope_value",
  "NDVI_diff_pos_eos_slope_doy", "SAVI_diff_pos_eos_slope_doy",
  # Growing season duration
  "NDVI_slope_gsd", "SAVI_slope_gsd",
  # Values of NDVI and SAVI at march and oct
  "NDVI_avg_value_03", "NDVI_avg_value_10",
  "SAVI_avg_value_03", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "SAVI_diff_pos_march_value",
  "NDVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "SAVI_diff_pos_oct_value",
  "NDVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy",
  # Canopy height
  "canopy_height")
```

# Correlations

```{r}
A_cols <- data_validation %>%
  select(starts_with("A")) %>%
  names()
```

## SatEmb and canopy height

```{r}
cor_A_CH <- cor(
  data_validation[, A_cols],
  data_validation[, "canopy_height"],
  use = "pairwise.complete.obs"
)
```

```{r}
cor_A_CH_df <- cor_A_CH %>%
  as.data.frame() %>%
  tibble::rownames_to_column("A_band") %>%
  pivot_longer(
    -A_band,
    names_to = "CH",
    values_to = "correlation"
  )
```

```{r}
thr <- 0.5

A_CH_strong <- cor_A_CH_df %>%
  filter(abs(correlation) >= thr) %>%
  arrange(A_band, desc(abs(correlation)))
```

```{r}
A_CH_strong
```

## SatEmb and S2

```{r}
A_cols  <- names(data_validation)[startsWith(names(data_validation), "A")]
S2_cols <- names(data_validation)[startsWith(names(data_validation), "S2_")]

S2_cols <- intersect(paste0("S2_", vars_RF_both), names(data_validation))

length(A_cols)
length(S2_cols)
```

```{r}
cor_A_S2 <- cor(
  data_validation[, A_cols],
  data_validation[, S2_cols],
  use = "pairwise.complete.obs"
)
```

```{r}
cor_A_S2_df <- cor_A_S2 %>%
  as.data.frame() %>%
  tibble::rownames_to_column("A_band") %>%
  pivot_longer(
    -A_band,
    names_to = "S2_var",
    values_to = "correlation"
  )
```

See the three most strongly correlated vars for each embedding band:

```{r}
top3_A_S2 <- cor_A_S2_df %>%
  group_by(A_band) %>%
  slice_max(order_by = abs(correlation), n = 3, with_ties = FALSE) %>%
  arrange(A_band, desc(abs(correlation)))
```

```{r}
top3_A_S2
```

See the stronger correlations:

```{r}
thr <- 0.5

A_S2_strong <- cor_A_S2_df %>%
  filter(abs(correlation) >= thr) %>%
  arrange(A_band, desc(abs(correlation)))
```

```{r}
A_S2_strong
```

## SatEmb and Landsat

```{r}
A_cols  <- names(data_validation)[startsWith(names(data_validation), "A")]
Landsat_cols <- names(data_validation)[startsWith(names(data_validation), "Landsat_")]

Landsat_cols <- intersect(paste0("Landsat_", vars_RF_both_Landsat), names(data_validation))

length(A_cols)
length(Landsat_cols)
```

```{r}
cor_A_Landsat <- cor(
  data_validation[, A_cols],
  data_validation[, Landsat_cols],
  use = "pairwise.complete.obs"
)
```

```{r}
cor_A_Landsat_df <- cor_A_Landsat %>%
  as.data.frame() %>%
  tibble::rownames_to_column("A_band") %>%
  pivot_longer(
    -A_band,
    names_to = "Landsat_var",
    values_to = "correlation"
  )
```

See the three most strongly correlated vars for each embedding band:

```{r}
top3_A_Landsat <- cor_A_Landsat_df %>%
  group_by(A_band) %>%
  slice_max(order_by = abs(correlation), n = 3, with_ties = FALSE) %>%
  arrange(A_band, desc(abs(correlation)))
```

```{r}
top3_A_Landsat
```

See the stronger correlations:

```{r}
thr <- 0.5

A_Landsat_strong <- cor_A_Landsat_df %>%
  filter(abs(correlation) >= thr) %>%
  arrange(A_band, desc(abs(correlation)))
```

```{r}
A_Landsat_strong
```

# Some trials or regressions

```{r}
ggplot(data_validation, aes(x = canopy_height, y = A07)) + geom_point() +
  geom_smooth(method = "lm")
ggplot(data_validation, aes(x = S2_NDVI_pos_value, y = A05)) + geom_point() +
  geom_smooth(method = "lm")
ggplot(data_validation, aes(x = Landsat_NDVI_pos_value, y = A05)) + geom_point() +
  geom_smooth(method = "lm")
ggplot(data_validation, aes(x = S2_NDMI_max, y = A61)) + geom_point() +
  geom_smooth(method = "lm")
ggplot(data_validation, aes(x = Landsat_NDMI_max, y = A61)) + geom_point() +
  geom_smooth(method = "lm")
```

```{r}
summary(lm(A07 ~ canopy_height, data = data_validation))
summary(lm(A05 ~ S2_NDVI_pos_value, data = data_validation))
summary(lm(A05 ~ Landsat_NDVI_pos_value, data = data_validation))
summary(lm(A61 ~ S2_NDMI_max, data = data_validation))
summary(lm(A61 ~ Landsat_NDMI_max, data = data_validation))
```

# PCA of A‑bands + RS loadings

```{r}
X_A <- data_validation[, A_cols]
complete_rows <- complete.cases(X_A)
X_A_complete <- X_A[complete_rows, ]
X_A_scaled <- scale(X_A_complete)
pca_A <- prcomp(X_A_scaled, center = FALSE, scale. = FALSE)
```

```{r}
summary(pca_A)$importance[2, 1:5]
```

```{r}
PC_scores <- pca_A$x[, 1:2]  # PC1 & PC2

RS_cols <- c(S2_cols, Landsat_cols)
RS_data_complete <- data_validation[complete_rows, RS_cols]

RS_cor <- cor(PC_scores, RS_data_complete, use = "pairwise.complete.obs")
```

```{r}
RS_loadings <- as.data.frame(t(RS_cor)) |>
  rownames_to_column("RS_var") |>
  rename(PC1 = PC1, PC2 = PC2)
```

```{r}
ggplot() +
  geom_point(
    aes(pca_A$x[,1], pca_A$x[,2]),
    alpha = 0.3, size = 1
  ) +
  geom_segment(
    data = RS_loadings,
    aes(x = 0, y = 0, xend = PC1 * 5, yend = PC2 * 5),
    arrow = arrow(length = unit(0.15, "cm")),
    color = "red", alpha = 0.7
  ) +
  geom_text(
    data = RS_loadings,
    aes(x = PC1 * 5, y = PC2 * 5, label = RS_var),
    size = 2, vjust = 1.2
  ) +
  theme_minimal() +
  labs(
    x = "PC1 (embeddings)",
    y = "PC2 (embeddings)",
    title = "Embedding PCA with Sentinel-2 & Landsat correlations"
  )
```

# Corr matrix 4 most imp bands

```{r}
top3_A_S2 %>% filter(A_band %in% c("A24", "A33", "A07", "A58"))
```


# Session info

```{r}
sessionInfo()
```

