---
title: "Script to validate points in ReSurvey database using RS data"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

This R script is used to make a first attempt to validate the points in the ReSurvey database using RS indicators (NDVI and NDMI).

# Load libraries

```{r}
library(tidyverse)
library(here)
library(sf)
library(gridExtra)
library(sf)
library(rnaturalearth)
library(readxl)
```

# Define printall function

```{r}
printall <- function(tibble) {
  print(tibble, width = Inf)
  }
```

# Read ReSurvey data with RS indicators

```{r}
db_resurv_RS<-read_tsv(
  here("data", "clean", "db_resurv_RS.csv"),
  col_types = cols(
    # Dynamically specify EUNIS columns as character
    .default = col_guess(),  # Default guessing for other columns
    EUNISa = col_character(),
    EUNISb = col_character(),
    EUNISc = col_character(),
    EUNISd = col_character(),
    EUNISa_1 = col_character(),
    EUNISa_2 = col_character(),
    EUNISa_3 = col_character(),
    EUNISa_4 = col_character(),
    EUNISb_1 = col_character(),
    EUNISb_2 = col_character(),
    EUNISb_3 = col_character(),
    EUNISb_4 = col_character(),
    EUNISc_1 = col_character(),
    EUNISc_2 = col_character(),
    EUNISc_3 = col_character(),
    EUNISc_4 = col_character(),
    EUNISd_1 = col_character(),
    EUNISd_2 = col_character(),
    EUNISd_3 = col_character(),
    EUNISd_4 = col_character(),
    EUNISa_1_descr = col_character(),
    EUNISb_1_descr = col_character(),
    EUNISc_1_descr = col_character(),
    EUNISd_1_descr = col_character(),
    EUNIS_assignation = col_character(),
    EUNISa_2_descr = col_character(),
    EUNISa_3_descr = col_character(),
    EUNISa_4_descr = col_character(),
    EUNISb_2_descr = col_character(),
    EUNISb_3_descr = col_character(),
    EUNISb_4_descr = col_character(),
    EUNISc_2_descr = col_character(),
    EUNISc_3_descr = col_character(),
    EUNISc_4_descr = col_character(),
    EUNISd_2_descr = col_character(),
    EUNISd_3_descr = col_character(),
    EUNISd_4_descr = col_character()
    )
  )
```

No parsing issues!

# Several EUNIS level 1 assigned

Number of rows where there is more than one EUNIS 1 assigned, and they are different among them. See what to do with these later! So far I take EUNISa_1.

```{r}
nrow(db_resurv_RS %>% 
       # Rows with more than one EUNIS 1 assigned
       filter(!is.na(EUNISb_1)) %>% 
       filter(EUNISa_1!=EUNISb_1 | EUNISb_1 != EUNISc_1 | EUNISa_1 != EUNISc_1))
```

# Tibble with selected columns

```{r}
db_resurv_RS_short <- db_resurv_RS %>%
  select(PlotObservationID, Country, RS_CODE, `ReSurvey site`, `ReSurvey plot`,
         `Manipulate (y/n)`, `Type of manipulation`, Lon_updated, Lat_updated,
         `Location method`, `Location uncertainty (m)`, EUNISa_1,
         EUNISa_1_descr, EUNISa_2, EUNISa_2_descr, EUNISa_3, EUNISa_3_descr,
         EUNISa_4, EUNISa_4_descr, date, year, biogeo, unit, year_RS, source,
         Lon_RS, Lat_RS, NDVI, NDMI, RS_data)
```

# Missing data checks

Do when all RS data is ready!

# Flag when year is different between RS data and ReSurvey db

```{r}
db_resurv_RS_short <- db_resurv_RS_short %>%
  mutate(year_diff = year != year_RS)
```

```{r}
db_resurv_RS_short %>% count(year_diff)
```

# Flag when coordinates are different between RS data and ReSurvey db

```{r}
db_resurv_RS_short <- db_resurv_RS_short %>%
  mutate(Lon_diff = Lon_updated != Lon_RS,
         Lat_diff = Lat_updated != Lat_RS)
```

```{r}
db_resurv_RS_short %>% count(Lon_diff)
db_resurv_RS_short %>% count(Lat_diff)
```

NDVI and NDMI would need to be calculated again for those?

# Distributions NDVI and NDMI for all bioregions

```{r}
ggplot(db_resurv_RS_short %>% filter(!is.na(EUNISa_1_descr)), aes(x = NDVI)) +
  facet_wrap(~ EUNISa_1_descr) +
  geom_density()
ggplot(db_resurv_RS_short %>% filter(!is.na(EUNISa_1_descr)), aes(x = NDMI)) +
  facet_wrap(~ EUNISa_1_descr) +
  geom_density()
```

# Violin + boxplot + points

```{r}
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

```{r warning=FALSE}
ggplot(data = filter(db_resurv_RS_short %>%
                       filter(RS_data == T & !is.na(EUNISa_1_descr))),
       aes(x = EUNISa_1_descr, y = NDVI, fill = EUNISa_1_descr)) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
  geom_point(aes(y = NDVI, color = EUNISa_1_descr),
             position = position_jitter(width = 0.15), size = 1, alpha = 0.25) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y=mean, geom="point", shape = 20, size = 1) +
  stat_summary(fun.data = function(x) data.frame(y = max(x) + 0.1,
                                                 label = length(x)),
               geom = "text", aes(label = ..label..), vjust = 0.5) +
  labs(y = "NDVI", x = "EUNISa_1_descr") +
  guides(fill = FALSE, color = FALSE) +
  theme_bw() + coord_flip()
ggsave(
  here("output", "figures", "NDVI.tiff"),
  width = 21, height = 29.7, units = "cm", dpi = 300)
ggplot(data = filter(db_resurv_RS_short %>%
                       filter(RS_data == T & !is.na(EUNISa_1_descr))),
       aes(x = EUNISa_1_descr, y = NDMI, fill = EUNISa_1_descr)) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
  geom_point(aes(y = NDMI, color = EUNISa_1_descr),
             position = position_jitter(width = 0.15), size = 1, alpha = 0.25) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y=mean, geom="point", shape = 20, size = 1) +
  stat_summary(fun.data = function(x) data.frame(y = max(x) + 0.1,
                                                 label = length(x)),
               geom = "text", aes(label = ..label..), vjust = 0.5) +
  labs(y = "NDMI", x = "EUNISa_1_descr") +
  guides(fill = FALSE, color = FALSE) +
  theme_bw() + coord_flip()
ggsave(
  here("output", "figures", "NDMI.tiff"),
  width = 21, height = 29.7, units = "cm", dpi = 300)
```


# Distributions NDVI and NDMI for each bioregion

```{r}
# Get unique values of EUNISa_1_descr, excluding NA
eunis_levels <- unique(db_resurv_RS_short$EUNISa_1_descr
                       [!is.na(db_resurv_RS_short$EUNISa_1_descr)])

# Generate plots using purrr::map
eunis_levels %>%
  set_names() %>%  # Set names to access plots by EUNISa_1_descr
  map(~
        ggplot() +
        # Density plot for each biogeo (filled)
        geom_density(data = filter(db_resurv_RS_short, EUNISa_1_descr == .x), 
                     aes(x = NDVI, fill = biogeo), alpha = 0.5) +
        # Density plot for NDVI in red with thicker line
        geom_density(data = filter(db_resurv_RS_short, EUNISa_1_descr == .x), 
                     aes(x = NDVI),
                     color = "red", alpha = 0.5, size = 1,
                     show.legend = FALSE) +
        labs(title = .x)
      ) %>%
  walk(print)
```

```{r}
eunis_levels %>%
  set_names() %>%  # Set names to access plots by EUNISa_1_descr
  map(~
        ggplot() +
        # Density plot for each biogeo (filled)
        geom_density(data = filter(db_resurv_RS_short, EUNISa_1_descr == .x), 
                     aes(x = NDMI, fill = biogeo), alpha = 0.5) +
        # Density plot for NDMI in red with thicker line
        geom_density(data = filter(db_resurv_RS_short, EUNISa_1_descr == .x), 
                     aes(x = NDMI),
                     color = "red", alpha = 0.5, size = 1,
                     show.legend = FALSE) +
        labs(title = .x)
      ) %>%
  walk(print)
```

# Violin + boxplot + points for each bioregion

```{r warning=FALSE}
eunis_levels %>%
  set_names() %>%  # Set names to access plots by EUNISa_1_descr
  map(~
        ggplot(data = filter(db_resurv_RS_short %>% filter(RS_data == T),
                             EUNISa_1_descr == .x),
               aes(x = biogeo, y = NDVI, fill = biogeo)) +
        geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
        geom_point(aes(y = NDVI, color = biogeo),
                   position = position_jitter(width = 0.15), size = 1, alpha = 0.25) +
        geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
        stat_summary(fun.y=mean, geom="point", shape = 20, size=2) +
        labs(y = "NDVI", x = "biogeo") +
        guides(fill = FALSE, color = FALSE) +
        labs(title = .x)
      ) %>%
  walk(print)
```

```{r warning=FALSE}
ggplot(data = filter(db_resurv_RS_short %>%
                       filter(RS_data == T & !is.na(EUNISa_1_descr))),
       aes(x = biogeo, y = NDVI, fill = biogeo)) +
  facet_wrap(~ EUNISa_1_descr) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
  geom_point(aes(y = NDVI, color = biogeo),
             position = position_jitter(width = 0.15), size = 1, alpha = 0.25) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y=mean, geom="point", shape = 20, size = 1) +
  stat_summary(fun.data = function(x) data.frame(y = max(x) + 0.1,
                                                 label = length(x)),
               geom = "text", aes(label = ..label..), vjust = 0.5) +
  labs(y = "NDVI", x = "biogeo") +
  guides(fill = FALSE, color = FALSE) +
  theme_bw()
ggsave(
  here("output", "figures", "NDVI_biogeo.tiff"),
  width = 29.7, height = 21, units = "cm", dpi = 300)
```

```{r warning=FALSE}
eunis_levels %>%
  set_names() %>%  # Set names to access plots by EUNISa_1_descr
  map(~
        ggplot(data = filter(db_resurv_RS_short %>% filter(RS_data == T),
                             EUNISa_1_descr == .x),
               aes(x = biogeo, y = NDMI, fill = biogeo)) +
        geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
        geom_point(aes(y = NDMI, color = biogeo),
                   position = position_jitter(width = 0.15), size = 1, alpha = 0.25) +
        geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
        labs(y = "NDMI", x = "biogeo") +
        guides(fill = FALSE, color = FALSE) +
        labs(title = .x)
      ) %>%
  walk(print)
```

```{r warning=FALSE}
ggplot(data = filter(db_resurv_RS_short %>%
                       filter(RS_data == T & !is.na(EUNISa_1_descr))),
       aes(x = biogeo, y = NDMI, fill = biogeo)) +
  facet_wrap(~ EUNISa_1_descr) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
  geom_point(aes(y = NDMI, color = biogeo),
             position = position_jitter(width = 0.15), size = 1, alpha = 0.25) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y=mean, geom="point", shape = 20, size = 1) +
  stat_summary(fun.data = function(x) data.frame(y = max(x) + 0.1,
                                                 label = length(x)),
               geom = "text", aes(label = ..label..), vjust = 0.5) +
  labs(y = "NDMI", x = "biogeo") +
  guides(fill = FALSE, color = FALSE) +
  theme_bw()
ggsave(
  here("output", "figures", "NDMI_biogeo.tiff"),
  width = 29.7, height = 21, units = "cm", dpi = 300)
```

# PCA

```{r}
PCA_all <- prcomp(~ NDVI + NDMI, data = db_resurv_RS_short)
```

```{r}
PCA_ALP <- prcomp(~ NDVI + NDMI, 
                    data = db_resurv_RS_short %>% filter(biogeo == "ALP"))
PCA_CON <- prcomp(~ NDVI + NDMI, 
                    data = db_resurv_RS_short %>% filter(biogeo == "CON"))
PCA_ATL <- prcomp(~ NDVI + NDMI, 
                    data = db_resurv_RS_short %>% filter(biogeo == "ATL"))
PCA_PANNONIA <- prcomp(~ NDVI + NDMI, 
                    data = db_resurv_RS_short %>% filter(biogeo == "PANNONIA"))
PCA_MED <- prcomp(~ NDVI + NDMI, 
                    data = db_resurv_RS_short %>% filter(biogeo == "MED"))
PCA_BOR <- prcomp(~ NDVI + NDMI, 
                    data = db_resurv_RS_short %>% filter(biogeo == "BOR"))
```

```{r}
plot(PCA_all)
biplot(PCA_all)
```

```{r}
# Define a function to generate PCA biplots
generate_pca_biplot <- function(pca_result, title) {
  ggplot(as.data.frame(pca_result$x), aes(x = PC1, y = PC2)) +
    geom_point() +
    geom_segment(data = as.data.frame(pca_result$rotation),
                 aes(x = 0, y = 0, xend = PC1, yend = PC2),
                 arrow = arrow(length = unit(0.2, "inches")),
                 color = "blue", size = 1) +
    geom_text(data = as.data.frame(pca_result$rotation),
              aes(x = PC1, y = PC2,
                  label = rownames(as.data.frame(pca_result$rotation))),
              vjust = 1, hjust = 1, color = "blue") +
    labs(title = title, x = "PC1", y = "PC2") +
    theme_minimal()
}
```

```{r}
# Function to generate variance explained bar plot
generate_variance_plot <- function(pca_result, title) {
  ggplot(
    data.frame(PC = names(summary(pca_result)$importance[2, ]),
               Variance = summary(pca_result)$importance[2, ]), 
    aes(x = PC, y = Variance)
  ) +
    geom_bar(stat = "identity") +
    labs(x = "Principal Component", y = "Variance Explained (%)", title = title) +
    theme_minimal()
}
```


```{r}
generate_pca_biplot(PCA_all, "PCA_all")
```

```{r}
generate_variance_plot(PCA_all, "PCA_all")
```

```{r}
pca_list <- list(PCA_ALP = PCA_ALP, PCA_CON = PCA_CON, PCA_ATL = PCA_ATL,
                 PCA_PANNONIA = PCA_PANNONIA, PCA_MED = PCA_MED,
                 PCA_BOR = PCA_BOR)
```

```{r}
pca_biplots <- lapply(names(pca_list), 
                      function(pca_name) {
                        generate_pca_biplot(pca_list[[pca_name]],
                                            pca_name)
                        })
grid.arrange(grobs = pca_biplots, ncol = 3)
```

```{r}
variance_plots <- lapply(names(pca_list), function(pca_name) {
  generate_variance_plot(pca_list[[pca_name]], pca_name)
})
grid.arrange(grobs = variance_plots, ncol = 3)
```

# Different years: Forests and grasslands

```{r}
db_resurv_RS_short %>% filter(EUNISa_1 == "T" & RS_data == T) %>%
  ggplot(aes(x = factor(year), y = NDVI, fill = biogeo)) +
  geom_boxplot(outlier.shape = NA)
db_resurv_RS_short %>% filter(EUNISa_1 == "R" & RS_data == T) %>%
  ggplot(aes(x = factor(year), y = NDVI, fill = biogeo)) +
  geom_boxplot(outlier.shape = NA)
```

# Save as shapefile

```{r}
# st_write(db_resurv_RS_short %>% filter(!is.na(Lon_updated)) %>%
#   st_as_sf(coords = c("Lon_updated", "Lat_updated"), crs = 4326),
#   here("..", "..", "..", "GIS", "shapefiles", "db_resurv_RS_short.shp"))
```

# Keep only "good" points

How many points with data about Location uncertainty?

```{r}
db_resurv_RS_short %>%
  mutate(loc_uncertainty = ifelse(is.na(`Location uncertainty (m)`), "N", "Y")) %>%
  ggplot(aes(x = loc_uncertainty)) + geom_bar()
```

Distribution of location uncertainty when it is available (i.e. no NA):

```{r}
db_resurv_RS_short %>%
  filter(!is.na(`Location uncertainty (m)`)) %>%
  ggplot(aes(x = `Location uncertainty (m)`)) + geom_histogram()
```

Distribution of location uncertainty when it is < 200 m:

```{r}
db_resurv_RS_short %>%
  filter(`Location uncertainty (m)` < 200) %>%
  ggplot(aes(x = `Location uncertainty (m)`)) + geom_histogram()
```

Distribution of location uncertainty when it is < 20 m:

```{r}
db_resurv_RS_short %>%
  filter(`Location uncertainty (m)` < 20) %>%
  ggplot(aes(x = `Location uncertainty (m)`)) + geom_histogram()
```

## Distributions with uncertainty < 20 m

```{r warning=FALSE}
ggplot(data = db_resurv_RS_short %>%
         filter(`Location uncertainty (m)` < 20) %>%
         filter(RS_data == T & !is.na(EUNISa_1_descr)),
       aes(x = EUNISa_1_descr, y = NDVI, fill = EUNISa_1_descr)) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
  geom_point(aes(y = NDVI, color = EUNISa_1_descr),
             position = position_jitter(width = 0.15), size = 1, alpha = 0.25) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y=mean, geom="point", shape = 20, size = 1) +
  stat_summary(fun.data = function(x) data.frame(y = max(x) + 0.1,
                                                 label = length(x)),
               geom = "text", aes(label = ..label..), vjust = 0.5) +
  labs(y = "NDVI", x = "EUNISa_1_descr") +
  guides(fill = FALSE, color = FALSE) +
  theme_bw() + coord_flip()
ggsave(
  here("output", "figures", "NDVI_20.tiff"),
  width = 21, height = 29.7, units = "cm", dpi = 300)
ggplot(data = db_resurv_RS_short %>%
         filter(`Location uncertainty (m)` < 20) %>%
         filter(RS_data == T & !is.na(EUNISa_1_descr)),
       aes(x = EUNISa_1_descr, y = NDMI, fill = EUNISa_1_descr)) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
  geom_point(aes(y = NDMI, color = EUNISa_1_descr),
             position = position_jitter(width = 0.15), size = 1, alpha = 0.25) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y=mean, geom="point", shape = 20, size = 1) +
  stat_summary(fun.data = function(x) data.frame(y = max(x) + 0.1,
                                                 label = length(x)),
               geom = "text", aes(label = ..label..), vjust = 0.5) +
  labs(y = "NDMI", x = "EUNISa_1_descr") +
  guides(fill = FALSE, color = FALSE) +
  theme_bw() + coord_flip()
ggsave(
  here("output", "figures", "NDMI_20.tiff"),
  width = 21, height = 29.7, units = "cm", dpi = 300)
```

## Distributions with uncertainty < 5 m

```{r warning=FALSE}
ggplot(data = db_resurv_RS_short %>%
         filter(`Location uncertainty (m)` < 5) %>%
         filter(RS_data == T & !is.na(EUNISa_1_descr)),
       aes(x = EUNISa_1_descr, y = NDVI, fill = EUNISa_1_descr)) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
  geom_point(aes(y = NDVI, color = EUNISa_1_descr),
             position = position_jitter(width = 0.15), size = 1, alpha = 0.25) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y=mean, geom="point", shape = 20, size = 1) +
  stat_summary(fun.data = function(x) data.frame(y = max(x) + 0.1,
                                                 label = length(x)),
               geom = "text", aes(label = ..label..), vjust = 0.5) +
  labs(y = "NDVI", x = "EUNISa_1_descr") +
  guides(fill = FALSE, color = FALSE) +
  theme_bw() + coord_flip()
ggsave(
  here("output", "figures", "NDVI_5.tiff"),
  width = 21, height = 29.7, units = "cm", dpi = 300)
ggplot(data = db_resurv_RS_short %>%
         filter(`Location uncertainty (m)` < 5) %>%
         filter(RS_data == T & !is.na(EUNISa_1_descr)),
       aes(x = EUNISa_1_descr, y = NDMI, fill = EUNISa_1_descr)) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
  geom_point(aes(y = NDMI, color = EUNISa_1_descr),
             position = position_jitter(width = 0.15), size = 1, alpha = 0.25) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y=mean, geom="point", shape = 20, size = 1) +
  stat_summary(fun.data = function(x) data.frame(y = max(x) + 0.1,
                                                 label = length(x)),
               geom = "text", aes(label = ..label..), vjust = 0.5) +
  labs(y = "NDMI", x = "EUNISa_1_descr") +
  guides(fill = FALSE, color = FALSE) +
  theme_bw() + coord_flip()
ggsave(
  here("output", "figures", "NDMI_5.tiff"),
  width = 21, height = 29.7, units = "cm", dpi = 300)
```

## Distributions with Location method = "Location with differential GPS"

```{r warning=FALSE}
ggplot(data = db_resurv_RS_short %>%
         filter(`Location method` == "Location with differential GPS") %>%
         filter(RS_data == T & !is.na(EUNISa_1_descr)),
       aes(x = EUNISa_1_descr, y = NDVI, fill = EUNISa_1_descr)) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
  geom_point(aes(y = NDVI, color = EUNISa_1_descr),
             position = position_jitter(width = 0.15), size = 1, alpha = 0.25) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y=mean, geom="point", shape = 20, size = 1) +
  stat_summary(fun.data = function(x) data.frame(y = max(x) + 0.1,
                                                 label = length(x)),
               geom = "text", aes(label = ..label..), vjust = 0.5) +
  labs(y = "NDVI", x = "EUNISa_1_descr") +
  guides(fill = FALSE, color = FALSE) +
  theme_bw() + coord_flip()
ggsave(
  here("output", "figures", "NDVI_differential_GPS.tiff"),
  width = 21, height = 29.7, units = "cm", dpi = 300)
ggplot(data = db_resurv_RS_short %>%
         filter(`Location method` == "Location with differential GPS") %>%
         filter(RS_data == T & !is.na(EUNISa_1_descr)),
       aes(x = EUNISa_1_descr, y = NDMI, fill = EUNISa_1_descr)) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
  geom_point(aes(y = NDMI, color = EUNISa_1_descr),
             position = position_jitter(width = 0.15), size = 1, alpha = 0.25) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y=mean, geom="point", shape = 20, size = 1) +
  stat_summary(fun.data = function(x) data.frame(y = max(x) + 0.1,
                                                 label = length(x)),
               geom = "text", aes(label = ..label..), vjust = 0.5) +
  labs(y = "NDMI", x = "EUNISa_1_descr") +
  guides(fill = FALSE, color = FALSE) +
  theme_bw() + coord_flip()
ggsave(
  here("output", "figures", "NDMI_differential_GPS.tiff"),
  width = 21, height = 29.7, units = "cm", dpi = 300)
```

```{r}
ggplot(data = filter(db_resurv_RS_short %>%
                       filter(`Location uncertainty (m)` < 50) %>%
                       filter(RS_data == T & !is.na(EUNISa_1_descr))),
       aes(x = biogeo, y = NDVI, fill = biogeo)) +
  facet_wrap(~ EUNISa_1_descr) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
  geom_point(aes(y = NDVI, color = biogeo),
             position = position_jitter(width = 0.15), size = 1, alpha = 0.25) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y=mean, geom="point", shape = 20, size = 1) +
  stat_summary(fun.data = function(x) data.frame(y = max(x) + 0.1,
                                                 label = length(x)),
               geom = "text", aes(label = ..label..), vjust = 0.5) +
  labs(y = "NDVI", x = "biogeo") +
  guides(fill = FALSE, color = FALSE) +
  theme_bw()
```

# Save csv for SONATA workshop

Points with uncertainty < 50 m, last resurvey for each plot.

```{r}
points_SONATA <- db_resurv_RS_short %>%
  filter(`Location uncertainty (m)` < 50 & RS_data == T &
           !is.na(EUNISa_1_descr) & Lon_diff == F & Lat_diff == F) %>%
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`) %>%
  mutate(max_year = max(year)) %>%
  filter(year == max_year) %>%
  # In some cases there is more than one row per group,
  # i.e. more than one row for the last year
  slice_sample(n = 1) %>%
  ungroup() %>%
  select(-max_year, -year_RS, -source, -Lon_RS, -Lat_RS, -RS_data, -year_diff,
         -Lon_diff, -Lat_diff)
```

Plot points in a map:

```{r}
# Load world boundaries
world <- ne_countries(scale = "medium", returnclass = "sf")

# Calculate the extent of the points
extent <- points_SONATA %>%
  summarise(
    lon_min = min(Lon_updated, na.rm = TRUE),
    lon_max = max(Lon_updated, na.rm = TRUE),
    lat_min = min(Lat_updated, na.rm = TRUE),
    lat_max = max(Lat_updated, na.rm = TRUE)
  )

# Add padding to the extent (adjust as needed)
padding <- 0.5  # Adjust padding to your preference
x_limits <- c(extent$lon_min - padding, extent$lon_max + padding)
y_limits <- c(extent$lat_min - padding, extent$lat_max + padding)

# Compute centroids for labeling
world_centroids <- world %>%
  mutate(centroid = st_centroid(geometry)) %>%
  mutate(lon = st_coordinates(centroid)[, 1],
         lat = st_coordinates(centroid)[, 2])

# Create the zoomed map
ggplot() +
  geom_sf(data = world, fill = "lightblue", color = "gray") +
  geom_point(data = points_SONATA %>%
               filter(`Location uncertainty (m)` < 20),
             aes(x = Lon_updated, y = Lat_updated, color = Country),
             size = 2) +
  geom_text(data = world_centroids, 
            aes(x = lon, y = lat, label = sovereignt),
            size = 3, color = "black") +  # Adjust size and color as needed
  coord_sf(xlim = x_limits, ylim = y_limits) +
  theme_minimal()
```

Save csv:

```{r}
write_tsv(points_SONATA,here("data", "clean","points_SONATA.csv"))
```

# Puntos Cordillera

## Read and merge RS data

Read sheets from Excel file sent by Gonzalo (and edited by me to create new sheets). Rename identifiers (Codigo and Plot_ID to id to be able to bind rows of both datasets).

```{r}
cordi1 <- read_excel(
  here("data", "edited","NDVI_NDMI_Pastizales_Bosques 2_AV.xlsx"),
  sheet = "pastizal_AV") %>%
  rename(id = "Releve_num") %>% mutate(id = as.character(id))
cordi2 <- read_excel(
  here("data", "edited","NDVI_NDMI_Pastizales_Bosques_AV.xlsx"), 
  sheet = "bosque_AV") %>%
  rename(id = "Plot_ID")
```

Merge both datasets and get max NDVI and NDMI for each point:

```{r}
cordi <- bind_rows(cordi1, cordi2) %>% select(-date, -mes) %>%
  group_by(id) %>%
  summarise(max_NDVI = max(NDVI, na.rm = TRUE),
            max_NDMI = max(NDMI, na.rm = TRUE))
```

## Read and merge shapefiles to get vegetation / habitat info

```{r}
alpine_grasslands <- st_read(
  here("..", "..", "..", "GIS", "shapefiles", "EVEREST",
       "AlpineGrasslands_WGS84.shp")) %>%
  rename(id = "Releve_num", hab = "Hábitat") %>% select (id, hab) %>% as_tibble() %>%
  mutate(id = as.character(id))
vegetation_types <- st_read(
  here("..", "..", "..", "GIS", "shapefiles", "EVEREST",
       "VegetationTypes_WGS84.shp")) %>%
  rename(id = "Plot_ID", hab = "Vegetation") %>% select(id, hab) %>% as_tibble()
```

```{r}
habitats <- bind_rows(alpine_grasslands, vegetation_types)
```


## Merge RS and habitat data

```{r}
cordi <- habitats %>%
  left_join(cordi)
```

See the different info on hab:

```{r}
cordi %>% distinct(hab)
```

Assign to EUNIS level 1:

```{r}
cordi <- cordi %>%
  mutate(EUNIS_1 = case_when(hab = str_detect(hab, "Pastizal|Cervunal|grassland|meadow") ~ "R",
                             hab = str_detect(hab, "forest") ~ "T",
                             hab = str_detect(hab, "Scrub|scrub|Shrubland|shrubland|shrub|Heathland") ~ "S",
                             hab = str_detect(hab, "Suelo") ~ "U",
                             TRUE ~ NA_character_))
```

## Plots NDVI and NDMI

```{r}
ggplot(data = cordi %>% filter(!is.na(max_NDVI)),
       aes(x = EUNIS_1, y = max_NDVI, fill = EUNIS_1)) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
  geom_point(aes(y = max_NDVI, color = EUNIS_1),
             position = position_jitter(width = 0.15), size = 3,
             alpha = 0.5) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y=mean, geom="point", shape = 20, size = 3) +
  stat_summary(fun.data = function(x) data.frame(y = max(x) + 0.1,
                                                 label = length(x)),
               geom = "text", aes(label = ..label..), vjust = 0.5) +
  guides(fill = FALSE, color = FALSE) + theme_bw() + coord_flip()
ggsave(here("output", "figures", "cordi_NDVI.tiff"),
       width = 21, height = 29.7, units = "cm", dpi = 300)
ggplot(data = cordi %>% filter(!is.na(max_NDMI)),
       aes(x = EUNIS_1, y = max_NDMI, fill = EUNIS_1)) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +
  geom_point(aes(y = max_NDMI, color = EUNIS_1),
             position = position_jitter(width = 0.15), size = 3,
             alpha = 0.5) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y=mean, geom="point", shape = 20, size = 3) +
  stat_summary(fun.data = function(x) data.frame(y = max(x) + 0.1,
                                                 label = length(x)),
               geom = "text", aes(label = ..label..), vjust = 0.5) +
  guides(fill = FALSE, color = FALSE) + theme_bw() + coord_flip()
ggsave(here("output", "figures", "cordi_NDMI.tiff"),
       width = 21, height = 29.7, units = "cm", dpi = 300)
```


# Session info

```{r}
sessionInfo()
```

