---
title: "Script to validate points in ReSurvey database using RS data (S2 + SatEmb + Landsat)"
subtitle: "K-fold cross validation of best models"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

This R script is used to validate the points in the ReSurvey database using either RS indicators from S2 or Landsat (indices + phenology + canopy height), or bands from the Google Satellite Embedding dataset: https://developers.google.com/earth-engine/datasets/catalog/GOOGLE_SATELLITE_EMBEDDING_V1_ANNUAL#description
Here, we perform a K-fold cross validation of the best models. 

# Load libraries

```{r}
library(tidyverse)
library(here)
library(parallel)
library(doParallel)
library(moreparty)
library(caret)
library(randomForest)
library(pROC)
library(rlang)
library(stringr)
library(beepr)
library(permimp)
library(purrr)
library(MLmetrics)
```

# Load functions

```{r}
knitr::purl("functions_20260122.Rmd", output = "functions_20260122.R")
source("functions_20260122.R")
```

# Load previously created objects (CHANGE TO BEST)

```{r}
# Load best model Landsat (and prediction) - Accuracy = 0.8212
model_Landsat <- get(load(here("objects", "RF_1500_20260126",
                               "rf_months_Landsat_refin_35_10_2x_model.Rdata")))
prediction_Landsat <- get(load(here("objects", "RF_1500_20260126", 
                                    "rf_months_Landsat_refin_35_10_2x_prediction.Rdata")))
# Load filtered data Landsat
filtered_data_months_Landsat_with_probs_2x <- get(load(here("objects", "RF_1500_20260126",
                                       "filtered_data_months_Landsat_with_probs_2x.Rdata")))
# Load best model S2 or SatEmb (and prediction) - Accuracy = 0.8218
model_SatEmb <- get(load(here("objects", "RF_1500_20260126", 
                              "rf_SatEmb_refin_35_10_2x_model.Rdata")))
prediction_SatEmb <- get (load(here("objects", "RF_1500_20260126",
                                    "rf_SatEmb_refin_35_10_2x_prediction.Rdata")))
# Load filtered data S2 or SatEmb
filtered_data_SatEmb_with_probs_2x <- get(load(here("objects", "RF_1500_20260126",
                                       "filtered_data_SatEmb_with_probs_2x.Rdata")))
```

# Read data

```{r}
data_validation_SatEmb <- read_tsv(here(
  "data", "clean","final_RS_data_SatEmb_20251010.csv")) %>%
  rename(EUNISa_1_descr = EUNIS_1_d)
data_validation_Landsat <- read_tsv(here(
  "data", "clean","final_RS_data_bands_Landsat_all_20251021.csv"))
```

No parsing issues!

```{r}
nrow(data_validation_Landsat)
nrow(data_validation_SatEmb)
```

# Filter out manipulated observations

## Read list of PlotObservationIDs to keep

```{r}
PlotObservationID_list <- read_csv(
  here("data", "clean","PlotObservationIDs_filtered_QRST_20260122.csv"))
```

## Filter out data to keep only those PlotObservationIDs

```{r}
data_validation_SatEmb <- data_validation_SatEmb %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
data_validation_Landsat <- data_validation_Landsat %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
```

```{r}
nrow(data_validation_Landsat)
nrow(data_validation_SatEmb)
```

# Define lists of predictor vars

```{r}
# Excluding EVI because many values outside of range -1, 1
vars_RF_months_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, and SAVI between march and oct
  "NDVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI and SAVI at march, pos and oct
  "NDVI_avg_value_03", "NDVI_pos_value", "NDVI_avg_value_10",
  "SAVI_avg_value_03", "SAVI_pos_value", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "SAVI_diff_pos_march_value",
  "NDVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "SAVI_diff_pos_oct_value",
  "NDVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy",
  # Canopy height
  "canopy_height")
```

```{r}
vars_RF_SatEmb <- colnames(
  data_validation_SatEmb)[startsWith(colnames(data_validation_SatEmb), "A")]
```

# Get filtered data

## Landsat

```{r}
filter_data_Landsat <- function(data, variables) {
  data %>%
    mutate(EUNISa_1 = as.factor(EUNISa_1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    # With Landsat data, many observations removed because of high/low EVI values
    # In Landsat_extraction, values of EVI outside of expected range were set as NA,
    # so better not to use EVI here
    # dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Select only variables needed
    dplyr::select(PlotObservationID, EUNISa_1, all_of(variables), Lctnmth,
           valid_1) %>%
    # Keep only rows with differences > 0 in variables
    filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0)) %>%
    # Select only GPS points
    dplyr::filter(Lctnmth == "Location with GPS" | 
                    Lctnmth == "Location with differential GPS")
}
```

```{r}
filtered_data_Landsat <- filter_data_Landsat(
  data_validation_Landsat %>%
    rough_validation_Landsat(), # Rough validation based on CH
  vars_RF_months_Landsat)

prob_cols_Landsat <- names(filtered_data_months_Landsat_with_probs_2x)[grepl("^prob_", names(filtered_data_months_Landsat_with_probs_2x))]

df_aux_Landsat_2x <- calculate_df_aux(filtered_data_months_Landsat_with_probs_2x,
                                  prob_cols_Landsat)

df_final_20_10_months_Landsat_2x <- df_aux_Landsat_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)

filtered_data_Landsat_cv <- filtered_data_Landsat %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_months_Landsat_2x$PlotObservationID)
```

## SatEmb

```{r}
filter_data_SatEmb <- function(data, variables) {
  data %>%
    mutate(EUNISa_1 = as.factor(EUNISa_1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Select only variables needed
    dplyr::select(PlotObservationID, EUNISa_1, all_of(variables), 
           Lctnmth, valid_1) %>%
    # Select only GPS points
    dplyr::filter(Lctnmth == "Location with GPS" |
                    Lctnmth == "Location with differential GPS")
  }
```

```{r}
filtered_data_SatEmb <- filter_data_SatEmb(
  data_validation_SatEmb %>%
    rough_validation_SatEmb(), # Rough validation based on CH
  vars_RF_SatEmb)

prob_cols_SatEmb <- names(filtered_data_SatEmb_with_probs_2x)[grepl("^prob_", names(filtered_data_SatEmb_with_probs_2x))]

df_aux_SatEmb_2x <- calculate_df_aux(filtered_data_SatEmb_with_probs_2x,
                                  prob_cols_SatEmb)

df_final_20_10_SatEmb_2x <- df_aux_SatEmb_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)

filtered_data_SatEmb_cv <- filtered_data_SatEmb %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_SatEmb_2x$PlotObservationID)
```

# k-fold cross-validation (4 folds)

```{r}
set.seed(123)
ctrl <- trainControl(
  method = "cv",
  number = 4,
  allowParallel = FALSE,
  savePredictions = "final",
  classProbs = TRUE,     # set TRUE if you also want probabilities (for ROC)
  summaryFunction = multiClassSummary,  # optional, for multi-class summaries
  verboseIter = TRUE      # <--- this gives progress output
)
```

## Landsat

```{r}
cv_Landsat <- caret::train(
  x = filtered_data_Landsat_cv[, vars_RF_months_Landsat],
  y = filtered_data_Landsat_cv $EUNISa_1,
  # Using cforest, not fastcforest
  method = "cforest",
  trControl = ctrl,
  tuneGrid = expand.grid(
    mtry = round(sqrt(length(vars_RF_months_Landsat)))
  ),
  controls = party::cforest_control(
    ntree = 1500
  )
)
```

### Save

```{r}
save(cv_Landsat, file = "objects/RF_1500_cv/cv_Landsat.Rdata")
```

## SatEmb

```{r}
cv_SatEmb <- caret::train(
  x = filtered_data_SatEmb_cv[, vars_RF_SatEmb],
  y = filtered_data_SatEmb_cv $EUNISa_1,
  # Using cforest, not fastcforest
  method = "cforest",
  trControl = ctrl,
  tuneGrid = expand.grid(
    mtry = round(sqrt(length(vars_RF_SatEmb)))
  ),
  controls = party::cforest_control(
    ntree = 1500
  )
)
```

### Save

```{r}
save(cv_SatEmb, file = "objects/RF_1500_cv/cv_SatEmb.Rdata")
```


ADD: Have a look at results

### Get an average ROC curve (macro‑average), varimp...

```{r}

```


