---
title: "Script to validate points in ReSurvey database using RS data (S2 + SatEmb + Landsat)"
subtitle: "Fitting Random Forest models with ntree = 1500, excluding manipulated observations"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

This R script is used to validate the points in the ReSurvey database using either RS indicators from S2 or Landsat (indices + phenology + canopy height), or bands from the Google Satellite Embedding dataset: https://developers.google.com/earth-engine/datasets/catalog/GOOGLE_SATELLITE_EMBEDDING_V1_ANNUAL#description
Here, we filter out manipulated observations to avoid using them in analyses.

# Load libraries

```{r}
library(tidyverse)
library(here)
library(parallel)
library(doParallel)
library(moreparty)
library(caret)
library(randomForest)
library(pROC)
library(rlang)
library(stringr)
library(beepr)
library(permimp)
library(purrr)
library(gghalves)
library(patchwork)
library(scales)
```

# Load functions

```{r}
knitr::purl("functions_20260122.Rmd", output = "functions_20260122.R")
source("functions_20260122.R")
```

# Load previously created objects

```{r}
# Define the folder path
folder_path <- here("objects", "RF_1500_20260128")

# List all .RData or .rda files in the folder
rdata_files <- list.files(folder_path, full.names = TRUE)

# Load each file
lapply(rdata_files, load, envir = .GlobalEnv)
```

# Read data

```{r}
data_validation_S2 <- read_tsv(here(
  "data", "clean","final_RS_data_bands_S2_all_20251021.csv"))
data_validation_SatEmb <- read_tsv(here(
  "data", "clean","final_RS_data_SatEmb_20251010.csv")) %>%
  rename(EUNISa_1_descr = EUNIS_1_d)
data_validation_Landsat <- read_tsv(here(
  "data", "clean","final_RS_data_bands_Landsat_all_20251021.csv"))
```

No parsing issues!

```{r}
nrow(data_validation_S2)
nrow(data_validation_Landsat)
nrow(data_validation_SatEmb)
```

# Filter out manipulated observations

## Read list of PlotObservationIDs to keep

```{r}
PlotObservationID_list <- read_csv(
  here("data", "clean","PlotObservationIDs_filtered_QRST_20260122.csv"))
```

## Filter out data to keep only those PlotObservationIDs

```{r}
data_validation_S2 <- data_validation_S2 %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
data_validation_SatEmb <- data_validation_SatEmb %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
data_validation_Landsat <- data_validation_Landsat %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
```

```{r}
nrow(data_validation_S2)
nrow(data_validation_Landsat)
nrow(data_validation_SatEmb)
```

# Define lists of predictor vars

## S2 and Landsat

### Moments with slope method

Not using threshold method cause it gave worse results earlier.

```{r}
vars_RF_slope <- c(
  # Min values of all indices
  "NDVI_min", "EVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, EVI and SAVI
  "NDVI_auc_slope", "EVI_auc_slope", "SAVI_auc_slope",
  # Values of NDVI, EVI and SAVI at sos, pos and eos
  "NDVI_sos_slope_value", "NDVI_pos_value", "NDVI_eos_slope_value",
  "EVI_sos_slope_value", "EVI_pos_value", "EVI_eos_slope_value", 
  "SAVI_sos_slope_value", "SAVI_pos_value", "SAVI_eos_slope_value",
  # Differences pos-sos in value and doy
  "NDVI_diff_pos_sos_slope_value", "EVI_diff_pos_sos_slope_value",
  "SAVI_diff_pos_sos_slope_value","NDVI_diff_pos_sos_slope_doy", 
  "EVI_diff_pos_sos_slope_doy", "SAVI_diff_pos_sos_slope_doy",
  # Differences pos-eos in value and doy
  "NDVI_diff_pos_eos_slope_value", "EVI_diff_pos_eos_slope_value",
  "SAVI_diff_pos_eos_slope_value", "NDVI_diff_pos_eos_slope_doy", 
  "EVI_diff_pos_eos_slope_doy", "SAVI_diff_pos_eos_slope_doy",
  # Growing season duration
  "NDVI_slope_gsd", "EVI_slope_gsd", "SAVI_slope_gsd",
  # Canopy height
  "canopy_height")
```

```{r}
# Excluding EVI because many values outside of range -1, 1
vars_RF_slope_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI and SAVI
  "NDVI_auc_slope", "SAVI_auc_slope",
  # Values of NDVI and SAVI at sos, pos and eos
  "NDVI_sos_slope_value", "NDVI_pos_value", "NDVI_eos_slope_value",
  "SAVI_sos_slope_value", "SAVI_pos_value", "SAVI_eos_slope_value",
  # Differences pos-sos in value and doy
  "NDVI_diff_pos_sos_slope_value","SAVI_diff_pos_sos_slope_value",
  "NDVI_diff_pos_sos_slope_doy", "SAVI_diff_pos_sos_slope_doy",
  # Differences pos-eos in value and doy
  "NDVI_diff_pos_eos_slope_value", "SAVI_diff_pos_eos_slope_value",
  "NDVI_diff_pos_eos_slope_doy", "SAVI_diff_pos_eos_slope_doy",
  # Growing season duration
  "NDVI_slope_gsd", "SAVI_slope_gsd",
  # Canopy height
  "canopy_height")
```

### Months method

```{r}
vars_RF_months <- c(
  # Min values of all indices
  "NDVI_min", "EVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, EVI and SAVI between march and oct
  "NDVI_auc_mar_oct", "EVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI, EVI and SAVI at march, pos and oct
  "NDVI_avg_value_03", "NDVI_pos_value", "NDVI_avg_value_10",
  "EVI_avg_value_03", "EVI_pos_value", "EVI_avg_value_10", 
  "SAVI_avg_value_03", "SAVI_pos_value", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "EVI_diff_pos_march_value",
  "SAVI_diff_pos_march_value","NDVI_diff_pos_march_doy", 
  "EVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "EVI_diff_pos_oct_value",
  "SAVI_diff_pos_oct_value", "NDVI_diff_pos_oct_doy", 
  "EVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy",
  # Canopy height
  "canopy_height")
```

```{r}
# Excluding EVI because many values outside of range -1, 1
vars_RF_months_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, and SAVI between march and oct
  "NDVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI and SAVI at march, pos and oct
  "NDVI_avg_value_03", "NDVI_pos_value", "NDVI_avg_value_10",
  "SAVI_avg_value_03", "SAVI_pos_value", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "SAVI_diff_pos_march_value",
  "NDVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "SAVI_diff_pos_oct_value",
  "NDVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy",
  # Canopy height
  "canopy_height")
```

### Both

Slope + months (Borja's suggestion).

```{r}
vars_RF_both <- c(
  # Min values of all indices
  "NDVI_min", "EVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, EVI and SAVI
  "NDVI_auc_slope", "EVI_auc_slope", "SAVI_auc_slope",
  # AUC of NDVI, EVI and SAVI between march and oct
  "NDVI_auc_mar_oct", "EVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI, EVI and SAVI at sos, pos and eos
  "NDVI_sos_slope_value", "NDVI_pos_value", "NDVI_eos_slope_value",
  "EVI_sos_slope_value", "EVI_pos_value", "EVI_eos_slope_value", 
  "SAVI_sos_slope_value", "SAVI_pos_value", "SAVI_eos_slope_value",
  # Differences pos-sos in value and doy
  "NDVI_diff_pos_sos_slope_value", "EVI_diff_pos_sos_slope_value",
  "SAVI_diff_pos_sos_slope_value","NDVI_diff_pos_sos_slope_doy", 
  "EVI_diff_pos_sos_slope_doy", "SAVI_diff_pos_sos_slope_doy",
  # Differences pos-eos in value and doy
  "NDVI_diff_pos_eos_slope_value", "EVI_diff_pos_eos_slope_value",
  "SAVI_diff_pos_eos_slope_value", "NDVI_diff_pos_eos_slope_doy", 
  "EVI_diff_pos_eos_slope_doy", "SAVI_diff_pos_eos_slope_doy",
  # Growing season duration
  "NDVI_slope_gsd", "EVI_slope_gsd", "SAVI_slope_gsd",
  # Values of NDVI, EVI and SAVI at march and oct
  "NDVI_avg_value_03", "NDVI_avg_value_10",
  "EVI_avg_value_03", "EVI_avg_value_10", 
  "SAVI_avg_value_03", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "EVI_diff_pos_march_value",
  "SAVI_diff_pos_march_value","NDVI_diff_pos_march_doy", 
  "EVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "EVI_diff_pos_oct_value",
  "SAVI_diff_pos_oct_value", "NDVI_diff_pos_oct_doy", 
  "EVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy",
  # Canopy height
  "canopy_height")
```

```{r}
# Excluding EVI because many values outside of range -1, 1
vars_RF_both_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI and SAVI
  "NDVI_auc_slope", "SAVI_auc_slope",
  # AUC of NDVI, and SAVI between march and oct
  "NDVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI and SAVI at sos, pos and eos
  "NDVI_sos_slope_value", "NDVI_pos_value", "NDVI_eos_slope_value",
  "SAVI_sos_slope_value", "SAVI_pos_value", "SAVI_eos_slope_value",
  # Differences pos-sos in value and doy
  "NDVI_diff_pos_sos_slope_value","SAVI_diff_pos_sos_slope_value",
  "NDVI_diff_pos_sos_slope_doy", "SAVI_diff_pos_sos_slope_doy",
  # Differences pos-eos in value and doy
  "NDVI_diff_pos_eos_slope_value", "SAVI_diff_pos_eos_slope_value",
  "NDVI_diff_pos_eos_slope_doy", "SAVI_diff_pos_eos_slope_doy",
  # Growing season duration
  "NDVI_slope_gsd", "SAVI_slope_gsd",
  # Values of NDVI and SAVI at march and oct
  "NDVI_avg_value_03", "NDVI_avg_value_10",
  "SAVI_avg_value_03", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "SAVI_diff_pos_march_value",
  "NDVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "SAVI_diff_pos_oct_value",
  "NDVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy",
  # Canopy height
  "canopy_height")
```

## SatEmb

```{r}
vars_RF_SatEmb <- colnames(
  data_validation_SatEmb)[startsWith(colnames(data_validation_SatEmb), "A")]
vars_RF_SatEmb_CH <- c(vars_RF_SatEmb, "canopy_height")
```

# Filter data to use in RF models

## Functions

```{r}
filter_data_S2 <- function(data, variables) {
  data %>%
    mutate(EUNISa_1 = as.factor(EUNISa_1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Select only variables needed
    dplyr::select(PlotObservationID, EUNISa_1, all_of(variables), Lctnmth,
           valid_1) %>%
    # Keep only rows with differences > 0 in variables
    filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0)) %>%
    # Select only GPS points
    dplyr::filter(Lctnmth == "Location with GPS" | 
                    Lctnmth == "Location with differential GPS")
}
```

```{r}
filter_data_Landsat <- function(data, variables) {
  data %>%
    mutate(EUNISa_1 = as.factor(EUNISa_1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    # With Landsat data, many observations removed because of high/low EVI values
    # In Landsat_extraction, values of EVI outside of expected range were set as NA,
    # so better not to use EVI here
    # dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Select only variables needed
    dplyr::select(PlotObservationID, EUNISa_1, all_of(variables), Lctnmth,
           valid_1) %>%
    # Keep only rows with differences > 0 in variables
    filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0)) %>%
    # Select only GPS points
    dplyr::filter(Lctnmth == "Location with GPS" | 
                    Lctnmth == "Location with differential GPS")
}
```

```{r}
filter_data_SatEmb <- function(data, variables) {
  data %>%
    mutate(EUNISa_1 = as.factor(EUNISa_1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Select only variables needed
    dplyr::select(PlotObservationID, EUNISa_1, all_of(variables), 
           Lctnmth, valid_1) %>%
    # Select only GPS points
    dplyr::filter(Lctnmth == "Location with GPS" |
                    Lctnmth == "Location with differential GPS")
  }
```

## Filtered data

Select only GPS points that have passed the rough validation in each case.

For S2 and Landsat, use only slope and months methods, as threshold method has many missing values.

```{r}
# S2
filtered_data_slope_S2 <- data_validation_S2 %>%
  # Rough validation based on CH and NDWI
  rough_validation_S2()  %>%
  # Data filtering
  filter_data_S2(vars_RF_slope) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value")

filtered_data_months_S2 <- data_validation_S2 %>%
  # Rough validation based on CH and NDWI
  rough_validation_S2() %>% 
  # Data filtering
  filter_data_S2(vars_RF_months) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value")

filtered_data_both_S2 <- data_validation_S2 %>%
  # Rough validation based on CH and NDWI
  rough_validation_S2() %>% 
  # Data filtering
  filter_data_S2(vars_RF_both) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value")

# Landsat
filtered_data_slope_Landsat <- data_validation_Landsat %>%
  # Rough validation based on NDWI
  rough_validation_Landsat() %>% 
  # Data filtering
  filter_data_Landsat(vars_RF_slope_Landsat) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value")

filtered_data_months_Landsat <- data_validation_Landsat %>%
  # Rough validation based on NDWI
  rough_validation_Landsat() %>% 
  # Data filtering
  filter_data_Landsat(vars_RF_months_Landsat) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value")

filtered_data_both_Landsat <- data_validation_Landsat %>%
  # Rough validation based on NDWI
  rough_validation_Landsat() %>% 
  # Data filtering
  filter_data_Landsat(vars_RF_both_Landsat) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value")

# SatEmb
filtered_data_SatEmb <- data_validation_SatEmb %>%
  # Rough validation based on CH
  rough_validation_SatEmb() %>% 
  # Join with S2 data to get NDWI_max info
  left_join(
    data_validation_S2 %>% select(PlotObservationID, NDWI_max)
  ) %>%
  # Rough validation based on NDWI (from S2)
  # Remove those with NDWI_max > 0.3, keep if NDWI_max is NA
  filter(NDWI_max <= 0.3 | is.na(NDWI_max)) %>%
  # Remove unneeded var
  select(-NDWI_max) %>%
  # Data filtering
  filter_data_SatEmb(vars_RF_SatEmb) %>% 
  # Join with S2 data to get NDVI_pos_value info
  left_join(
    data_validation_S2 %>% select(PlotObservationID, NDVI_pos_value)
  ) %>%
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value") %>%
  # Remove unneeded var
  select(-NDVI_pos_value) 
filtered_data_SatEmb_CH <- data_validation_SatEmb %>%
  # Rough validation based on CH
  rough_validation_SatEmb() %>% 
  # Join with S2 data to get NDWI_max info
  left_join(
    data_validation_S2 %>% select(PlotObservationID, NDWI_max)
  ) %>%
  # Rough validation based on NDWI (from S2)
  # Remove those with NDWI_max > 0.3, keep if NDWI_max is NA
  filter(NDWI_max <= 0.3 | is.na(NDWI_max)) %>%
  # Remove unneeded var
  select(-NDWI_max) %>%
  # Data filtering
  filter_data_SatEmb(vars_RF_SatEmb_CH) %>% 
  # Join with S2 data to get NDVI_pos_value info
  left_join(
    data_validation_S2 %>% select(PlotObservationID, NDVI_pos_value)
  ) %>%
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value") %>%
  # Remove unneeded vars
  select(-NDVI_pos_value) 
```

## (Appendix 2) Plots distribution max NDVI

```{r}
plot_perc_S2 <- data_validation_S2 %>%
  # Rough validation based on CH and NDWI
  rough_validation_S2() %>% 
  # Data filtering
  filter_data_S2(vars_RF_months) %>% 
  distr_plot("NDVI_pos_value", "Maximum NDVI", grey = T)
```

```{r}
plot_perc_Landsat <- data_validation_Landsat %>%
  # Rough validation based on NDWI
  rough_validation_Landsat() %>% 
  # Data filtering
  filter_data_Landsat(vars_RF_months_Landsat) %>%
  distr_plot("NDVI_pos_value", "Maximum NDVI", grey = T)
```

```{r}
plot_perc_Landsat + plot_perc_S2 +
  plot_annotation(tag_levels = "A", tag_suffix = ")")
ggsave(here("output", "figures", "manuscript", "appendix2.jpeg"), 
       dpi = 300, width = 8, height = 5)
```

# Fit RF models (level 1)

## Without refinement

### N points per category

```{r}
bind_rows(
  # S2
  filtered_data_slope_S2 %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "S2", method = "slope") %>% 
    select(sensor, method, Q, R, S, T),
  filtered_data_months_S2 %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "S2", method = "months") %>% 
    select(sensor, method, Q, R, S, T),
  filtered_data_both_S2 %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "S2", method = "both") %>% 
    select(sensor, method, Q, R, S, T),
  # # Landsat
  filtered_data_slope_Landsat %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>%
    mutate(sensor = "Landsat", method = "slope") %>% 
    select(sensor, method, Q, R, S, T),
  filtered_data_months_Landsat %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>%
    mutate(sensor = "Landsat", method = "months") %>% 
    select(sensor, method, Q, R, S, T),
  filtered_data_both_Landsat %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>%
    mutate(sensor = "Landsat", method = "both") %>% 
    select(sensor, method, Q, R, S, T),
  # SatEmb
  filtered_data_SatEmb %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "SatEmb", method ="SatEmb only") %>% 
    select(sensor, method, Q, R, S, T),
  filtered_data_SatEmb_CH %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "SatEmb", method ="SatEmb + CH") %>% 
    select(sensor, method, Q, R, S, T)
)
```

```{r}
bind_rows(
  # S2
  filtered_data_slope_S2 %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "S2", method = "slope") %>% 
    select(sensor, method, Q, R, S, T),
  filtered_data_months_S2 %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "S2", method = "months") %>% 
    select(sensor, method, Q, R, S, T),
  filtered_data_both_S2 %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "S2", method = "both") %>% 
    select(sensor, method, Q, R, S, T),
  # # Landsat
  filtered_data_slope_Landsat %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>%
    mutate(sensor = "Landsat", method = "slope") %>% 
    select(sensor, method, Q, R, S, T),
  filtered_data_months_Landsat %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>%
    mutate(sensor = "Landsat", method = "months") %>% 
    select(sensor, method, Q, R, S, T),
  filtered_data_both_Landsat %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>%
    mutate(sensor = "Landsat", method = "both") %>% 
    select(sensor, method, Q, R, S, T),
  # SatEmb
  filtered_data_SatEmb %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "SatEmb", method ="SatEmb only") %>% 
    select(sensor, method, Q, R, S, T),
  filtered_data_SatEmb_CH %>% count(EUNISa_1) %>%
    pivot_wider(names_from = EUNISa_1, values_from = n) %>% 
    mutate(sensor = "SatEmb", method ="SatEmb + CH") %>% 
    select(sensor, method, Q, R, S, T)
) %>%
  pivot_longer(cols = 3:6, names_to = "EUNIS", values_to = "Count") %>%
  ggplot(aes(x = EUNIS, y = Count, fill = EUNIS, group = EUNIS)) +
  geom_col() +
  facet_wrap(~ sensor + method) + theme_bw() +
  # ggtitle("Reduction in number of observations with filtering steps") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  
```

### Split into training and test data sets

Split data:

```{r}
# S2
# Slope
set.seed(123)
split_data_slope_S2 <- split_train_test(filtered_data_slope_S2)

train_data_slope_S2 <- split_data_slope_S2$train
test_data_slope_S2 <- split_data_slope_S2$test

# Check overlap between training and testing datasets
sum(train_data_slope_S2$PlotObservationID %in% test_data_slope_S2$PlotObservationID)

# Months
set.seed(123)
split_data_months_S2 <- split_train_test(filtered_data_months_S2)

train_data_months_S2 <- split_data_months_S2$train
test_data_months_S2 <- split_data_months_S2$test

# Check overlap between training and testing datasets
sum(train_data_months_S2$PlotObservationID %in% test_data_months_S2$PlotObservationID)

# Both
set.seed(123)
split_data_both_S2 <- split_train_test(filtered_data_both_S2)

train_data_both_S2 <- split_data_both_S2$train
test_data_both_S2 <- split_data_both_S2$test

# Check overlap between training and testing datasets
sum(train_data_both_S2$PlotObservationID %in% test_data_both_S2$PlotObservationID)

# Landsat
# Slope
set.seed(123)
split_data_slope_Landsat <- split_train_test(filtered_data_slope_Landsat)

train_data_slope_Landsat <- split_data_slope_Landsat$train
test_data_slope_Landsat <- split_data_slope_Landsat$test

# Check overlap between training and testing datasets
sum(train_data_slope_Landsat$PlotObservationID %in% test_data_slope_Landsat$PlotObservationID)

# Months
set.seed(123)
split_data_months_Landsat <- split_train_test(filtered_data_months_Landsat)

train_data_months_Landsat <- split_data_months_Landsat$train
test_data_months_Landsat <- split_data_months_Landsat$test

# Check overlap between training and testing datasets
sum(train_data_months_Landsat$PlotObservationID %in% test_data_months_Landsat$PlotObservationID)

# Both
set.seed(123)
split_data_both_Landsat <- split_train_test(filtered_data_both_Landsat)

train_data_both_Landsat <- split_data_both_Landsat$train
test_data_both_Landsat <- split_data_both_Landsat$test

# Check overlap between training and testing datasets
sum(train_data_both_Landsat$PlotObservationID %in% test_data_both_Landsat$PlotObservationID)

# SatEmb
set.seed(123)
split_data_SatEmb <- split_train_test(filtered_data_SatEmb)

train_data_SatEmb <- split_data_SatEmb$train
test_data_SatEmb <- split_data_SatEmb$test

# Check overlap between training and testing datasets
sum(train_data_SatEmb$PlotObservationID %in% test_data_SatEmb$PlotObservationID)

# SatEmb_CH
set.seed(123)
split_data_SatEmb_CH <- split_train_test(filtered_data_SatEmb_CH)

train_data_SatEmb_CH <- split_data_SatEmb_CH$train
test_data_SatEmb_CH <- split_data_SatEmb_CH$test

# Check overlap between training and testing datasets
sum(train_data_SatEmb_CH$PlotObservationID %in% test_data_SatEmb_CH$PlotObservationID)
```

### Fit models

```{r eval=FALSE, include=FALSE}
# S2
rf_slope_S2 <- run_rf_parallel(vars_RF_slope, train_data_slope_S2, "EUNISa_1",
                      ntree = 1500)
rf_months_S2 <- run_rf_parallel(vars_RF_months, train_data_months_S2, "EUNISa_1",
                       ntree = 1500)
rf_both_S2 <- run_rf_parallel(vars_RF_both, train_data_both_S2, "EUNISa_1",
                       ntree = 1500)

# Landsat
rf_slope_Landsat <- run_rf_parallel(vars_RF_slope_Landsat, train_data_slope_Landsat, 
                           "EUNISa_1", ntree = 1500)
rf_months_Landsat <- run_rf_parallel(vars_RF_months_Landsat, train_data_months_Landsat,
                            "EUNISa_1", ntree = 1500)
rf_both_Landsat <- run_rf_parallel(vars_RF_both_Landsat, train_data_both_Landsat,
                            "EUNISa_1", ntree = 1500)

# SatEmb
rf_SatEmb <- run_rf_parallel(vars_RF_SatEmb, train_data_SatEmb, "EUNISa_1",
                    ntree = 1500)
rf_SatEmb_CH <- run_rf_parallel(vars_RF_SatEmb_CH, train_data_SatEmb_CH, "EUNISa_1",
                       ntree = 1500)
```

```{r}
models_no_refin <- c("rf_slope_S2", "rf_months_S2", "rf_both_S2", 
                     "rf_slope_Landsat", "rf_months_Landsat", "rf_both_Landsat",
                     "rf_SatEmb", "rf_SatEmb_CH")
```

Print the time:

```{r}
for (model_name in models_no_refin) {
  obj <- get(model_name)
  cat("Time for", model_name, ":\n")
  print(obj$time)
}
```

Save:

```{r eval=FALSE, include=FALSE}
for (model_name in models_no_refin) {
  file_path <- paste0("objects/RF_1500_20260128/", model_name, "_model.Rdata")
  save(list = model_name, file = file_path)
}
```

### Predictions

```{r eval=FALSE, include=FALSE}
for (model_name in models_no_refin) {
  # Derive test data name by replacing "rf_" with "test_data_"
  test_data_name <- sub("^rf_", "test_data_", model_name)
  
  # Create prediction variable name
  prediction_name <- paste0(model_name, "_prediction")
  
  # Get model and test data
  model <- get(model_name)$model
  test_data <- get(test_data_name)
  
  # Run prediction
  prediction <- predict(model, newdata = test_data, 
                        type = "response")
  
  # Assign prediction to a variable
  assign(prediction_name, prediction)
}
```

```{r}
predictions_no_refin <- c("rf_slope_S2_prediction", "rf_months_S2_prediction",
                          "rf_both_S2_prediction", 
                          "rf_slope_Landsat_prediction", 
                          "rf_months_Landsat_prediction", 
                          "rf_both_Landsat_prediction",
                          "rf_SatEmb_prediction", "rf_SatEmb_CH_prediction")
```

Save:

```{r eval=FALSE, include=FALSE}
for (prediction_name in predictions_no_refin) {
  file_path <- paste0("objects/RF_1500_20260128/", prediction_name, ".Rdata")
  save(list = prediction_name, file = file_path)
}
```

### Confusion matrices

```{r}
for (prediction_name in predictions_no_refin) {
  # Derive test data name by replacing "_prediction" with "" 
  # and "rf_" with "test_data_"
  test_data_name <- paste0("test_data_", sub("^rf_", "", sub("_prediction$", "",
                                                             prediction_name)))
  
  # Get prediction and test data
  prediction <- get(prediction_name)
  test_data <- get(test_data_name)
  
  # Compute confusion matrix
  cat("Confusion Matrix for", sub("_prediction$", "", prediction_name), ":\n")
  print(confusionMatrix(prediction, test_data$EUNISa_1))
  cat("\n")
}
```

#### Plots

```{r}
for (prediction_name in predictions_no_refin) {
  # Derive name of test data
  test_data_name <- paste0("test_data_", sub("^rf_", "", sub("_prediction$", "",
                                                             prediction_name)))
  
  # Obtain objects
  prediction <- get(prediction_name)
  test_data <- get(test_data_name)

  # Personalized title
  title <- paste(#"Confusion Matrix for", 
    sub("_prediction$", "",
                                              prediction_name))

  # Plot
  print(plot_confusion_matrix(prediction, test_data$EUNISa_1, title))
}
```

## Refinement step 1

Based on PAs "Morralla protocol".

### Compute predictions for all data and combine with original data

```{r eval=FALSE, include=FALSE}
# S2, slope
filtered_data_slope_S2_with_probs <- bind_cols(
  filtered_data_slope_S2 %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_slope_S2$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_slope_S2,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_slope_S2$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_slope_S2,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# S2, months
filtered_data_months_S2_with_probs <- bind_cols(
  filtered_data_months_S2 %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_months_S2$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_months_S2,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_months_S2$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_months_S2,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# S2, both
filtered_data_both_S2_with_probs <- bind_cols(
  filtered_data_both_S2 %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_both_S2$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_both_S2,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_both_S2$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_both_S2,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# Landsat, slope
filtered_data_slope_Landsat_with_probs <- bind_cols(
  filtered_data_slope_Landsat %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_slope_Landsat$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_slope_Landsat,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_slope_Landsat$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_slope_Landsat,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# Landsat, months
filtered_data_months_Landsat_with_probs <- bind_cols(
  filtered_data_months_Landsat %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_months_Landsat$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_months_Landsat,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_months_Landsat$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_months_Landsat,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# Landsat, both
filtered_data_both_Landsat_with_probs <- bind_cols(
  filtered_data_both_Landsat %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_both_Landsat$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_both_Landsat,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_both_Landsat$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_both_Landsat,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# SatEmb
filtered_data_SatEmb_with_probs <- bind_cols(
  filtered_data_SatEmb %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_SatEmb$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_SatEmb,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_SatEmb$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_SatEmb,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# SatEmb_CH
filtered_data_SatEmb_CH_with_probs <- bind_cols(
  filtered_data_SatEmb_CH %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_SatEmb_CH$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_SatEmb_CH,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_SatEmb_CH$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_SatEmb_CH,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)
```

```{r}
filtered_data_with_probs <- c("filtered_data_slope_S2_with_probs",
                              "filtered_data_months_S2_with_probs",
                              "filtered_data_both_S2_with_probs",
                              "filtered_data_slope_Landsat_with_probs", 
                              "filtered_data_months_Landsat_with_probs",
                              "filtered_data_both_Landsat_with_probs",
                              "filtered_data_SatEmb_with_probs", 
                              "filtered_data_SatEmb_CH_with_probs")
```

Save:

```{r eval=FALSE, include=FALSE}
for (filtered_data_name in filtered_data_with_probs) {
  file_path <- paste0("objects/RF_1500_20260128/", filtered_data_name, ".Rdata")
  save(list = filtered_data_name, file = file_path)
}
```

### Detect probability columns

```{r}
prob_cols_slope_S2 <- names(filtered_data_slope_S2_with_probs)[grepl("^prob_", names(filtered_data_slope_S2_with_probs))]
prob_cols_months_S2 <- names(filtered_data_months_S2_with_probs)[grepl("^prob_", names(filtered_data_months_S2_with_probs))]
prob_cols_both_S2 <- names(filtered_data_both_S2_with_probs)[grepl("^prob_", names(filtered_data_both_S2_with_probs))]
prob_cols_slope_Landsat <- names(filtered_data_slope_Landsat_with_probs)[grepl("^prob_", names(filtered_data_slope_Landsat_with_probs))]
prob_cols_months_Landsat <- names(filtered_data_months_Landsat_with_probs)[grepl("^prob_", names(filtered_data_months_Landsat_with_probs))]
prob_cols_both_Landsat <- names(filtered_data_both_Landsat_with_probs)[grepl("^prob_", names(filtered_data_both_Landsat_with_probs))]
prob_cols_SatEmb <- names(filtered_data_SatEmb_with_probs)[grepl("^prob_", names(filtered_data_SatEmb_with_probs))]
prob_cols_SatEmb_CH <- names(filtered_data_SatEmb_CH_with_probs)[grepl("^prob_", names(filtered_data_SatEmb_CH_with_probs))]
```

### Calculate auxiliary metrics

```{r}
df_aux_slope_S2 <- calculate_df_aux(filtered_data_slope_S2_with_probs,
                                     prob_cols_slope_S2)
df_aux_months_S2 <- calculate_df_aux(filtered_data_months_S2_with_probs,
                                     prob_cols_months_S2)
df_aux_both_S2 <- calculate_df_aux(filtered_data_both_S2_with_probs,
                                     prob_cols_both_S2)
df_aux_slope_Landsat <- calculate_df_aux(filtered_data_slope_Landsat_with_probs,
                                     prob_cols_slope_Landsat)
df_aux_months_Landsat <- calculate_df_aux(filtered_data_months_Landsat_with_probs,
                                     prob_cols_months_Landsat)
df_aux_both_Landsat <- calculate_df_aux(filtered_data_both_Landsat_with_probs,
                                     prob_cols_both_Landsat)
df_aux_SatEmb <- calculate_df_aux(filtered_data_SatEmb_with_probs,
                                  prob_cols_SatEmb)
df_aux_SatEmb_CH <- calculate_df_aux(filtered_data_SatEmb_CH_with_probs,
                                     prob_cols_SatEmb_CH)
```

### Apply filtering function

Applying "Morralla" protocol once, removing negatives until 20% of the class and, if there is room, uncertain positives until 10% of the rest.

```{r}
df_final_20_10_slope_S2 <- df_aux_slope_S2 %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20_pos10)
df_final_20_10_months_S2 <- df_aux_months_S2 %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20_pos10)
df_final_20_10_both_S2 <- df_aux_both_S2 %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20_pos10)
df_final_20_10_slope_Landsat <- df_aux_slope_Landsat %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20_pos10)
df_final_20_10_months_Landsat <- df_aux_months_Landsat %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20_pos10)
df_final_20_10_both_Landsat <- df_aux_both_Landsat %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20_pos10)
df_final_20_10_SatEmb <- df_aux_SatEmb %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20_pos10)
df_final_20_10_SatEmb_CH <- df_aux_SatEmb_CH %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20_pos10)
```

```{r}
filtered_data_20_10_slope_S2_probs <- filtered_data_slope_S2 %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_slope_S2$PlotObservationID)
filtered_data_20_10_months_S2_probs <- filtered_data_months_S2 %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_months_S2$PlotObservationID)
filtered_data_20_10_both_S2_probs <- filtered_data_both_S2 %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_both_S2$PlotObservationID)
filtered_data_20_10_slope_Landsat_probs <- filtered_data_slope_Landsat %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_slope_Landsat$PlotObservationID)
filtered_data_20_10_months_Landsat_probs <- filtered_data_months_Landsat %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_months_Landsat$PlotObservationID)
filtered_data_20_10_both_Landsat_probs <- filtered_data_both_Landsat %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_both_Landsat$PlotObservationID)
filtered_data_20_10_SatEmb_probs <- filtered_data_SatEmb %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_SatEmb$PlotObservationID)
filtered_data_20_10_SatEmb_CH_probs <- filtered_data_SatEmb_CH %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_SatEmb_CH$PlotObservationID)
```

### Compare classes before and after

```{r}
compare_classes(filtered_data_slope_S2, filtered_data_20_10_slope_S2_probs)
compare_classes(filtered_data_months_S2, filtered_data_20_10_months_S2_probs)
compare_classes(filtered_data_both_S2, filtered_data_20_10_both_S2_probs)
compare_classes(filtered_data_slope_Landsat, filtered_data_20_10_slope_Landsat_probs)
compare_classes(filtered_data_months_Landsat, filtered_data_20_10_months_Landsat_probs)
compare_classes(filtered_data_both_Landsat, filtered_data_20_10_both_Landsat_probs)
compare_classes(filtered_data_SatEmb, filtered_data_20_10_SatEmb_probs)
compare_classes(filtered_data_SatEmb_CH, filtered_data_20_10_SatEmb_CH_probs)
```

### Split into training and testing

```{r}
# S2
# Slope
set.seed(123)
split_data_20_10_slope_S2 <- split_train_test(filtered_data_20_10_slope_S2_probs)

train_data_20_10_slope_S2 <- split_data_20_10_slope_S2$train
test_data_20_10_slope_S2 <- split_data_20_10_slope_S2$test

# Check overlap between training and testing datasets
sum(train_data_20_10_slope_S2$PlotObservationID %in% test_data_20_10_slope_S2$PlotObservationID)

# Months
set.seed(123)
split_data_20_10_months_S2 <- split_train_test(filtered_data_20_10_months_S2_probs)

train_data_20_10_months_S2 <- split_data_20_10_months_S2$train
test_data_20_10_months_S2 <- split_data_20_10_months_S2$test

# Check overlap between training and testing datasets
sum(train_data_20_10_months_S2$PlotObservationID %in% test_data_20_10_months_S2$PlotObservationID)

# Both
set.seed(123)
split_data_20_10_both_S2 <- split_train_test(filtered_data_20_10_both_S2_probs)

train_data_20_10_both_S2 <- split_data_20_10_both_S2$train
test_data_20_10_both_S2 <- split_data_20_10_both_S2$test

# Check overlap between training and testing datasets
sum(train_data_20_10_both_S2$PlotObservationID %in% test_data_20_10_both_S2$PlotObservationID)

# Landsat
# Slope
set.seed(123)
split_data_20_10_slope_Landsat <- split_train_test(filtered_data_20_10_slope_Landsat_probs)

train_data_20_10_slope_Landsat <- split_data_20_10_slope_Landsat$train
test_data_20_10_slope_Landsat <- split_data_20_10_slope_Landsat$test

# Check overlap between training and testing datasets
sum(train_data_20_10_slope_Landsat$PlotObservationID %in% test_data_20_10_slope_Landsat$PlotObservationID)

# Months
set.seed(123)
split_data_20_10_months_Landsat <- split_train_test(filtered_data_20_10_months_Landsat_probs)

train_data_20_10_months_Landsat <- split_data_20_10_months_Landsat$train
test_data_20_10_months_Landsat <- split_data_20_10_months_Landsat$test

# Check overlap between training and testing datasets
sum(train_data_20_10_months_Landsat$PlotObservationID %in% test_data_20_10_months_Landsat$PlotObservationID)

# Both
set.seed(123)
split_data_20_10_both_Landsat <- split_train_test(filtered_data_20_10_both_Landsat_probs)

train_data_20_10_both_Landsat <- split_data_20_10_both_Landsat$train
test_data_20_10_both_Landsat <- split_data_20_10_both_Landsat$test

# Check overlap between training and testing datasets
sum(train_data_20_10_both_Landsat$PlotObservationID %in% test_data_20_10_both_Landsat$PlotObservationID)

# SatEmb
set.seed(123)
split_data_20_10_SatEmb <- split_train_test(filtered_data_20_10_SatEmb_probs)

train_data_20_10_SatEmb <- split_data_20_10_SatEmb$train
test_data_20_10_SatEmb <- split_data_20_10_SatEmb$test

# Check overlap between training and testing datasets
sum(train_data_20_10_SatEmb$PlotObservationID %in% test_data_20_10_SatEmb$PlotObservationID)

# SatEmb_CH
set.seed(123)
split_data_20_10_SatEmb_CH <- split_train_test(filtered_data_20_10_SatEmb_CH_probs)

train_data_20_10_SatEmb_CH <- split_data_20_10_SatEmb_CH$train
test_data_20_10_SatEmb_CH <- split_data_20_10_SatEmb_CH$test

# Check overlap between training and testing datasets
sum(train_data_20_10_SatEmb_CH$PlotObservationID %in% test_data_20_10_SatEmb_CH$PlotObservationID)
```

### Fit models

```{r eval=FALSE, include=FALSE}
rf_slope_S2_refin_20_10 <- run_rf_parallel(vars_RF_slope, 
                                  train_data_20_10_slope_S2, "EUNISa_1",
                                  ntree = 1500)
rf_months_S2_refin_20_10 <- run_rf_parallel(vars_RF_months,
                                   train_data_20_10_months_S2, "EUNISa_1",
                                   ntree = 1500)
rf_both_S2_refin_20_10 <- run_rf_parallel(vars_RF_both,
                                   train_data_20_10_both_S2, "EUNISa_1",
                                   ntree = 1500)
rf_slope_Landsat_refin_20_10 <- run_rf_parallel(vars_RF_slope_Landsat,
                                       train_data_20_10_slope_Landsat, 
                                       "EUNISa_1", ntree = 1500)
rf_months_Landsat_refin_20_10 <- run_rf_parallel(vars_RF_months_Landsat,
                                        train_data_20_10_months_Landsat, 
                                        "EUNISa_1", ntree = 1500)
rf_both_Landsat_refin_20_10 <- run_rf_parallel(vars_RF_both_Landsat,
                                        train_data_20_10_both_Landsat, 
                                        "EUNISa_1", ntree = 1500)
rf_SatEmb_refin_20_10 <- run_rf_parallel(vars_RF_SatEmb, train_data_20_10_SatEmb, 
                                "EUNISa_1", ntree = 1500)
rf_SatEmb_CH_refin_20_10 <- run_rf_parallel(vars_RF_SatEmb_CH, 
                                   train_data_20_10_SatEmb_CH, "EUNISa_1",
                                   ntree = 1500)
```

```{r}
models_refin_20_10 <- c("rf_slope_S2_refin_20_10", "rf_months_S2_refin_20_10",
                        "rf_both_S2_refin_20_10", 
                        "rf_slope_Landsat_refin_20_10", 
                        "rf_months_Landsat_refin_20_10",
                        "rf_both_Landsat_refin_20_10",
                        "rf_SatEmb_refin_20_10", "rf_SatEmb_CH_refin_20_10")
```

Save:

```{r eval=FALSE, include=FALSE}
for (model_name in models_refin_20_10) {
  file_path <- paste0("objects/RF_1500_20260128/", model_name, "_model.Rdata")
  save(list = model_name, file = file_path)
}
```

### Predictions

```{r eval=FALSE, include=FALSE}
for (model_name in models_refin_20_10) {
  # Derive test data name
  test_data_name <- paste0(gsub("_refin_20_10$", "", 
                                sub("^rf_", "test_data_20_10_", model_name)))

  # Create prediction variable name
  prediction_name <- paste0(model_name, "_prediction")
  
  # Get model and test data
  model <- get(model_name)$model
  test_data <- get(test_data_name)
  
  # Run prediction
  prediction <- predict(model, newdata = test_data, 
                        type = "response")
  
  # Assign prediction to a variable
  assign(prediction_name, prediction)
}
```

```{r}
predictions_refin_20_10 <- c("rf_slope_S2_refin_20_10_prediction",
                        "rf_months_S2_refin_20_10_prediction",
                        "rf_both_S2_refin_20_10_prediction",
                        "rf_slope_Landsat_refin_20_10_prediction",
                        "rf_months_Landsat_refin_20_10_prediction",
                        "rf_both_Landsat_refin_20_10_prediction",
                        "rf_SatEmb_refin_20_10_prediction",
                        "rf_SatEmb_CH_refin_20_10_prediction")
```

Save:

```{r eval=FALSE, include=FALSE}
for (prediction_name in predictions_refin_20_10) {
  file_path <- paste0("objects/RF_1500_20260128/", prediction_name, ".Rdata")
  save(list = prediction_name, file = file_path)
  }
```

### Confusion matrices

```{r}
for (prediction_name in predictions_refin_20_10) {
  # Derive test data name
  test_data_name <- paste0(gsub("_refin_20_10$", "",
                              sub("^rf_", "test_data_20_10_", 
                                  sub("_prediction$", "", prediction_name))))
  
  # Get prediction and test data
  prediction <- get(prediction_name)
  test_data <- get(test_data_name)
  
  # Compute confusion matrix
  cat("Confusion Matrix for", sub("_prediction$", "", prediction_name), ":\n")
  print(confusionMatrix(prediction, test_data$EUNISa_1))
  cat("\n")
}
```

#### Plots

```{r}
for (prediction_name in predictions_refin_20_10) {
  # Derive test data name
  test_data_name <- paste0(gsub("_refin_20_10$", "",
                              sub("^rf_", "test_data_20_10_", 
                                  sub("_prediction$", "", prediction_name))))
  
  # Obtain objects
  prediction <- get(prediction_name)
  test_data <- get(test_data_name)

  # Personalized title
  title <- paste("Confusion Matrix for", sub("_prediction$", "",
                                              prediction_name))

  # Plot
  print(plot_confusion_matrix(prediction, test_data$EUNISa_1, title))
}
```

## Refinement step 2

Then refitting the model and applying "Morralla" protocol again, removing only negatives until 20% of the class.

#### Compute predictions for all data and combine with original data

```{r eval=FALSE, include=FALSE}
# S2, slope
filtered_data_slope_S2_with_probs_2x <- bind_cols(
  filtered_data_20_10_slope_S2_probs %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_slope_S2_refin_20_10$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_20_10_slope_S2_probs,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_slope_S2_refin_20_10$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_20_10_slope_S2_probs,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# S2, months
filtered_data_months_S2_with_probs_2x <- bind_cols(
  filtered_data_20_10_months_S2_probs %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_months_S2_refin_20_10$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_20_10_months_S2_probs,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_months_S2_refin_20_10$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_20_10_months_S2_probs,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# S2, both
filtered_data_both_S2_with_probs_2x <- bind_cols(
  filtered_data_20_10_both_S2_probs %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_both_S2_refin_20_10$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_20_10_both_S2_probs,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_both_S2_refin_20_10$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_20_10_both_S2_probs,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# Landsat, slope
filtered_data_slope_Landsat_with_probs_2x <- bind_cols(
  filtered_data_20_10_slope_Landsat_probs %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_slope_Landsat_refin_20_10$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_20_10_slope_Landsat_probs,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_slope_Landsat_refin_20_10$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_20_10_slope_Landsat_probs,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# Landsat, months
filtered_data_months_Landsat_with_probs_2x <- bind_cols(
  filtered_data_20_10_months_Landsat_probs %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_months_Landsat_refin_20_10$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_20_10_months_Landsat_probs,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_months_Landsat_refin_20_10$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_20_10_months_Landsat_probs,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# Landsat, both
filtered_data_both_Landsat_with_probs_2x <- bind_cols(
  filtered_data_20_10_both_Landsat_probs %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_both_Landsat_refin_20_10$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_20_10_both_Landsat_probs,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_both_Landsat_refin_20_10$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_20_10_both_Landsat_probs,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# SatEmb
filtered_data_SatEmb_with_probs_2x <- bind_cols(
  filtered_data_20_10_SatEmb_probs %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_SatEmb_refin_20_10$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_20_10_SatEmb_probs,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_SatEmb_refin_20_10$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_20_10_SatEmb_probs,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)

# SatEmb_CH
filtered_data_SatEmb_CH_with_probs_2x <- bind_cols(
  filtered_data_20_10_SatEmb_CH_probs %>% select(PlotObservationID, EUNISa_1),
  as.data.frame(predict(rf_SatEmb_CH_refin_20_10$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_20_10_SatEmb_CH_probs,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_SatEmb_CH_refin_20_10$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_20_10_SatEmb_CH_probs,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNISa_1, prob_Q = EUNISa_1.Q,
         prob_R = EUNISa_1.R, prob_S = EUNISa_1.S, prob_T = EUNISa_1.T)
```

```{r}
filtered_data_with_probs_2x <- c("filtered_data_slope_S2_with_probs_2x",
                              "filtered_data_months_S2_with_probs_2x",
                              "filtered_data_both_S2_with_probs_2x",
                              "filtered_data_slope_Landsat_with_probs_2x", 
                              "filtered_data_months_Landsat_with_probs_2x",
                              "filtered_data_both_Landsat_with_probs_2x",
                              "filtered_data_SatEmb_with_probs_2x", 
                              "filtered_data_SatEmb_CH_with_probs_2x")
```

Save:

```{r eval=FALSE, include=FALSE}
for (filtered_data_name in filtered_data_with_probs_2x) {
  file_path <- paste0("objects/RF_1500_20260128/", filtered_data_name, ".Rdata")
  save(list = filtered_data_name, file = file_path)
}
```

#### Calculate auxiliary metrics

```{r}
df_aux_slope_S2_2x <- calculate_df_aux(filtered_data_slope_S2_with_probs_2x,
                                     prob_cols_slope_S2)
df_aux_months_S2_2x <- calculate_df_aux(filtered_data_months_S2_with_probs_2x,
                                     prob_cols_months_S2)
df_aux_both_S2_2x <- calculate_df_aux(filtered_data_both_S2_with_probs_2x,
                                     prob_cols_both_S2)
df_aux_slope_Landsat_2x <- calculate_df_aux(filtered_data_slope_Landsat_with_probs_2x,
                                     prob_cols_slope_Landsat)
df_aux_months_Landsat_2x <- calculate_df_aux(filtered_data_months_Landsat_with_probs_2x,
                                     prob_cols_months_Landsat)
df_aux_both_Landsat_2x <- calculate_df_aux(filtered_data_both_Landsat_with_probs_2x,
                                     prob_cols_both_Landsat)
df_aux_SatEmb_2x <- calculate_df_aux(filtered_data_SatEmb_with_probs_2x,
                                  prob_cols_SatEmb)
df_aux_SatEmb_CH_2x <- calculate_df_aux(filtered_data_SatEmb_CH_with_probs_2x,
                                     prob_cols_SatEmb_CH)
```

#### Apply filtering function

```{r}
df_final_20_10_slope_S2_2x <- df_aux_slope_S2_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)
df_final_20_10_months_S2_2x <- df_aux_months_S2_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)
df_final_20_10_both_S2_2x <- df_aux_both_S2_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)
df_final_20_10_slope_Landsat_2x <- df_aux_slope_Landsat_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)
df_final_20_10_months_Landsat_2x <- df_aux_months_Landsat_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)
df_final_20_10_both_Landsat_2x <- df_aux_both_Landsat_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)
df_final_20_10_SatEmb_2x <- df_aux_SatEmb_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)
df_final_20_10_SatEmb_CH_2x <- df_aux_SatEmb_CH_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)
```

```{r}
filtered_data_20_10_slope_S2_probs_2x <- filtered_data_slope_S2 %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_slope_S2_2x$PlotObservationID)
filtered_data_20_10_months_S2_probs_2x <- filtered_data_months_S2 %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_months_S2_2x$PlotObservationID)
filtered_data_20_10_both_S2_probs_2x <- filtered_data_both_S2 %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_both_S2_2x$PlotObservationID)
filtered_data_20_10_slope_Landsat_probs_2x <- filtered_data_slope_Landsat %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_slope_Landsat_2x$PlotObservationID)
filtered_data_20_10_months_Landsat_probs_2x <- filtered_data_months_Landsat %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_months_Landsat_2x$PlotObservationID)
filtered_data_20_10_both_Landsat_probs_2x <- filtered_data_both_Landsat %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_both_Landsat_2x$PlotObservationID)
filtered_data_20_10_SatEmb_probs_2x <- filtered_data_SatEmb %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_SatEmb_2x$PlotObservationID)
filtered_data_20_10_SatEmb_CH_probs_2x <- filtered_data_SatEmb_CH %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_SatEmb_CH_2x$PlotObservationID)
```

#### (Appendix 3) Compare classes before and after

```{r}
compare_classes(filtered_data_slope_S2, filtered_data_20_10_slope_S2_probs,
                filtered_data_20_10_slope_S2_probs_2x)
compare_classes(filtered_data_months_S2, filtered_data_20_10_months_S2_probs,
                filtered_data_20_10_months_S2_probs_2x)
compare_classes(filtered_data_both_S2, filtered_data_20_10_both_S2_probs,
                filtered_data_20_10_both_S2_probs_2x)
compare_classes(filtered_data_slope_Landsat,
                filtered_data_20_10_slope_Landsat_probs,
                filtered_data_20_10_slope_Landsat_probs_2x)
compare_classes(filtered_data_months_Landsat, 
                filtered_data_20_10_months_Landsat_probs,
                filtered_data_20_10_months_Landsat_probs_2x)
compare_classes(filtered_data_both_Landsat, 
                filtered_data_20_10_both_Landsat_probs,
                filtered_data_20_10_both_Landsat_probs_2x)
compare_classes(filtered_data_SatEmb, filtered_data_20_10_SatEmb_probs,
                filtered_data_20_10_SatEmb_probs_2x)
compare_classes(filtered_data_SatEmb_CH, filtered_data_20_10_SatEmb_CH_probs,
                filtered_data_20_10_SatEmb_CH_probs_2x)
```

```{r}
(
  compare_classes(filtered_data_months_Landsat, 
                  filtered_data_20_10_months_Landsat_probs,
                  filtered_data_20_10_months_Landsat_probs_2x,
                  return = "plot") +
  compare_classes(filtered_data_SatEmb, 
                  filtered_data_20_10_SatEmb_probs,
                  filtered_data_20_10_SatEmb_probs_2x,
                  return = "plot")
) +
  plot_annotation(tag_levels = "A", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  scale_fill_manual(
    breaks = c("before", "step1", "step2"),
    labels = c("No refinement", "Refinement step 1", "Refinement step 2"),
    values = c(
      "before" = "#edf8b1", 
      "step1"  = "#7fcdbb",
      "step2"  = "#2c7fb8"
    )
  ) &
  scale_y_continuous(labels = label_number(scale = 1e-3, suffix = "k")) &
  theme(legend.position = "bottom", legend.title = element_blank())
ggsave(here("output", "figures", "manuscript", "appendix3.jpeg"), 
       dpi = 300, width = 6, height = 4)
```




#### Split into training and testing

```{r}
# S2
# Slope
set.seed(123)
split_data_20_10_2x_slope_S2 <- split_train_test(filtered_data_20_10_slope_S2_probs_2x)

train_data_20_10_2x_slope_S2 <- split_data_20_10_2x_slope_S2$train
test_data_20_10_2x_slope_S2 <- split_data_20_10_2x_slope_S2$test

# Check overlap between training and testing datasets
sum(train_data_20_10_2x_slope_S2$PlotObservationID %in% test_data_20_10_2x_slope_S2$PlotObservationID)

# Months
set.seed(123)
split_data_20_10_2x_months_S2 <- split_train_test(filtered_data_20_10_months_S2_probs_2x)

train_data_20_10_2x_months_S2 <- split_data_20_10_2x_months_S2$train
test_data_20_10_2x_months_S2 <- split_data_20_10_2x_months_S2$test

# Check overlap between training and testing datasets
sum(train_data_20_10_2x_months_S2$PlotObservationID %in% test_data_20_10_2x_months_S2$PlotObservationID)

# Both
set.seed(123)
split_data_20_10_2x_both_S2 <- split_train_test(filtered_data_20_10_both_S2_probs_2x)

train_data_20_10_2x_both_S2 <- split_data_20_10_2x_both_S2$train
test_data_20_10_2x_both_S2 <- split_data_20_10_2x_both_S2$test

# Check overlap between training and testing datasets
sum(train_data_20_10_2x_both_S2$PlotObservationID %in% test_data_20_10_2x_both_S2$PlotObservationID)

# Landsat
# Slope
set.seed(123)
split_data_20_10_2x_slope_Landsat <- split_train_test(filtered_data_20_10_slope_Landsat_probs_2x)

train_data_20_10_2x_slope_Landsat <- split_data_20_10_2x_slope_Landsat$train
test_data_20_10_2x_slope_Landsat <- split_data_20_10_2x_slope_Landsat$test

# Check overlap between training and testing datasets
sum(train_data_20_10_2x_slope_Landsat$PlotObservationID %in% test_data_20_10_2x_slope_Landsat$PlotObservationID)

# Months
set.seed(123)
split_data_20_10_2x_months_Landsat <- split_train_test(filtered_data_20_10_months_Landsat_probs_2x)

train_data_20_10_2x_months_Landsat <- split_data_20_10_2x_months_Landsat$train
test_data_20_10_2x_months_Landsat <- split_data_20_10_2x_months_Landsat$test

# Check overlap between training and testing datasets
sum(train_data_20_10_2x_months_Landsat$PlotObservationID %in% test_data_20_10_2x_months_Landsat$PlotObservationID)

# Both
set.seed(123)
split_data_20_10_2x_both_Landsat <- split_train_test(filtered_data_20_10_both_Landsat_probs_2x)

train_data_20_10_2x_both_Landsat <- split_data_20_10_2x_both_Landsat$train
test_data_20_10_2x_both_Landsat <- split_data_20_10_2x_both_Landsat$test

# Check overlap between training and testing datasets
sum(train_data_20_10_2x_both_Landsat$PlotObservationID %in% test_data_20_10_2x_both_Landsat$PlotObservationID)

# SatEmb
set.seed(123)
split_data_20_10_2x_SatEmb <- split_train_test(filtered_data_20_10_SatEmb_probs_2x)

train_data_20_10_2x_SatEmb <- split_data_20_10_2x_SatEmb$train
test_data_20_10_2x_SatEmb <- split_data_20_10_2x_SatEmb$test

# Check overlap between training and testing datasets
sum(train_data_20_10_2x_SatEmb$PlotObservationID %in% test_data_20_10_2x_SatEmb$PlotObservationID)

# SatEmb_CH
set.seed(123)
split_data_20_10_2x_SatEmb_CH <- split_train_test(filtered_data_20_10_SatEmb_CH_probs)

train_data_20_10_2x_SatEmb_CH <- split_data_20_10_2x_SatEmb_CH$train
test_data_20_10_2x_SatEmb_CH <- split_data_20_10_2x_SatEmb_CH$test

# Check overlap between training and testing datasets
sum(train_data_20_10_2x_SatEmb_CH$PlotObservationID %in% test_data_20_10_2x_SatEmb_CH$PlotObservationID)
```

#### Fit models

```{r eval=FALSE, include=FALSE}
rf_slope_S2_refin_20_10_2x <- run_rf_parallel(vars_RF_slope, 
                                  train_data_20_10_2x_slope_S2, "EUNISa_1",
                                  ntree = 1500)
rf_months_S2_refin_20_10_2x <- run_rf_parallel(vars_RF_months,
                                   train_data_20_10_2x_months_S2, "EUNISa_1",
                                   ntree = 1500)
rf_both_S2_refin_20_10_2x <- run_rf_parallel(vars_RF_both,
                                   train_data_20_10_2x_both_S2, "EUNISa_1",
                                   ntree = 1500)
rf_slope_Landsat_refin_20_10_2x <- run_rf_parallel(vars_RF_slope_Landsat,
                                       train_data_20_10_2x_slope_Landsat, 
                                       "EUNISa_1", ntree = 1500)
rf_months_Landsat_refin_20_10_2x <- run_rf_parallel(vars_RF_months_Landsat,
                                        train_data_20_10_2x_months_Landsat, 
                                        "EUNISa_1", ntree = 1500)
rf_both_Landsat_refin_20_10_2x <- run_rf_parallel(vars_RF_both_Landsat,
                                        train_data_20_10_2x_both_Landsat, 
                                        "EUNISa_1", ntree = 1500)
rf_SatEmb_refin_20_10_2x <- run_rf_parallel(vars_RF_SatEmb, train_data_20_10_2x_SatEmb, 
                                "EUNISa_1", ntree = 1500)
rf_SatEmb_CH_refin_20_10_2x <- run_rf_parallel(vars_RF_SatEmb_CH, 
                                   train_data_20_10_2x_SatEmb_CH, "EUNISa_1",
                                   ntree = 1500)
```

```{r}
models_refin_20_10_2x <- c("rf_slope_S2_refin_20_10_2x", 
                     "rf_months_S2_refin_20_10_2x",
                     "rf_both_S2_refin_20_10_2x", 
                     "rf_slope_Landsat_refin_20_10_2x", 
                     "rf_months_Landsat_refin_20_10_2x",
                     "rf_both_Landsat_refin_20_10_2x",
                     "rf_SatEmb_refin_20_10_2x", 
                     "rf_SatEmb_CH_refin_20_10_2x")
```

Save:

```{r eval=FALSE, include=FALSE}
for (model_name in models_refin_20_10_2x) {
  file_path <- paste0("objects/RF_1500_20260128/", model_name, "_model.Rdata")
  save(list = model_name, file = file_path)
}
```

#### Predictions

```{r eval=FALSE, include=FALSE}
for (model_name in models_refin_20_10_2x) {
  # Derive test data name
  test_data_name <- paste0(gsub("_refin_20_10_2x$", "", 
                                sub("^rf_", "test_data_20_10_2x_", model_name)))

  # Create prediction variable name
  prediction_name <- paste0(model_name, "_prediction")
  
  # Get model and test data
  model <- get(model_name)$model
  test_data <- get(test_data_name)
  
  # Run prediction
  prediction <- predict(model, newdata = test_data, 
                        type = "response")
  
  # Assign prediction to a variable
  assign(prediction_name, prediction)
}
```

```{r}
predictions_refin_20_10_2x <- c("rf_slope_S2_refin_20_10_2x_prediction",
                        "rf_months_S2_refin_20_10_2x_prediction",
                        "rf_both_S2_refin_20_10_2x_prediction",
                        "rf_slope_Landsat_refin_20_10_2x_prediction",
                        "rf_months_Landsat_refin_20_10_2x_prediction",
                        "rf_both_Landsat_refin_20_10_2x_prediction",
                        "rf_SatEmb_refin_20_10_2x_prediction",
                        "rf_SatEmb_CH_refin_20_10_2x_prediction")
```

Save:

```{r eval=FALSE, include=FALSE}
for (prediction_name in predictions_refin_20_10_2x) {
  file_path <- paste0("objects/RF_1500_20260128/", prediction_name, ".Rdata")
  save(list = prediction_name, file = file_path)
  }
```

#### Confusion matrices

```{r}
for (prediction_name in predictions_refin_20_10_2x) {
  # Derive test data name
  test_data_name <- paste0(gsub("_refin_20_10_2x$", "",
                              sub("^rf_", "test_data_20_10_2x_", 
                                  sub("_prediction$", "", prediction_name))))
  
  # Get prediction and test data
  prediction <- get(prediction_name)
  test_data <- get(test_data_name)
  
  # Compute confusion matrix
  cat("Confusion Matrix for", sub("_prediction$", "", prediction_name), ":\n")
  print(confusionMatrix(prediction, test_data$EUNISa_1))
  cat("\n")
}
```

##### Plots

```{r}
for (prediction_name in predictions_refin_20_10_2x) {
  # Derive test data name
  test_data_name <- paste0(gsub("_refin_20_10_2x$", "",
                              sub("^rf_", "test_data_20_10_2x_", 
                                  sub("_prediction$", "", prediction_name))))
  
  # Obtain objects
  prediction <- get(prediction_name)
  test_data <- get(test_data_name)

  # Personalized title
  title <- paste("Confusion Matrix for", sub("_prediction$", "",
                                              prediction_name))

  # Plot
  print(plot_confusion_matrix(prediction, test_data$EUNISa_1, title))
}
```

# N obs

## N obs with metrics extracted

```{r}
data_validation_Landsat %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  nrow()
data_validation_S2 %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  nrow()
data_validation_SatEmb %>%
  filter(if_any(c(2:65), ~ !is.na(.))) %>%
  nrow()
```

## N obs after data filtering

```{r}
data_validation_Landsat %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  # Data filtering
  # Remove rows with missing values in predictors
  dplyr::filter(if_all(all_of(vars_RF_slope_Landsat), ~ !is.na(.))) %>%
  # Remove all rows with wrong values of indices (not between -1 and 1)
  dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
  # With Landsat data, many observations removed because of high/low EVI values
  # In Landsat_extraction, values of EVI outside of expected range were set as NA,
  # so better not to use EVI here
  # dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
  dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
  dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
  dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
  # Select only variables needed
  dplyr::select(PlotObservationID, EUNISa_1, all_of(vars_RF_slope_Landsat), 
                Lctnmth) %>%
  # Keep only rows with differences > 0 in variables
  filter(if_all(any_of(vars_RF_slope_Landsat[grepl("diff", 
                                                   vars_RF_slope_Landsat)]), 
                ~ .x > 0)) %>%
  nrow()
data_validation_Landsat %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  # Data filtering
  # Remove rows with missing values in predictors
  dplyr::filter(if_all(all_of(vars_RF_months_Landsat), ~ !is.na(.))) %>%
  # Remove all rows with wrong values of indices (not between -1 and 1)
  dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
  # With Landsat data, many observations removed because of high/low EVI values
  # In Landsat_extraction, values of EVI outside of expected range were set as NA,
  # so better not to use EVI here
  # dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
  dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
  dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
  dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
  # Select only variables needed
  dplyr::select(PlotObservationID, EUNISa_1, all_of(vars_RF_months_Landsat), 
                Lctnmth) %>%
  # Keep only rows with differences > 0 in variables
  filter(if_all(any_of(vars_RF_months_Landsat[grepl("diff", 
                                                   vars_RF_months_Landsat)]), 
                ~ .x > 0)) %>%
  nrow()
data_validation_S2 %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  # Data filtering
  # Remove rows with missing values in predictors
  dplyr::filter(if_all(all_of(vars_RF_slope), ~ !is.na(.))) %>%
  # Remove all rows with wrong values of indices (not between -1 and 1)
  dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
  dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
  dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
  dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
  dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
  # Select only variables needed
  dplyr::select(PlotObservationID, EUNISa_1, all_of(vars_RF_slope), Lctnmth) %>%
  # Keep only rows with differences > 0 in variables
  filter(if_all(any_of(vars_RF_slope[grepl("diff", vars_RF_slope)]), 
                ~ .x > 0)) %>%
  nrow()
data_validation_S2 %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  # Data filtering
  # Remove rows with missing values in predictors
  dplyr::filter(if_all(all_of(vars_RF_months), ~ !is.na(.))) %>%
  # Remove all rows with wrong values of indices (not between -1 and 1)
  dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
  dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
  dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
  dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
  dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
  # Select only variables needed
  dplyr::select(PlotObservationID, EUNISa_1, all_of(vars_RF_months), Lctnmth) %>%
  # Keep only rows with differences > 0 in variables
  filter(if_all(any_of(vars_RF_months[grepl("diff", vars_RF_months)]), 
                ~ .x > 0)) %>%
  nrow()
data_validation_SatEmb %>%
  filter(if_any(c(2:65), ~ !is.na(.))) %>%
  # Data filtering
  # Remove rows with missing values in predictors
  dplyr::filter(if_all(all_of(vars_RF_SatEmb), ~ !is.na(.))) %>%
  # Select only variables needed
  dplyr::select(PlotObservationID, EUNISa_1, all_of(vars_RF_SatEmb), 
                Lctnmth) %>%
  nrow()
```

## N obs after rough validation

```{r}
data_validation_Landsat %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>% 
  # Data filtering
  # Remove rows with missing values in predictors
  dplyr::filter(if_all(all_of(vars_RF_slope_Landsat), ~ !is.na(.))) %>%
  # Remove all rows with wrong values of indices (not between -1 and 1)
  dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
  # With Landsat data, many observations removed because of high/low EVI values
  # In Landsat_extraction, values of EVI outside of expected range were set as NA,
  # so better not to use EVI here
  # dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
  dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
  dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
  dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
  # Select only variables needed
  dplyr::select(PlotObservationID, EUNISa_1, all_of(vars_RF_slope_Landsat), 
                Lctnmth) %>%
  # Keep only rows with differences > 0 in variables
  filter(if_all(any_of(vars_RF_slope_Landsat[grepl("diff", 
                                                   vars_RF_slope_Landsat)]), 
                ~ .x > 0)) %>%
  # Rough validation based on NDWI
  rough_validation_Landsat() %>%
  nrow()
data_validation_Landsat %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.)))%>% 
  # Data filtering
  # Remove rows with missing values in predictors
  dplyr::filter(if_all(all_of(vars_RF_months_Landsat), ~ !is.na(.))) %>%
  # Remove all rows with wrong values of indices (not between -1 and 1)
  dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
  # With Landsat data, many observations removed because of high/low EVI values
  # In Landsat_extraction, values of EVI outside of expected range were set as NA,
  # so better not to use EVI here
  # dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
  dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
  dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
  dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
  # Select only variables needed
  dplyr::select(PlotObservationID, EUNISa_1, all_of(vars_RF_months_Landsat), 
                Lctnmth) %>%
  # Keep only rows with differences > 0 in variables
  filter(if_all(any_of(vars_RF_months_Landsat[grepl("diff", 
                                                   vars_RF_months_Landsat)]), 
                ~ .x > 0)) %>%
  # Rough validation based on NDWI
  rough_validation_Landsat() %>%
  nrow()
data_validation_S2 %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  # Data filtering
  # Remove rows with missing values in predictors
  dplyr::filter(if_all(all_of(vars_RF_slope), ~ !is.na(.))) %>%
  # Remove all rows with wrong values of indices (not between -1 and 1)
  dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
  dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
  dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
  dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
  dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
  # Select only variables needed
  dplyr::select(PlotObservationID, EUNISa_1, all_of(vars_RF_slope), Lctnmth) %>%
  # Keep only rows with differences > 0 in variables
  filter(if_all(any_of(vars_RF_slope[grepl("diff", vars_RF_slope)]), 
                ~ .x > 0)) %>%
  # Rough validation based on CH and NDWI
  rough_validation_S2() %>%
  nrow()
data_validation_S2 %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  # Data filtering
  # Remove rows with missing values in predictors
  dplyr::filter(if_all(all_of(vars_RF_months), ~ !is.na(.))) %>%
  # Remove all rows with wrong values of indices (not between -1 and 1)
  dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
  dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
  dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
  dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
  dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
  # Select only variables needed
  dplyr::select(PlotObservationID, EUNISa_1, all_of(vars_RF_months), Lctnmth) %>%
  # Keep only rows with differences > 0 in variables
  filter(if_all(any_of(vars_RF_months[grepl("diff", vars_RF_months)]), 
                ~ .x > 0)) %>%
  # Rough validation based on CH and NDWI
  rough_validation_S2() %>%
  nrow()
data_validation_SatEmb %>%
  filter(if_any(c(2:65), ~ !is.na(.))) %>%
  # Data filtering
  # Remove rows with missing values in predictors
  dplyr::filter(if_all(all_of(vars_RF_SatEmb), ~ !is.na(.))) %>%
  # Select only variables needed
  dplyr::select(PlotObservationID, EUNISa_1, all_of(vars_RF_SatEmb_CH), 
                Lctnmth) %>%
  # Rough validation based on CH
  rough_validation_SatEmb() %>% 
  # Join with S2 data to get NDWI_max info
  left_join(
    data_validation_S2 %>% select(PlotObservationID, NDWI_max)
  ) %>%
  # Rough validation based on NDWI (from S2)
  # Remove those with NDWI_max > 0.3, keep if NDWI_max is NA
  filter(NDWI_max <= 0.3 | is.na(NDWI_max)) %>%
  # Remove unneeded var
  select(-NDWI_max) %>%
  nrow()
```

## N obs GPS-lcoated

```{r}
data_validation_Landsat %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  # Rough validation based on NDWI
  rough_validation_Landsat() %>% 
  filter_data_Landsat(vars_RF_slope_Landsat) %>% 
  nrow()
data_validation_Landsat %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  # Rough validation based on NDWI
  rough_validation_Landsat() %>% 
  filter_data_Landsat(vars_RF_months_Landsat) %>% 
  nrow()
data_validation_S2 %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  # Rough validation based on CH and NDWI
  rough_validation_S2() %>%
  filter_data_S2(vars_RF_slope) %>%
  nrow()
data_validation_S2 %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  # Rough validation based on CH and NDWI
  rough_validation_S2() %>%
  filter_data_S2(vars_RF_months) %>%
  nrow()
data_validation_SatEmb %>%
  filter(if_any(c(2:65), ~ !is.na(.))) %>%
  # Rough validation based on CH
  rough_validation_SatEmb()%>%
  # Join with S2 data to get NDWI_max info
  left_join(
    data_validation_S2 %>% select(PlotObservationID, NDWI_max)
  ) %>%
  # Rough validation based on NDWI (from S2)
  # Remove those with NDWI_max > 0.3, keep if NDWI_max is NA
  filter(NDWI_max <= 0.3 | is.na(NDWI_max)) %>%
  # Remove unneeded var
  select(-NDWI_max) %>%
  filter_data_SatEmb(vars_RF_SatEmb) %>% 
  nrow()
```

## N obs after NDVI-filtering

```{r}
data_validation_Landsat %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  # Rough validation based on NDWI
  rough_validation_Landsat() %>% 
  filter_data_Landsat(vars_RF_slope_Landsat) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value") %>% 
  nrow()
data_validation_Landsat %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  # Rough validation based on NDWI
  rough_validation_Landsat() %>% 
  filter_data_Landsat(vars_RF_months_Landsat) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value") %>% 
  nrow()
data_validation_S2 %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  # Rough validation based on CH and NDWI
  rough_validation_S2() %>%
  filter_data_S2(vars_RF_slope) %>%
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value") %>% 
  nrow()
data_validation_S2 %>%
  filter(if_any(c(2:86, 91:132), ~ !is.na(.))) %>%
  # Rough validation based on CH and NDWI
  rough_validation_S2() %>%
  filter_data_S2(vars_RF_months) %>%
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value") %>% 
  nrow()
data_validation_SatEmb %>%
  filter(if_any(c(2:65), ~ !is.na(.))) %>%
  # Rough validation based on CH
  rough_validation_SatEmb()%>%
  # Join with S2 data to get NDWI_max info
  left_join(
    data_validation_S2 %>% select(PlotObservationID, NDWI_max)
  ) %>%
  # Rough validation based on NDWI (from S2)
  # Remove those with NDWI_max > 0.3, keep if NDWI_max is NA
  filter(NDWI_max <= 0.3 | is.na(NDWI_max)) %>%
  # Remove unneeded var
  select(-NDWI_max) %>%
  filter_data_SatEmb(vars_RF_SatEmb) %>%
  # Join with S2 data to get NDVI_pos_value info
  left_join(
    data_validation_S2 %>% select(PlotObservationID, NDVI_pos_value)
  ) %>%
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value") %>%
  # Remove unneeded var
  select(-NDVI_pos_value) %>%
  nrow()
```

# Session info

```{r}
sessionInfo()
```
