---
title: "Script to work with Landsat bands derived from GEE"
subtitle: "Read and manipulation data, calculate indices"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

# Load libraries

```{r}
library(signal)
library(tidyverse)
library(here)
library(lubridate)
library(dtplyr)
library(sf)
library(knitr)
library(mgcv)
library(future)
library(furrr)
library(progressr)
library(pracma)
```

# Load functions

```{r}
knitr::purl("functions_20260225.Rmd", output = "functions_20260225.R")
source("functions_20260225.R")
```

# Supress package messages

```{r}
suppressPackageStartupMessages({
  library(mgcv)
  library(nlme)
})
```

# Load previously created objects

```{r}
load(file = "objects/data_RS_Landsat_bands_indices.Rdata")
load(file = "objects/GAM_data_Landsat.Rdata")
load(file = "objects/smoothed_data_Landsat.Rdata")
load(file = "objects/monthly_avg_indices_Landsat.Rdata")
```

# Load Resurvey db

```{r}
ReSurvey_data <- read_csv(
  here("data", "clean", "ReSurvey_QRST_data_V2025.2_20260304.csv")) %>%
  select(PlotObservationID, EUNIS1, EUNIS2) %>%
  mutate(PlotObservationID = factor(PlotObservationID),
         EUNIS1 = factor(EUNIS1), EUNIS2 = factor(EUNIS2))
```

# Read files with band data

I got these files using the GEE code prepared by Bea.

These files contain all observations in the ReSurvey database from all years. In order to avoid computation problems in GEE, biogeographical units that contain more than 4500 points have been subdivided in ArcGIS.

## From the observation year

```{r message=FALSE}
# Set the folder path
folder_path <- "C:/Data/MOTIVATE/MOTIVATE_RS_data/validation/Landsat/obs_year"

# List CSV files
csv_files <- list.files(folder_path, full.names = TRUE, recursive = TRUE)

# Read one file to get column names
sample_file <- csv_files[1]
sample_data <- read_csv(sample_file, n_max = 0)

# Read and process each file
data_list <- lapply(csv_files, function(file) {
  info <- extract_info(basename(file)) # Use only the filename
  
  # Read the file
  df <- read_csv(file) %>%
    mutate(biogeo = info$biogeo, unit = info$unit)
  
  return(df)
  })

# Combine all data
data_RS_Landsat_bands_obs <- bind_rows(data_list) %>%
  rename(PlotObservationID = PltObID) %>%
  select(-UNIT)

# View the resulting tibble
print(data_RS_Landsat_bands_obs)

# Counts per biogeo and unit
print(data_RS_Landsat_bands_obs %>% count(biogeo, unit), n = 100)
```

## From the previous year

```{r message=FALSE}
# Set the folder path
folder_path <- "C:/Data/MOTIVATE/MOTIVATE_RS_data/validation/Landsat/prev_year"

# List CSV files
csv_files <- list.files(folder_path, full.names = TRUE, recursive = TRUE)

# Read one file to get column names
sample_file <- csv_files[1]
sample_data <- read_csv(sample_file, n_max = 0)

# Read and process each file
data_list <- lapply(csv_files, function(file) {
  info <- extract_info(basename(file)) # Use only the filename
  
  # Read the file
  df <- read_csv(file) %>%
    mutate(biogeo = info$biogeo, unit = info$unit)
  
  return(df)
  })

# Combine all data
data_RS_Landsat_bands_prev <- bind_rows(data_list) %>%
  rename(PlotObservationID = PltObID) %>%
  select(-UNIT)

# View the resulting tibble
print(data_RS_Landsat_bands_prev)

# Counts per biogeo and unit
print(data_RS_Landsat_bands_prev %>% count(biogeo, unit), n = 100)
```

## From the posterior year

```{r message=FALSE}
# Set the folder path
folder_path <- "C:/Data/MOTIVATE/MOTIVATE_RS_data/validation/Landsat/post_year"

# List CSV files
csv_files <- list.files(folder_path, full.names = TRUE, recursive = TRUE)

# Read one file to get column names
sample_file <- csv_files[1]
sample_data <- read_csv(sample_file, n_max = 0)

# Read and process each file
data_list <- lapply(csv_files, function(file) {
  info <- extract_info(basename(file)) # Use only the filename
  
  # Read the file
  df <- read_csv(file) %>%
    mutate(biogeo = info$biogeo, unit = info$unit)
  
  return(df)
  })

# Combine all data
data_RS_Landsat_bands_post <- bind_rows(data_list) %>%
  rename(PlotObservationID = PltObID) %>%
  select(-UNIT)

# View the resulting tibble
print(data_RS_Landsat_bands_post)

# Counts per biogeo and unit
print(data_RS_Landsat_bands_post %>% count(biogeo, unit), n = 100)
```

# Read file with band data for points outside polygons

Some points from the ReSurvey database fall outside the bioregion polygons. These points were manually assigned to the closest biogeo and unit in ArcGIS (those far out were assigned biogeo = OUT and unit = OUT). Data for these points has been extracted with a different GEE script.

## From the observation year

```{r}
data_RS_Landsat_bands_out_obs <- read_csv(
  "C:/Data/MOTIVATE/MOTIVATE_RS_data/validation/Landsat_out/AllPoints_Landsat_Annual.csv") %>%
  rename(biogeo = BIOGEO, unit = UNIT, PlotObservationID = PltObID)
```

Check that column names are the same:

```{r}
# Columns in data_RS_Landsat_bands_obs but not in data_RS_Landsat_bands_out
setdiff(names(data_RS_Landsat_bands_obs), names(data_RS_Landsat_bands_out_obs))

# Columns in data_RS_Landsat_bands_out but not in data_RS_Landsat_bands_obs
setdiff(names(data_RS_Landsat_bands_out_obs), names(data_RS_Landsat_bands_obs))
```

## From the previous year

```{r}
data_RS_Landsat_bands_out_prev <- read_csv(
  "C:/Data/MOTIVATE/MOTIVATE_RS_data/validation/Landsat_out/AllPoints_Landsat_PREVIOUS_YEAR.csv") %>%
  rename(biogeo = BIOGEO, unit = UNIT, PlotObservationID = PltObID)
```

Check that column names are the same:

```{r}
# Columns in data_RS_Landsat_bands_prev but not in data_RS_Landsat_bands_out_prev
setdiff(names(data_RS_Landsat_bands_prev), names(data_RS_Landsat_bands_out_prev))

# Columns in data_RS_Landsat_bands_out_prev but not in data_RS_Landsat_bands_prev
setdiff(names(data_RS_Landsat_bands_out_prev), names(data_RS_Landsat_bands_prev))
```

## From the posterior year

```{r}
data_RS_Landsat_bands_out_post <- read_csv(
  "C:/Data/MOTIVATE/MOTIVATE_RS_data/validation/Landsat_out/AllPoints_Landsat_POSTERIOR_YEAR.csv") %>%
  rename(biogeo = BIOGEO, unit = UNIT, PlotObservationID = PltObID)
```

Check that column names are the same:

```{r}
# Columns in data_RS_Landsat_bands_post but not in data_RS_Landsat_bands_out_post
setdiff(names(data_RS_Landsat_bands_post), names(data_RS_Landsat_bands_out_post))

# Columns in data_RS_Landsat_bands_out_post but not in data_RS_Landsat_bands_post
setdiff(names(data_RS_Landsat_bands_out_post), names(data_RS_Landsat_bands_post))
```

# Join all data from inside polygons

```{r}
data_RS_Landsat_bands <- bind_rows(data_RS_Landsat_bands_obs,
                                   data_RS_Landsat_bands_prev,
                                   data_RS_Landsat_bands_post) %>%
  mutate(image_year = case_when(!is.na(image_year) ~ image_year,  
                                TRUE ~ year(date)
                                ),
         point_year = case_when(!is.na(point_year) ~ point_year,  
                                TRUE ~ year),
         year_cat = case_when(image_year == point_year ~ "obs",
                              image_year < point_year ~ "prev",
                              image_year > point_year ~ "post",
                              TRUE ~ NA_character_)
         )
```

# Join all data from outside polygons

```{r}
data_RS_Landsat_bands_out <- bind_rows(data_RS_Landsat_bands_out_obs,
                                       data_RS_Landsat_bands_out_prev,
                                       data_RS_Landsat_bands_out_post) %>%
  mutate(image_year = case_when(!is.na(image_year) ~ image_year,  
                                TRUE ~ year(date)
                                ),
         point_year = case_when(!is.na(point_year) ~ point_year,  
                                TRUE ~ year),
         year_cat = case_when(image_year == point_year ~ "obs",
                              image_year < point_year ~ "prev",
                              image_year > point_year ~ "post",
                              TRUE ~ NA_character_)
         )
```

# Join all data

```{r}
data_RS_Landsat_bands_all <- bind_rows(data_RS_Landsat_bands, 
                                       data_RS_Landsat_bands_out) %>% 
  select(-location)
```

# Some checks

Check that the year in the date of the images is not different to the sampling year:

```{r}
data_RS_Landsat_bands %>% dplyr::filter(year != year(date) & year_cat == "obs")
```

Check how many different images are for each observation, date and time:

```{r}
data_RS_Landsat_bands %>% group_by(PlotObservationID, date, time_utc) %>%
  summarise(n_images = n_distinct(image_id), .groups = "drop") %>%
  count(n_images)
```

# Average the bands

When there is more than one image for each point and day, average the values of the bands:

```{r eval=FALSE, include=FALSE}
band_summary <- data_RS_Landsat_bands %>%
  group_by(PlotObservationID, date) %>%
  summarise(
    n_images = n_distinct(image_id),
    Blue = if (n_images > 1) mean(Blue, na.rm = TRUE) else first(Blue),
    Green  = if (n_images > 1) mean(Green,  na.rm = TRUE) else first(Green),
    NIR  = if (n_images > 1) mean(NIR,  na.rm = TRUE) else first(NIR),
    Red  = if (n_images > 1) mean(Red,  na.rm = TRUE) else first(Red),
    SWIR1 = if (n_images > 1) mean(SWIR1,  na.rm = TRUE) else first(SWIR1),
    .groups = "drop"
  ) 

# Calculate how many different DOYs for each PlotObservationID
n_days <- band_summary %>%
  mutate(DOY = yday(date)) %>%
  group_by(PlotObservationID) %>%
  summarise(n_days = n_distinct(DOY), .groups = "drop")

# Join back to original data
data_RS_Landsat_bands_updated <- data_RS_Landsat_bands %>%
  # Remove old band values
  select(-Blue, -Green, -NIR, -Red, -SWIR1) %>%
  # Join band_summary
  left_join(band_summary, by = c("PlotObservationID", "date")) %>%
  # Keep one row per group
  distinct(PlotObservationID, date, .keep_all = TRUE) %>%
  # Remove unwanted columns
  select(-`system:index`, -image_id, -.geo, -time_utc, -timestamp) %>%
  # Join
  left_join(n_days)
```

# Calculate indices

```{r eval=FALSE, include=FALSE}
# Calculate indices
data_RS_Landsat_bands_indices <- data_RS_Landsat_bands_updated %>%
  # Set PlotObservationID as factor
  mutate(PlotObservationID = factor(PlotObservationID)) %>%
  # Rename the bands
  rename(blue = Blue, green = Green, red = Red, NIR = NIR, SWIR = SWIR1) %>%
  # Scale the bands
  mutate(blue = blue / 10000, green = green / 10000, red = red / 10000,
         NIR = NIR / 10000, SWIR = SWIR / 10000) %>%
  # Create column that combines the day of the month and the time
  mutate(
    date = as.POSIXct(date),
    # Normalize the dates to a fixed year (2000)
    # so that seasonal patterns across different years can be compared visually
    day_month = as.POSIXct(format(date, "2000-%m-%d"))) %>%
  # Create column with DOY
  mutate(DOY = yday(date)) %>%
  # Calculate NDVI
  mutate(NDVI = (NIR - red) / (NIR + red),
         EVI = (NIR - red) * 2.5 / (NIR + 6 * red - 7.5 * blue + 1),
         SAVI = (NIR - red) * 1.5 / (NIR + red + 0.5),
         NDMI = (NIR - SWIR) / (NIR + SWIR),
         NDWI = (green - NIR) / (green + NIR)) %>%
  # Setting values of indices outside expected ranges (errors) to NA
  mutate(EVI = if_else(EVI > 1 | EVI < -1, NA, EVI)) # 3066726  values of EVI as NA!
```

Save:

```{r eval=FALSE, include=FALSE}
save(data_RS_Landsat_bands_indices, file = "objects/data_RS_Landsat_bands_indices.Rdata")
```

Plot n_daytime:

```{r}
data_RS_Landsat_bands_indices %>%
  group_by(PlotObservationID) %>%
  summarise(n_days = first(n_days)) %>% ungroup() %>%
  ggplot(aes(x = n_days)) + geom_histogram(color = "black", fill = "white") +
  geom_vline(xintercept = 10, color = "red", linewidth = 1) +
  theme_minimal()
```

# Compute phenological metrics from models fitted to time series data

## Calculation

Apply the function with batch processing

```{r eval=FALSE, message=FALSE, include=FALSE}
plan(multisession, workers = 24)

ids <- unique(data_RS_Landsat_bands_indices$PlotObservationID)
batches <- split(ids, ceiling(seq_along(ids) / 50))  # batches of 50

start_total <- Sys.time()

GAM_data <- map_dfr(seq_along(batches), function(i) {
  batch_ids <- batches[[i]]
  total_batches <- length(batches)
  batch_file <- file.path("objects/GAM_batches_Landsat", paste0("batch_", i, ".rds"))

  if (file.exists(batch_file)) {
    message("✅ Batch  ", i, " of ", total_batches, 
            " already processed. Loading from file.")
    return(readRDS(batch_file))
  }

  message("🔄 Processing batch  ", i, " of ", total_batches, " with ",
          length(batch_ids), " IDs...")

  start_batch <- Sys.time()

  result <- data_RS_Landsat_bands_indices %>%
    filter(PlotObservationID %in% batch_ids) %>%
    group_split(PlotObservationID) %>%
    set_names(map_chr(., ~ as.character(unique(.x$PlotObservationID)))) %>%
    future_map_dfr(~ compute_metrics_models(df = .,
                                            index_cols = c("NDVI", "EVI", "SAVI")),
                   .progress = TRUE)

  end_batch <- Sys.time()
  duration <- round(difftime(end_batch, start_batch, units = "mins"), 2)
  message("⏱️ Batch time ", i, ": ", duration, " minutes")

  message("💾 Saving batch ", i, " to file...")
  saveRDS(result, batch_file)
  message("✅ Batch ", i, " saved.") 

  result
})

end_total <- Sys.time()
total_time <- round(difftime(end_total, start_total, units = "mins"), 2)
message("⏱️ Total time: ", total_time, " minutes")
```

```{r}
plan(sequential)
```

## Save

Look:

```{r}
GAM_data
```

Save as an object:

```{r eval=FALSE, include=FALSE}
save(GAM_data, file = "objects/GAM_data_Landsat.Rdata")
```

Add year to GAM_data:

```{r}
GAM_data <- GAM_data %>%
  left_join(data_RS_Landsat_bands_indices %>%
              select(PlotObservationID, year) %>%
              distinct(),
            by = "PlotObservationID")
```

## Extract average values of indices per month

Extract average values of indices per month and AUC between March and October;

```{r eval=FALSE, include=FALSE}
monthly_avg_indices <- extract_monthly_avg_indices(GAM_data)
```

Save as an object:

```{r eval=FALSE, include=FALSE}
save(monthly_avg_indices, file = "objects/monthly_avg_indices_Landsat.Rdata")
```

## Assess time series quality

For the time series to be acceptable, it should have a reasonable number of time points, and these points should be distributed along almost all months (could be ok to miss the winter months).

In GAM data, check how many time points are there for each PlotObservationID, how many months, and which months are missing.

```{r}
ts_quality <- GAM_data %>%
  # Filter only NDVI (all indices will have the same time points)
  dplyr::filter(index == "NDVI") %>%
  # Get month from DOY
  mutate(month = month(ymd("2020-01-01") + days(DOY - 1))) %>%
  # For each PlotObservationID
  group_by(PlotObservationID) %>%
  # Get the number of time points (days) and the number of months
  summarise(
    n_days = n_distinct(DOY),
    n_months = n_distinct(month),
    .groups = "drop"
  ) %>%
  left_join(GAM_data %>%
              # Filter only NDVI
              dplyr::filter(index == "NDVI") %>%
              # Get month from DOY
              mutate(month = month(ymd("2020-01-01") + days(DOY - 1))) %>%
              # Get unique values of PlotObservationID and month
              distinct(PlotObservationID, month) %>%
              # Add 1 as value
              mutate(value = 1) %>%
              # Reshape to wide format and add zeros when month is missing
              pivot_wider(
                names_from = month,
                names_prefix = "month",
                values_from = value,
                values_fill = 0),
            by = "PlotObservationID")
```

Histograms time points and n months:

```{r}
ggplot(ts_quality, aes(x = n_days)) +
  geom_histogram(color = "black", fill = "white") +
  xlab("Number of time points (days) in the Landsat time series") +
  theme_minimal()
ggplot(ts_quality, aes(x = n_months)) +
  geom_histogram(color = "black", fill = "white") +
  xlab("Number of months in the Landsat time series") +
  theme_minimal()
```

Count how many PlotObservationIDs have missing data (value 0) for each month:

```{r}
obs_missing_month <- ts_quality %>%
  summarise(across(starts_with("month"), ~ sum(.x == 0))) %>%
  pivot_longer(cols = everything(), names_to = "month", values_to = "nobs_missing")

ggplot(obs_missing_month %>%
         mutate(month = factor(month, levels = paste0("month", 1:12))), 
       aes(x = month, y = nobs_missing)) + geom_bar(stat = "identity") +
  ylab("Number of PlotObservationID with missing data") +
  ggtitle("Missing data in Landsat time series") +
  theme_minimal()
```

Add quality flag:

```{r}
ts_quality_flag <- ts_quality %>%
  rowwise() %>%
  mutate(
    #  If 2 consecutive months of the period March-October are missing
    # quality_flag = 0
    quality_flag = {
      months <- c_across(month3:month10)
      if (any(months[-length(months)] == 0 & months[-1] == 0)) 0 else 1
    }
  ) %>%
  ungroup()
```

```{r}
ts_quality_flag %>% count(quality_flag)
```

## Boxplot comparing moments for different indices

```{r}
GAM_data %>% 
  select(PlotObservationID, index, sos_slope, pos, eos_slope) %>% 
  distinct() %>%
  pivot_longer(cols = c(sos_slope, pos, eos_slope),
               names_to = "moment", values_to = "value") %>%
  ggplot(aes(x = moment, y = value, fill = index)) + geom_boxplot() +
  theme_minimal()
```

## Plot fit and moments for each PlotObservationID

### Quality = 1

```{r}
# Get unique IDs with quality_flag == 1
ids_q1 <- ts_quality_flag %>%
  dplyr::filter(quality_flag == 1) %>%
  mutate(PlotObservationID = droplevels(PlotObservationID)) %>%
  pull(PlotObservationID)
GAM_data_ids_q1 <- GAM_data %>%
  # Join to get biogeo and unit
  left_join(data_RS_Landsat_bands_indices %>%
              select(PlotObservationID, biogeo, unit) %>%
              distinct()) %>%
  # Join to get EUNIS info
   left_join(ReSurvey_data %>%
               select(PlotObservationID, EUNIS1, EUNIS2)) %>%
  # Join to get original values of indices
  left_join(data_RS_Landsat_bands_indices %>%
              select(PlotObservationID, DOY, NDVI, EVI, SAVI) %>%
              pivot_longer(cols = c(NDVI, EVI, SAVI), names_to = "index", 
                           values_to = "value_orig")) %>%
  # Join to get ts_quality data
  left_join(ts_quality_flag %>% select(PlotObservationID, quality_flag)) %>%
  # Keep only those with quality_flag == 1
  dplyr::filter(quality_flag == 1)
```

```{r eval=FALSE, include=FALSE}
# Get unique PlotObservationIDs
unique_ids1 <- ids_q1[1:10]

# Create and store plots in a list
ts_plots_q1 <- map(unique_ids1, function(id) {
  plot_data <- GAM_data_ids_q1 %>%
    mutate(PlotObservationID = as.character(PlotObservationID)) %>%
    dplyr::filter(PlotObservationID == id) 
  
  # Extract metadata for title
  metadata <- plot_data %>%
    select(biogeo, unit, EUNIS1, EUNIS2, quality_flag) %>%
    distinct()
  
  ggplot() +
    # Raw data points
    geom_point(data = plot_data,aes(x = DOY, y = value_orig), alpha = 0.5) +
    geom_line(data = plot_data, aes(x = DOY, y = value), 
              size = 0.5, color = "blue") +
    geom_vline(data = plot_data %>% distinct(index, sos_slope),
               aes(xintercept = sos_slope, group = index),
               linetype = "dashed", size = 0.5, color = "red") +
    geom_vline(data = plot_data %>% distinct(index, pos),
               aes(xintercept = pos, group = index),
               linetype = "dotted", size = 0.5, color = "blue") +
    geom_vline(data = plot_data %>% distinct(index, eos_slope),
               aes(xintercept = eos_slope, group = index),
               linetype = "dashed", size = 0.5, color = "red") +
    facet_grid(cols = vars(index)) +
    labs(
      title = glue::glue("{id} | {metadata$biogeo}{metadata$unit} | {metadata$EUNIS1} | {metadata$EUNIS2} | Quality: {metadata$quality_flag}"),
      x = "Day of Year",
      y = "Index Value"
    ) +
    theme_minimal() + theme(legend.position = "top")
})

# Name the list by PlotObservationID
names(ts_plots_q1) <- unique_ids1

# Display the first plot
ts_plots_q1[1]
```

Save each plot to a file:

```{r eval=FALSE, include=FALSE}
walk2(ts_plots_q1, seq_along(ts_plots_q1), ~ ggsave(
  filename = paste0("C:/Analyses/MOTIVATE_validation/output/figures/phenology/ts_q1_Landsat/ts_plots_q1", .y, ".jpeg"),
  plot = .x,
  width = 8,
  height = 5
))
```

### Quality = 0

```{r}
# Get unique IDs with quality_flag == 0
ids_q0 <- ts_quality_flag %>%
  dplyr::filter(quality_flag == 0) %>%
  mutate(PlotObservationID = droplevels(PlotObservationID)) %>%
  pull(PlotObservationID)
GAM_data_ids_q0 <- GAM_data %>%
  # Join to get biogeo and unit
  left_join(data_RS_Landsat_bands_indices %>%
              select(PlotObservationID, biogeo, unit) %>%
              distinct()) %>%
  # Join to get EUNIS info
   left_join(ReSurvey_data %>%
               select(PlotObservationID, EUNIS1, EUNIS2)) %>%
  # Join to get original values of indices
  left_join(data_RS_Landsat_bands_indices %>%
              select(PlotObservationID, DOY, NDVI, EVI, SAVI) %>%
              pivot_longer(cols = c(NDVI, EVI, SAVI), names_to = "index", 
                           values_to = "value_orig")) %>%
  # Join to get ts_quality data
  left_join(ts_quality_flag %>%
              select(PlotObservationID, n_months, quality_flag)) %>%
  # Keep only those with quality_flag == 0
  dplyr::filter(quality_flag == 0)
```

```{r eval=FALSE, include=FALSE}
# Get unique PlotObservationIDs
unique_ids0 <- ids_q0[1:10]

# Create and store plots in a list
ts_plots_q0 <- map(unique_ids0, function(id) {
  plot_data <- GAM_data_ids_q0 %>%
    mutate(PlotObservationID = as.character(PlotObservationID)) %>%
    dplyr::filter(PlotObservationID == id) 
  
  # Extract metadata for title
  metadata <- plot_data %>%
    select(biogeo, unit, EUNIS1, EUNIS2, quality_flag) %>%
    distinct()
  
  ggplot() +
    # Raw data points
    geom_point(data = plot_data,aes(x = DOY, y = value_orig), alpha = 0.5) +
    geom_line(data = plot_data, aes(x = DOY, y = value), 
              size = 0.5, color = "blue") +
    geom_vline(data = plot_data %>% distinct(index, sos_slope),
               aes(xintercept = sos_slope, group = index),
               linetype = "dashed", size = 0.5, color = "red") +
    geom_vline(data = plot_data %>% distinct(index, pos),
               aes(xintercept = pos, group = index),
               linetype = "dotted", size = 0.5, color = "blue") +
    geom_vline(data = plot_data %>% distinct(index, eos_slope),
               aes(xintercept = eos_slope, group = index),
               linetype = "dashed", size = 0.5, color = "red") +
    facet_grid(cols = vars(index)) +
    labs(
      title = glue::glue("{id} | {metadata$biogeo}{metadata$unit} | {metadata$EUNIS1} | {metadata$EUNIS2} | Quality: {metadata$quality_flag}"),
      x = "Day of Year",
      y = "Index Value"
    ) +
    theme_minimal() + theme(legend.position = "top")
})

# Name the list by PlotObservationID
names(ts_plots_q0) <- unique_ids0

# Display the first plot
ts_plots_q0[1]
```

Save each plot to a file:

```{r eval=FALSE, include=FALSE}
walk2(ts_plots_q0, seq_along(ts_plots_q0), ~ ggsave(
  filename = paste0("C:/Analyses/MOTIVATE_validation/output/figures/phenology/ts_q0_Landsat/ts_plots_q0", .y, ".jpeg"),
  plot = .x,
  width = 8,
  height = 5
))
```

# Smooth the time series of NDMI and NDWI

Apply the function with batch processing.

```{r eval=FALSE, message=FALSE, include=FALSE}
plan(multisession, workers = 24)

ids <- unique(data_RS_Landsat_bands_indices$PlotObservationID)
batches <- split(ids, ceiling(seq_along(ids) / 50))  # batches of 50

start_total <- Sys.time()

smoothed_data <- map_dfr(seq_along(batches), function(i) {
  batch_ids <- batches[[i]]
  total_batches <- length(batches)
  batch_file <- file.path("objects/smoothing_batches_Landsat", paste0("batch_", i, ".rds"))

  if (file.exists(batch_file)) {
    message("✅ Batch  ", i, " of ", total_batches, 
            " already processed. Loading from file.")
    return(readRDS(batch_file))
  }

  message("🔄 Processing batch  ", i, " of ", total_batches, " with ",
          length(batch_ids), " IDs...")

  start_batch <- Sys.time()

  result <- data_RS_Landsat_bands_indices %>%
    filter(PlotObservationID %in% batch_ids) %>%
    group_split(PlotObservationID) %>%
    set_names(map_chr(., ~ as.character(unique(.x$PlotObservationID)))) %>%
    future_map_dfr(~ compute_unweighted_fit(df = .,
                                            index_cols = c("NDMI", "NDWI")),
                   .progress = TRUE)

  end_batch <- Sys.time()
  duration <- round(difftime(end_batch, start_batch, units = "mins"), 2)
  message("⏱️ Batch time ", i, ": ", duration, " minutes")

  message("💾 Saving batch ", i, " to file...")
  saveRDS(result, batch_file)
  message("✅ Batch ", i, " saved.") 

  result
})

end_total <- Sys.time()
total_time <- round(difftime(end_total, start_total, units = "mins"), 2)
message("⏱️ Total time: ", total_time, " minutes")
```

```{r}
plan(sequential)
```

## Save

Look:

```{r}
smoothed_data
```

Save as an object:

```{r eval=FALSE, include=FALSE}
save(smoothed_data, file = "objects/smoothed_data_Landsat.Rdata")
```

## TBD if needed: Plot fit and moments for each PlotObservationID

```{r}
smoothed_data_ids <- smoothed_data %>%
  # Join to get biogeo and unit
  left_join(data_RS_Landsat_bands_indices %>%
              select(PlotObservationID, biogeo, unit) %>%
              distinct()) %>%
  # Join to get EUNIS info
   left_join(ReSurvey_data %>%
               select(PlotObservationID, EUNIS1,EUNIS2)) %>%
  mutate(PlotObservationID = as.character(PlotObservationID))
```

```{r eval=FALSE, include=FALSE}
# Get unique PlotObservationIDs
unique_ids <- unique(smoothed_data_ids$PlotObservationID)[1:10]

# Create and store plots in a list
ts_plots_NDMI_NDWI<- map(unique_ids, function(id) {
  plot_data <- smoothed_data_ids %>% 
    dplyr::filter(PlotObservationID == id)
  
  # Extract metadata for title
  metadata <- plot_data %>%
    select(biogeo, unit, EUNIS1, EUNIS2) %>%
    distinct()
  
  ggplot() +
    # Raw data points
     geom_point(data = data_RS_Landsat_bands_indices %>%
                  select(PlotObservationID, DOY, NDMI, NDWI) %>%
                  pivot_longer(cols = c(NDMI, NDWI), names_to = "index",
                               values_to = "value") %>%
                  filter(PlotObservationID == id),
                aes(x = DOY, y = value), alpha = 0.6) +
    geom_line(data = plot_data, aes(x = DOY, y = value),
              size = 0.5, color = "blue") +
    facet_grid(cols = vars(index)) +
    labs(
      title = glue::glue("{id} | {metadata$biogeo}{metadata$unit} | {metadata$EUNIS1} | {metadata$EUNIS2}"),
      x = "Day of Year",
      y = "Index Value"
    ) +
    theme_minimal() + theme(legend.position = "top")
})

# Name the list by PlotObservationID
names(ts_plots_NDMI_NDWI) <- unique_ids

# Display the first plot
print(ts_plots_NDMI_NDWI)
```

Save each plot to a file:

```{r eval=FALSE, include=FALSE}
walk2(ts_plots_NDMI_NDWI, seq_along(ts_plots_NDMI_NDWI), ~ ggsave(
  filename = paste0("C:/Analyses/MOTIVATE_validation/output/figures/phenology/ts_NDMI_NDWI_Landsat/ts_plots_NDVI_NDMI",
                    .y, ".jpeg"),
  plot = .x,
  width = 8,
  height = 5
))
```

# Get indices data (max. and min.)

These maximum and minimum values are from the smoothed time series. For values in DOY 1–50 and DOY 315–end, the GAM smoothing function replaced the original values with the mean base value of observations during each of these respective periods (this has now been done for all indices, i.e. NDVI, EVI, SAVI, NDMI and NDWI).

```{r}
final_indices_data <- GAM_data %>%
  group_by(PlotObservationID, index) %>%
  summarise(max = max(value), min = min(value)) %>%
  ungroup() %>%
  pivot_wider(names_from = index, values_from = c(max, min),
              names_glue = "{index}_{.value}") %>%
  full_join(
    smoothed_data %>%
      group_by(PlotObservationID, index) %>%
      summarise(max = max(value), min = min(value)) %>%
      ungroup() %>%
      pivot_wider(names_from = index, values_from = c(max, min),
                  names_glue = "{index}_{.value}")
    )
```

# Get phenology data

Use GAM iter_3 to get dates of the moments, values at those moments and AUC (time-integrated indices) between SOS and EOS:

```{r}
# Join to get values at SOS, POS, EOS and auc
final_phenology_data <- GAM_data %>%
  mutate(
    stage = case_when(
      DOY == sos_slope ~ "sos",
      DOY == pos ~ "pos",
      DOY == eos_slope ~ "eos",
      TRUE ~ NA_character_
    )
  ) %>%
  dplyr::filter(!is.na(stage)) %>%
  select(PlotObservationID, index, stage, doy = DOY, value) %>% 
  # Remove duplicate rows
  distinct() %>%
  pivot_wider(
    names_from = c(index, stage),
    values_from = c(doy, value),
    names_glue = "{index}_{stage}_{.value}"
  ) %>%
  # Convert list cols to regular numeric cols
  mutate(
    NDVI_sos_value = map_dbl(NDVI_sos_value, 1),
    NDVI_pos_value = map_dbl(NDVI_pos_value, 1),
    NDVI_eos_value = map_dbl(NDVI_eos_value, 1),
    EVI_sos_value = map_dbl(EVI_sos_value, 1),
    EVI_pos_value = map_dbl(EVI_pos_value, 1),
    EVI_eos_value = map_dbl(EVI_eos_value, 1),
    SAVI_sos_value = map_dbl(SAVI_sos_value, 1),
    SAVI_pos_value = map_dbl(SAVI_pos_value, 1),
    SAVI_eos_value = map_dbl(SAVI_eos_value, 1)
  ) %>%
  full_join(GAM_data %>%
              distinct(PlotObservationID, index, auc_slope) %>%
              rename(auc = auc_slope) %>%
              pivot_wider(names_from = index, values_from = c(auc),
                          names_glue = "{index}_{.value}"))
```

# Get values at POS-2 and POS+2

Alternative to use SOS/EOS or March/October: Use values at POS - 2 months and POS + 2 months.

```{r}
pos_pm60_df <- get_pos_minus_plus60(GAM_data, extrapolate = TRUE)
```

# Join indices and phenology data

```{r}
final_RS_data <- full_join(
  # Indices data (max and min)
  final_indices_data,
  # Average values of indices per month
  monthly_avg_indices %>%
    pivot_wider(names_from = index, values_from = c(avg_value_01:auc_mar_oct),
                names_glue = "{index}_{.value}")
  ) %>%
  full_join(
    # Phenology data
    final_phenology_data 
    ) %>%
   full_join(
     # Values at POS - 60 days and POS + 60 dAYS
     pos_pm60_df %>%
       select(-year, -pos, -target_doy_minus60, - target_doy_plus60) %>%
       pivot_wider(names_from = index, 
                   values_from = c(pos_minus60_value, pos_plus60_value, auc_pos2),
                   names_glue = "{index}_{.value}")
   ) %>%
  # Sort cols in alphabetical order
  select(PlotObservationID, sort(names(.)[names(.) != "PlotObservationID"]))
```

# Add EUNIS codes

```{r}
final_RS_data <- final_RS_data %>% left_join(ReSurvey_data) %>%
  filter(!is.na(EUNIS1))
  # 1 PlotObservationID with EUNIS1 as NA because it had multiple
  # assignations by the Expert system
```

```{r}
data_RS_Landsat_bands_indices <- data_RS_Landsat_bands_indices %>%
  left_join(ReSurvey_data) %>%
  filter(!is.na(EUNIS1))
  # 1 PlotObservationID with EUNIS1 as NA because it had multiple
  # assignations by the Expert system
```

# Monthly spectrophenology per habitat type

```{r}
# Prepare the data
data_monthly_EUNIS1 <- data_RS_Landsat_bands_indices %>%
  mutate(month = month(date, label = TRUE, abbr = TRUE, locale="EN-us")) %>%
  group_by(month, EUNIS1) %>%
  summarise(
    mean_NDVI = mean(NDVI, na.rm = TRUE),
    sd_NDVI = sd(NDVI, na.rm = TRUE),
    n_NDVI = sum(!is.na(NDVI)),
    mean_EVI = mean(EVI, na.rm = TRUE),
    sd_EVI = sd(EVI, na.rm = TRUE),
    n_EVI = sum(!is.na(EVI)),
    mean_SAVI = mean(SAVI, na.rm = TRUE),
    sd_SAVI = sd(SAVI, na.rm = TRUE),
    n_SAVI = sum(!is.na(SAVI)),
    .groups = "drop"
  )

data_monthly_EUNIS1 <- data_monthly_EUNIS1 %>%
  # Add label with n
  left_join(data_monthly_EUNIS1 %>%
              group_by(EUNIS1) %>%
              summarise(n_total = sum(n_NDVI, na.rm = TRUE)), 
            by = "EUNIS1") %>%
  mutate(EUNIS1_label = paste0(EUNIS1, " (n = ", n_total, ")"))

data_monthly_EUNIS2 <- data_RS_Landsat_bands_indices %>%
  mutate(month = month(date, label = TRUE, abbr = TRUE, locale="EN-us")) %>%
  group_by(month, EUNIS1, EUNIS2) %>%
  summarise(
    mean_NDVI = mean(NDVI, na.rm = TRUE),
    sd_NDVI = sd(NDVI, na.rm = TRUE),
    n_NDVI = sum(!is.na(NDVI)),
    mean_EVI = mean(EVI, na.rm = TRUE),
    sd_EVI = sd(EVI, na.rm = TRUE),
    n_EVI = sum(!is.na(EVI)),
    mean_SAVI = mean(SAVI, na.rm = TRUE),
    sd_SAVI = sd(SAVI, na.rm = TRUE),
    n_SAVI = sum(!is.na(SAVI)),
    .groups = "drop"
  )

data_monthly_EUNIS2 <- data_monthly_EUNIS2 %>%
  # Add label with n
  left_join(data_monthly_EUNIS2 %>%
              group_by(EUNIS2) %>%
              summarise(n_total = sum(n_NDVI, na.rm = TRUE)), 
            by = "EUNIS2") %>%
  mutate(EUNIS2_label = paste0(EUNIS2, " (n = ", n_total, ")"))
```

## EUNIS level 1

```{r}
# Plots

# EUNIS1
ggplot(data_monthly_EUNIS1, 
       aes(x = month, y = mean_NDVI, color = EUNIS1, 
           group = EUNIS1)) +
  geom_point() +
  geom_line(aes(group = EUNIS1)) +
  labs(
    title = "Monthly NDVI by Habitat Type",
    x = "Month",
    y = "NDVI",
    color = "Habitat (EUNIS1)"
  ) +
  theme_minimal()
ggsave(
  here("output", "figures", "monthly_spectrophenology_Landsat", "EUNIS1_NDVI.jpeg"),
  dpi = 300, width = 6, height = 2.5)

ggplot(data_monthly_EUNIS1, 
       aes(x = month, y = mean_EVI, color = EUNIS1, 
           group = EUNIS1)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Monthly EVI by Habitat Type",
    x = "Month",
    y = "EVI",
    color = "Habitat (EUNIS1)"
  ) +
  theme_minimal()
ggsave(
  here("output", "figures", "monthly_spectrophenology_Landsat", "EUNIS1_EVI.jpeg"),
  dpi = 300, width = 6, height = 2.5)

ggplot(data_monthly_EUNIS1, 
       aes(x = month, y = mean_SAVI, color = EUNIS1, 
           group = EUNIS1)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Monthly SAVI by Habitat Type",
    x = "Month",
    y = "SAVI",
    color = "Habitat (EUNIS1)"
  ) +
  theme_minimal()
ggsave(
  here("output", "figures", "monthly_spectrophenology_Landsat", "EUNIS1_SAVI.jpeg"),
  dpi = 300, width = 6, height = 2.5)
```

## EUNIS level 2

### Q

```{r}
# NDVI
ggplot(data_monthly_EUNIS2 %>% 
         dplyr::filter(EUNIS1 == "Q" & !is.na(EUNIS2)) %>%
         # Remove those with EUNIS level 2 that does not match current classif
         dplyr::filter(EUNIS2 != "Qa" & EUNIS2 != "Qb"), 
       aes(x = month, y = mean_NDVI, color = EUNIS2, group = EUNIS2)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Monthly NDVI by Habitat Type",
    x = "Month",
    y = "NDVI",
    color = "Habitat (EUNIS2)"
  ) +
  theme_minimal()
ggsave(
  here("output", "figures", "monthly_spectrophenology_Landsat", "EUNIS2_Q_NDVI.jpeg"),
  dpi = 300, width = 7, height = 2.5)

# EVI
ggplot(data_monthly_EUNIS2 %>% 
         dplyr::filter(EUNIS1 == "Q" & !is.na(EUNIS2)) %>%
         # Remove those with EUNIS level 2 that does not match current classif
         dplyr::filter(EUNIS2 != "Qa" & EUNIS2 != "Qb"), 
       aes(x = month, y = mean_EVI, color = EUNIS2, group = EUNIS2)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Monthly EVI by Habitat Type",
    x = "Month",
    y = "EVI",
    color = "Habitat (EUNIS2)"
  ) +
  theme_minimal()
ggsave(
  here("output", "figures", "monthly_spectrophenology_Landsat", "EUNIS2_Q_EVI.jpeg"),
  dpi = 300, width = 7, height = 2.5)

# SAVI
ggplot(data_monthly_EUNIS2 %>%
         dplyr::filter(EUNIS1 == "Q" & !is.na(EUNIS2)) %>%
         # Remove those with EUNIS level 2 that does not match current classif
         dplyr::filter(EUNIS2 != "Qa" & EUNIS2 != "Qb"), 
       aes(x = month, y = mean_SAVI, color = EUNIS2, group = EUNIS2)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Monthly SAVI by Habitat Type",
    x = "Month",
    y = "SAVI",
    color = "Habitat (EUNIS2)"
  ) +
  theme_minimal()
ggsave(
  here("output", "figures", "monthly_spectrophenology_Landsat", "EUNIS2_Q_SAVI.jpeg"),
  dpi = 300, width = 7, height = 2.5)
```

### R

```{r}
ggplot(data_monthly_EUNIS2 %>% 
         dplyr::filter(EUNIS1 == "R" & !is.na(EUNIS2)), 
       aes(x = month, y = mean_NDVI, color = EUNIS2, group = EUNIS2)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Monthly NDVI by Habitat Type",
    x = "Month",
    y = "NDVI",
    color = "Habitat (EUNIS2)"
  ) +
  theme_minimal()
ggsave(
  here("output", "figures", "monthly_spectrophenology_Landsat", "EUNIS2_R_NDVI.jpeg"),
  dpi = 300, width = 7, height = 2.5)

ggplot(data_monthly_EUNIS2 %>% 
         dplyr::filter(EUNIS1 == "R" & !is.na(EUNIS2)), 
       aes(x = month, y = mean_EVI, color = EUNIS2, group = EUNIS2)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Monthly EVI by Habitat Type",
    x = "Month",
    y = "EVI",
    color = "Habitat (EUNIS2)"
  ) +
  theme_minimal()
ggsave(
  here("output", "figures", "monthly_spectrophenology_Landsat", "EUNIS2_R_EVI.jpeg"),
  dpi = 300, width = 7, height = 2.5)

ggplot(data_monthly_EUNIS2 %>% 
         dplyr::filter(EUNIS1 == "R" & !is.na(EUNIS2)), 
       aes(x = month, y = mean_SAVI, color = EUNIS2, group = EUNIS2)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Monthly SAVI by Habitat Type",
    x = "Month",
    y = "SAVI",
    color = "Habitat (EUNIS2)"
  ) +
  theme_minimal()
ggsave(
  here("output", "figures", "monthly_spectrophenology_Landsat", "EUNIS2_R_SAVI.jpeg"),
  dpi = 300, width = 7, height = 2.5)
```

### S

```{r}
ggplot(data_monthly_EUNIS2 %>% 
         dplyr::filter(EUNIS1 == "S" & !is.na(EUNIS2)) %>%
         # Remove those with EUNIS level 2 that does not match current classif
         dplyr::filter(EUNIS2 != "Sa" & EUNIS2 != "Sb"), 
       aes(x = month, y = mean_NDVI, color = EUNIS2, group = EUNIS2)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Monthly NDVI by Habitat Type",
    x = "Month",
    y = "NDVI",
    color = "Habitat (EUNIS2)"
  ) +
  theme_minimal()
ggsave(
  here("output", "figures", "monthly_spectrophenology_Landsat", "EUNIS2_S_NDVI.jpeg"),
  dpi = 300, width = 8, height = 2.5)

ggplot(data_monthly_EUNIS2 %>% 
         dplyr::filter(EUNIS1 == "S" & !is.na(EUNIS2)) %>%
         # Remove those with EUNIS level 2 that does not match current classif
         dplyr::filter(EUNIS2 != "Sa" & EUNIS2 != "Sb"), 
       aes(x = month, y = mean_EVI, color = EUNIS2, group = EUNIS2)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Monthly EVI by Habitat Type",
    x = "Month",
    y = "EVI",
    color = "Habitat (EUNIS2)"
  ) +
  theme_minimal()
ggsave(
  here("output", "figures", "monthly_spectrophenology_Landsat", "EUNIS2_S_EVI.jpeg"),
  dpi = 300, width = 8, height = 2.5)

ggplot(data_monthly_EUNIS2 %>% 
         dplyr::filter(EUNIS1 == "S" & !is.na(EUNIS2)) %>%
         # Remove those with EUNIS level 2 that does not match current classif
         dplyr::filter(EUNIS2 != "Sa" & EUNIS2 != "Sb"), 
       aes(x = month, y = mean_SAVI, color = EUNIS2, group = EUNIS2)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Monthly SAVI by Habitat Type",
    x = "Month",
    y = "SAVI",
    color = "Habitat (EUNIS2)"
  ) +
  theme_minimal()
ggsave(
  here("output", "figures", "monthly_spectrophenology_Landsat", "EUNIS2_S_SAVI.jpeg"),
  dpi = 300, width = 8, height = 2.5)
```

### T

```{r}
ggplot(data_monthly_EUNIS2 %>% 
         dplyr::filter(EUNIS1 == "T" & !is.na(EUNIS2)), 
       aes(x = month, y = mean_NDVI, color = EUNIS2, group = EUNIS2)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Monthly NDVI by Habitat Type",
    x = "Month",
    y = "NDVI",
    color = "Habitat (EUNIS2)"
  ) +
  theme_minimal()
ggsave(
  here("output", "figures", "monthly_spectrophenology_Landsat", "EUNIS2_T_NDVI.jpeg"),
  dpi = 300, width = 6, height = 2.5)

ggplot(data_monthly_EUNIS2 %>% 
         dplyr::filter(EUNIS1 == "T" & !is.na(EUNIS2)), 
       aes(x = month, y = mean_EVI, color = EUNIS2, group = EUNIS2)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Monthly EVI by Habitat Type",
    x = "Month",
    y = "EVI",
    color = "Habitat (EUNIS2)"
  ) +
  theme_minimal()
ggsave(
  here("output", "figures", "monthly_spectrophenology_Landsat", "EUNIS2_T_EVI.jpeg"),
  dpi = 300, width = 6, height = 2.5)

ggplot(data_monthly_EUNIS2 %>% 
         dplyr::filter(EUNIS1 == "T" & !is.na(EUNIS2)), 
       aes(x = month, y = mean_SAVI, color = EUNIS2, group = EUNIS2)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Monthly SAVI by Habitat Type",
    x = "Month",
    y = "SAVI",
    color = "Habitat (EUNIS2)"
  ) +
  theme_minimal()
ggsave(
  here("output", "figures", "monthly_spectrophenology_Landsat", "EUNIS2_T_SAVI.jpeg"),
  dpi = 300, width = 6, height = 2.5)
```

# Calculate other phenological metrics

```{r}
final_RS_data <- final_RS_data %>%
  mutate(
    # with slope method
    # Growing season duration
    NDVI_gsd = NDVI_eos_doy - NDVI_sos_doy,
    EVI_gsd = NDVI_eos_doy - NDVI_sos_doy,
    SAVI_gsd = SAVI_eos_doy - SAVI_sos_doy,
    # Difference in value between pos and sos
    NDVI_diff_pos_sos_value = NDVI_pos_value - NDVI_sos_value,
    EVI_diff_pos_sos_value = EVI_pos_value - EVI_sos_value,
    SAVI_diff_pos_sos_value = SAVI_pos_value - SAVI_sos_value,
    # Difference in value between pos and eos
    NDVI_diff_pos_eos_value = NDVI_pos_value - NDVI_eos_value,
    EVI_diff_pos_eos_value = EVI_pos_value - EVI_eos_value,
    SAVI_diff_pos_eos_value = SAVI_pos_value - SAVI_eos_value,
    # Difference in doy between pos and sos
    NDVI_diff_pos_sos_doy = NDVI_pos_doy - NDVI_sos_doy,
    EVI_diff_pos_sos_doy = EVI_pos_doy - EVI_sos_doy,
    SAVI_diff_pos_sos_doy = SAVI_pos_doy - SAVI_sos_doy,
    # Absolute difference in doy between pos and eos
    NDVI_diff_pos_eos_doy = abs(NDVI_pos_doy - NDVI_eos_doy),
    EVI_diff_pos_eos_doy = abs(EVI_pos_doy - EVI_eos_doy),
    SAVI_diff_pos_eos_doy = abs(SAVI_pos_doy - SAVI_eos_doy),
    # With months method
    # Difference in value between pos and march
    NDVI_diff_pos_march_value = NDVI_pos_value - NDVI_avg_value_03,
    EVI_diff_pos_march_value = EVI_pos_value - EVI_avg_value_03,
    SAVI_diff_pos_march_value = SAVI_pos_value - SAVI_avg_value_03,
    # Difference in value between pos and oct
    NDVI_diff_pos_oct_value = NDVI_pos_value - NDVI_avg_value_10,
    EVI_diff_pos_oct_value = EVI_pos_value - EVI_avg_value_10,
    SAVI_diff_pos_oct_value = SAVI_pos_value - SAVI_avg_value_10,
    # Difference in doy between pos and march
    NDVI_diff_pos_march_doy = NDVI_pos_doy - 75, 
    EVI_diff_pos_march_doy = EVI_pos_doy - 75,
    SAVI_diff_pos_march_doy = SAVI_pos_doy - 75,
    # Difference in doy between pos and oct
    NDVI_diff_pos_oct_doy = abs(NDVI_pos_doy - 285),
    EVI_diff_pos_oct_doy = abs(EVI_pos_doy - 285),
    SAVI_diff_pos_oct_doy = abs(SAVI_pos_doy - 285),
     # With POS-2 and POS+2 method
    # Difference in value between pos and pos_minus60
    NDVI_diff_pos_minus60_value = NDVI_pos_value - NDVI_pos_minus60_value,
    EVI_diff_pos_minus60_value = EVI_pos_value - EVI_pos_minus60_value,
    SAVI_diff_pos_minus60_value = SAVI_pos_value - SAVI_pos_minus60_value,
    # Difference in value between pos and pos_plus60
    NDVI_diff_pos_plus60_value = NDVI_pos_value - NDVI_pos_plus60_value,
    EVI_diff_pos_plus60_value = EVI_pos_value - EVI_pos_plus60_value,
    SAVI_diff_pos_plus60_value = SAVI_pos_value - SAVI_pos_plus60_value
  )
```

## Checks

### Slope method

```{r}
# Growing season duration should be positive
nrow(final_RS_data %>% 
       dplyr::filter(NDVI_gsd <= 0 | EVI_gsd <= 0 | 
                       SAVI_gsd <= 0))
# Difference in value between pos and sos should be positive
nrow(final_RS_data %>%
       dplyr::filter(NDVI_diff_pos_sos_value <= 0))
nrow(final_RS_data %>%
       dplyr::filter(EVI_diff_pos_sos_value <= 0))
nrow(final_RS_data %>%
       dplyr::filter(SAVI_diff_pos_sos_value <= 0))
# Difference in value between pos and eos should be positive
nrow(final_RS_data %>%
       dplyr::filter(NDVI_diff_pos_eos_value <= 0))
nrow(final_RS_data %>%
       dplyr::filter(EVI_diff_pos_eos_value <= 0))
nrow(final_RS_data %>%
       dplyr::filter(SAVI_diff_pos_eos_value <= 0))
# Difference in doy between pos and sos should be positive
nrow(final_RS_data %>%
       dplyr::filter(NDVI_diff_pos_sos_doy <= 0 |
                       EVI_diff_pos_sos_doy <= 0 |
                       SAVI_diff_pos_sos_doy <= 0))
# Difference in doy between eos and pos should be positive
nrow(final_RS_data %>%
       dplyr::filter(NDVI_diff_pos_eos_doy <= 0 |
                       EVI_diff_pos_eos_doy <= 0 |
                       SAVI_diff_pos_eos_doy <= 0))
```

### POS-2 and POS+2 method

```{r}
# Difference in value between pos and pos_minus60 should be positive
nrow(final_RS_data %>%
       dplyr::filter(NDVI_diff_pos_minus60_value <= 0))
nrow(final_RS_data %>%
       dplyr::filter(EVI_diff_pos_minus60_value <= 0))
nrow(final_RS_data %>%
       dplyr::filter(SAVI_diff_pos_minus60_value <= 0))
# Difference in value between pos and pos_plus60 should be positive
nrow(final_RS_data %>%
       dplyr::filter(NDVI_diff_pos_plus60_value <= 0))
nrow(final_RS_data %>%
       dplyr::filter(EVI_diff_pos_plus60_value <= 0))
nrow(final_RS_data %>%
       dplyr::filter(SAVI_diff_pos_plus60_value <= 0))
```

## Further checks (EVI)

### Slope method

```{r}
# Difference in value between pos and sos should be positive
nrow(final_RS_data %>%
       dplyr::filter(EVI_diff_pos_sos_value <= 0))
# Difference in value between pos and eos should be positive
nrow(final_RS_data %>%
       dplyr::filter(EVI_diff_pos_eos_value <= 0))
```

# Add some columns needed

```{r}
final_RS_data <- final_RS_data %>%
  left_join(
    data_RS_Landsat_bands_indices %>%
      distinct(PlotObservationID, year, biogeo, unit)
    )
```

# Add canopy height data

Read the data:

```{r}
data_RS_CH <- read_csv(
  "C:/Data/MOTIVATE/MOTIVATE_RS_data/validation/Canopy_Height_1m/CanopyHeight_1m.csv")
```

Join:

```{r}
final_RS_data <- final_RS_data %>%
  left_join(data_RS_CH %>% 
              dplyr::select(PltObID, canopy_height) %>%
              mutate(PltObID = factor(PltObID)) %>%
              rename(PlotObservationID = PltObID))
```

# Save to clean data

```{r}
write_tsv(final_RS_data,
          here("data", "clean","final_RS_data_bands_Landsat_all_20260309.csv"))
```

# Session info

```{r}
sessionInfo()
```
