---
title: "Script to validate points in ReSurvey database using RS data (S2 + SatEmb + Landsat)"
subtitle: "K-fold cross validation of best models"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

This R script is used to validate the points in the ReSurvey database using either RS indicators from S2 or Landsat (indices + phenology + canopy height), or bands from the Google Satellite Embedding dataset: https://developers.google.com/earth-engine/datasets/catalog/GOOGLE_SATELLITE_EMBEDDING_V1_ANNUAL#description
Here, we perform a K-fold cross validation of the best models. 

# Load libraries

```{r}
library(tidyverse)
library(here)
library(parallel)
library(doParallel)
library(moreparty)
library(caret)
library(randomForest)
library(pROC)
library(rlang)
library(stringr)
library(beepr)
library(permimp)
library(purrr)
library(MLmetrics)
library(patchwork)
```

# Load functions

```{r}
knitr::purl("functions_20260122.Rmd", output = "functions_20260122.R")
source("functions_20260122.R")
```

# Load previously created objects

```{r}
# Load best model Landsat (and prediction) - Accuracy = 0.730 (without k-fold)
model_Landsat <- get(load(here("objects", "RF_1500_20260128",
                               "rf_months_Landsat_refin_20_10_2x_model.Rdata")))
prediction_Landsat <- get(load(here("objects", "RF_1500_20260128", 
                                    "rf_months_Landsat_refin_20_10_2x_prediction.Rdata")))
# Load filtered data Landsat
filtered_data_months_Landsat_with_probs_2x <- get(load(here("objects", "RF_1500_20260128",
                                       "filtered_data_months_Landsat_with_probs_2x.Rdata")))
# Load best model S2 or SatEmb (and prediction) - Accuracy = 0.7413 (without k-fold)
model_SatEmb <- get(load(here("objects", "RF_1500_20260128", 
                              "rf_SatEmb_refin_20_10_2x_model.Rdata")))
prediction_SatEmb <- get (load(here("objects", "RF_1500_20260128",
                                    "rf_SatEmb_refin_20_10_2x_prediction.Rdata")))
# Load filtered data S2 or SatEmb
filtered_data_SatEmb_with_probs_2x <- get(load(here("objects", "RF_1500_20260128",
                                       "filtered_data_SatEmb_with_probs_2x.Rdata")))
```

```{r}
# Load cross-validation objects
cv_SatEmb <- get(load(here("objects", "RF_1500_cv", "cv_SatEmb.Rdata")))
cv_Landsat <- get(load(here("objects", "RF_1500_cv", "cv_Landsat.Rdata")))
```

```{r}
# Load varimp objects
varimp_SatEmb <- get(load(here("objects", "RF_1500_cv", "cv_SatEmb_varimp.Rdata")))
varimp_Landsat <- get(load(here("objects", "RF_1500_cv", "cv_Landsat_varimp.Rdata")))
```


# Read data

```{r}
data_validation_SatEmb <- read_tsv(here(
  "data", "clean","final_RS_data_SatEmb_20251010.csv")) %>%
  rename(EUNISa_1_descr = EUNIS_1_d)
data_validation_Landsat <- read_tsv(here(
  "data", "clean","final_RS_data_bands_Landsat_all_20251021.csv"))
```

No parsing issues!

```{r}
nrow(data_validation_Landsat)
nrow(data_validation_SatEmb)
```

# Filter out manipulated observations

## Read list of PlotObservationIDs to keep

```{r}
PlotObservationID_list <- read_csv(
  here("data", "clean","PlotObservationIDs_filtered_QRST_20260122.csv"))
```

## Filter out data to keep only those PlotObservationIDs

```{r}
data_validation_SatEmb <- data_validation_SatEmb %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
data_validation_Landsat <- data_validation_Landsat %>%
  filter(PlotObservationID %in% PlotObservationID_list$PlotObservationID)
```

```{r}
nrow(data_validation_Landsat)
nrow(data_validation_SatEmb)
```

# Define lists of predictor vars

```{r}
# Excluding EVI because many values outside of range -1, 1
vars_RF_months_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, and SAVI between march and oct
  "NDVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI and SAVI at march, pos and oct
  "NDVI_avg_value_03", "NDVI_pos_value", "NDVI_avg_value_10",
  "SAVI_avg_value_03", "SAVI_pos_value", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "SAVI_diff_pos_march_value",
  "NDVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "SAVI_diff_pos_oct_value",
  "NDVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy",
  # Canopy height
  "canopy_height")
```

```{r}
vars_RF_SatEmb <- colnames(
  data_validation_SatEmb)[startsWith(colnames(data_validation_SatEmb), "A")]
```

# Get filtered data

## Landsat

```{r}
filter_data_Landsat <- function(data, variables) {
  data %>%
    mutate(EUNISa_1 = as.factor(EUNISa_1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    # With Landsat data, many observations removed because of high/low EVI values
    # In Landsat_extraction, values of EVI outside of expected range were set as NA,
    # so better not to use EVI here
    # dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Select only variables needed
    dplyr::select(PlotObservationID, EUNISa_1, all_of(variables), Lctnmth,
           valid_1) %>%
    # Keep only rows with differences > 0 in variables
    filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0)) %>%
    # Select only GPS points
    dplyr::filter(Lctnmth == "Location with GPS" | 
                    Lctnmth == "Location with differential GPS")
}
```

```{r}
# Get filtered data
filtered_data_Landsat <- data_validation_Landsat %>%
  # Rough validation based on NDWI
  rough_validation_Landsat() %>% 
  # Data filtering
  filter_data_Landsat(vars_RF_months_Landsat) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value")

# Detect probability columns
prob_cols_Landsat <- names(
  filtered_data_months_Landsat_with_probs_2x
  )[grepl("^prob_", names(filtered_data_months_Landsat_with_probs_2x))]

# calculate auxiliary metrics
df_aux_Landsat_2x <- calculate_df_aux(filtered_data_months_Landsat_with_probs_2x,
                                  prob_cols_Landsat)

# Apply filtering function
df_final_20_10_months_Landsat_2x <- df_aux_Landsat_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)

# Get data for cross-validation
filtered_data_Landsat_cv <- filtered_data_Landsat %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_months_Landsat_2x$PlotObservationID)
```

## SatEmb

```{r}
filter_data_SatEmb <- function(data, variables) {
  data %>%
    mutate(EUNISa_1 = as.factor(EUNISa_1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Select only variables needed
    dplyr::select(PlotObservationID, EUNISa_1, all_of(variables), 
           Lctnmth, valid_1) %>%
    # Select only GPS points
    dplyr::filter(Lctnmth == "Location with GPS" |
                    Lctnmth == "Location with differential GPS")
}
```

```{r}
data_validation_S2 <- read_tsv(here(
  "data", "clean","final_RS_data_bands_S2_all_20251021.csv"))
```

```{r}
# Get filtered data
filtered_data_SatEmb <- data_validation_SatEmb %>%
  # Rough validation based on CH
  rough_validation_SatEmb() %>% 
  # Join with S2 data to get NDWI_max info
  left_join(
    data_validation_S2 %>% select(PlotObservationID, NDWI_max)
  ) %>%
  # Rough validation based on NDWI (from S2)
  # Remove those with NDWI_max > 0.3, keep if NDWI_max is NA
  filter(NDWI_max <= 0.3 | is.na(NDWI_max)) %>%
  # Remove unneeded var
  select(-NDWI_max) %>%
  # Data filtering
  filter_data_SatEmb(vars_RF_SatEmb) %>% 
  # Join with S2 data to get NDVI_pos_value info
  left_join(
    data_validation_S2 %>% select(PlotObservationID, NDVI_pos_value)
  ) %>%
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value") %>%
  # Remove unneeded var
  select(-NDVI_pos_value) 

# Detect probability columns
prob_cols_SatEmb <- names(
  filtered_data_SatEmb_with_probs_2x
  )[grepl("^prob_", names(filtered_data_SatEmb_with_probs_2x))]

# calculate auxiliary metrics
df_aux_SatEmb_2x <- calculate_df_aux(filtered_data_SatEmb_with_probs_2x,
                                  prob_cols_SatEmb)

# Apply filtering function
df_final_20_10_SatEmb_2x <- df_aux_SatEmb_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)

# Get data for cross-validation
filtered_data_SatEmb_cv <- filtered_data_SatEmb %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_SatEmb_2x$PlotObservationID)
```

# k-fold cross-validation (4 folds)

```{r}
set.seed(123)
ctrl <- trainControl(
  method = "cv",
  number = 4, # Creates stratified folds when y is a factor, preserving class proportions
  allowParallel = FALSE,
  savePredictions = "final",
  classProbs = TRUE,     # set TRUE if you also want probabilities (for ROC)
  summaryFunction = multiClassSummary,  # optional, for multi-class summaries
  verboseIter = TRUE      # <--- this gives progress output
)
```

## SatEmb

```{r eval=FALSE, include=FALSE}
cv_SatEmb <- caret::train(
  x = filtered_data_SatEmb_cv[, vars_RF_SatEmb],
  y = filtered_data_SatEmb_cv $EUNISa_1,
  # Using cforest, not fastcforest
  method = "cforest",
  trControl = ctrl,
  tuneGrid = expand.grid(
    mtry = round(sqrt(length(vars_RF_SatEmb)))
  ),
  controls = party::cforest_control(
    ntree = 1500
  )
)
```

### Save

```{r eval=FALSE, include=FALSE}
save(cv_SatEmb, file = "objects/RF_1500_cv/cv_SatEmb.Rdata")
```

### Results

#### Global Accuracy and Kappa

```{r}
res_SatEmb <- cv_SatEmb$results
res_SatEmb[, c("Accuracy", "Kappa", "AccuracySD", "KappaSD")]
```

#### Balanced accuracy per class (one-vs-all)

```{r}
pred_df_SatEmb <- cv_SatEmb$pred   # todas las predicciones CV
cm_SatEmb <- confusionMatrix(pred_df_SatEmb$pred, pred_df_SatEmb$obs)

by_class_SatEmb <- as.data.frame(cm_SatEmb$byClass)
balanced_acc_SatEmb <- by_class_SatEmb %>%
  mutate(Balanced_Accuracy = (Sensitivity + Specificity)/2) %>%
  select(Sensitivity, Specificity, Balanced_Accuracy)

balanced_acc_SatEmb$Class <- rownames(by_class_SatEmb)
balanced_acc_SatEmb <- balanced_acc_SatEmb %>% select(Class, everything())
balanced_acc_SatEmb
```

#### Balanced accuracy macro (average over classes)

```{r}
balanced_acc_macro_SatEmb <- mean(balanced_acc_SatEmb$Balanced_Accuracy)
balanced_acc_macro_SatEmb
```

#### AUC one-vs-all per class + AUC macro

```{r}
classes_SatEmb <- levels(pred_df_SatEmb$obs)

auc_one_vs_all <- function(df, cls) {
  bin_obs <- as.integer(df$obs == cls)
  prob_cls <- df[[cls]]
  roc_obj <- roc(bin_obs, prob_cls, quiet = TRUE)
  auc(roc_obj)
}

auc_by_class_SatEmb <- sapply(classes_SatEmb, function(cls) auc_one_vs_all(pred_df_SatEmb, cls))
auc_by_class_SatEmb

auc_macro_SatEmb <- mean(auc_by_class_SatEmb)
auc_macro_SatEmb
```

#### Average confusion matrix

```{r}
cm_SatEmb$table
```

#### F1-score per class

```{r}
as.data.frame(cm_SatEmb$byClass)[, c("Precision", "Recall", "F1")]
```

#### Variable importance

```{r eval=FALSE, include=FALSE}
varimp_SatEmb <- varImp(cv_SatEmb)
```

```{r}
print(varimp_SatEmb, top = 64)
```

##### Save

```{r eval=FALSE, include=FALSE}
save(varimp_SatEmb, file = "objects/RF_1500_cv/cv_SatEmb_varimp.Rdata")
```

##### (Appendix 7) Plot

```{r}
vi_SatEmb_caret_df <- as.data.frame(varimp_SatEmb$importance)
vi_SatEmb_caret_df$Variable <- rownames(vi_SatEmb_caret_df)
colnames(vi_SatEmb_caret_df)[1] <- "ScaledImportance"
vi_SatEmb_caret_df <- vi_SatEmb_caret_df %>% arrange(desc(ScaledImportance))
ggplot(vi_SatEmb_caret_df, 
       aes(x = reorder(Variable, ScaledImportance), y = ScaledImportance)) +
  geom_col(fill = "#2c7fb8") +
  coord_flip() +
  labs(x = "Predictor", y = "Importance (scaled)") +
  theme_minimal()
ggsave(here("output", "figures", "manuscript", "appendix7.jpeg"), 
       dpi = 300, width = 6, height = 10)
```

#### ROC one-vs-all for each class (aggregated over folds)

```{r}
# Compute a ROC object for each class (one-vs-all)
roc_list_SatEmb <- lapply(classes_SatEmb, function(cls) {
  # Binary labels: 1 for cls, 0 for others
  bin_obs <- as.integer(pred_df_SatEmb$obs == cls)
  # Probability for the class
  prob_cls <- pred_df_SatEmb[[cls]]
  # Compute ROC; set direction = "<" if needed depending on your scores
  pROC::roc(response = bin_obs, predictor = prob_cls, quiet = TRUE)
})
names(roc_list_SatEmb) <- classes_SatEmb
```

```{r}
# Extract ROC curve data (TPR vs FPR) per class
roc_df_SatEmb <- lapply(names(roc_list_SatEmb), function(cls) {
  coords <- pROC::coords(roc_list_SatEmb[[cls]], ret = c("specificity", "sensitivity"), transpose = FALSE)
  data.frame(Class = cls,
             FPR = 1 - coords$specificity,
             TPR = coords$sensitivity)
}) %>% bind_rows()

roc_plot_SatEmb <- ggplot(roc_df_SatEmb, aes(x = FPR, y = TPR, color = Class)) +
  geom_line(linewidth = 0.75) +
  #scale_color_manual(values = cols) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  labs(# title = "ROC curves (one-vs-all) by class",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 12)
roc_plot_SatEmb
```

## Landsat

```{r eval=FALSE, include=FALSE}
cv_Landsat <- caret::train(
  x = filtered_data_Landsat_cv[, vars_RF_months_Landsat],
  y = filtered_data_Landsat_cv $EUNISa_1,
  # Using cforest, not fastcforest
  method = "cforest",
  trControl = ctrl,
  tuneGrid = expand.grid(
    mtry = round(sqrt(length(vars_RF_months_Landsat)))
  ),
  controls = party::cforest_control(
    ntree = 1500
  )
)
```

### Save

```{r eval=FALSE, include=FALSE}
save(cv_Landsat, file = "objects/RF_1500_cv/cv_Landsat.Rdata")
```

### Results

#### Global Accuracy and Kappa

```{r}
res_Landsat <- cv_Landsat$results
res_Landsat[, c("Accuracy", "Kappa", "AccuracySD", "KappaSD")]
```

#### Balanced accuracy per class (one-vs-all)

```{r}
pred_df_Landsat <- cv_Landsat$pred   # todas las predicciones CV
cm_Landsat <- confusionMatrix(pred_df_Landsat$pred, pred_df_Landsat$obs)

by_class_Landsat <- as.data.frame(cm_Landsat$byClass)
balanced_acc_Landsat <- by_class_Landsat %>%
  mutate(Balanced_Accuracy = (Sensitivity + Specificity)/2) %>%
  select(Sensitivity, Specificity, Balanced_Accuracy)

balanced_acc_Landsat$Class <- rownames(by_class_Landsat)
balanced_acc_Landsat <- balanced_acc_Landsat %>% select(Class, everything())
balanced_acc_Landsat
```

#### Balanced accuracy macro (average over classes)

```{r}
balanced_acc_macro_Landsat <- mean(balanced_acc_Landsat$Balanced_Accuracy)
balanced_acc_macro_Landsat
```

#### AUC one-vs-all per class + AUC macro

```{r}
classes_Landsat <- levels(pred_df_Landsat$obs)

auc_one_vs_all <- function(df, cls) {
  bin_obs <- as.integer(df$obs == cls)
  prob_cls <- df[[cls]]
  roc_obj <- roc(bin_obs, prob_cls, quiet = TRUE)
  auc(roc_obj)
}

auc_by_class_Landsat <- sapply(classes_Landsat, function(cls) auc_one_vs_all(pred_df_Landsat, cls))
auc_by_class_Landsat

auc_macro_Landsat <- mean(auc_by_class_Landsat)
auc_macro_Landsat
```

#### Average confusion matrix

```{r}
cm_Landsat$table
```

#### F1-score per class

```{r}
as.data.frame(cm_Landsat$byClass)[, c("Precision", "Recall", "F1")]
```

#### Variable importance

```{r eval=FALSE, include=FALSE}
varimp_Landsat <- varImp(cv_Landsat)
```

```{r}
print(varimp_Landsat, top = 30)
```

##### Save

```{r eval=FALSE, include=FALSE}
save(varimp_Landsat, file = "objects/RF_1500_cv/cv_Landsat_varimp.Rdata")
```

##### (Appendix 6) Plot

```{r}
vi_Landsat_caret_df <- as.data.frame(varimp_Landsat$importance)
vi_Landsat_caret_df$Variable <- rownames(vi_Landsat_caret_df)
colnames(vi_Landsat_caret_df)[1] <- "ScaledImportance"
vi_Landsat_caret_df <- vi_Landsat_caret_df %>% arrange(desc(ScaledImportance))
ggplot(vi_Landsat_caret_df, 
       aes(x = reorder(Variable, ScaledImportance), y = ScaledImportance)) +
  geom_col(fill = "#2c7fb8") +
  coord_flip() +
  labs(x = "Predictor", y = "Importance (scaled)") +
  theme_minimal()
ggsave(here("output", "figures", "manuscript", "appendix6.jpeg"), 
       dpi = 300, width = 6, height = 4)
```

#### ROC one-vs-all for each class (aggregated over folds)

```{r}
# Compute a ROC object for each class (one-vs-all)
roc_list_Landsat <- lapply(classes_Landsat, function(cls) {
  # Binary labels: 1 for cls, 0 for others
  bin_obs <- as.integer(pred_df_Landsat$obs == cls)
  # Probability for the class
  prob_cls <- pred_df_Landsat[[cls]]
  # Compute ROC; set direction = "<" if needed depending on your scores
  pROC::roc(response = bin_obs, predictor = prob_cls, quiet = TRUE)
})
names(roc_list_Landsat) <- classes_Landsat
```

```{r}
# Extract ROC curve data (TPR vs FPR) per class
roc_df_Landsat <- lapply(names(roc_list_Landsat), function(cls) {
  coords <- pROC::coords(roc_list_Landsat[[cls]], ret = c("specificity", "sensitivity"), transpose = FALSE)
  data.frame(Class = cls,
             FPR = 1 - coords$specificity,
             TPR = coords$sensitivity)
}) %>% bind_rows()

roc_plot_Landsat <- ggplot(roc_df_Landsat, aes(x = FPR, y = TPR, color = Class)) +
  geom_line(linewidth = 0.75) +
  #scale_color_manual(values = cols) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  labs(# title = "ROC curves (one-vs-all) by class",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 12)
roc_plot_Landsat
```

# (Appendix 4) ROC curves

```{r}
roc_plot_Landsat + roc_plot_SatEmb +
  plot_annotation(tag_levels = "A", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom", legend.title = element_blank())
ggsave(here("output", "figures", "manuscript", "appendix4.jpeg"), 
       dpi = 300, width = 9, height = 4)
```

# Session info

```{r}
sessionInfo()
```


