---
title: "Script to validate points in ReSurvey database using RS data (Landsat)"
subtitle: "K-fold cross validation of all models"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

This R script is used to validate the points in the ReSurvey database using RS indicators from Landsat (indices + phenology).
Here, we perform a K-fold cross validation of all the models in 41_validation_models_onlyLandsat.Rmd. 

# Load libraries

```{r}
library(parallel)
library(doParallel)
library(foreach)
library(partykit)
library(caret)
library(here)
library(readr)
library(progressr)
library(party)
library(permimp)
library(tibble)
```

# Load functions

```{r}
knitr::purl("functions_20260225.Rmd", output = "functions_20260225.R")
source("functions_20260225.R")
```

# Load previously created objects

```{r}
lapply(list.files(here("objects", "RF_1500_20260318"), full.names = TRUE), load,
       envir = .GlobalEnv)
```

```{r}
# Load cross-validation objects
lapply(list.files(here("objects", "RF_1500_cv"), full.names = TRUE), load,
       envir = .GlobalEnv)
```

# Read data

```{r}
data_validation_Landsat <- read_tsv(here(
  "data", "clean","final_RS_data_bands_Landsat_all_20260309.csv"))
```

No parsing issues!

```{r}
nrow(data_validation_Landsat)
```

# Load Resurvey db

```{r}
ReSurvey_data <- read_csv(
  here("data", "clean", "ReSurvey_QRST_data_V2025.2_20260304.csv")) %>%
  select(PlotObservationID, EUNIS1, EUNIS2, `Location method`) %>%
  rename(Lctnmth = `Location method`)
```

# Define lists of predictor vars

## Moments with slope method

```{r}
# Excluding EVI because many values outside of range -1, 1
# Not using canopy height!
vars_RF_slope_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI and SAVI
  "NDVI_auc", "SAVI_auc",
  # Values of NDVI and SAVI at sos, pos and eos
  "NDVI_sos_value", "NDVI_pos_value", "NDVI_eos_value",
  "SAVI_sos_value", "SAVI_pos_value", "SAVI_eos_value",
  # Differences pos-sos in value and doy
  "NDVI_diff_pos_sos_value","SAVI_diff_pos_sos_value",
  "NDVI_diff_pos_sos_doy", "SAVI_diff_pos_sos_doy",
  # Differences pos-eos in value and doy
  "NDVI_diff_pos_eos_value", "SAVI_diff_pos_eos_value",
  "NDVI_diff_pos_eos_doy", "SAVI_diff_pos_eos_doy",
  # Growing season duration
  "NDVI_gsd", "SAVI_gsd")
```

## Months method

```{r}
# Excluding EVI because many values outside of range -1, 1
# Not using canopy height!
vars_RF_months_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, and SAVI between march and oct
  "NDVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI and SAVI at march, pos and oct
  "NDVI_avg_value_03", "NDVI_pos_value", "NDVI_avg_value_10",
  "SAVI_avg_value_03", "SAVI_pos_value", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "SAVI_diff_pos_march_value",
  "NDVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "SAVI_diff_pos_oct_value",
  "NDVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy")
```

# Get filtered data

## Function

```{r}
filter_data_Landsat <- function(data, variables) {
  data %>%
    mutate(EUNIS1 = as.factor(EUNIS1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    # With Landsat data, many observations removed because of high/low EVI values
    # In Landsat_extraction, values of EVI outside of expected range were set as NA,
    # so better not to use EVI here
    # dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Select only variables needed
    dplyr::select(PlotObservationID, EUNIS1, all_of(variables), Lctnmth,
           valid_1) %>%
    # Keep only rows with differences > 0 in variables
    filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0)) %>%
    # Select only GPS points
    dplyr::filter(Lctnmth == "Location with GPS" | 
                    Lctnmth == "Location with differential GPS")
}
```

## Filtered data

Select only GPS points that have passed the rough validation in each case.

```{r}
filtered_data_slope_Landsat <- data_validation_Landsat %>%
  # Rough validation based on NDWI
  rough_validation_Landsat() %>% 
  # Join ReSurvey data to get Location method
  left_join(ReSurvey_data %>% select(PlotObservationID, Lctnmth)) %>%
  # Data filtering
  filter_data_Landsat(vars_RF_slope_Landsat) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value")

filtered_data_months_Landsat <- data_validation_Landsat %>%
  # Rough validation based on NDWI
  rough_validation_Landsat() %>% 
  # Join ReSurvey data to get Location method
  left_join(ReSurvey_data %>% select(PlotObservationID, Lctnmth)) %>%
  # Data filtering
  filter_data_Landsat(vars_RF_months_Landsat) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value")
```

# k-fold cross-validation 

Using 4 folds.

## Functions

### Function to run CV

```{r}
run_fastcforest_cv <- function(
    data, vars_RF, response_var,
    ntree = 500, k = 4
) {
  library(party)
  library(moreparty)
  library(doParallel)
  library(foreach)
  library(parallel)
  library(caret)

  # Create folds
  set.seed(123)
  folds <- createFolds(data[[response_var]], k = k)

  # Parallel cluster for CV
  ncores <- max(1, detectCores() - 1)
  cl <- makeCluster(ncores)
  registerDoParallel(cl)

  # Export needed objects
  clusterExport(cl, c("folds", "data", "vars_RF", "response_var"), envir = environment())

  # Load required packages inside workers
  clusterEvalQ(cl, {
    library(party)
    library(moreparty)
  })

  results <- list()

  for (i in seq_along(folds)) {

    cat("\n========== Starting fold", i, "of", k, "==========\n")
    t_start <- Sys.time()

    fold_result <- foreach(j = i, .combine = rbind,
                           .packages = c("party", "moreparty")) %dopar% {

      test_idx  <- folds[[j]]
      train_idx <- setdiff(seq_len(nrow(data)), test_idx)

      train_data <- data[train_idx, ]
      test_data  <- data[test_idx, ]

      # Build formula
      fmla <- reformulate(vars_RF, response = response_var)

      # ✅ Fast parallel conditional forest from moreparty
      model <- moreparty::fastcforest(
        formula  = fmla,
        data     = train_data,
        controls = party::cforest_unbiased(
          mtry  = round(sqrt(length(vars_RF))),
          ntree = ntree
        ),
        parallel = TRUE
      )

      pred <- predict(model, newdata = test_data, type = "response")

      data.frame(
        obs  = test_data[[response_var]],
        pred = pred,
        fold = j
      )
    }

    t_end <- Sys.time()
    cat(
      "✅ Finished fold", i, "in",
      round(difftime(t_end, t_start, units = "mins"), 2),
      "minutes\n"
    )

    results[[i]] <- fold_result
  }

  stopCluster(cl)

  do.call(rbind, results)
}
```

### Function to compute results from CV

```{r}
compute_rf_cv_metrics <- function(cv_results) {
  library(caret)
  library(pROC)
  library(dplyr)
  
  # Ensure pred is factor with same levels as obs
  cv_results$obs  <- factor(cv_results$obs)
  cv_results$pred <- factor(cv_results$pred, levels = levels(cv_results$obs))
  
  classes <- levels(cv_results$obs)
  folds   <- sort(unique(cv_results$fold))
  
  ## ---------------------------
  ## 1. PER-FOLD METRICS
  ## ---------------------------
  fold_metrics <- lapply(folds, function(f) {
    df <- cv_results %>% filter(fold == f)
    
    cm <- confusionMatrix(df$pred, df$obs)
    
    list(
      fold = f,
      accuracy = cm$overall["Accuracy"],
      kappa    = cm$overall["Kappa"]
    )
  })
  
  fold_metrics_df <- bind_rows(lapply(fold_metrics, as.data.frame))
  
  ## ---------------------------
  ## 2. OVERALL CONFUSION MATRIX
  ## ---------------------------
  overall_cm <- confusionMatrix(cv_results$pred, cv_results$obs)
  
  ## ---------------------------
  ## 3. CLASS-WISE METRICS
  ## ---------------------------
  class_metrics <- data.frame(
    class = classes,
    users_accuracy     = overall_cm$byClass[, "Pos Pred Value"],
    producers_accuracy = overall_cm$byClass[, "Sensitivity"],
    balanced_accuracy  = overall_cm$byClass[, "Balanced Accuracy"]
  )
  
  ## ---------------------------
  ## 4. FINAL OUTPUT STRUCTURE
  ## ---------------------------
  list(
    overall_confusion_matrix = overall_cm,
    
    accuracy_mean_sd = c(
      mean = mean(fold_metrics_df$accuracy),
      sd   = sd(fold_metrics_df$accuracy)
    ),
    
    kappa_mean_sd = c(
      mean = mean(fold_metrics_df$kappa),
      sd   = sd(fold_metrics_df$kappa)
    ),
    
    per_fold_metrics = fold_metrics_df,
    
    per_class_metrics = class_metrics
  )
}
```

## Without refinement

### Landsat, slope

```{r eval=FALSE, include=FALSE}
cv_slope_Landsat <- run_fastcforest_cv(
  data = filtered_data_slope_Landsat,
  vars_RF = vars_RF_slope_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_slope_Landsat,
     file = "objects/RF_1500_cv/cv_slope_Landsat.Rdata") 
```

#### Results

```{r}
cv_slope_Landsat_results <-
  compute_rf_cv_metrics(cv_slope_Landsat)
cv_slope_Landsat_results
```

### Landsat, months

```{r eval=FALSE, include=FALSE}
cv_months_Landsat <- run_fastcforest_cv(
  data = filtered_data_months_Landsat,
  vars_RF = vars_RF_months_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_months_Landsat,
     file = "objects/RF_1500_cv/cv_months_Landsat.Rdata") 
```

#### Results

```{r}
cv_months_Landsat_results <-
  compute_rf_cv_metrics(cv_months_Landsat)
cv_months_Landsat_results
```

## Refinement step 1

### Landsat, slope

```{r eval=FALSE, include=FALSE}
cv_slope_Landsat_step1 <- run_fastcforest_cv(
  data = filtered_data_20_10_slope_Landsat_probs,
  vars_RF = vars_RF_slope_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_slope_Landsat_step1, 
     file = "objects/RF_1500_cv/cv_slope_Landsat_step1.Rdata") 
```

#### Results

```{r}
cv_slope_Landsat_step1_results <-
  compute_rf_cv_metrics(cv_slope_Landsat_step1)
cv_slope_Landsat_step1_results
```

### Landsat, months

```{r eval=FALSE, include=FALSE}
cv_months_Landsat_step1 <- run_fastcforest_cv(
  data = filtered_data_20_10_months_Landsat_probs,
  vars_RF = vars_RF_months_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_months_Landsat_step1, 
     file = "objects/RF_1500_cv/cv_months_Landsat_step1.Rdata") 
```

#### Results

```{r}
cv_months_Landsat_step1_results <-
  compute_rf_cv_metrics(cv_months_Landsat_step1)
cv_months_Landsat_step1_results
```

## Refinement step 2

### Landsat, slope

```{r eval=FALSE, include=FALSE}
cv_slope_Landsat_step2 <- run_fastcforest_cv(
  data = filtered_data_20_10_slope_Landsat_probs_2x,
  vars_RF = vars_RF_slope_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_slope_Landsat_step2, 
     file = "objects/RF_1500_cv/cv_slope_Landsat_step2.Rdata") 
```

#### Results

```{r}
cv_slope_Landsat_step2_results <-
  compute_rf_cv_metrics(cv_slope_Landsat_step2)
cv_slope_Landsat_step2_results
```

### Landsat, months

```{r eval=FALSE, include=FALSE}
cv_months_Landsat_step2 <- run_fastcforest_cv(
  data = filtered_data_20_10_months_Landsat_probs_2x,
  vars_RF = vars_RF_months_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_months_Landsat_step2, file = "objects/RF_1500_cv/cv_months_Landsat_step2.Rdata") 
```

#### Results

```{r}
cv_months_Landsat_step2_results <-
  compute_rf_cv_metrics(cv_months_Landsat_step2)
cv_months_Landsat_step2_results
```

# Session info

```{r}
sessionInfo()
```


