---
title: "Script to validate points in ReSurvey database using RS data (S2 + SatEmb + Landsat)"
subtitle: "K-fold cross validation of all models"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

This R script is used to validate the points in the ReSurvey database using either RS indicators from S2 or Landsat (indices + phenology + canopy height), or bands from the Google Satellite Embedding dataset: https://developers.google.com/earth-engine/datasets/catalog/GOOGLE_SATELLITE_EMBEDDING_V1_ANNUAL#description
Here, we perform a K-fold cross validation of all the models in 39_validation_models.Rmd. 

# Load libraries

```{r}
library(tidyverse)
library(here)
library(parallel)
library(doParallel)
library(moreparty)
library(caret)
library(randomForest)
library(pROC)
library(rlang)
library(stringr)
library(beepr)
library(permimp)
library(purrr)
library(MLmetrics)
library(patchwork)
```

# Load functions

```{r}
knitr::purl("functions_20260225.Rmd", output = "functions_20260225.R")
source("functions_20260225.R")
```

# Load previously created objects

```{r eval=FALSE, include=FALSE}
lapply(list.files(here("objects", "RF_1500_20260309"), full.names = TRUE), load,
       envir = .GlobalEnv)
```

```{r eval=FALSE, include=FALSE}
# Load cross-validation objects
lapply(list.files(here("objects", "RF_1500_cv"), full.names = TRUE), load,
       envir = .GlobalEnv)
```

```{r eval=FALSE, include=FALSE}
# Load varimp objects
lapply(list.files(here("objects", "RF_1500_varmip"), full.names = TRUE), load,
       envir = .GlobalEnv)
```

# Read data

```{r}
data_validation_S2 <- read_tsv(here(
  "data", "clean","final_RS_data_bands_S2_all_20260310.csv"))
data_validation_SatEmb <- read_tsv(here(
  "data", "clean","final_RS_data_SatEmb_20260309.csv"))
data_validation_Landsat <- read_tsv(here(
  "data", "clean","final_RS_data_bands_Landsat_all_20260309.csv"))
```

No parsing issues!

```{r}
nrow(data_validation_S2)
nrow(data_validation_SatEmb)
nrow(data_validation_Landsat)
```

# Load Resurvey db

```{r}
ReSurvey_data <- read_csv(
  here("data", "clean", "ReSurvey_QRST_data_V2025.2_20260304.csv")) %>%
  select(PlotObservationID, EUNIS1, EUNIS2, `Location method`) %>%
  rename(Lctnmth = `Location method`)
```

# Define lists of predictor vars

## S2 and Landsat

### Moments with slope method

Not using threshold method cause it gave worse results earlier.

```{r}
vars_RF_slope <- c(
  # Min values of all indices
  "NDVI_min", "EVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, EVI and SAVI
  "NDVI_auc", "EVI_auc", "SAVI_auc",
  # Values of NDVI, EVI and SAVI at sos, pos and eos
  "NDVI_sos_value", "NDVI_pos_value", "NDVI_eos_value",
  "EVI_sos_value", "EVI_pos_value", "EVI_eos_value", 
  "SAVI_sos_value", "SAVI_pos_value", "SAVI_eos_value",
  # Differences pos-sos in value and doy
  "NDVI_diff_pos_sos_value", "EVI_diff_pos_sos_value",
  "SAVI_diff_pos_sos_value","NDVI_diff_pos_sos_doy", 
  "EVI_diff_pos_sos_doy", "SAVI_diff_pos_sos_doy",
  # Differences pos-eos in value and doy
  "NDVI_diff_pos_eos_value", "EVI_diff_pos_eos_value",
  "SAVI_diff_pos_eos_value", "NDVI_diff_pos_eos_doy", 
  "EVI_diff_pos_eos_doy", "SAVI_diff_pos_eos_doy",
  # Growing season duration
  "NDVI_gsd", "EVI_gsd", "SAVI_gsd",
  # Canopy height
  "canopy_height")
```

```{r}
# Excluding EVI because many values outside of range -1, 1
vars_RF_slope_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI and SAVI
  "NDVI_auc", "SAVI_auc",
  # Values of NDVI and SAVI at sos, pos and eos
  "NDVI_sos_value", "NDVI_pos_value", "NDVI_eos_value",
  "SAVI_sos_value", "SAVI_pos_value", "SAVI_eos_value",
  # Differences pos-sos in value and doy
  "NDVI_diff_pos_sos_value","SAVI_diff_pos_sos_value",
  "NDVI_diff_pos_sos_doy", "SAVI_diff_pos_sos_doy",
  # Differences pos-eos in value and doy
  "NDVI_diff_pos_eos_value", "SAVI_diff_pos_eos_value",
  "NDVI_diff_pos_eos_doy", "SAVI_diff_pos_eos_doy",
  # Growing season duration
  "NDVI_gsd", "SAVI_gsd",
  # Canopy height
  "canopy_height")
```

### Months method

```{r}
vars_RF_months <- c(
  # Min values of all indices
  "NDVI_min", "EVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, EVI and SAVI between march and oct
  "NDVI_auc_mar_oct", "EVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI, EVI and SAVI at march, pos and oct
  "NDVI_avg_value_03", "NDVI_pos_value", "NDVI_avg_value_10",
  "EVI_avg_value_03", "EVI_pos_value", "EVI_avg_value_10", 
  "SAVI_avg_value_03", "SAVI_pos_value", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "EVI_diff_pos_march_value",
  "SAVI_diff_pos_march_value","NDVI_diff_pos_march_doy", 
  "EVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "EVI_diff_pos_oct_value",
  "SAVI_diff_pos_oct_value", "NDVI_diff_pos_oct_doy", 
  "EVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy",
  # Canopy height
  "canopy_height")
```

```{r}
# Excluding EVI because many values outside of range -1, 1
vars_RF_months_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, and SAVI between march and oct
  "NDVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI and SAVI at march, pos and oct
  "NDVI_avg_value_03", "NDVI_pos_value", "NDVI_avg_value_10",
  "SAVI_avg_value_03", "SAVI_pos_value", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "SAVI_diff_pos_march_value",
  "NDVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "SAVI_diff_pos_oct_value",
  "NDVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy",
  # Canopy height
  "canopy_height")
```

### POS-2 and POS+2 method

```{r}
vars_RF_2 <- c(
  # Min values of all indices
  "NDVI_min", "EVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, EVI and SAVI between pos_minus60 and pos_plus_60
  "NDVI_auc_pos2", "EVI_auc_pos2", "SAVI_auc_pos2",
  # Values of NDVI, EVI and SAVI at pos_minus60, pos and pos_plus_60
  "NDVI_pos_minus60_value", "NDVI_pos_value", "NDVI_pos_plus60_value",
  "EVI_pos_minus60_value", "EVI_pos_value", "EVI_pos_plus60_value", 
  "SAVI_pos_minus60_value", "SAVI_pos_value", "SAVI_pos_plus60_value",
  # Differences pos-pos_minus60 in value
  "NDVI_diff_pos_minus60_value", "EVI_diff_pos_minus60_value",
  "SAVI_diff_pos_minus60_value",
  # Differences pos-pos_plus_60 in value
  "NDVI_diff_pos_plus60_value", "EVI_diff_pos_plus60_value",
  "SAVI_diff_pos_plus60_value",
  # Canopy height
  "canopy_height")
```

```{r}
vars_RF_2_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI and SAVI between pos_minus60 and pos_plus_60
  "NDVI_auc_pos2", "SAVI_auc_pos2",
  # Values of NDVI and SAVI at pos_minus60, pos and pos_plus_60
  "NDVI_pos_minus60_value", "NDVI_pos_value", "NDVI_pos_plus60_value",
  "SAVI_pos_minus60_value", "SAVI_pos_value", "SAVI_pos_plus60_value",
  # Differences pos-pos_minus60 in value
  "NDVI_diff_pos_minus60_value", "SAVI_diff_pos_minus60_value",
  # Differences pos-pos_plus_60 in value
  "NDVI_diff_pos_plus60_value", "SAVI_diff_pos_plus60_value",
  # Canopy height
  "canopy_height")
```

## SatEmb

```{r}
vars_RF_SatEmb <- colnames(
  data_validation_SatEmb)[startsWith(colnames(data_validation_SatEmb), "A")]
vars_RF_SatEmb_CH <- c(vars_RF_SatEmb, "canopy_height")
```

# Get filtered data

## Functions

```{r}
filter_data_S2 <- function(data, variables) {
  data %>%
    mutate(EUNIS1 = as.factor(EUNIS1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Select only variables needed
    dplyr::select(PlotObservationID, EUNIS1, all_of(variables), Lctnmth,
           valid_1) %>%
    # Keep only rows with differences > 0 in variables
    filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0)) %>%
    # Select only GPS points
    dplyr::filter(Lctnmth == "Location with GPS" | 
                    Lctnmth == "Location with differential GPS")
}
```

```{r}
filter_data_Landsat <- function(data, variables) {
  data %>%
    mutate(EUNIS1 = as.factor(EUNIS1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    # With Landsat data, many observations removed because of high/low EVI values
    # In Landsat_extraction, values of EVI outside of expected range were set as NA,
    # so better not to use EVI here
    # dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Select only variables needed
    dplyr::select(PlotObservationID, EUNIS1, all_of(variables), Lctnmth,
           valid_1) %>%
    # Keep only rows with differences > 0 in variables
    filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0)) %>%
    # Select only GPS points
    dplyr::filter(Lctnmth == "Location with GPS" | 
                    Lctnmth == "Location with differential GPS")
}
```

```{r}
filter_data_SatEmb <- function(data, variables) {
  data %>%
    mutate(EUNIS1 = as.factor(EUNIS1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Select only variables needed
    dplyr::select(PlotObservationID, EUNIS1, all_of(variables), 
           Lctnmth, valid_1) %>%
    # Select only GPS points
    dplyr::filter(Lctnmth == "Location with GPS" |
                    Lctnmth == "Location with differential GPS")
  }
```

## Filtered data

```{r}
# S2
filtered_data_slope_S2 <- data_validation_S2 %>%
  # Rough validation based on CH and NDWI
  rough_validation_S2()  %>%
  # Join ReSurvey data to get Location method
  left_join(ReSurvey_data %>% select(PlotObservationID, Lctnmth)) %>%
  # Data filtering
  filter_data_S2(vars_RF_slope) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value")

filtered_data_months_S2 <- data_validation_S2 %>%
  # Rough validation based on CH and NDWI
  rough_validation_S2() %>% 
  # Join ReSurvey data to get Location method
  left_join(ReSurvey_data %>% select(PlotObservationID, Lctnmth)) %>%
  # Data filtering
  filter_data_S2(vars_RF_months) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value")

filtered_data_2_S2 <- data_validation_S2 %>%
  # Rough validation based on CH and NDWI
  rough_validation_S2() %>% 
  # Join ReSurvey data to get Location method
  left_join(ReSurvey_data %>% select(PlotObservationID, Lctnmth)) %>%
  # Data filtering
  filter_data_S2(vars_RF_2) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value")

# Landsat
filtered_data_slope_Landsat <- data_validation_Landsat %>%
  # Rough validation based on NDWI
  rough_validation_Landsat() %>% 
  # Join ReSurvey data to get Location method
  left_join(ReSurvey_data %>% select(PlotObservationID, Lctnmth)) %>%
  # Data filtering
  filter_data_Landsat(vars_RF_slope_Landsat) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value")

filtered_data_months_Landsat <- data_validation_Landsat %>%
  # Rough validation based on NDWI
  rough_validation_Landsat() %>% 
  # Join ReSurvey data to get Location method
  left_join(ReSurvey_data %>% select(PlotObservationID, Lctnmth)) %>%
  # Data filtering
  filter_data_Landsat(vars_RF_months_Landsat) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value")

filtered_data_2_Landsat <- data_validation_Landsat %>%
  # Rough validation based on NDWI
  rough_validation_Landsat() %>% 
  # Join ReSurvey data to get Location method
  left_join(ReSurvey_data %>% select(PlotObservationID, Lctnmth)) %>%
  # Data filtering
  filter_data_Landsat(vars_RF_2_Landsat) %>% 
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value")

# SatEmb
filtered_data_SatEmb <- data_validation_SatEmb %>%
  # Join ReSurvey data to get EUNIS1, EUNIS2 and Location method
  left_join(ReSurvey_data) %>%
  # Rough validation based on CH
  rough_validation_SatEmb() %>% 
  # Join with S2 data to get NDWI_max info
  left_join(
    data_validation_S2 %>% select(PlotObservationID, NDWI_max)
  ) %>%
  # Rough validation based on NDWI (from S2)
  # Remove those with NDWI_max > 0.3, keep if NDWI_max is NA
  filter(NDWI_max <= 0.3 | is.na(NDWI_max)) %>%
  # Remove unneeded var
  select(-NDWI_max) %>%
  # Data filtering
  filter_data_SatEmb(vars_RF_SatEmb) %>% 
  # Join with S2 data to get NDVI_pos_value info
  left_join(
    data_validation_S2 %>% select(PlotObservationID, NDVI_pos_value)
  ) %>%
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value") %>%
  # Remove unneeded var
  select(-NDVI_pos_value) 

filtered_data_SatEmb_CH <- data_validation_SatEmb %>%
  # Join ReSurvey data to get EUNIS1, EUNIS2 and Location method
  left_join(ReSurvey_data) %>%
  # Rough validation based on CH
  rough_validation_SatEmb() %>% 
  # Join with S2 data to get NDWI_max info
  left_join(
    data_validation_S2 %>% select(PlotObservationID, NDWI_max)
  ) %>%
  # Rough validation based on NDWI (from S2)
  # Remove those with NDWI_max > 0.3, keep if NDWI_max is NA
  filter(NDWI_max <= 0.3 | is.na(NDWI_max)) %>%
  # Remove unneeded var
  select(-NDWI_max) %>%
  # Data filtering
  filter_data_SatEmb(vars_RF_SatEmb_CH) %>% 
  # Join with S2 data to get NDVI_pos_value info
  left_join(
    data_validation_S2 %>% select(PlotObservationID, NDVI_pos_value)
  ) %>%
  # From GPS points, filter based on NDVI percentiles
  filter_by_group_percentiles("NDVI_pos_value") %>%
  # Remove unneeded vars
  select(-NDVI_pos_value) 
```

# k-fold cross-validation 

Using 4 folds.

```{r}
set.seed(123)
ctrl <- trainControl(
  method = "cv",
  number = 4, # Creates stratified folds when y is a factor, preserving class proportions
  allowParallel = FALSE,
  savePredictions = "final",
  classProbs = TRUE,     # set TRUE if you also want probabilities (for ROC)
  summaryFunction = multiClassSummary,  # optional, for multi-class summaries
  verboseIter = TRUE      # <--- this gives progress output
)
```

## Without refinement

### Run CV

#### S2

```{r eval=FALSE, include=FALSE}
set.seed(123)
cv_slope_S2 <- caret::train(
  x = filtered_data_slope_S2[, vars_RF_slope],
  y = filtered_data_slope_S2 $EUNIS1,
  # Using cforest, not fastcforest
  method = "cforest",
  trControl = ctrl,
  tuneGrid = expand.grid(
    mtry = round(sqrt(length(vars_RF_slope)))
  ),
  controls = party::cforest_control(
    ntree = 1500,
    teststat = "quad" # Faster (not used here when run!)
  )
)
```

```{r eval=FALSE, include=FALSE}
save(cv_slope_S2, file = "objects/RF_1500_cv/cv_slope_S2.Rdata") # DONE
```

```{r eval=FALSE, include=FALSE}
set.seed(123)
cv_months_S2 <- caret::train(
  x = filtered_data_months_S2[, vars_RF_months],
  y = filtered_data_months_S2 $EUNIS1,
  # Using cforest, not fastcforest
  method = "cforest",
  trControl = ctrl,
  tuneGrid = expand.grid(
    mtry = round(sqrt(length(vars_RF_months)))
  ),
  controls = party::cforest_control(
    ntree = 1500,
    teststat = "quad" # Faster
  )
)
```

```{r eval=FALSE, include=FALSE}
save(cv_months_S2, file = "objects/RF_1500_cv/cv_months_S2.Rdata") # DONE
```

```{r eval=FALSE, include=FALSE}
# set.seed(123)
# cv_2_S2 <- caret::train(
#   x = filtered_data_2_S2[, vars_RF_2],
#   y = filtered_data_2_S2 $EUNIS1,
#   # Using cforest, not fastcforest
#   method = "cforest",
#   trControl = ctrl,
#   tuneGrid = expand.grid(
#     mtry = round(sqrt(length(vars_RF_2)))
#   ),
#   controls = party::cforest_control(
#     ntree = 1500,
#     teststat = "quad" # Faster
#   )
# )
```

```{r eval=FALSE, include=FALSE}
# save(cv_2_S2, file = "objects/RF_1500_cv/cv_2_S2.Rdata")
```

#### Landsat

```{r eval=FALSE, include=FALSE}
set.seed(123)
cv_slope_Landsat <- caret::train(
  x = filtered_data_slope_Landsat[, vars_RF_slope_Landsat],
  y = filtered_data_slope_Landsat $EUNIS1,
  # Using cforest, not fastcforest
  method = "cforest",
  trControl = ctrl,
  tuneGrid = expand.grid(
    mtry = round(sqrt(length(vars_RF_slope_Landsat)))
  ),
  controls = party::cforest_control(
    ntree = 1500,
    teststat = "quad" # Faster
  )
)
```

```{r eval=FALSE, include=FALSE}
save(cv_slope_Landsat, file = "objects/RF_1500_cv/cv_slope_Landsat.Rdata") # RUNNING
```

```{r eval=FALSE, include=FALSE}
set.seed(123)
cv_months_Landsat <- caret::train(
  x = filtered_data_months_Landsat[, vars_RF_months_Landsat],
  y = filtered_data_months_Landsat $EUNIS1,
  # Using cforest, not fastcforest
  method = "cforest",
  trControl = ctrl,
  tuneGrid = expand.grid(
    mtry = round(sqrt(length(vars_RF_months_Landsat)))
  ),
  controls = party::cforest_control(
    ntree = 1500,
    teststat = "quad" # Faster
  )
)
```

```{r eval=FALSE, include=FALSE}
save(cv_months_Landsat, file = "objects/RF_1500_cv/cv_months_Landsat.Rdata") # RUNNING
```

```{r eval=FALSE, include=FALSE}
# set.seed(123)
# cv_2_Landsat <- caret::train(
#   x = filtered_data_2_Landsat[, vars_RF_2_Landsat],
#   y = filtered_data_2_Landsat $EUNIS1,
#   # Using cforest, not fastcforest
#   method = "cforest",
#   trControl = ctrl,
#   tuneGrid = expand.grid(
#     mtry = round(sqrt(length(vars_RF_2_Landsat)))
#   ),
#   controls = party::cforest_control(
#     ntree = 1500,
#     teststat = "quad" # Faster
#   )
# )
```

```{r eval=FALSE, include=FALSE}
# save(cv_2_Landsat, file = "objects/RF_1500_cv/cv_2_Landsat.Rdata")
```

#### SatEmb

```{r eval=FALSE, include=FALSE}
set.seed(123)
cv_SatEmb <- caret::train(
  x = filtered_data_SatEmb[, vars_RF_SatEmb],
  y = filtered_data_SatEmb $EUNIS1,
  # Using cforest, not fastcforest
  method = "cforest",
  trControl = ctrl,
  tuneGrid = expand.grid(
    mtry = round(sqrt(length(vars_RF_SatEmb)))
  ),
  controls = party::cforest_control(
    ntree = 1500,
    teststat = "quad" # Faster
  )
)
```

```{r eval=FALSE, include=FALSE}
save(cv_SatEmb, file = "objects/RF_1500_cv/cv_SatEmb.Rdata")
```

### Results

#### Global Accuracy and Kappa

```{r}
cv_slope_S2$results[, c("Accuracy", "Kappa", "AccuracySD", "KappaSD")]
cv_months_S2$results[, c("Accuracy", "Kappa", "AccuracySD", "KappaSD")]
cv_2_S2$results[, c("Accuracy", "Kappa", "AccuracySD", "KappaSD")]
cv_slope_Landsat$results[, c("Accuracy", "Kappa", "AccuracySD", "KappaSD")]
cv_months_Landsat$results[, c("Accuracy", "Kappa", "AccuracySD", "KappaSD")]
cv_2_Landsat$results[, c("Accuracy", "Kappa", "AccuracySD", "KappaSD")]
cv_SatEmb$results[, c("Accuracy", "Kappa", "AccuracySD", "KappaSD")]
```

#### Balanced accuracy per class (one-vs-all)

```{r}
pred_df_slope_S2 <- cv_slope_S2$pred
pred_df_months_S2 <- cv_months_S2$pred
pred_df_2_S2 <- cv_2_S2$pred
pred_df_slope_Landsat <- cv_slope_Landsat$pred
pred_df_months_Landsat <- cv_months_Landsat$pred
pred_df_2_Landsat <- cv_2_Landsat$pred
pred_df_SatEmb <- cv_SatEmb$pred

cm_slope_S2 <- confusionMatrix(pred_df_slope_S2$pred, 
                               pred_df_slope_S2$obs)
cm_months_S2 <- confusionMatrix(pred_df_months_S2$pred,
                                pred_df_months_S2$obs)
cm_2_S2 <- confusionMatrix(pred_df_2_S2$pred, 
                           pred_df_2_S2$obs)
cm_slope_Landsat <- confusionMatrix(pred_df_slope_Landsat$pred, 
                                    pred_df_slope_Landsat$obs)
cm_months_Landsat <- confusionMatrix(pred_df_months_Landsat$pred, 
                                     pred_df_months_Landsat$obs)
cm_2_Landsat <- confusionMatrix(pred_df_2_Landsat$pred, 
                                pred_df_2_Landsat$obs)
cm_SatEmb <- confusionMatrix(pred_df_SatEmb$pred, 
                             pred_df_SatEmb$obs)

by_class_slope_S2 <- as.data.frame(cm_slope_S2$byClass)
by_class_months_s2 <- as.data.frame(cm_months_S2$byClass)
by_class_2_S2 <- as.data.frame(cm_2_S2$byClass)
by_class_slope_Landsat <- as.data.frame(cm_slope_Landsat$byClass)
by_class_months_Landsat <- as.data.frame(cm_months_Landsat$byClass)
by_class_2_Landsat <- as.data.frame(cm_2_Landsat$byClass)
by_class_SatEmb <- as.data.frame(cm_SatEmb$byClass)

balanced_acc_slope_S2 <- by_class_slope_S2 %>%
  mutate(Balanced_Accuracy = (Sensitivity + Specificity)/2) %>%
  select(Sensitivity, Specificity, Balanced_Accuracy)
balanced_acc_months_S2 <- by_class_months_S2 %>%
  mutate(Balanced_Accuracy = (Sensitivity + Specificity)/2) %>%
  select(Sensitivity, Specificity, Balanced_Accuracy)
balanced_acc_2_S2 <- by_class_2_S2 %>%
  mutate(Balanced_Accuracy = (Sensitivity + Specificity)/2) %>%
  select(Sensitivity, Specificity, Balanced_Accuracy)
balanced_acc_slope_Landsat <- by_class_slope_Landsat %>%
  mutate(Balanced_Accuracy = (Sensitivity + Specificity)/2) %>%
  select(Sensitivity, Specificity, Balanced_Accuracy)
balanced_acc_months_Landsat <- by_class_months_Landsat %>%
  mutate(Balanced_Accuracy = (Sensitivity + Specificity)/2) %>%
  select(Sensitivity, Specificity, Balanced_Accuracy)
balanced_acc_2_Landsat <- by_class_2_Landsat %>%
  mutate(Balanced_Accuracy = (Sensitivity + Specificity)/2) %>%
  select(Sensitivity, Specificity, Balanced_Accuracy)
balanced_acc_SatEmb <- by_class_SatEmb %>%
  mutate(Balanced_Accuracy = (Sensitivity + Specificity)/2) %>%
  select(Sensitivity, Specificity, Balanced_Accuracy)

balanced_acc_slope_S2$Class <- rownames(by_class_slope_S2)
balanced_acc_slope_S2 <- balanced_acc_slope_S2 %>% select(Class, everything())
balanced_acc_months_S2$Class <- rownames(by_class_months_s2)
balanced_acc_months_S2 <- balanced_acc_months_S2 %>% select(Class, everything())
balanced_acc_2_S2$Class <- rownames(by_class_2_S2)
balanced_acc_2_S2 <- balanced_acc_2_S2 %>% select(Class, everything())
balanced_acc_slope_Landsat$Class <- rownames(by_class_slope_Landsat)
balanced_acc_slope_Landsat <- balanced_acc_slope_Landsat %>% select(Class, everything())
balanced_acc_months_Landsat$Class <- rownames(by_class_months_Landsat)
balanced_acc_months_Landsat <- balanced_acc_months_Landsat %>% select(Class, everything())
balanced_acc_2_Landsat$Class <- rownames(by_class_2_Landsat)
balanced_acc_2_Landsat <- balanced_acc_2_Landsat %>% select(Class, everything())
balanced_acc_SatEmb$Class <- rownames(by_class_SatEmb)
balanced_acc_SatEmb <- balanced_acc_SatEmb %>% select(Class, everything())
balanced_acc_slope_S2
balanced_acc_months_S2
balanced_acc_2_S2
balanced_acc_slope_Landsat
balanced_acc_months_Landsat
balanced_acc_2_Landsat
balanced_acc_SatEmb
```

#### Balanced accuracy macro (average over classes)

```{r}
balanced_acc_macro_slope_S2 <- mean(balanced_acc_slope_S2$Balanced_Accuracy)
balanced_acc_macro_months_S2 <- mean(balanced_acc_months_S2$Balanced_Accuracy)
balanced_acc_macro_2_S2 <- mean(balanced_acc_2_S2$Balanced_Accuracy)
balanced_acc_macro_slope_Landsat <- mean(balanced_acc_slope_Landsat$Balanced_Accuracy)
balanced_acc_macro_months_Landsat <- mean(balanced_acc_months_Landsat$Balanced_Accuracy)
balanced_acc_macro_2_Landsat <- mean(balanced_acc_2_Landsat$Balanced_Accuracy)
balanced_acc_macro_SatEmb <- mean(balanced_acc_SatEmb$Balanced_Accuracy)
balanced_acc_macro_slope_S2
balanced_acc_macro_months_S2
balanced_acc_macro_2_S2
balanced_acc_macro_slope_Landsat
balanced_acc_macro_months_Landsat
balanced_acc_macro_2_Landsat
balanced_acc_macro_SatEmb
```

#### AUC one-vs-all per class + AUC macro

```{r}
classes_slope_S2 <- levels(pred_df_slope_S2$obs)
classes_months_S2 <- levels(pred_df_months_S2$obs)
classes_2_S2 <- levels(pred_df_2_S2$obs)
classes_slope_Landsat <- levels(pred_df_slope_Landsat$obs)
classes_months_Landsat <- levels(pred_df_months_Landsat$obs)
classes_2_Landsat <- levels(pred_df_2_Landsat$obs)
classes_SatEmb <- levels(pred_df_SatEmb$obs)


auc_one_vs_all <- function(df, cls) {
  bin_obs <- as.integer(df$obs == cls)
  prob_cls <- df[[cls]]
  roc_obj <- roc(bin_obs, prob_cls, quiet = TRUE)
  auc(roc_obj)
}

auc_by_class_slope_S2 <- sapply(classes_slope_S2, 
                              function(cls) auc_one_vs_all(pred_df_slope_S2, cls))
auc_by_class_months_s2 <- sapply(classes_months_S2, 
                              function(cls) auc_one_vs_all(pred_df_months_S2, cls))
auc_by_class_2_S2 <- sapply(classes_2_S2, 
                              function(cls) auc_one_vs_all(pred_df_2_S2, cls))
auc_by_class_slope_Landsat <- sapply(classes_slope_Landsat, 
                              function(cls) auc_one_vs_all(pred_df_slope_Landsat, cls))
auc_by_class_months_Landsat <- sapply(classes_months_Landsat, 
                              function(cls) auc_one_vs_all(pred_df_months_Landsat, cls))
auc_by_class_2_Landsat <- sapply(classes_2_Landsat, 
                              function(cls) auc_one_vs_all(pred_df_2_Landsat, cls))
auc_by_class_SatEmb <- sapply(classes_SatEmb, 
                              function(cls) auc_one_vs_all(pred_df_SatEmb, cls))
auc_by_class_slope_S2
auc_by_class_months_s2
auc_by_class_2_S2
auc_by_class_slope_Landsat
auc_by_class_months_Landsat
auc_by_class_2_Landsat
auc_by_class_SatEmb

auc_macro_slope_S2 <- mean(auc_by_class_slope_S2)
auc_macro_months_S2 <- mean(auc_by_class_months_s2)
auc_macro_2_S2 <- mean(auc_by_class_2_S2)
auc_macro_slope_Landsat <- mean(auc_by_class_slope_Landsat)
auc_macro_months_Landsat <- mean(auc_by_class_months_Landsat)
auc_macro_2_Landsat <- mean(auc_by_class_2_Landsat)
auc_macro_SatEmb <- mean(auc_by_class_SatEmb)
auc_macro_slope_S2
auc_macro_months_S2
auc_macro_2_S2
auc_macro_slope_Landsat
auc_macro_months_Landsat
auc_macro_2_Landsat
auc_macro_SatEmb
```

#### Average confusion matrix

```{r}
cm_slope_S2$table
cm_months_S2$table
cm_2_S2$table
cm_slope_Landsat$table
cm_months_Landsat$table
cm_2_Landsat$table
cm_SatEmb$table
```

#### F1-score per class

```{r}
as.data.frame(cm_slope_S2$byClass)[, c("Precision", "Recall", "F1")]
as.data.frame(cm_months_S2$byClass)[, c("Precision", "Recall", "F1")]
as.data.frame(cm_2_S2$byClass)[, c("Precision", "Recall", "F1")]
as.data.frame(cm_slope_Landsat$byClass)[, c("Precision", "Recall", "F1")]
as.data.frame(cm_months_Landsat$byClass)[, c("Precision", "Recall", "F1")]
as.data.frame(cm_2_Landsat$byClass)[, c("Precision", "Recall", "F1")]
as.data.frame(cm_SatEmb$byClass)[, c("Precision", "Recall", "F1")]
```

#### Variable importance

```{r eval=FALSE, include=FALSE}
varimp_slope_S2 <- varImp(cv_slope_S2)
varimp_months_S2 <- varImp(cv_months_S2)
varimp_2_S2 <- varImp(cv_2_S2)
varimp_slope_Landsat <- varImp(cv_slope_Landsat)
varimp_months_Landsat <- varImp(cv_months_Landsat)
varimp_2_Landsat <- varImp(cv_2_Landsat)
varimp_SatEmb <- varImp(cv_SatEmb)
```

```{r}
print(varimp_slope_S2, top = 64)
print(varimp_months_S2, top = 64)
print(varimp_2_S2, top = 64)
print(varimp_slope_Landsat, top = 64)
print(varimp_months_Landsat, top = 64)
print(varimp_2_Landsat, top = 64)
print(varimp_SatEmb, top = 64)
```

##### Save

```{r eval=FALSE, include=FALSE}
save(varimp_slope_S2, file = "objects/RF_1500_cv/cv_slope_S2_varimp.Rdata")
save(varimp_months_S2, file = "objects/RF_1500_cv/cv_months_S2_varimp.Rdata")
save(varimp_2_S2, file = "objects/RF_1500_cv/cv_2_S2_varimp.Rdata")
save(varimp_slope_Landsat, file = "objects/RF_1500_cv/cv_slope_Landsat_varimp.Rdata")
save(varimp_months_Landsat, file = "objects/RF_1500_cv/cv_months_Landsat_varimp.Rdata")
save(varimp_2_Landsat, file = "objects/RF_1500_cv/cv_2_Landsat_varimp.Rdata")
save(varimp_SatEmb, file = "objects/RF_1500_cv/cv_SatEmb_varimp.Rdata")
```

#### ROC one-vs-all for each class (aggregated over folds)

```{r}
# Compute a ROC object for each class (one-vs-all)
roc_list_slope_S2 <- lapply(classes_slope_S2, function(cls) {
  # Binary labels: 1 for cls, 0 for others
  bin_obs <- as.integer(pred_df_slope_S2$obs == cls)
  # Probability for the class
  prob_cls <- pred_df_slope_S2[[cls]]
  # Compute ROC; set direction = "<" if needed depending on your scores
  pROC::roc(response = bin_obs, predictor = prob_cls, quiet = TRUE)
})
names(roc_list_slope_S2) <- classes_slope_S2
# Compute a ROC object for each class (one-vs-all)
roc_list_months_S2 <- lapply(classes_months_S2, function(cls) {
  # Binary labels: 1 for cls, 0 for others
  bin_obs <- as.integer(pred_df_months_S2$obs == cls)
  # Probability for the class
  prob_cls <- pred_df_months_S2[[cls]]
  # Compute ROC; set direction = "<" if needed depending on your scores
  pROC::roc(response = bin_obs, predictor = prob_cls, quiet = TRUE)
})
names(roc_list_months_S2) <- classes_months_S2
# Compute a ROC object for each class (one-vs-all)
roc_list_2_S2 <- lapply(classes_2_S2, function(cls) {
  # Binary labels: 1 for cls, 0 for others
  bin_obs <- as.integer(pred_df_2_S2$obs == cls)
  # Probability for the class
  prob_cls <- pred_df_2_S2[[cls]]
  # Compute ROC; set direction = "<" if needed depending on your scores
  pROC::roc(response = bin_obs, predictor = prob_cls, quiet = TRUE)
})
names(roc_list_2_S2) <- classes_2_S2
# Compute a ROC object for each class (one-vs-all)
roc_list_slope_Landsat <- lapply(classes_slope_Landsat, function(cls) {
  # Binary labels: 1 for cls, 0 for others
  bin_obs <- as.integer(pred_df_slope_Landsat$obs == cls)
  # Probability for the class
  prob_cls <- pred_df_slope_Landsat[[cls]]
  # Compute ROC; set direction = "<" if needed depending on your scores
  pROC::roc(response = bin_obs, predictor = prob_cls, quiet = TRUE)
})
names(roc_list_slope_Landsat) <- classes_slope_Landsat
# Compute a ROC object for each class (one-vs-all)
roc_list_months_Landsat <- lapply(classes_months_Landsat, function(cls) {
  # Binary labels: 1 for cls, 0 for others
  bin_obs <- as.integer(pred_df_months_Landsat$obs == cls)
  # Probability for the class
  prob_cls <- pred_df_months_Landsat[[cls]]
  # Compute ROC; set direction = "<" if needed depending on your scores
  pROC::roc(response = bin_obs, predictor = prob_cls, quiet = TRUE)
})
names(roc_list_months_Landsat) <- classes_months_Landsat
# Compute a ROC object for each class (one-vs-all)
roc_list_2_Landsat <- lapply(classes_2_Landsat, function(cls) {
  # Binary labels: 1 for cls, 0 for others
  bin_obs <- as.integer(pred_df_2_Landsat$obs == cls)
  # Probability for the class
  prob_cls <- pred_df_2_Landsat[[cls]]
  # Compute ROC; set direction = "<" if needed depending on your scores
  pROC::roc(response = bin_obs, predictor = prob_cls, quiet = TRUE)
})
names(roc_list_2_Landsat) <- classes_2_Landsat
# Compute a ROC object for each class (one-vs-all)
roc_list_SatEmb <- lapply(classes_SatEmb, function(cls) {
  # Binary labels: 1 for cls, 0 for others
  bin_obs <- as.integer(pred_df_SatEmb$obs == cls)
  # Probability for the class
  prob_cls <- pred_df_SatEmb[[cls]]
  # Compute ROC; set direction = "<" if needed depending on your scores
  pROC::roc(response = bin_obs, predictor = prob_cls, quiet = TRUE)
})
names(roc_list_SatEmb) <- classes_SatEmb
```

```{r}
# Extract ROC curve data (TPR vs FPR) per class
roc_df_slope_S2 <- lapply(names(roc_list_slope_S2), function(cls) {
  coords <- pROC::coords(roc_list_slope_S2[[cls]], 
                         ret = c("specificity", "sensitivity"), 
                         transpose = FALSE)
  data.frame(Class = cls,
             FPR = 1 - coords$specificity,
             TPR = coords$sensitivity)
}) %>% bind_rows()
# Extract ROC curve data (TPR vs FPR) per class
roc_df_months_S2 <- lapply(names(roc_list_months_S2), function(cls) {
  coords <- pROC::coords(roc_list_months_S2[[cls]], 
                         ret = c("specificity", "sensitivity"),
                         transpose = FALSE)
  data.frame(Class = cls,
             FPR = 1 - coords$specificity,
             TPR = coords$sensitivity)
}) %>% bind_rows()
# Extract ROC curve data (TPR vs FPR) per class
roc_df_2_S2 <- lapply(names(roc_list_2_S2), function(cls) {
  coords <- pROC::coords(roc_list_2_S2[[cls]], 
                         ret = c("specificity", "sensitivity"), 
                         transpose = FALSE)
  data.frame(Class = cls,
             FPR = 1 - coords$specificity,
             TPR = coords$sensitivity)
}) %>% bind_rows()
# Extract ROC curve data (TPR vs FPR) per class
roc_df_slope_Landsat <- lapply(names(roc_list_slope_Landsat), function(cls) {
  coords <- pROC::coords(roc_list_slope_Landsat[[cls]], 
                         ret = c("specificity", "sensitivity"), 
                         transpose = FALSE)
  data.frame(Class = cls,
             FPR = 1 - coords$specificity,
             TPR = coords$sensitivity)
}) %>% bind_rows()
# Extract ROC curve data (TPR vs FPR) per class
roc_df_months_Landsat <- lapply(names(roc_list_months_Landsat), function(cls) {
  coords <- pROC::coords(roc_list_months_Landsat[[cls]], 
                         ret = c("specificity", "sensitivity"), 
                         transpose = FALSE)
  data.frame(Class = cls,
             FPR = 1 - coords$specificity,
             TPR = coords$sensitivity)
}) %>% bind_rows()
# Extract ROC curve data (TPR vs FPR) per class
roc_df_2_Landsat <- lapply(names(roc_list_2_Landsat), function(cls) {
  coords <- pROC::coords(roc_list_2_Landsat[[cls]],
                         ret = c("specificity", "sensitivity"),
                         transpose = FALSE)
  data.frame(Class = cls,
             FPR = 1 - coords$specificity,
             TPR = coords$sensitivity)
}) %>% bind_rows()
# Extract ROC curve data (TPR vs FPR) per class
roc_df_SatEmb <- lapply(names(roc_list_SatEmb), function(cls) {
  coords <- pROC::coords(roc_list_SatEmb[[cls]], ret = c("specificity", "sensitivity"), transpose = FALSE)
  data.frame(Class = cls,
             FPR = 1 - coords$specificity,
             TPR = coords$sensitivity)
}) %>% bind_rows()

roc_plot_slope_S2 <- ggplot(roc_df_slope_S2, 
                            aes(x = FPR, y = TPR, color = Class)) +
  geom_line(linewidth = 0.75) +
  #scale_color_manual(values = cols) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  labs(# title = "ROC curves (one-vs-all) by class",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 12)
roc_plot_months_S2 <- ggplot(roc_df_months_S2, 
                          aes(x = FPR, y = TPR, color = Class)) +
  geom_line(linewidth = 0.75) +
  #scale_color_manual(values = cols) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  labs(# title = "ROC curves (one-vs-all) by class",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 12)
roc_plot_2_S2 <- ggplot(roc_df_2_S2, 
                          aes(x = FPR, y = TPR, color = Class)) +
  geom_line(linewidth = 0.75) +
  #scale_color_manual(values = cols) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  labs(# title = "ROC curves (one-vs-all) by class",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 12)
roc_plot_slope_Landsat <- ggplot(roc_df_slope_Landsat, 
                          aes(x = FPR, y = TPR, color = Class)) +
  geom_line(linewidth = 0.75) +
  #scale_color_manual(values = cols) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  labs(# title = "ROC curves (one-vs-all) by class",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 12)
roc_plot_months_Landsat <- ggplot(roc_df_months_Landsat, 
                          aes(x = FPR, y = TPR, color = Class)) +
  geom_line(linewidth = 0.75) +
  #scale_color_manual(values = cols) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  labs(# title = "ROC curves (one-vs-all) by class",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 12)
roc_plot_2_Landsat <- ggplot(roc_df_2_Landsat, 
                          aes(x = FPR, y = TPR, color = Class)) +
  geom_line(linewidth = 0.75) +
  #scale_color_manual(values = cols) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  labs(# title = "ROC curves (one-vs-all) by class",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 12)
roc_plot_SatEmb <- ggplot(roc_df_SatEmb, 
                          aes(x = FPR, y = TPR, color = Class)) +
  geom_line(linewidth = 0.75) +
  #scale_color_manual(values = cols) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  labs(# title = "ROC curves (one-vs-all) by class",
       x = "False Positive Rate (1 - Specificity)",
       y = "True Positive Rate (Sensitivity)") +
  theme_minimal(base_size = 12)
roc_plot_slope_S2
roc_plot_months_S2
roc_plot_2_S2
roc_plot_slope_Landsat
roc_plot_months_Landsat
roc_plot_2_Landsat
roc_plot_SatEmb
```

# Plots ROC curves

## CHANGE (Appendix 5) ROC curves Landsat

```{r}
roc_plot_monhts_Landsat + roc_plot_SatEmb +
  plot_annotation(tag_levels = "A", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom", legend.title = element_blank())
ggsave(here("output", "figures", "manuscript", "appendix4.jpeg"), 
       dpi = 300, width = 9, height = 4)
```

# Plots variable importance 

## CHANGE (Appendix 6) Best model Landsat

```{r}
vi_Landsat_caret_df <- as.data.frame(varimp_months_Landsat$importance)
vi_Landsat_caret_df$Variable <- rownames(vi_Landsat_caret_df)
colnames(vi_Landsat_caret_df)[1] <- "ScaledImportance"
vi_Landsat_caret_df <- vi_Landsat_caret_df %>% arrange(desc(ScaledImportance))
ggplot(vi_Landsat_caret_df, 
       aes(x = reorder(Variable, ScaledImportance), y = ScaledImportance)) +
  geom_col(fill = "#2c7fb8") +
  coord_flip() +
  labs(x = "Predictor", y = "Importance (scaled)") +
  theme_minimal()
ggsave(here("output", "figures", "manuscript", "appendix6.jpeg"), 
       dpi = 300, width = 6, height = 10)
```

## CHANGE (Appendix 7) Best model SatEmb / S2

```{r}
vi_SatEmb_caret_df <- as.data.frame(varimp_SatEmb$importance)
vi_SatEmb_caret_df$Variable <- rownames(vi_SatEmb_caret_df)
colnames(vi_SatEmb_caret_df)[1] <- "ScaledImportance"
vi_SatEmb_caret_df <- vi_SatEmb_caret_df %>% arrange(desc(ScaledImportance))
ggplot(vi_SatEmb_caret_df, 
       aes(x = reorder(Variable, ScaledImportance), y = ScaledImportance)) +
  geom_col(fill = "#2c7fb8") +
  coord_flip() +
  labs(x = "Predictor", y = "Importance (scaled)") +
  theme_minimal()
ggsave(here("output", "figures", "manuscript", "appendix7.jpeg"), 
       dpi = 300, width = 6, height = 10)
```

# Session info

```{r}
sessionInfo()
```


