---
title: "Script to get validated points from the ReSurvey database using best RF models"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

This R script is used get validated points from the ReSurvey database using the best RF models.

# Load libraries

```{r}
library(tidyverse)
library(here)
library(rnaturalearth)
library(sf)
library(patchwork)
library(scales)
```

# Load functions

```{r}
knitr::purl("functions_20260225.Rmd", output = "functions_20260225.R")
source("functions_20260225.R")
```

# Load previously created objects

```{r}
# Define the folder path
folder_path <- here("objects", "validated_points")

# List all .RData or .rda files in the folder
rdata_files <- list.files(folder_path, full.names = TRUE)

# Load each file
lapply(rdata_files, load, envir = .GlobalEnv)
```

# Load best models

Model with SatEmb for predicting where there is SatEmb data:

```{r}
# Load best model S2 or SatEmb (and prediction) - Accuracy = 0.837 (without k-fold)
rf_SatEmb <- get(load(here("objects", "RF_1500_20260309",
                           "rf_SatEmb_refin_20_10_2x_model.Rdata")))$model
```

Model with months, Landsat for predicting where there is no SatEmb data:

```{r}
# Load best model Landsat (and prediction) - Accuracy = 0.872 (without k-fold)
rf_months_Landsat <- get(load(here("objects", "RF_1500_20260309",
                               "rf_months_Landsat_refin_20_10_2x_model.Rdata")))$model
```

# Read data

Read data validation SatEmb and Landsat:

```{r}
data_validation_SatEmb <- read_tsv(here(
  "data", "clean","final_RS_data_SatEmb_20260309.csv"))
data_validation_Landsat <- read_tsv(here(
  "data", "clean","final_RS_data_bands_Landsat_all_20260309.csv"))
```

No parsing issues!

```{r}
nrow(data_validation_Landsat)
nrow(data_validation_SatEmb)
```

Read original db:

```{r}
data_V2025.2 <- read_tsv(
  "C:/Data/MOTIVATE/MOTIVATE_data/Data/EVA and ReSurvey Data/2025 data - V2025.2/Core data/200_Motivate20250415_notJUICE_header_with_precise_coordinates.csv")
```

# Define lists of predictor variables

```{r}
# Excluding EVI because many values outside of range -1, 1
vars_RF_months_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, and SAVI between march and oct
  "NDVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI and SAVI at march, pos and oct
  "NDVI_avg_value_03", "NDVI_pos_value", "NDVI_avg_value_10",
  "SAVI_avg_value_03", "SAVI_pos_value", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "SAVI_diff_pos_march_value",
  "NDVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "SAVI_diff_pos_oct_value",
  "NDVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy",
  # Canopy height
  "canopy_height")
vars_RF_SatEmb <- colnames(
  data_validation_SatEmb)[startsWith(colnames(data_validation_SatEmb), "A")]
```

# Merge both datasets

```{r}
data_validation <- data_validation_Landsat %>% 
  full_join(data_validation_SatEmb) %>%
  # Add SatEmb T/F
  mutate(SatEmb = ifelse(is.na(A00), FALSE, TRUE))
```

```{r}
data_validation %>% count(SatEmb)
```

# Apply similar filtering as to prepare data for models

First, we were predicting in all data, even with missing values in predictors and values out ot the expected ranges!

```{r}
nrow(data_validation)
```

```{r}
data_validation_filtered <- data_validation %>%
  # Data filtering
  # Remove rows with missing values in predictors
  dplyr::filter(if_all(all_of(vars_RF_months_Landsat), ~ !is.na(.))) %>%
  # Remove all rows with wrong values of indices (not between -1 and 1)
  dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
  # With Landsat data, many observations removed because of high/low EVI values
  # In Landsat_extraction, values of EVI outside of expected range were set as NA,
  # so better not to use EVI here
  # dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
  dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
  dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
  dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
  # Keep only rows with differences > 0 in variables
  filter(if_all(any_of(vars_RF_months_Landsat[grepl("diff", 
                                                   vars_RF_months_Landsat)]), 
                ~ .x > 0))
```

```{r}
nrow(data_validation_filtered)
```

```{r}
# No missing values in SatEmb variables
data_validation_filtered %>%
  filter(if_any(starts_with("A"), ~ is.na(.x)) & SatEmb == T)
```

# Predictions from best models 

Predict in data without NAs! 

Model with SatEmb for predicting where there is SatEmb data:

```{r eval=FALSE, include=FALSE}
prediction_SatEmb_resp <- predict(rf_SatEmb,
                                  newdata = data_validation %>%
                                    filter(SatEmb == T) %>%
                                    filter(if_all(all_of(vars_RF_SatEmb), ~ !is.na(.))), 
                                  type = "response")
```

```{r eval=FALSE, include=FALSE}
prediction_SatEmb_prob <- predict(rf_SatEmb,
                                  newdata = data_validation %>%
                                    filter(SatEmb == T) %>%
                                    filter(if_all(all_of(vars_RF_SatEmb), ~ !is.na(.))), 
                                  type = "prob")
```

Model with months, Landsat for predicting where there is no SatEmb data:

```{r eval=FALSE, include=FALSE}
prediction_Landsat_resp <- predict(rf_months_Landsat,
                                   newdata = data_validation_filtered, 
                                   type = "response")
```

```{r eval=FALSE, include=FALSE}
prediction_Landsat_prob <- predict(rf_months_Landsat,
                                   newdata = data_validation_filtered, 
                                   type = "prob")
```

```{r}
predictions <- c("prediction_SatEmb_resp", "prediction_SatEmb_prob",
                 "prediction_Landsat_resp", "prediction_Landsat_prob")
```

Save:

```{r eval=FALSE, include=FALSE}
for (prediction_name in predictions) {
  file_path <- paste0("objects/validated_points/", prediction_name, ".Rdata")
  save(list = prediction_name, file = file_path)
}
```

Add predictions:

```{r}
data_validated <- data_validation %>%
  left_join(
    # Predictions from model with SatEmb
    cbind(
      data.frame(
        # PlotObservationIDs
        PlotObservationID = data_validation %>% 
          filter(SatEmb == T)  %>%
          filter(if_all(all_of(vars_RF_SatEmb), ~ !is.na(.))) %>% 
          pull(PlotObservationID),
        # Prediction response
        pred_SatEmb_resp = prediction_SatEmb_resp
        ),
      # Prediction probabilities
      do.call(rbind, prediction_SatEmb_prob) %>% as.data.frame() %>%
        rename(pred_SatEmb_Q = EUNIS1.Q, pred_SatEmb_R = EUNIS1.R,
               pred_SatEmb_S = EUNIS1.S, pred_SatEmb_T = EUNIS1.T)
      )
  ) %>%
  left_join(
    # Predictions from model with months, Landsat
    cbind(
      data.frame(
        # PlotObservationIDs
        PlotObservationID = data_validation_filtered %>% 
          pull(PlotObservationID),
        # Prediction response
        pred_Landsat_resp = prediction_Landsat_resp
        ),
      # Prediction probabilities
      do.call(rbind, prediction_Landsat_prob) %>% as.data.frame() %>%
        rename(pred_Landsat_Q = EUNIS1.Q, pred_Landsat_R = EUNIS1.R,
               pred_Landsat_S = EUNIS1.S, pred_Landsat_T = EUNIS1.T)
      )
  ) %>%
  # Validated or not (SatEmb)
  mutate(valid_SatEmb = ifelse(pred_SatEmb_resp == EUNIS1, TRUE, 
                               ifelse(is.na(pred_SatEmb_resp), NA, FALSE))) %>%
  # Validated or not (Landsat)
  mutate(valid_Landsat = ifelse(pred_Landsat_resp == EUNIS1, TRUE, 
                                ifelse(is.na(pred_Landsat_resp), NA, FALSE)))
```

# Counts

```{r}
data_validated %>% count(valid_SatEmb)
data_validated %>% count(valid_Landsat)
```

# HERE: Data cleaning

```{r}
data_validated_clean <- data_validated %>%
  # Keep only needed cols
  select(
    # These three cols are supposed to uniquely define each resurvey plot
    RS_CODE, RSrvyst, RSrvypl,
    # The three previous cols + this one are supposed to uniquely define
    # each resurvey observation
    RSrvyob,
    # Unique identifier for each observation
    PlotObservationID,
    Dtofrcr, year,
    biogeo, unit,
    Lctnmth,
    EUNIS1, EUNIS2,
    SatEmb, 
    pred_SatEmb_resp, pred_SatEmb_Q, pred_SatEmb_R, pred_SatEmb_S, pred_SatEmb_T,
    pred_Landsat_resp, pred_Landsat_Q, pred_Landsat_R, pred_Landsat_S, pred_Landsat_T,
    valid_SatEmb, valid_Landsat, 
    ) %>%
  # Rename some cols as in original db
  rename(`ReSurvey site` = RSrvyst, `ReSurvey plot` = RSrvypl, 
         `ReSurvey observation` = RSrvyob, `Date of recording` = Dtofrcr,
         `Location method` = Lctnmth ) %>%
  # Create new valid col
  mutate(valid = case_when(
    # If there is no SatEmb info, use Landsat validation
    SatEmb == F ~ valid_Landsat,
    # If there is SatEmb info, use SatEmb validation
    SatEmb == T ~ valid_SatEmb
    )) %>%
  # Create new cols with predictions for each habitat type
  mutate(
    pred_Q = case_when(
      # If there is no SatEmb info, use Landsat prediction
      SatEmb == F ~ pred_Landsat_Q,
      # If there is SatEmb info, use SatEmb prediction
      SatEmb == T ~ pred_SatEmb_Q),
    pred_R = case_when(
      # If there is no SatEmb info, use Landsat prediction
      SatEmb == F ~ pred_Landsat_R,
      # If there is SatEmb info, use SatEmb prediction
      SatEmb == T ~ pred_SatEmb_R),
    pred_S = case_when(
      # If there is no SatEmb info, use Landsat prediction
      SatEmb == F ~ pred_Landsat_S,
      # If there is SatEmb info, use SatEmb prediction
      SatEmb == T ~ pred_SatEmb_S),
    pred_T = case_when(
      # If there is no SatEmb info, use Landsat prediction
      SatEmb == F ~ pred_Landsat_T,
      # If there is SatEmb info, use SatEmb prediction
      SatEmb == T ~ pred_SatEmb_T)
    )
```

```{r}
data_validated_clean %>% count(valid)
```


How many NAs for each row?

```{r}
data_validated_clean %>% 
  summarise(across(everything(), ~ sum(is.na(.))))
```

# Get info from original db

## Read original db

```{r}
data_V2025.2 <- read_tsv(
  "C:/Data/MOTIVATE/MOTIVATE_data/Data/EVA and ReSurvey Data/2025 data - V2025.2/Core data/200_Motivate20250415_notJUICE_header_with_precise_coordinates.csv")
```

## Get info for cols where there were NAs

```{r}
data_validated_clean <- rows_patch( # Updates only the NAs
  data_validated_clean,
  data_V2025.2 %>%
    # Select cols where there were NAs
    select(RS_CODE, `ReSurvey site`, `ReSurvey plot`, `ReSurvey observation`, 
           PlotObservationID, `Date of recording`, `Location method`),
  by = "PlotObservationID",
  unmatched = "ignore" # Only patches matching rows and ignores the rest
  )
```

## Add coordinates

```{r}
data_validated_clean <- data_validated_clean %>%
  left_join(
    data_V2025.2 %>% 
      # Update coordinates
      mutate(Lon_updated = ifelse(is.na(Lon_prec), Longitude, Lon_prec),
             Lat_updated = ifelse(is.na(Lat_prec), Latitude, Lat_prec)) %>%
      # Select columns needed
      select(PlotObservationID, Lon_updated, Lat_updated)
    )
```

How many NAs for each row?

```{r}
data_validated_clean %>% 
  summarise(across(everything(), ~ sum(is.na(.))))
```

# More data cleaning

```{r}
data_validated_clean <- data_validated_clean %>%
  # Convert to date
  mutate(`Date of recording` = dmy(`Date of recording`))
```

# CHECK: Obs with more than one row

```{r}
data_validated_clean %>%
  # These 4 cols uniquely identify each obs
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`, `ReSurvey observation`) %>%
  # Filter obs with more than 1 row
  filter(n() > 1) %>%
  arrange(RS_CODE, `ReSurvey site`, `ReSurvey plot`, `ReSurvey observation`)
```

For each unique combination of RS_CODE, ReSurvey site, ReSurvey plot and ReSurvey observation, check whether any of the other columns (except PlotObservationID) differ across the rows in that group:

```{r}
data_validated_clean %>%
  group_by(
    RS_CODE, `ReSurvey site`, `ReSurvey plot`, `ReSurvey observation`
  ) %>%
  filter(n() > 1) %>%
  group_modify(~ {
    
    # Exclude columns that should NOT be checked
    cols_to_check <- setdiff(names(.x), "PlotObservationID")
    
    # Identify columns where at least one value differs
    diffs <- map_lgl(.x[cols_to_check], ~ n_distinct(.x, na.rm = FALSE) > 1)
    
    tibble(
      differing_columns = paste(names(diffs)[diffs], collapse = ", ")
      )
    }) %>%
  ungroup() %>%
  # Keep only groups where at least one difference exists
  filter(differing_columns != "")
```

# Add info on nr resurveys (n different years)

Some have only 1 year, either because the first year(s) were prior to Landsat, or because they were still not resurveyed. 

```{r}
data_validated_clean <- data_validated_clean %>%
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`) %>%
  mutate(
    n_resurveys = n_distinct(year, na.rm = TRUE),   # count unique years
    first_year  = min(year, na.rm = TRUE),          # earliest year
    last_year   = max(year, na.rm = TRUE)           # latest year
  ) %>%
  ungroup()
```

# Get only latest obs for each plot

```{r}
data_validated_clean_latest <- data_validated_clean %>%
  # Keep only rows from the last year
  filter(year == last_year) %>%
  # These 3 cols are supposed to uniquely identify each plot
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`) %>%
  # If more than one row in the same year, keep the latest date
  #slice_max(`Date of recording`, n = 1, with_ties = FALSE) %>%
  ungroup()
```

## Obs with more than one row

```{r}
data_validated_clean_latest %>%
  # These 3 cols are supposed to uniquely identify each plot
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`) %>%
  # Filter obs with more than 1 row
  filter(n() > 1) %>%
  ungroup() %>%
  arrange(RS_CODE, `ReSurvey site`, `ReSurvey plot`)
```

# Write csvs

```{r}
write_csv(data_validated_clean,
          here("data", "clean","data_validated_clean_20260316.csv"))
write_csv(data_validated_clean_latest,
          here("data", "clean","data_validated_clean_latest_20260316.csv"))
```

# Counts of validated / not-validated obs

## General

```{r}
data_validated_clean %>%
  count(valid) %>%
  mutate(percent = n / sum(n) * 100)
```

```{r}
data_validated_clean %>%
  filter(!is.na(valid)) %>%
  count(valid) %>%
  mutate(percent = n / sum(n) * 100)
```

## By habitat

```{r}
data_validated_clean %>%
  filter(EUNIS1 == "Q") %>%
  filter(!is.na(valid)) %>%
  count(valid) %>%
  mutate(percent = n / sum(n) * 100)
```

```{r}
data_validated_clean %>%
  filter(EUNIS1 == "R") %>%
  filter(!is.na(valid)) %>%
  count(valid) %>%
  mutate(percent = n / sum(n) * 100)
```

```{r}
data_validated_clean %>%
  filter(EUNIS1 == "S") %>%
  filter(!is.na(valid)) %>%
  count(valid) %>%
  mutate(percent = n / sum(n) * 100)
```

```{r}
data_validated_clean %>%
  filter(EUNIS1 == "T") %>%
  filter(!is.na(valid)) %>%
  count(valid) %>%
  mutate(percent = n / sum(n) * 100)
```

# HERE: Validation of plots

Extract EUNIS assignations for each plot (sometimes different among different observations of the same plot):

```{r}
eunis_wide <- data_validated_clean %>%
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`, n_resurveys) %>%
  reframe(
    EUNISa_vals = unique(EUNIS1)) %>%
  unnest(EUNISa_vals) %>%
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`, n_resurveys) %>%
  mutate(EUNIS_idx = row_number()) %>%
  pivot_wider(
    names_from = EUNIS_idx,
    values_from = EUNISa_vals,
    names_prefix = "EUNIS1_"
  ) %>%
  # Categorize as "Habitat change" if there was more than one EUNIS assigned
  # to different observations within a plot --> HERE
  mutate(EUNIS = case_when(
    is.na(EUNISa_1_2) & is.na(EUNISa_1_3) & is.na(EUNISa_1_4) ~ EUNISa_1_1,
    TRUE ~ "Change"
    )) %>%
  ungroup() %>%
  # Remove unneeded cols
  select(-EUNISa_1_1, -EUNISa_1_2, -EUNISa_1_3, -EUNISa_1_4)
```

Extract the latest coordinate pair per plot (sometimes coordinates are different among different observations of the same plot):

```{r}
coords_latest <- data_validated_clean %>%
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`, n_resurveys) %>%
  slice_max(order_by = year, n = 1, with_ties = FALSE) %>%
  select(
    RS_CODE, `ReSurvey site`, `ReSurvey plot`, n_resurveys,
    Lon_updated, Lat_updated
  ) %>%
  ungroup()
```


```{r}
data_validated_plots <- data_validated_clean %>%
  left_join(
    eunis_wide,
    by = c("RS_CODE", "ReSurvey site", "ReSurvey plot", "n_resurveys")
  ) %>%
  group_by(RS_CODE, `ReSurvey site`, `ReSurvey plot`, n_resurveys, EUNIS) %>%
  summarise(
    n_obs = n(),
    n_true = sum(valid == TRUE, na.rm = TRUE),
    n_false = sum(valid == FALSE, na.rm = TRUE),
    n_na = sum(is.na(valid)),
    .groups = "drop"
  ) %>%
  mutate(
    share_true = ifelse((n_true + n_false) > 0,
                        n_true / (n_true + n_false), NA_real_),
    share_false = ifelse((n_true + n_false) > 0,
                         n_false / (n_true + n_false), NA_real_)
  ) %>%
  mutate(
    valid_status = case_when(
      # --- ALL cases (keep first so they never get overridden) ---
      # All observations validated
      n_true  == n_obs ~ "all_TRUE",
      # All observations not validated
      n_false == n_obs ~ "all_FALSE",
      # All observations NA
      n_na    == n_obs ~ "all_NA",
      
      # Equal number of observations validated and not validated
      n_true == n_false & n_na == 0 ~ "TRUE_FALSE",

      # All three present --> keep as some observations validated and some not validated
      # --- TRUE + FALSE + NA -> treat as TRUE_FALSE
      n_true == n_false & n_na > 0 ~ "TRUE_FALSE",
      
      # --- More cases (dominance) ---
      # TRUE dominates FALSE
      share_true > share_false & n_false > 0 ~ "more_TRUE",
      # FALSE dominates TRUE
      share_false > share_true & n_true > 0 ~ "more_FALSE",

      # Some observations validated and some NA --> keep as all validated
      # TRUE + NA  --> treat as all_TRUE
      n_true  > 0 & n_na > 0 & n_false == 0 ~ "all_TRUE",

      # Some observations not validated and some NA --> keep as all not validated
      # FALSE + NA --> treat as all_FALSE
      n_false > 0 & n_na > 0 & n_true  == 0 ~ "all_FALSE",
      
      # Fallback (shouldn’t happen, but safe to keep)
      TRUE ~ "mixed"
    ),
    n_resurveys_equal_n_obs = n_resurveys == n_obs
  ) %>%
  # Less restrictive validity status
  mutate(valid_status_new = case_when(
    valid_status == "TRUE_FALSE" ~ "tie",
    valid_status %in% c("all_FALSE", "more_FALSE") ~ "not_valid",
    valid_status %in% c("all_TRUE", "more_TRUE") ~ "valid",
    valid_status == "all_NA" ~ NA,
    TRUE ~ "other"
  )) %>%
  # Add coordinates (if coordinates are different among different observations 
  # of the same plot, extract the latest coordinate pair per plot, i.e. from
  # the last observation year)
  left_join(
    coords_latest,
    by = c("RS_CODE", "ReSurvey site", "ReSurvey plot", "n_resurveys")
  )
```

```{r}
data_validated_plots %>% count(valid_status)
data_validated_plots %>% count(valid_status_new)
```

# Counts of validated / not-validated plots

## General

```{r}
data_validated_plots %>%
  count(valid_status_new) %>%
  mutate(percent = n / sum(n) * 100)
```

```{r}
data_validated_plots %>%
  filter(!is.na(valid_status_new)) %>%
  count(valid_status_new) %>%
  mutate(percent = n / sum(n) * 100)
```

## By habitat

```{r}
data_validated_plots %>%
  filter(EUNIS == "Q") %>%
  filter(!is.na(valid_status_new)) %>%
  count(valid_status_new) %>%
  mutate(percent = n / sum(n) * 100)
```

```{r}
data_validated_plots %>%
  filter(EUNIS == "R") %>%
  filter(!is.na(valid_status_new)) %>%
  count(valid_status_new) %>%
  mutate(percent = n / sum(n) * 100)
```

```{r}
data_validated_plots %>%
  filter(EUNIS == "S") %>%
  filter(!is.na(valid_status_new)) %>%
  count(valid_status_new) %>%
  mutate(percent = n / sum(n) * 100)
```

```{r}
data_validated_plots %>%
  filter(EUNIS == "T") %>%
  filter(!is.na(valid_status_new)) %>%
  count(valid_status_new) %>%
  mutate(percent = n / sum(n) * 100)
```

```{r}
data_validated_plots %>%
  filter(EUNIS == "Change") %>%
  filter(!is.na(valid_status_new)) %>%
  count(valid_status_new) %>%
  mutate(percent = n / sum(n) * 100)
```

# Plots

## Obs level

```{r}
plot1 <- data_validated_clean %>%
  count(EUNIS1, valid, name = "n") %>% 
  ggplot(aes(x = EUNIS1, y = n, fill = valid)) +
  geom_col(position = position_stack()) +
  xlab(NULL) + ylab("Number of observations") +
  scale_y_continuous(labels = label_number(scale = 1e-3, suffix = "k")) +
  scale_fill_manual(values = c("FALSE" = "#D55E00", "TRUE" = "#0072B2",
                               "NA" = "#828282"),
                     name = "Validated") +
  theme_minimal() + theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot1
plot1_noNAs <- data_validated_clean %>%
  filter(!is.na(valid)) %>%
  count(EUNIS1, valid, name = "n") %>% 
  ggplot(aes(x = EUNIS1, y = n, fill = valid)) +
  geom_col(position = position_stack()) +
  xlab(NULL) + ylab("Number of observations") +
  scale_y_continuous(labels = label_number(scale = 1e-3, suffix = "k")) +
  scale_fill_manual(values = c("FALSE" = "#D55E00", "TRUE" = "#0072B2"),
                     name = "Validated") +
  theme_minimal() + theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot1_noNAs
```

## Plot level

```{r}
plot2 <- data_validated_plots %>%
  count(EUNIS, valid_status_new, name = "n") %>% 
  ggplot(aes(x = EUNIS, y = n, fill = valid_status_new)) +
  geom_col(position = position_stack()) +
  xlab(NULL) + ylab("Number of plots") +
  scale_y_continuous(labels = label_number(scale = 1e-3, suffix = "k")) +
  scale_fill_manual(values = c("valid" = "#D55E00", "not_valid" = "#0072B2",
                               "NA" = "#828282", tie = "#009E73"),
                     name = "Validated") +
  theme_minimal() + theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot2
plot2_noNAs <- data_validated_plots %>%
  filter(!is.na(valid_status_new)) %>%
  count(EUNIS, valid_status_new, name = "n") %>% 
  ggplot(aes(x = EUNIS, y = n, fill = valid_status_new)) +
  geom_col(position = position_stack()) +
  xlab(NULL) + ylab("Number of plots") +
  scale_y_continuous(labels = label_number(scale = 1e-3, suffix = "k")) +
  scale_fill_manual(values = c("valid" = "#D55E00", "not_valid" = "#0072B2",
                               "NA" = "#828282", tie = "#009E73"),
                     name = "Validated") +
  theme_minimal() + theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot2_noNAs
```

# Map

## Obs level

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
world_3035 <- st_transform(world, 3035)
points_3035 <- st_as_sf(
  data_validated_clean,
  coords = c("Lon_updated", "Lat_updated"),
  crs = 4326
) |>
  st_transform(3035)

europe_bbox_3035 <- st_bbox(
  c(
    xmin = 2500000,
    ymin = 1400000,
    xmax = 6500000,
    ymax = 5500000
  ),
  crs = st_crs(3035)
)
world_3035_crop <- st_crop(world_3035, europe_bbox_3035)
points_3035_crop <- st_crop(points_3035, europe_bbox_3035)
```

```{r}
map1 <- ggplot() +
  geom_sf(data = world_3035_crop, fill = "lightgrey", color = "gray") +
  geom_sf(
    data = points_3035_crop,
    aes(color = valid),
    size = 0.05
  ) +
  facet_wrap(~ EUNIS1) +
  scale_color_manual(values = c("FALSE" = "#D55E00", "TRUE"  = "#0072B2",
                                "NA" = "#828282"),
                     name = "Validated") +
  theme_minimal() +
  theme(legend.position = "none")
map1
map1_noNAs <- ggplot() +
  geom_sf(data = world_3035_crop, fill = "lightgrey", color = "gray") +
  geom_sf(
    data = points_3035_crop %>% filter(!is.na(valid)),
    aes(color = valid),
    size = 0.05
  ) +
  facet_wrap(~ EUNIS1) +
  scale_color_manual(values = c("FALSE" = "#D55E00", "TRUE"  = "#0072B2"),
                     name = "Validated") +
  theme_minimal() +
  theme(legend.position = "none")
map1_noNAs
```

## Plot level

```{r}
plots_3035 <- st_as_sf(
  data_validated_plots,
  coords = c("Lon_updated", "Lat_updated"),
  crs = 4326
) |>
  st_transform(3035)
plots_3035_crop <- st_crop(plots_3035, europe_bbox_3035)
```

```{r}
map2 <- ggplot() +
  geom_sf(data = world_3035_crop, fill = "lightgrey", color = "gray") +
  geom_sf(
    data = plots_3035_crop,
    aes(color = valid_status_new),
    size = 0.05
  ) +
  facet_wrap(~ EUNIS) +
  scale_color_manual(values = c("valid" = "#D55E00", "not_valid"  = "#0072B2",
                     "NA" = "#828282", tie = "#009E73"), name = "Validated") +
  theme_minimal() +
  theme(legend.position = "none")
map2
map2_noNAs <- ggplot() +
  geom_sf(data = world_3035_crop, fill = "lightgrey", color = "gray") +
  geom_sf(
    data = plots_3035_crop %>% filter(!is.na(valid_status_new)),
    aes(color = valid_status_new),
    size = 0.05
  ) +
  facet_wrap(~ EUNIS) +
  scale_color_manual(values = c("valid" = "#D55E00", "not_valid"  = "#0072B2",
                                "NA" = "#828282", tie = "#009E73"),
                     name = "Validated") +
  theme_minimal() +
  theme(legend.position = "none")
map2_noNAs
```

# (Fig 2) Combined (obs level)

```{r}
map1 + plot1  +
  plot_layout(widths = c(4, 1)) +
  plot_annotation(tag_levels = "A", tag_suffix = ")") +
  theme(plot.margin = margin(1, 1, 1, 1))
ggsave(here("output", "figures", "manuscript", "fig2.jpeg"), 
       dpi = 300, width = 5, height = 4)
map1_noNAs + plot1_noNAs +
  plot_layout(widths = c(4, 1)) +
  plot_annotation(tag_levels = "A", tag_suffix = ")") +
  theme(plot.margin = margin(1, 1, 1, 1))
ggsave(here("output", "figures", "manuscript", "fig2_noNAs.jpeg"), 
       dpi = 300, width = 5, height = 4)
```

# (Fig 3) Combined (plot level)

```{r}
plot2_inset <- plot2 +
  theme(
    aspect.ratio = 3/2,        # height / width  → tall & thin (try 3/1, 4/1, 5/1)
    plot.margin  = margin(0, 0 , 0, 0),
    axis.text.x  = element_text(angle = 45, size = 6, vjust = 1, hjust = 1),
    legend.position = "none"
  ) +
  coord_cartesian(clip = "off", expand = FALSE) +
  scale_x_discrete(expand = expansion(mult = c(0.02, 0.05)))

map2_with_inset <- map2 +
  inset_element(
    plot2_inset,
    left   = 0.65,  # move a bit right → narrower
    right  = 0.965, # reduce width
    bottom = -0.2,  # keep low
    top    = 0.62,  # increase height
    align_to = "panel"
  )

map2_with_inset +
  plot_annotation(tag_levels = "A", tag_suffix = ")") +
  theme(plot.margin = margin(1, 1, 1, 1))

ggsave(here("output", "figures", "manuscript", "fig3.jpeg"), 
       dpi = 300, width = 5, height = 4)
```


```{r}
plot2_noNAs_inset <- plot2_noNAs +
  theme(
    aspect.ratio = 3/2,        # height / width  → tall & thin (try 3/1, 4/1, 5/1)
    plot.margin  = margin(0, 0 , 0, 0),
    axis.text.x  = element_text(angle = 45, size = 6, vjust = 1, hjust = 1),
    legend.position = "none"
  ) +
  coord_cartesian(clip = "off", expand = FALSE) +
  scale_x_discrete(expand = expansion(mult = c(0.02, 0.05)))

map2_noNAs_with_inset <- map2_noNAs +
  inset_element(
    plot2_noNAs_inset,
    left   = 0.65,  # move a bit right → narrower
    right  = 0.965, # reduce width
    bottom = -0.2,  # keep low
    top    = 0.62,  # increase height
    align_to = "panel"
  )

map2_noNAs_with_inset +
  plot_annotation(tag_levels = "A", tag_suffix = ")") +
  theme(plot.margin = margin(1, 1, 1, 1))

ggsave(here("output", "figures", "manuscript", "fig3_noNAs.jpeg"), 
       dpi = 300, width = 5, height = 4)
```

# (Fig 1) Plots most influential variables

In these plots, include only obs where valid == T or valid == F, not obs where valid is NA. I.e. include obs where we had data from all predictors, and where able to get a prediction from the RF models. 

```{r}
# Landsat
plot1_Landsat <- data_validated_clean %>%
  filter(!is.na(valid)) %>%
  left_join(data_validation %>% 
              select(PlotObservationID, canopy_height)) %>%
  distr_plot("canopy_height", "Canopy height (m)", grey = F)
plot2_Landsat <- data_validated_clean %>%
  filter(!is.na(valid)) %>%
  left_join(data_validation %>% 
              select(PlotObservationID, SAVI_diff_pos_oct_doy)) %>%
  distr_plot("SAVI_diff_pos_oct_doy", 
             "Difference in SAVI DOY between POS and October", grey = F)
plot3_Landsat <- data_validated_clean %>%
  filter(!is.na(valid)) %>%
  left_join(data_validation %>% 
              select(PlotObservationID, SAVI_diff_pos_march_value)) %>%
  distr_plot("SAVI_diff_pos_march_value", 
             "Difference in SAVI value between POS and March", grey = F)
plot4_Landsat <- data_validated_clean %>%
  filter(!is.na(valid)) %>%
  left_join(data_validation %>% 
              select(PlotObservationID, NDVI_min)) %>%
  distr_plot("NDVI_min", 
             "Minimum value of NDVI", grey = F)
# SatEmb
plot1_SatEmb <- data_validated_clean %>%
  filter(SatEmb == T) %>%
  filter(!is.na(valid)) %>%
  left_join(data_validation %>% 
              select(PlotObservationID, A24)) %>%
  distr_plot("A24", "Band A24", grey = F)
plot2_SatEmb <- data_validated_clean %>%
  filter(SatEmb == T) %>%
  filter(!is.na(valid)) %>%
  left_join(data_validation %>% 
              select(PlotObservationID, A33)) %>%
  distr_plot("A33", "Band A33", grey = F)
plot3_SatEmb <- data_validated_clean %>%
  filter(SatEmb == T) %>%
  filter(!is.na(valid)) %>%
  left_join(data_validation %>% 
              select(PlotObservationID, A07)) %>%
  distr_plot("A07", "Band A07", grey = F)
plot4_SatEmb <- data_validated_clean %>%
  filter(SatEmb == T) %>%
  filter(!is.na(valid)) %>%
  left_join(data_validation %>% 
              select(PlotObservationID, A58)) %>%
  distr_plot("A58", "Band A58", grey = F)
plot1_Landsat + plot2_Landsat + plot3_Landsat + plot4_Landsat +
  plot1_SatEmb + plot2_SatEmb + plot3_SatEmb + plot4_SatEmb +
  plot_layout(ncol = 2, nrow = 4) +
  plot_annotation(tag_levels = "A", tag_suffix = ")")
ggsave(here("output", "figures", "manuscript", "fig1.jpeg"), 
       dpi = 300, width = 8, height = 12)
```

# (Appendix 1) Plot GPS observations used for models

```{r}
# Excluding EVI because many values outside of range -1, 1
vars_RF_months_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, and SAVI between march and oct
  "NDVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI and SAVI at march, pos and oct
  "NDVI_avg_value_03", "NDVI_pos_value", "NDVI_avg_value_10",
  "SAVI_avg_value_03", "SAVI_pos_value", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "SAVI_diff_pos_march_value",
  "NDVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "SAVI_diff_pos_oct_value",
  "NDVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy",
  # Canopy height
  "canopy_height")
```

```{r}
filter_data_Landsat <- function(data, variables) {
  data %>%
    mutate(EUNIS1 = as.factor(EUNIS1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    # With Landsat data, many observations removed because of high/low EVI values
    # In Landsat_extraction, values of EVI outside of expected range were set as NA,
    # so better not to use EVI here
    # dplyr::filter(EVI_max <= 1 & EVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Select only variables needed
    dplyr::select(PlotObservationID, EUNIS1, all_of(variables), Lctnmth,
           valid_1) %>%
    # Keep only rows with differences > 0 in variables
    filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0))
}
```

```{r}
data_models <- filter_data_Landsat(
  data_validation_Landsat %>%
    rough_validation_Landsat(), # Rough validation based on NDWI
  vars_RF_months_Landsat) %>%
  left_join(data_V2025.2 %>%
              # Update coordinates
              mutate(Lon_updated = ifelse(is.na(Lon_prec), Longitude, Lon_prec),
                     Lat_updated = ifelse(is.na(Lat_prec), Latitude, Lat_prec)) %>%
              select(PlotObservationID, `Location method`, Lon_updated, Lat_updated)) %>%
  mutate(GPS = case_when(`Location method` %in% 
                           c("Location with GPS", 
                             "Location with differential GPS") ~ TRUE,
                         !`Location method` %in% 
                           c("Location with GPS", 
                             "Location with differential GPS") ~ FALSE,
                         is.na(`Location method`) ~ FALSE))
```

```{r}
points_3035 <- st_as_sf(data_models,
                        coords = c("Lon_updated", "Lat_updated"),
                        crs = 4326
                        ) |>
  st_transform(3035)

points_3035_crop <- st_crop(points_3035, europe_bbox_3035)
```

```{r}
map2 <- ggplot() +
  geom_sf(data = world_3035_crop, fill = "lightgrey", color = "gray") +
  geom_sf(
    data = points_3035_crop,
    aes(color = GPS),
    size = 0.1
  ) +
  facet_wrap(~ EUNIS1) +
  scale_color_manual(values = c("FALSE" = "#D55E00", "TRUE"  = "#0072B2"),
                     name = "Recorded with GPS") +
  theme_minimal() + theme(legend.position = "none")
map2
ggsave(here("output", "figures", "manuscript", "appendix1.jpeg"), 
       dpi = 300, width = 6, height = 6)
```

# Session info

```{r}
sessionInfo()
```

