---
title: "Spatial validation of plot locations in vegetation databases using remote sensing"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

This R script is used to validate the points in the ReSurvey database using RS indicators from Landsat (indices + phenology).

# Load libraries

```{r}
library(tidyverse)
library(parallel)
library(doParallel)
library(foreach)
library(partykit)
library(caret)
library(here)
library(readr)
library(progressr)
library(party)
library(permimp)
library(tibble)
library(moreparty)
library(patchwork)
library(scales)
library(sf)
library(rnaturalearth)
library(cowplot)
```

# Load previously created objects

```{r}
# Load cross-validation objects
lapply(list.files(here("objects", "RF_1500_cv_20260609"), full.names = TRUE), load,
       envir = .GlobalEnv)
```

```{r}
# Define the folder path
folder_path <- here("objects", "validated_points_20260702")

# List all .RData or .rda files in the folder
rdata_files <- list.files(folder_path, full.names = TRUE)

# Load each file
lapply(rdata_files, load, envir = .GlobalEnv)
```

# Read initial dataset

```{r}
initial_dataset <- read_tsv(here(
  "data", "clean","initial_dataset.csv"))
```

No parsing issues!

```{r}
nrow(initial_dataset)
```

# Define lists of predictor vars

## Moments with slope method

```{r}
vars_RF_slope_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI and SAVI
  "NDVI_auc", "SAVI_auc",
  # Values of NDVI and SAVI at sos, pos and eos
  "NDVI_sos_value", "NDVI_pos_value", "NDVI_eos_value",
  "SAVI_sos_value", "SAVI_pos_value", "SAVI_eos_value",
  # Differences pos-sos in value and doy
  "NDVI_diff_pos_sos_value","SAVI_diff_pos_sos_value",
  "NDVI_diff_pos_sos_doy", "SAVI_diff_pos_sos_doy",
  # Differences pos-eos in value and doy
  "NDVI_diff_pos_eos_value", "SAVI_diff_pos_eos_value",
  "NDVI_diff_pos_eos_doy", "SAVI_diff_pos_eos_doy",
  # Growing season duration
  "NDVI_gsd", "SAVI_gsd")
```

## Months method

```{r}
vars_RF_months_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, and SAVI between march and oct
  "NDVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI and SAVI at march, pos and oct
  "NDVI_avg_value_03", "NDVI_pos_value", "NDVI_avg_value_10",
  "SAVI_avg_value_03", "SAVI_pos_value", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "SAVI_diff_pos_march_value",
  "NDVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "SAVI_diff_pos_oct_value",
  "NDVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy")
```

# Filter data to use in RF models

## Functions

### Function to get full datasets

```{r}
get_full_datasets <- function(data, variables) {
  data %>%
    mutate(EUNIS1 = as.factor(EUNIS1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Keep only rows with differences > 0 in variables
    dplyr::filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0)) %>%
    # Filter based on max NDWI
    mutate(valid_1_NDWI = case_when(NDWI_max > 0.3 ~ "wrong", 
                                    TRUE ~ NA_character_),
           valid_1_count = rowSums(across(c(valid_1_NDWI), ~ . == "wrong"), 
                                   na.rm = TRUE),
           valid_1 = if_else(valid_1_count > 0,
                             "At least 1 rule broken", "No rules broken so far")
           ) %>%
    # Keep only rows with no rules broken
    dplyr::filter(valid_1 == "No rules broken so far") %>%
    # Remove unneeded cols
    select(-starts_with("valid"))
  }
```

### Function to get training subsets

```{r}
get_training_subsets <- function(data, variables) {
  data %>%
    mutate(EUNIS1 = as.factor(EUNIS1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Keep only rows with differences > 0 in variables
    dplyr::filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0)) %>%
    # Filter based on max NDWI
    mutate(valid_1_NDWI = case_when(NDWI_max > 0.3 ~ "wrong", 
                                    TRUE ~ NA_character_),
           valid_1_count = rowSums(across(c(valid_1_NDWI), ~ . == "wrong"), 
                                   na.rm = TRUE),
           valid_1 = if_else(valid_1_count > 0,
                             "At least 1 rule broken", "No rules broken so far")
           ) %>%
    # Keep only rows with no rules broken
    dplyr::filter(valid_1 == "No rules broken so far") %>%
    # Remove unneeded cols
    select(-starts_with("valid")) %>%
    # Select only GPS points
    dplyr::filter(`Location method` == "Location with GPS" | 
                    `Location method` == "Location with differential GPS") %>%
    # Filter based on NDVI percentiles
    filter_by_group_percentiles("NDVI_pos_value")
  }
```

### Function for NDVI-based filtering

```{r}
# data: your tibble/data.frame
# y_var: name of the numeric column as a string, e.g., "NDVI_pos_value"
filter_by_group_percentiles <- function(data, y_var) {
  cats <- c("T","Q","R","S")

  # Compute thresholds per group for the selected column
  qs <- data %>%
    dplyr::group_by(EUNIS1) %>%
    dplyr::summarise(
      p10 = stats::quantile(.data[[y_var]], probs = 0.10, na.rm = TRUE),
      p90 = stats::quantile(.data[[y_var]], probs = 0.90, na.rm = TRUE),
      .groups = "drop"
    )

  # Join thresholds and filter rows according to the rule
  d_filtered <- data %>%
    dplyr::left_join(qs, by = "EUNIS1") %>%
    dplyr::filter(
      dplyr::case_when(
        EUNIS1 == "T" ~ .data[[y_var]] >= p10,        # keep >= 10% for T
        EUNIS1 %in% c("Q","R","S") ~ .data[[y_var]] <= p90,  # keep <= 90% for Q/R/S
        TRUE ~ TRUE
      )
    ) %>%
    dplyr::select(-p10, -p90)  # drop helper columns

  return(d_filtered)
}
```

## Get full datasets and training subsets

```{r}
full_dataset_slope <- initial_dataset %>%
  get_full_datasets(vars_RF_slope_Landsat)
full_dataset_months <- initial_dataset %>%
  get_full_datasets(vars_RF_months_Landsat)  
training_subset_slope <- initial_dataset %>%
  get_training_subsets(vars_RF_slope_Landsat)
training_subset_months <- initial_dataset %>%
  get_training_subsets(vars_RF_months_Landsat)
```

## N obs

```{r}
nrow(full_dataset_slope)
nrow(full_dataset_months)
nrow(training_subset_slope)
nrow(training_subset_months)
```

# Random forest models with k-fold cross-validation 

Using 4 folds.

## Functions

### Function to run CV

```{r}
run_fastcforest_cv <- function(
    data, vars_RF, response_var,
    ntree = 500, k = 4
) {
  library(party)
  library(moreparty)
  library(doParallel)
  library(foreach)
  library(parallel)
  library(caret)

  # Create folds
  set.seed(123)
  folds <- createFolds(data[[response_var]], k = k)

  # Parallel cluster for CV
  ncores <- max(1, detectCores() - 1)
  cl <- makeCluster(ncores)
  registerDoParallel(cl)

  # Export needed objects
  clusterExport(cl, c("folds", "data", "vars_RF", "response_var"), envir = environment())

  # Load required packages inside workers
  clusterEvalQ(cl, {
    library(party)
    library(moreparty)
  })

  results <- list()

  for (i in seq_along(folds)) {

    cat("\n========== Starting fold", i, "of", k, "==========\n")
    t_start <- Sys.time()

    fold_result <- foreach(j = i, .combine = rbind,
                           .packages = c("party", "moreparty")) %dopar% {

      test_idx  <- folds[[j]]
      train_idx <- setdiff(seq_len(nrow(data)), test_idx)

      train_data <- data[train_idx, ]
      test_data  <- data[test_idx, ]

      # Build formula
      fmla <- reformulate(vars_RF, response = response_var)

      # ✅ Fast parallel conditional forest from moreparty
      model <- moreparty::fastcforest(
        formula  = fmla,
        data     = train_data,
        controls = party::cforest_unbiased(
          mtry  = round(sqrt(length(vars_RF))),
          ntree = ntree
        ),
        parallel = TRUE
      )

      pred <- predict(model, newdata = test_data, type = "response")

      data.frame(
        obs  = test_data[[response_var]],
        pred = pred,
        fold = j
      )
    }

    t_end <- Sys.time()
    cat(
      "✅ Finished fold", i, "in",
      round(difftime(t_end, t_start, units = "mins"), 2),
      "minutes\n"
    )

    results[[i]] <- fold_result
  }

  stopCluster(cl)

  do.call(rbind, results)
}
```

### Function to compute results from CV

```{r}
compute_rf_cv_metrics <- function(cv_results) {
  library(caret)
  library(pROC)
  library(dplyr)
  
  # Ensure pred is factor with same levels as obs
  cv_results$obs  <- factor(cv_results$obs)
  cv_results$pred <- factor(cv_results$pred, levels = levels(cv_results$obs))
  
  classes <- levels(cv_results$obs)
  folds   <- sort(unique(cv_results$fold))
  
  ## ---------------------------
  ## 1. PER-FOLD METRICS
  ## ---------------------------
  fold_metrics <- lapply(folds, function(f) {
    df <- cv_results %>% filter(fold == f)
    
    cm <- confusionMatrix(df$pred, df$obs)
    
    list(
      fold = f,
      accuracy = cm$overall["Accuracy"],
      kappa    = cm$overall["Kappa"]
    )
  })
  
  fold_metrics_df <- bind_rows(lapply(fold_metrics, as.data.frame))
  
  ## ---------------------------
  ## 2. OVERALL CONFUSION MATRIX
  ## ---------------------------
  overall_cm <- confusionMatrix(cv_results$pred, cv_results$obs)
  
  ## ---------------------------
  ## 3. CLASS-WISE METRICS
  ## ---------------------------
  class_metrics <- data.frame(
    class = classes,
    users_accuracy     = overall_cm$byClass[, "Pos Pred Value"],
    producers_accuracy = overall_cm$byClass[, "Sensitivity"],
    balanced_accuracy  = overall_cm$byClass[, "Balanced Accuracy"]
  )
  
  ## ---------------------------
  ## 4. FINAL OUTPUT STRUCTURE
  ## ---------------------------
  list(
    overall_confusion_matrix = overall_cm,
    
    accuracy_mean_sd = c(
      mean = mean(fold_metrics_df$accuracy),
      sd   = sd(fold_metrics_df$accuracy)
    ),
    
    kappa_mean_sd = c(
      mean = mean(fold_metrics_df$kappa),
      sd   = sd(fold_metrics_df$kappa)
    ),
    
    per_fold_metrics = fold_metrics_df,
    
    per_class_metrics = class_metrics
  )
}
```

## Function for fitting RF models on full dataset

```{r}
run_rf_parallel <- function(vars_RF, train_data, response_var, ntree = 500) {
  # Build formula locally (no NSE)
  fmla <- reformulate(vars_RF, response = response_var)

  # Start cluster
  n_cores <- max(1, parallel::detectCores() - 1)
  cl <- parallel::makeCluster(n_cores)
  doParallel::registerDoParallel(cl)

  # Make sure packages are available on workers
  parallel::clusterEvalQ(cl, {
    library(party)
  })

  # Export the *data* to workers under a simple name
  local_data <- train_data
  parallel::clusterExport(cl, varlist = c("local_data"), envir = environment())

  # Fit in parallel
  execution_time <- system.time({
    rf_model <- fastcforest(
      formula  = fmla,
      data     = local_data,   # use the exported name
      controls = party::cforest_control(
        mtry  = max(1, round(sqrt(length(vars_RF)))),
        ntree = ntree,
        # Decision trees were developed using subsampling without replacement,
        # which is appropriate to use when predictors vary in their scale of
        # measurement (Strobl et al., 2007).
        replace = FALSE
      ),
      parallel = TRUE
    )
  })

  parallel::stopCluster(cl)

  list(model = rf_model, time = execution_time)
}
```

## Without refinement

### Landsat, slope

#### CV

```{r eval=FALSE, include=FALSE}
cv_slope_Landsat <- run_fastcforest_cv(
  data = training_subset_slope,
  vars_RF = vars_RF_slope_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_slope_Landsat,
     file = "objects/RF_1500_cv_20260609/cv_slope_Landsat.Rdata") 
```

##### Results

```{r}
cv_slope_Landsat_results <-
  compute_rf_cv_metrics(cv_slope_Landsat)
cv_slope_Landsat_results
```

#### Fit model on full dataset

```{r eval=FALSE, include=FALSE}
RF_slope_Landsat <- run_rf_parallel(vars_RF_slope_Landsat, 
                                    training_subset_slope, 
                                    "EUNIS1", ntree = 1500)
```

```{r eval=FALSE, include=FALSE}
save(RF_slope_Landsat,
     file = "objects/RF_1500_cv_20260609/RF_slope_Landsat.Rdata") 
```

### Landsat, months

#### CV

```{r eval=FALSE, include=FALSE}
cv_months_Landsat <- run_fastcforest_cv(
  data = training_subset_months,
  vars_RF = vars_RF_months_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_months_Landsat,
     file = "objects/RF_1500_cv_20260609/cv_months_Landsat.Rdata") 
```

##### Results

```{r}
cv_months_Landsat_results <-
  compute_rf_cv_metrics(cv_months_Landsat)
cv_months_Landsat_results
```

#### Fit model on full dataset

```{r eval=FALSE, include=FALSE}
RF_months_Landsat <- run_rf_parallel(vars_RF_months_Landsat, 
                                    training_subset_slope, 
                                    "EUNIS1", ntree = 1500)
```

```{r eval=FALSE, include=FALSE}
save(RF_months_Landsat,
     file = "objects/RF_1500_cv_20260609/RF_months_Landsat.Rdata") 
```

## Refinement step 1

### Compute predictions for all data and combine with original data

```{r eval=FALSE, include=FALSE}
# Slope
training_subset_slope_with_probs <- bind_cols(
  training_subset_slope %>% select(PlotObservationID, EUNIS1),
  as.data.frame(predict(RF_slope_Landsat$model, 
                        # Predicting in all data
                                          newdata = training_subset_slope,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(RF_slope_Landsat$model, 
                         # Predicting in all
                         newdata = training_subset_slope,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNIS1, prob_Q = EUNIS1.Q,
         prob_R = EUNIS1.R, prob_S = EUNIS1.S, prob_T = EUNIS1.T)

# Months
training_subset_months_with_probs <- bind_cols(
  training_subset_months %>% select(PlotObservationID, EUNIS1),
  as.data.frame(predict(RF_months_Landsat$model, 
                        # Predicting in all data
                                          newdata = training_subset_months,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(RF_months_Landsat$model, 
                         # Predicting in all data (training + testing)
                         newdata = training_subset_months,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNIS1, prob_Q = EUNIS1.Q,
         prob_R = EUNIS1.R, prob_S = EUNIS1.S, prob_T = EUNIS1.T)
```

### Detect probability columns

```{r eval=FALSE, include=FALSE}
prob_cols_slope_Landsat <- names(training_subset_slope_with_probs)[grepl("^prob_", names(training_subset_slope_with_probs))]
prob_cols_months_Landsat <- names(training_subset_months_with_probs)[grepl("^prob_", names(training_subset_months_with_probs))]
```

### Calculate auxiliary metrics

```{r}
calculate_df_aux <- function(data, prob_cols) {
  data %>%
    rowwise() %>%
    mutate(
      # Probability of reference (original) class
      p_ref = get(paste0("prob_", original_EUNIS)),
      # Maximum probability between all classes
      max_prob = max(c_across(all_of(prob_cols)), na.rm = TRUE),
      # Second maximum probability
      second_max = sort(c_across(all_of(prob_cols)), decreasing = TRUE)[2],
      # Difference between original class and second best
      delta = p_ref - second_max,
      # TRUE if original class is not the most probable
      neg_flag = p_ref < max_prob,
      # TRUE if difference is small (uncertainty)
      small_flag = delta < 0.1,
      # How much worse is the original class with respect to the best class
      gap_neg = max_prob - p_ref
    ) %>%
    ungroup()
}
```

```{r eval=FALSE, include=FALSE}
df_aux_slope_Landsat <- calculate_df_aux(training_subset_slope_with_probs,
                                     prob_cols_slope_Landsat)
df_aux_months_Landsat <- calculate_df_aux(training_subset_months_with_probs,
                                     prob_cols_months_Landsat)
```

### Apply filtering function

Removing negatives until 20% of the class and, if there is room, uncertain positives until 10% of the rest.

```{r}
filter_probs_neg20_pos10 <- function(gdf) {
  # Removes negatives until 20% of the class
  cap_total <- ceiling(0.20 * nrow(gdf))  # max 20% of the class

  neg <- gdf %>% dplyr::filter(neg_flag) %>% arrange(desc(gap_neg))
  drop_neg <- head(neg, cap_total)
  cap_left <- cap_total - nrow(drop_neg)

  drop_pos <- tibble()
  if (cap_left > 0) {
    remaining <- nrow(gdf) - nrow(drop_neg)
    # If there is room, removes positives with doubts (small_flag = TRUE)
    # until 10% of the rest
    pos_cap <- ceiling(0.10 * remaining)

    drop_pos <- gdf %>%
      dplyr::filter(!neg_flag & small_flag) %>%
      arrange(delta) %>%
      slice_head(n = min(cap_left, pos_cap))
  }

  drop_ids <- c(drop_neg$PlotObservationID, drop_pos$PlotObservationID)
  gdf %>% dplyr::filter(!PlotObservationID %in% drop_ids) %>%
    select(-max_prob, -second_max, -delta, -neg_flag, -small_flag, -gap_neg)
}
```

```{r eval=FALSE, include=FALSE}
df_final_20_10_slope_Landsat <- df_aux_slope_Landsat %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20_pos10)
df_final_20_10_months_Landsat <- df_aux_months_Landsat %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20_pos10)
```

```{r eval=FALSE, include=FALSE}
filtered_data_20_10_slope_Landsat_probs <- training_subset_slope %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_slope_Landsat$PlotObservationID)
filtered_data_20_10_months_Landsat_probs <- training_subset_months %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_months_Landsat$PlotObservationID)
```

### Save

```{r}
filtered_data_with_probs <- c("filtered_data_20_10_slope_Landsat_probs", 
                              "filtered_data_20_10_months_Landsat_probs")
```

Save:

```{r eval=FALSE, include=FALSE}
for (filtered_data_name in filtered_data_with_probs) {
  file_path <- paste0("objects/RF_1500_cv_20260609/", filtered_data_name, ".Rdata")
  save(list = filtered_data_name, file = file_path)
}
```

### Fit models

#### Landsat, slope

##### CV

```{r eval=FALSE, include=FALSE}
cv_slope_Landsat_step1 <- run_fastcforest_cv(
  data = filtered_data_20_10_slope_Landsat_probs,
  vars_RF = vars_RF_slope_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_slope_Landsat_step1, 
     file = "objects/RF_1500_cv_20260609/cv_slope_Landsat_step1.Rdata") 
```

##### Results

```{r}
cv_slope_Landsat_step1_results <-
  compute_rf_cv_metrics(cv_slope_Landsat_step1)
cv_slope_Landsat_step1_results
```

##### Fit model on full dataset

```{r eval=FALSE, include=FALSE}
RF_slope_Landsat_step1 <- run_rf_parallel(vars_RF_slope_Landsat, 
                                          filtered_data_20_10_slope_Landsat_probs, 
                                          "EUNIS1", ntree = 1500)
```

```{r eval=FALSE, include=FALSE}
save(RF_slope_Landsat_step1,
     file = "objects/RF_1500_cv_20260609/RF_slope_Landsat_step1.Rdata") 
```

#### Landsat, months

##### CV

```{r eval=FALSE, include=FALSE}
cv_months_Landsat_step1 <- run_fastcforest_cv(
  data = filtered_data_20_10_months_Landsat_probs,
  vars_RF = vars_RF_months_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_months_Landsat_step1, 
     file = "objects/RF_1500_cv_20260609/cv_months_Landsat_step1.Rdata") 
```

##### Results

```{r}
cv_months_Landsat_step1_results <-
  compute_rf_cv_metrics(cv_months_Landsat_step1)
cv_months_Landsat_step1_results
```

##### Fit model on full dataset

```{r eval=FALSE, include=FALSE}
RF_months_Landsat_step1 <- run_rf_parallel(vars_RF_months_Landsat, 
                                          filtered_data_20_10_months_Landsat_probs, 
                                          "EUNIS1", ntree = 1500)
```

```{r eval=FALSE, include=FALSE}
save(RF_months_Landsat_step1,
     file = "objects/RF_1500_cv_20260609/RF_months_Landsat_step1.Rdata") 
```

## Refinement step 2

#### Compute predictions for all data and combine with original data

```{r eval=FALSE, include=FALSE}
# Landsat, slope
filtered_data_slope_Landsat_with_probs_2x <- bind_cols(
  filtered_data_20_10_slope_Landsat_probs %>% select(PlotObservationID, EUNIS1),
  as.data.frame(predict(RF_slope_Landsat_step1$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_20_10_slope_Landsat_probs,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(RF_slope_Landsat_step1$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_20_10_slope_Landsat_probs,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNIS1, prob_Q = EUNIS1.Q,
         prob_R = EUNIS1.R, prob_S = EUNIS1.S, prob_T = EUNIS1.T)

# Landsat, months
filtered_data_months_Landsat_with_probs_2x <- bind_cols(
  filtered_data_20_10_months_Landsat_probs %>% select(PlotObservationID, EUNIS1),
  as.data.frame(predict(RF_months_Landsat_step1$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_20_10_months_Landsat_probs,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(RF_months_Landsat_step1$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_20_10_months_Landsat_probs,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNIS1, prob_Q = EUNIS1.Q,
         prob_R = EUNIS1.R, prob_S = EUNIS1.S, prob_T = EUNIS1.T)
```

### Detect probability columns

```{r eval=FALSE, include=FALSE}
prob_cols_slope_Landsat <- names(filtered_data_slope_Landsat_with_probs_2x)[grepl("^prob_", names(filtered_data_slope_Landsat_with_probs_2x))]
prob_cols_months_Landsat <- names(filtered_data_months_Landsat_with_probs_2x)[grepl("^prob_", names(filtered_data_months_Landsat_with_probs_2x))]
```

#### Calculate auxiliary metrics

```{r eval=FALSE, include=FALSE}
df_aux_slope_Landsat_2x <- calculate_df_aux(filtered_data_slope_Landsat_with_probs_2x,
                                     prob_cols_slope_Landsat)
df_aux_months_Landsat_2x <- calculate_df_aux(filtered_data_months_Landsat_with_probs_2x,
                                     prob_cols_months_Landsat)
```

#### Apply filtering function

```{r}
filter_probs_neg20 <- function(gdf) {
  # Removes negatives until 20% of the class
  cap_total <- ceiling(0.20 * nrow(gdf))  # max 20% of the class

  neg <- gdf %>% filter(neg_flag) %>% arrange(desc(gap_neg))
  drop_neg <- head(neg, cap_total)
  cap_left <- cap_total - nrow(drop_neg)

  drop_ids <- c(drop_neg$PlotObservationID)
  gdf %>% filter(!PlotObservationID %in% drop_ids) %>%
    select(-max_prob, -second_max, -delta, -neg_flag, -small_flag, -gap_neg)
}
```

```{r eval=FALSE, include=FALSE}
df_final_20_10_slope_Landsat_2x <- df_aux_slope_Landsat_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)
df_final_20_10_months_Landsat_2x <- df_aux_months_Landsat_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)
```

```{r eval=FALSE, include=FALSE}
filtered_data_20_10_slope_Landsat_probs_2x <- filtered_data_20_10_slope_Landsat_probs %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_slope_Landsat_2x$PlotObservationID)
filtered_data_20_10_months_Landsat_probs_2x <- filtered_data_20_10_months_Landsat_probs %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_months_Landsat_2x$PlotObservationID)
```

### Save

```{r}
filtered_data_with_probs_2x <- c("filtered_data_20_10_slope_Landsat_probs_2x", 
                              "filtered_data_20_10_months_Landsat_probs_2x")
```

Save:

```{r eval=FALSE, include=FALSE}
for (filtered_data_name in filtered_data_with_probs_2x) {
  file_path <- paste0("objects/RF_1500_cv_20260609/", filtered_data_name, ".Rdata")
  save(list = filtered_data_name, file = file_path)
}
```

### Fit models

#### Landsat, slope

##### CV

```{r eval=FALSE, include=FALSE}
cv_slope_Landsat_step2 <- run_fastcforest_cv(
  data = filtered_data_20_10_slope_Landsat_probs_2x,
  vars_RF = vars_RF_slope_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_slope_Landsat_step2, 
     file = "objects/RF_1500_cv_20260609/cv_slope_Landsat_step2.Rdata") 
```

##### Results

```{r}
cv_slope_Landsat_step2_results <-
  compute_rf_cv_metrics(cv_slope_Landsat_step2)
cv_slope_Landsat_step2_results
```

##### Fit model on full dataset

```{r eval=FALSE, include=FALSE}
RF_slope_Landsat_step2 <- run_rf_parallel(vars_RF_slope_Landsat, 
                                          filtered_data_20_10_slope_Landsat_probs_2x, 
                                          "EUNIS1", ntree = 1500)
```

```{r eval=FALSE, include=FALSE}
save(RF_slope_Landsat_step2,
     file = "objects/RF_1500_cv_20260609/RF_slope_Landsat_step2.Rdata") 
```

#### Landsat, months --> BEST MODEL

##### CV

```{r eval=FALSE, include=FALSE}
cv_months_Landsat_step2 <- run_fastcforest_cv(
  data = filtered_data_20_10_months_Landsat_probs_2x,
  vars_RF = vars_RF_months_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_months_Landsat_step2, file = "objects/RF_1500_cv_20260609/cv_months_Landsat_step2.Rdata") 
```

##### (Table 2) Results

```{r}
cv_months_Landsat_step2_results <-
  compute_rf_cv_metrics(cv_months_Landsat_step2)
cv_months_Landsat_step2_results
```

##### Fit model on full dataset

```{r eval=FALSE, include=FALSE}
RF_months_Landsat_step2 <- run_rf_parallel(vars_RF_months_Landsat, 
                                          filtered_data_20_10_months_Landsat_probs_2x, 
                                          "EUNIS1", ntree = 1500)
```

```{r eval=FALSE, include=FALSE}
save(RF_months_Landsat_step2,
     file = "objects/RF_1500_cv_20260609/RF_months_Landsat_step2.Rdata") 
```

# Variable importance from best model

```{r}
compute_varimp <- function(model, nperm = 100) {

  # Measure execution time
  execution_time <- system.time({
    varimp_result <- permimp(model, conditional = FALSE, progressBar = TRUE)
  })

  return(list(varimp = varimp_result, time = execution_time))
}
```

```{r eval=FALSE, include=FALSE}
varimp_months_Landsat_step2 <-
  compute_varimp(RF_months_Landsat_step2$model, nperm = 100)
```

```{r eval=FALSE, include=FALSE}
save(varimp_months_Landsat_step2,
     file = "objects/RF_1500_cv_20260609/varimp_months_Landsat_step2.Rdata")
```

# ROC curves from best model

```{r}
compute_roc_level1 <- function(model, test_data) {
  # Measure execution time
  execution_time <- system.time({
    # Step 1: Predict probabilities
    probabilities <- predict(model, newdata = test_data, type = "prob")
    
    # Step 2: Convert list of matrices to a proper data frame
    prob_matrix <- t(sapply(probabilities, as.vector))
    prob_df <- as.data.frame(prob_matrix)
    colnames(prob_df) <- c("Q", "R", "S", "T")
    
    # Step 3: Prepare actual class labels
    actual <- factor(test_data$EUNIS1, levels = c("Q", "R", "S", "T"))
    
    # Step 4: Binarize actual labels
    actual_bin <- model.matrix(~ actual - 1)
    colnames(actual_bin) <- gsub("actual", "", colnames(actual_bin))
    
    # Step 5: Compute ROC data for each class
    roc_data <- lapply(levels(actual), function(class) {
      roc_obj <- roc(actual_bin[, class], prob_df[[class]])
      auc_val <- round(auc(roc_obj), 3)
      data.frame(
        FPR = 1 - roc_obj$specificities,
        TPR = roc_obj$sensitivities,
        Class = paste0(class, " (AUC = ", auc_val, ")")
      )
    }) %>% bind_rows()
  })
  
  # Return both ROC data and execution time
  return(list(roc = roc_data, time = execution_time))
}
```

```{r eval=FALSE, include=FALSE}
rocdata_months_Landsat_step2 <-
  compute_roc_level1(RF_months_Landsat_step2$model,
                     filtered_data_20_10_months_Landsat_probs_2x)
```

```{r eval=FALSE, include=FALSE}
save(rocdata_months_Landsat_step2,
     file = "objects/RF_1500_cv_20260609/rocdata_months_Landsat_step2.Rdata")
```

# Validation

## Get predictions from best models 

Predict in full dataset.

Model with months, Landsat for predicting EVERYWHERE:

```{r eval=FALSE, include=FALSE}
prediction_Landsat_resp <- predict(RF_months_Landsat_step2$model,
                                   newdata = full_dataset_months, 
                                   type = "response")
```

```{r eval=FALSE, include=FALSE}
prediction_Landsat_prob <- predict(RF_months_Landsat_step2$model,
                                   newdata = full_dataset_months, 
                                   type = "prob")
```

```{r}
predictions <- c("prediction_Landsat_resp", "prediction_Landsat_prob")
```

Save:

```{r eval=FALSE, include=FALSE}
for (prediction_name in predictions) {
  file_path <- paste0("objects/validated_points_20260630/", prediction_name, ".Rdata")
  save(list = prediction_name, file = file_path)
}
```

Add predictions:

```{r}
data_validated <- initial_dataset %>%
  left_join(
    # Predictions from model with months, Landsat
    cbind(
      data.frame(
        # PlotObservationIDs
        PlotObservationID = full_dataset_months %>% 
          pull(PlotObservationID),
        # Prediction response
        pred_Landsat_resp = prediction_Landsat_resp
        ),
      # Prediction probabilities
      do.call(rbind, prediction_Landsat_prob) %>% as.data.frame() %>%
        rename(pred_Landsat_Q = EUNIS1.Q, pred_Landsat_R = EUNIS1.R,
               pred_Landsat_S = EUNIS1.S, pred_Landsat_T = EUNIS1.T)
      )
  ) %>%
  # Validated or not (Landsat)
  mutate(valid_Landsat = ifelse(pred_Landsat_resp == EUNIS1, TRUE, 
                                ifelse(is.na(pred_Landsat_resp), NA, FALSE)))
```

## Data preparation

```{r}
data_validated_clean <- data_validated %>%
  # Keep only needed cols
  select(PlotObservationID, plotID, year, EUNIS1, Lon_updated, Lat_updated, 
         pred_Landsat_resp, pred_Landsat_Q, pred_Landsat_R, pred_Landsat_S, 
         pred_Landsat_T, valid_Landsat) %>%
  # Add info on nr resurveys (n different years)
  # Some have only 1 year, either because the first year(s) were prior to 
  # Landsat, or because they were still not resurveyed. 
  group_by(plotID) %>%
  mutate(
    n_resurveys = n_distinct(year, na.rm = TRUE),   # count unique years
    first_year  = min(year, na.rm = TRUE),          # earliest year
    last_year   = max(year, na.rm = TRUE)           # latest year
  ) %>%
  ungroup()
```

## Write csv

```{r}
write_csv(data_validated_clean,
          here("data", "clean","data_validated_clean_20260702.csv"))
```

## Counts of validated / not-validated obs

### General

#### From all obs

```{r}
data_validated_clean %>%
  count(valid_Landsat) %>%
  mutate(percent = n / sum(n) * 100)
# Here, NAs are those obs not in the full dataset
```

#### From full dataset

```{r}
data_validated_clean %>%
  filter(!is.na(valid_Landsat)) %>% # This keeps only those obs in the full dataset!
  count(valid_Landsat) %>%
  mutate(percent = n / sum(n) * 100)
```

### By habitat, from full dataset

```{r}
data_validated_clean %>%
  filter(EUNIS1 == "Q") %>%
  filter(!is.na(valid_Landsat)) %>%
  count(valid_Landsat) %>%
  mutate(percent = n / sum(n) * 100)
```

```{r}
data_validated_clean %>%
  filter(EUNIS1 == "R") %>%
  filter(!is.na(valid_Landsat)) %>%
  count(valid_Landsat) %>%
  mutate(percent = n / sum(n) * 100)
```

```{r}
data_validated_clean %>%
  filter(EUNIS1 == "S") %>%
  filter(!is.na(valid_Landsat)) %>%
  count(valid_Landsat) %>%
  mutate(percent = n / sum(n) * 100)
```

```{r}
data_validated_clean %>%
  filter(EUNIS1 == "T") %>%
  filter(!is.na(valid_Landsat)) %>%
  count(valid_Landsat) %>%
  mutate(percent = n / sum(n) * 100)
```

## Validation of plots

Check that only one EUNIS1 is assigned to each plot and year:

```{r}
data_validated_clean %>%
  group_by(plotID, n_resurveys, year) %>%
  reframe(
    EUNISa_vals = unique(EUNIS1)) %>%
  unnest(EUNISa_vals) %>%
  group_by(plotID, n_resurveys, year) %>%
  mutate(EUNIS_idx = row_number()) %>%
  pivot_wider(
    names_from = EUNIS_idx,
    values_from = EUNISa_vals,
    names_prefix = "EUNIS1_"
  ) %>%
  ungroup() %>%
  # Number of different EUNIS1 assigned
  mutate(n_EUNIS1 = rowSums(!is.na(select(., starts_with("EUNIS1_"))))) %>%
  filter(n_EUNIS1 > 1) %>%
  nrow()
```

Extract EUNIS assignations for each plot (sometimes different among years, i.e. different observations of the same plot):

```{r}
eunis_wide <- data_validated_clean %>%
  group_by(plotID, n_resurveys) %>%
  reframe(
    EUNISa_vals = unique(EUNIS1)) %>%
  unnest(EUNISa_vals) %>%
  group_by(plotID, n_resurveys) %>%
  mutate(EUNIS_idx = row_number()) %>%
  pivot_wider(
    names_from = EUNIS_idx,
    values_from = EUNISa_vals,
    names_prefix = "EUNIS1_"
  ) %>%
  ungroup() %>%
  # Number of different EUNIS1 assigned
  mutate(n_EUNIS1 = rowSums(!is.na(select(., starts_with("EUNIS1_"))))) %>%
  # Categorize as "Habitat change" if there was more than one EUNIS assigned
  # to different observations within a plot
  mutate(EUNIS = case_when(
    is.na(EUNIS1_2) & is.na(EUNIS1_3) & is.na(EUNIS1_4) ~ EUNIS1_1,
    TRUE ~ "Change"
    )) %>%
  ungroup() %>%
  # Remove unneeded cols
  select(-EUNIS1_1, -EUNIS1_2, -EUNIS1_3, -EUNIS1_4)
```

Check that the number of EUNIS assignations is never larger than the number of resurveys:

```{r}
eunis_wide %>% filter(n_EUNIS1 > n_resurveys) %>% nrow()
```

Extract the latest coordinate pair per plot (sometimes coordinates are different among different observations of the same plot):

```{r}
coords_latest <- data_validated_clean %>%
  group_by(plotID, n_resurveys) %>%
  slice_max(order_by = year, n = 1, with_ties = FALSE) %>%
  select(
    plotID, n_resurveys, Lon_updated, Lat_updated
  ) %>%
  ungroup()
```

```{r}
PlotObservationIDs_full_dataset <- full_dataset_months %>% pull(PlotObservationID)
data_validated_clean <- data_validated_clean %>%
  mutate(in_full_dataset = ifelse(
    PlotObservationID %in% PlotObservationIDs_full_dataset, TRUE, FALSE))
```

```{r}
data_validated_plots <- data_validated_clean %>%
  left_join(
    eunis_wide,
    by = c("plotID", "n_resurveys")
  ) %>%
  group_by(plotID, n_resurveys, EUNIS) %>%
  summarise(
    # Total number of observations for that plot
    n_obs = n(), 
    # Number of validated observations, not validated and NA
    n_val = sum(valid_Landsat == TRUE, na.rm = TRUE),
    n_notval = sum(valid_Landsat == FALSE, na.rm = TRUE),
    n_na = sum(is.na(valid_Landsat)),
    # N obs in full dataset
    n_in_full = sum(in_full_dataset),
    .groups = "drop"
  ) %>%
  mutate(
    # Calculate shares (proportions) of validated and not validated observations 
    # (only if there are any non-NA observations, otherwise set as NA)
    share_val = ifelse((n_val + n_notval) > 0,
                        n_val / (n_val + n_notval), NA_real_),
    share_notval = ifelse((n_val + n_notval) > 0,
                         n_notval / (n_val + n_notval), NA_real_),
    # Calculate proportion of NA observations
    prop_na = n_na/n_obs,
    # Calculate proportion of obs in full dataset
    prop_in_full = n_in_full / n_obs
  )
```

```{r}
cat("\nPlots with no observations in full dataset",
    data_validated_plots %>% filter(prop_in_full == 0) %>% nrow()) # --> Plot cannot be validated - NA
cat("\nPlots with all observations in full dataset",
    data_validated_plots %>% filter(prop_in_full == 1) %>% nrow()) # --> Plot can be validated
cat("\nPlots with some observations in full dataset and some not",
    data_validated_plots %>% filter(prop_in_full > 0 & prop_in_full < 1) %>% nrow()) # --> Plot cannot be validated - NA
```

```{r}
data_validated_plots <- data_validated_plots %>%
  mutate(valid_status = case_when(
    prop_in_full == 0 ~ NA,                           # Plots with no observations in full dataset
    prop_in_full > 0 & prop_in_full < 1 ~ NA,         # Plots with some observations in full dataset and some not
    prop_in_full == 1 & share_val == 1 ~ "valid",     # Plots with all observations in full dataset
    prop_in_full == 1 & share_val != 1 ~ "not_valid"  # Plots with all observations in full dataset
  )) %>%
  # Add coordinates (if coordinates are different among different observations 
  # of the same plot, extract the latest coordinate pair per plot, i.e. from
  # the last observation year)
  left_join(coords_latest, by = c("plotID", "n_resurveys"))
```

## Counts of validated / not-validated plots

### General

#### All plots

```{r}
data_validated_plots %>%
  count(valid_status) %>%
  mutate(percent = n / sum(n) * 100)
```

#### Plots with all obs in full dataset

```{r}
data_validated_plots %>%
  filter(!is.na(valid_status)) %>%
  count(valid_status) %>%
  mutate(percent = n / sum(n) * 100)
```

### By habitat, plots with all obs in full dataset

```{r}
data_validated_plots %>%
  filter(EUNIS == "Q") %>%
  filter(!is.na(valid_status)) %>%
  count(valid_status) %>%
  mutate(percent = n / sum(n) * 100)
```

```{r}
data_validated_plots %>%
  filter(EUNIS == "R") %>%
  filter(!is.na(valid_status)) %>%
  count(valid_status) %>%
  mutate(percent = n / sum(n) * 100)
```

```{r}
data_validated_plots %>%
  filter(EUNIS == "S") %>%
  filter(!is.na(valid_status)) %>%
  count(valid_status) %>%
  mutate(percent = n / sum(n) * 100)
```

```{r}
data_validated_plots %>%
  filter(EUNIS == "T") %>%
  filter(!is.na(valid_status)) %>%
  count(valid_status) %>%
  mutate(percent = n / sum(n) * 100)
```

```{r}
data_validated_plots %>%
  filter(EUNIS == "Change") %>%
  filter(!is.na(valid_status)) %>%
  count(valid_status) %>%
  mutate(percent = n / sum(n) * 100)
```

## Write csv of observations in validated plots

This csv contains only information about observations in validated plots (i.e. plots with all observations validated).

```{r}
data_obs_in_val_plots <- data_validated_clean %>% 
  # Filter validated plots
  filter(plotID %in% 
           c(data_validated_plots %>% 
           filter(valid_status == "valid") %>% 
           pull(plotID))) %>%
  # For each plotID, create a column plotID_num (numeric identifier)
  group_by(plotID) %>%
  mutate(plotID_num = cur_group_id()) %>%
  ungroup()
```

```{r}
write_csv(data_obs_in_val_plots,
          here("data", "clean","data_obs_in_val_plots_20260702.csv"))
```

# Figures

## (Fig 3) Plots most influential variables

```{r}
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

```{r}
distr_plot <- function(data,
                       y_vars,
                       y_labels,
                       grey = TRUE,
                       p_low = 0.10,
                       p_high = 0.90,
                       low_groups = c("T"),
                       high_groups = c("Q", "R", "S"),
                       grey_color = "grey60",
                       print_plots = FALSE) {
  stopifnot(length(y_vars) == length(y_labels))

  # Ensure y_vars/y_labels can be length-1 strings
  if (!is.vector(y_vars))   y_vars   <- c(y_vars)
  if (!is.vector(y_labels)) y_labels <- c(y_labels)

  library(ggplot2)
  library(dplyr)
  library(rlang)
  library(stringr)

  base_data <- data %>%
    dplyr::filter(EUNIS1 %in% union(low_groups, high_groups))

  plots <- vector("list", length(y_vars))

  for (i in seq_along(y_vars)) {
    y_var   <- y_vars[[i]]
    y_label <- y_labels[[i]]

    if (grey) {
      df_flag <- base_data %>%
        group_by(EUNIS1) %>%
        mutate(
          q_low  = quantile(!!sym(y_var), p_low,  na.rm = TRUE),
          q_high = quantile(!!sym(y_var), p_high, na.rm = TRUE),
          grey_flag = case_when(
            EUNIS1 %in% low_groups  ~ (!!sym(y_var)) <= q_low,
            EUNIS1 %in% high_groups ~ (!!sym(y_var)) >= q_high,
            TRUE ~ FALSE
          ),
          grey_flag = coalesce(grey_flag, FALSE)
        ) %>%
        ungroup()
    }

    p <- ggplot(
      data = base_data,
      aes(x = EUNIS1, y = !!sym(y_var), fill = EUNIS1)
    ) +
      # Violin
      geom_flat_violin(position = position_nudge(x = 0.2, y = 0), alpha = 0.8) +

      # Points (with or without greying)
      {
        if (grey) {
          list(
            geom_point(
              data = df_flag %>% dplyr::filter(!grey_flag),
              aes(y = !!sym(y_var), color = EUNIS1),
              position = position_jitter(width = 0.15),
              size = 0.05, alpha = 0.25
            ),
            geom_point(
              data = df_flag %>% dplyr::filter(grey_flag),
              aes(y = !!sym(y_var)),
              color = grey_color,
              position = position_jitter(width = 0.15),
              size = 0.05, alpha = 0.6
            )
          )
        } else {
          geom_point(
            aes(y = !!sym(y_var), color = EUNIS1),
            position = position_jitter(width = 0.15),
            size = 0.05, alpha = 0.25
          )
        }
      } +

      # BOX PLOT LAST (on top) — only change for rule #1
      geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.5) +

      labs(y = y_label, x = "EUNIS level 1") +
      scale_x_discrete(labels = function(x) str_wrap(x, width = 15)) +
      guides(fill = "none", color = "none") +
      theme_bw() + coord_flip()  +
      scale_color_manual(values = c(
        "#0072B2", # blue
        "#D55E00", # vermillion
        "#009E73", # bluish green
        "#CC79A7"  # reddish purple
        )) +
      scale_fill_manual(values = c(
        "#0072B2", # blue
        "#D55E00", # vermillion
        "#009E73", # bluish green
        "#CC79A7"  # reddish purple
        ))
    
    if (print_plots) print(p)
    plots[[i]] <- p
  }

  # Return a single ggplot if length == 1 (so you can use it with grid.arrange directly)
  if (length(plots) == 1L) {
    return(plots[[1L]])
  } else {
    return(plots)  # list of ggplots
  }
}
```

These plots include all observations in the full dataset.

```{r}
plot1_Landsat <- data_validated_clean %>%
  mutate(EUNIS1 = factor(EUNIS1, levels = c("Q", "R", "S", "T"))) %>%
  filter(!is.na(valid_Landsat)) %>%
  left_join(initial_dataset %>% 
              select(PlotObservationID, NDMI_max)) %>%
  distr_plot("NDMI_max", "Maximum NDMI", grey = F) +
  theme(plot.margin = margin(2, 2, 2, 2)) +
  scale_x_discrete(limits = c("T", "S", "R", "Q"))
plot2_Landsat <- data_validated_clean %>%
  mutate(EUNIS1 = factor(EUNIS1, levels = c("Q", "R", "S", "T"))) %>%
  filter(!is.na(valid_Landsat)) %>%
  left_join(initial_dataset %>% 
              select(PlotObservationID, SAVI_avg_value_03)) %>%
  distr_plot("SAVI_avg_value_03", 
             "SAVI value at greening", grey = F) +
  theme(axis.title.y = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank()) + 
  theme(plot.margin = margin(2, 2, 2, 2)) +
  scale_x_discrete(limits = c("T", "S", "R", "Q"))
plot3_Landsat <- data_validated_clean %>%
  mutate(EUNIS1 = factor(EUNIS1, levels = c("Q", "R", "S", "T"))) %>%
  filter(!is.na(valid_Landsat)) %>%
  left_join(initial_dataset %>% 
              select(PlotObservationID, NDWI_min)) %>%
  distr_plot("NDWI_min", 
             "Minimum NDWI", grey = F) + 
  theme(plot.margin = margin(2, 2, 2, 2)) +
  scale_x_discrete(limits = c("T", "S", "R", "Q"))
plot4_Landsat <- data_validated_clean %>%
  mutate(EUNIS1 = factor(EUNIS1, levels = c("Q", "R", "S", "T"))) %>%
  filter(!is.na(valid_Landsat)) %>%
  left_join(initial_dataset %>% 
              select(PlotObservationID, SAVI_diff_pos_march_value)) %>%
  distr_plot("SAVI_diff_pos_march_value", 
             "SAVI difference in value peak - greening", 
             grey = F) +
  theme(plot.margin = margin(2, 2, 2, 2)) +
  theme(axis.title.y = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank()) + 
  scale_x_discrete(limits = c("T", "S", "R", "Q"))
plot5_Landsat <- data_validated_clean %>%
  mutate(EUNIS1 = factor(EUNIS1, levels = c("Q", "R", "S", "T"))) %>%
  filter(!is.na(valid_Landsat)) %>%
  left_join(initial_dataset %>% 
              select(PlotObservationID, NDVI_diff_pos_march_value)) %>%
  distr_plot("NDVI_diff_pos_march_value", 
             "NDVI difference in value peak - greening", 
             grey = F) +
  theme(plot.margin = margin(2, 2, 2, 2)) +
  scale_x_discrete(limits = c("T", "S", "R", "Q"))
plot6_Landsat <- data_validated_clean %>%
  mutate(EUNIS1 = factor(EUNIS1, levels = c("Q", "R", "S", "T"))) %>%
  filter(!is.na(valid_Landsat)) %>%
  left_join(initial_dataset %>% 
              select(PlotObservationID, NDVI_auc_mar_oct)) %>%
  distr_plot("NDVI_auc_mar_oct", 
             "NDVI AUC greening-browning", grey = F) +
  theme(plot.margin = margin(2, 2, 2, 2)) +
  theme(axis.title.y = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank()) + 
  scale_x_discrete(limits = c("T", "S", "R", "Q"))
plot7_Landsat <- data_validated_clean %>%
  mutate(EUNIS1 = factor(EUNIS1, levels = c("Q", "R", "S", "T"))) %>%
  filter(!is.na(valid_Landsat)) %>%
  left_join(initial_dataset %>% 
              select(PlotObservationID, SAVI_auc_mar_oct)) %>%
  distr_plot("SAVI_auc_mar_oct", 
             "SAVI AUC greening-browning", grey = F)  +
  theme(plot.margin = margin(2, 2, 2, 2)) +
  scale_x_discrete(limits = c("T", "S", "R", "Q"))

plot_for_legend <- data_validated_clean %>%
  mutate(EUNIS1 = factor(EUNIS1, levels = c("Q", "R", "S", "T"))) %>%
  filter(!is.na(valid_Landsat)) %>%
  left_join(initial_dataset %>% 
              select(PlotObservationID, NDMI_max)) %>%
  distr_plot("NDMI_max", "Maximum NDMI", grey = FALSE) +
  guides(fill = guide_legend(), color = guide_legend()) +
  labs(fill = "EUNIS level 1", color = "EUNIS level 1") +
  theme(legend.position = "right") +
  scale_fill_manual(values = c("#0072B2", "#D55E00", "#009E73", "#CC79A7"),
                    breaks = c("Q", "R", "S", "T"), 
                    labels = c("Q" = "Q - Wetlands","R" = "R - Grasslands", 
                               "S" = "S - Shrublands", "T" = "T - Forests")) +
  scale_color_manual(values = c("#0072B2", "#D55E00", "#009E73", "#CC79A7"),
                     breaks = c("Q", "R", "S", "T"),
                     labels = c("Q" = "Q - Wetlands","R" = "R - Grasslands", 
                               "S" = "S - Shrublands", "T" = "T - Forests")) +
  guides(fill = guide_legend(
    override.aes = list(shape = 22, size = 8, 
                        fill = c("#0072B2", "#D55E00", "#009E73", "#CC79A7"),
                        color = NA, linetype = 0, alpha = 1)),
    color = "none") +
  theme(legend.text = element_text(size = 12), 
        legend.title = element_text(size = 13)) +
  scale_x_discrete(limits = c("Q", "R", "S", "T"))
legend <- patchwork::wrap_elements(get_legend(plot_for_legend))
```

```{r}
fig3 <- (
    plot1_Landsat + theme(plot.margin = margin(2,2,2,2)) |
      plot2_Landsat + theme(plot.margin = margin(2,2,2,2))
    ) /
  (plot3_Landsat + theme(plot.margin = margin(2,2,2,2)) |
     plot4_Landsat + theme(plot.margin = margin(2,2,2,2))
   ) /
  (plot5_Landsat + theme(plot.margin = margin(2,2,2,2)) |
     plot6_Landsat + theme(plot.margin = margin(2,2,2,2))
   ) /
  (plot7_Landsat + theme(plot.margin = margin(2,2,2,2)) |
    legend
  ) +
  plot_layout(heights = c(1,1,1,1)) &
  theme(plot.margin = margin(2,2,2,2))
fig3
ggsave(here("output", "figures", "manuscript", "fig3.jpeg"), plot = fig3,
       dpi = 300, width = 7, height = 11)
```

## (Fig 4) Maps + graph validated plots

```{r}
plot_plots <- data_validated_plots %>%
  mutate(EUNIS = forcats::fct_relevel(EUNIS, "Change", after = Inf)) %>%
  count(EUNIS, valid_status, name = "n") %>% 
  ggplot(aes(x = EUNIS, y = n, fill = valid_status)) +
  geom_col(position = position_stack()) +
  xlab(NULL) + ylab("Number of plots") +
  scale_y_continuous(labels = label_number(scale = 1e-3, suffix = "k")) +
  scale_fill_manual(values = c("not_valid" = "#D55E00", "valid" = "#0072B2",
                               "NA" = "#828282"),
                    labels = c("not_valid" = "Non-validated",
                               "valid"     = "Validated",
                               "NA"        = "NA"),
                    name = "Validated") +
  scale_x_discrete(
    breaks = c("Q", "R", "S", "T", "Change"),
    labels = function(x) {
      x <- ifelse(x == "Q", "Q - Wetlands", x)
      x <- ifelse(x == "R", "R - Grasslands", x)
      x <- ifelse(x == "S", "S - Shrublands", x)
      x <- ifelse(x == "T", "T - Forests", x)
      x <- ifelse(x == "Change", "Habitat change", x)
      x
      }
    ) +
  theme_minimal() + theme(legend.position = c(0.9, 0.9)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
plot_plots
```

```{r}
europe_bbox_3035 <- st_bbox(
  c(
    xmin = 2500000,
    ymin = 1400000,
    xmax = 6500000,
    ymax = 5500000
  ),
  crs = st_crs(3035)
)

world <- ne_countries(scale = "medium", returnclass = "sf")
world_3035 <- st_transform(world, 3035)

world_3035_crop <- st_crop(world_3035, europe_bbox_3035)

plots_3035 <- st_as_sf(
  data_validated_plots,
  coords = c("Lon_updated", "Lat_updated"),
  crs = 4326
) |>
  st_transform(3035)
```

```{r}
# 1) Create a regular grid over your bbox
cell_km <- 50
grid_3035 <- st_make_grid(
  europe_bbox_3035,
  cellsize = cell_km * 1000,
  what = "polygons",
  square = TRUE
) |> 
  (\(g) st_sf(grid_id = seq_len(length(g)), geometry = g))()

# 2) Keep only grid cells that overlap your area of interest (optional)
grid_3035 <- st_intersection(grid_3035, st_as_sf(world_3035_crop))

plots_3035_crop <- st_crop(plots_3035, europe_bbox_3035)

# 3) Spatial join: assign each point to a grid cell
# (st_join on polygons brings grid_id onto points)
pts_with_grid <- st_join(
  plots_3035_crop %>% # points, filter only validated / invalidated
    filter(valid_status == "not_valid" | valid_status == "valid"),
  grid_3035["grid_id"],           # only keep grid_id column from grid
  left = FALSE                    # drop points outside the grid
)

# 4) Aggregate by grid_id and EUNIS category
agg <- pts_with_grid |>
  st_drop_geometry() |>
  mutate(
    # ensure EUNIS "Change" is last if you facet later (optional)
    EUNIS = fct_relevel(EUNIS, "Change", after = Inf),
    is_valid = case_when(
      valid_status == "valid" ~ 1L,
      TRUE ~ 0L
    )
  ) |>
  group_by(grid_id, EUNIS) |>
  summarise(
    n_total = n(),
    n_valid = sum(is_valid, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    prop_valid = if_else(n_total > 0, n_valid / n_total, NA_real_)
  )

# 5) Join back to grid polygons
grid_agg <- grid_3035 |>
  inner_join(agg, by = "grid_id")

map_grid_plots <- ggplot() +
  geom_sf(data = world_3035_crop, fill = "grey95", color = "grey80", linewidth = 0.2) +
  geom_sf(data = grid_agg, aes(fill = prop_valid), color = NA) +   # no borders for speed
  facet_wrap(
    ~ EUNIS,
    nrow = 3,
    ncol = 2,
    labeller = labeller(EUNIS = c("Q" = "Q - Wetlands", "R" = "R - Grasslands",
                                   "S" = "S - Shrublands", "T" = "T - Forests",
                                   "Change" = "Habitat change"))) +
  scale_fill_gradientn(
    colours = c("#D55E00", "#F0E442", "#0072B2"),    # orange → yellow → blue
    values = rescale(c(0, 0.5, 1)),
    limits = c(0, 1),
    oob = squish,
    na.value = "grey90",
    name = "Proportion validated"
  ) +
  theme_minimal() +
  theme(
    legend.position = "top",
    legend.justification = c("center", "top"),
    # Tighten the legend itself
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
    # Pull the legend box closer to the panels (reduce bottom gap)
    legend.box.margin = margin(t = 0, r = 0, b = -6, l = 0),
    # Keep a little top plot margin so nothing clips
    plot.margin = margin(t = 6, r = 0, b = 5, l = 0),
    panel.spacing = unit(6, "pt"),
    legend.key.size = unit(0.3, "lines"),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 7)
  ) +
  coord_sf(expand = FALSE) +
  guides(
  fill = guide_colorbar(
    barwidth = unit(4, "cm"),   # increase width
    barheight = unit(0.3, "cm"),  # keep it thin
    title.vjust = 0.8
  )
)
map_grid_plots
```

```{r}
plot_plots_inset <- plot_plots +
  theme(
    aspect.ratio = 3/2,        # height / width  → tall & thin (try 3/1, 4/1, 5/1)
    plot.margin  = margin(0, 0 , 0, 0),
    axis.text.x  = element_text(angle = 45, size = 8, vjust = 1, hjust = 1),
    axis.text.y  = element_text(size = 8, vjust = 1, hjust = 1),
    axis.title.y = element_text(size = 8),
    legend.title = element_blank(),
    legend.text  = element_text(size = 8),
    legend.key.size = unit(6, "pt"),

    # 🔹 Put legend inside, top-right of inset
    legend.position = c(1.4, 0.98),
    legend.justification = c(1, 1),

    # 🔹 Semi-transparent background helps readability
    legend.background = element_rect(fill = alpha("white", 0.6), color = NA)
  ) +
  coord_cartesian(clip = "off", expand = FALSE)

map_plots_with_inset <- map_grid_plots +
  inset_element(
    plot_plots_inset,
    left   = 0.45,
    right  = 0.97,
    bottom = -0.05,
    top    = 0.3, 
    align_to = "panel"
  )

map_plots_with_inset +
  theme(plot.margin = margin(1, 1, 1, 1))

ggsave(here("output", "figures", "manuscript", "fig4.jpeg"), 
       dpi = 300, width = 5, height = 8)
```

# Appendices

## (Appendix 4) Plot GPS observations used for models

```{r}
data_GPS <- get_full_datasets(initial_dataset, vars_RF_months_Landsat) %>%
  mutate(GPS = case_when(`Location method` %in% 
                           c("Location with GPS", 
                             "Location with differential GPS") ~ TRUE,
                         !`Location method` %in% 
                           c("Location with GPS", 
                             "Location with differential GPS") ~ FALSE,
                         is.na(`Location method`) ~ FALSE))
```

```{r}
points_3035 <- st_as_sf(data_GPS,
                        coords = c("Lon_updated", "Lat_updated"),
                        crs = 4326
                        ) |>
  st_transform(3035)

points_3035_crop <- st_crop(points_3035, europe_bbox_3035)
```

```{r}
# 1) Create a regular grid over your bbox
cell_km <- 50
grid_3035 <- st_make_grid(
  europe_bbox_3035,
  cellsize = cell_km * 1000,
  what = "polygons",
  square = TRUE
) |> 
  (\(g) st_sf(grid_id = seq_len(length(g)), geometry = g))()

# 2) Keep only grid cells that overlap your area of interest (optional)
grid_3035 <- st_intersection(grid_3035, st_as_sf(world_3035_crop))

points_3035_crop <- st_crop(points_3035_crop, europe_bbox_3035)

# 3) Spatial join: assign each point to a grid cell
# (st_join on polygons brings grid_id onto points)
pts_with_grid <- st_join(
  points_3035_crop,
  grid_3035["grid_id"],           # only keep grid_id column from grid
  left = FALSE                    # drop points outside the grid
)

# 4) Aggregate by grid_id and EUNIS category
agg <- pts_with_grid |>
  st_drop_geometry() |>
  group_by(grid_id, EUNIS1) |>
  summarise(
    n_total = n(),
    n_GPS = sum(GPS, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    prop_GPS = if_else(n_total > 0, n_GPS / n_total, NA_real_)
  )

# 5) Join back to grid polygons
grid_agg <- grid_3035 |>
  inner_join(agg, by = "grid_id")

map_grid_GPS <- ggplot() +
  geom_sf(data = world_3035_crop, fill = "grey95", color = "grey80", linewidth = 0.2) +
  geom_sf(data = grid_agg, aes(fill = prop_GPS), color = NA) +   # no borders for speed
  facet_wrap(
    ~ EUNIS1,
    nrow = 2,
    ncol = 2,
    labeller = labeller(EUNIS1 = c("Q" = "Q - Wetlands", "R" = "R - Grasslands",
                                   "S" = "S - Shrublands", "T" = "T - Forests"))) +
  scale_fill_gradientn(
    colours = c("#009E73", "#D3D3D3", "#CC79A7"),  # green → light grey → purple
    values = rescale(c(0, 0.5, 1)),
    limits = c(0, 1),
    oob = squish,
    na.value = "grey90",
    name = "Proportion GNSS"
  ) +
  theme_minimal() +
  theme(
    legend.position = "top",
    legend.justification = c("center", "top"),
    # Tighten the legend itself
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
    # Pull the legend box closer to the panels (reduce bottom gap)
    legend.box.margin = margin(t = 0, r = 0, b = -6, l = 0),
    # Keep a little top plot margin so nothing clips
    plot.margin = margin(t = 6, r = 0, b = 5, l = 0),
    panel.spacing = unit(6, "pt"),
    legend.key.size = unit(0.3, "lines"),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 7)
  ) +
  coord_sf(expand = FALSE) +
  guides(
  fill = guide_colorbar(
    barwidth = unit(4, "cm"),   # increase width
    barheight = unit(0.3, "cm"),  # keep it thin
    title.vjust = 0.8
  )
)
map_grid_GPS
ggsave(here("output", "figures", "manuscript", "appendix4.jpeg"), 
       dpi = 300, width = 5, height = 6)
```

## (Appendix 5) Plots distribution max NDVI

```{r}
plot_perc_Landsat <- full_dataset_months %>%
    # Select only GPS points
    dplyr::filter(`Location method` == "Location with GPS" | 
                    `Location method` == "Location with differential GPS") %>%
  distr_plot("NDVI_pos_value", "Maximum NDVI", grey = T) +
  scale_x_discrete(limits = c("T", "S", "R", "Q"),
                   labels = c("T" = "T - Forests", "S" = "S - Shrublands",
                              "R" = "R - Grasslands", "Q" = "Q - Wetlands"))
plot_perc_Landsat
ggsave(here("output", "figures", "manuscript", "appendix5.jpeg"), 
       dpi = 300, width = 5, height = 4)
```

## (Appendix 6) Compare classes before and after

```{r}
compare_classes <- function(
  data_before, data_step1, data_step2 = NULL,
  class_col = "EUNIS1",
  x_lab = NULL,
  title = NULL,
  return = c("plot", "both", "counts")
) {
  return <- match.arg(return)

  # Helper: safe counting that works with empty data frames
  count_safe <- function(df, out_name) {
    if (is.null(df) || nrow(df) == 0) {
      tibble(!!class_col := character(), !!out_name := integer())
    } else {
      if (!(class_col %in% names(df))) {
        abort(paste0("Column '", class_col, "' is missing in one of the inputs."))
      }
      df %>% count(.data[[class_col]], name = out_name)
    }
  }

  # Build counts table (before + step1; optionally step2)
  counts <- full_join(
    count_safe(data_before, "before"),
    count_safe(data_step1,  "step1"),
    by = class_col
  )

  has_step2 <- !is.null(data_step2)
  if (has_step2) {
    counts <- full_join(
      counts,
      count_safe(data_step2, "step2"),
      by = class_col
    )
  } else {
    # Ensure the column exists for consistent downstream handling (as NA)
    counts <- counts %>% mutate(step2 = NA_integer_)
  }

  # Replace missing counts with zero for plotting
  counts <- counts %>%
    mutate(
      before = replace_na(before, 0L),
      step1  = replace_na(step1,  0L),
      step2  = replace_na(step2,  0L)
    )

  # Prepare long format depending on availability of step2
  if (has_step2) {
    counts_long <- counts %>%
      pivot_longer(
        cols = c(before, step1, step2),
        names_to = "state",
        values_to = "n"
      ) %>%
      mutate(state = factor(state, levels = c("before", "step1", "step2")))
    legend_title <- "State"
  } else {
    counts_long <- counts %>%
      select(any_of(c(class_col, "before", "step1"))) %>%
      pivot_longer(
        cols = c(before, step1),
        names_to = "state",
        values_to = "n"
      ) %>%
      mutate(state = factor(state, levels = c("before", "step1")))
    legend_title <- "State"
  }

    # Plot
  p <- ggplot(counts_long, aes(x = .data[[class_col]], y = n, fill = state)) +
    geom_col(position = position_dodge(width = 0.85),
             width = 0.8, color = "black", linewidth = 0.25) +
    labs(
      title = title,
      x = x_lab,
      y = "Number of plot observations",
      fill = legend_title
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 0, hjust = 1),
      panel.grid.minor = element_blank()
    ) +
    scale_x_discrete(
    breaks = c("Q", "R", "S", "T"),
    
    labels = function(x) {
      x <- ifelse(x == "Q", "Q - Wetlands", x)
      x <- ifelse(x == "R", "R - Grasslands", x)
      x <- ifelse(x == "S", "S - Shrublands", x)
      x <- ifelse(x == "T", "T - Forests", x)
      x
      }
    ) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  # Optionally attach counts to the plot as an attribute for quick retrieval
  attr(p, "counts") <- counts

  # Return mode
  if (return == "plot") {
    return(p)
  } else if (return == "counts") {
    return(counts)
  } else {
    # both
    return(list(plot = p, counts = counts))
  }
}
```

```{r}
(compare_classes(training_subset_slope, 
                  filtered_data_20_10_slope_Landsat_probs,
                  filtered_data_20_10_slope_Landsat_probs_2x,
                  return = "plot") +
  compare_classes(training_subset_months, 
                  filtered_data_20_10_months_Landsat_probs,
                  filtered_data_20_10_months_Landsat_probs_2x,
                  return = "plot")
 ) +
  plot_annotation(tag_levels = "A", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  scale_fill_manual(
    breaks = c("before", "step1", "step2"),
    labels = c("No refinement", "Refinement step 1", "Refinement step 2"),
    values = c(
      "before" = "#edf8b1", 
      "step1"  = "#7fcdbb",
      "step2"  = "#2c7fb8"
    )
  ) &
  scale_y_continuous(limits = c(0, 45000),
                  labels = label_number(scale = 1e-3, suffix = "k")) &
  theme(legend.position = "bottom", legend.title = element_blank()) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
ggsave(here("output", "figures", "manuscript", "appendix6.jpeg"), 
       dpi = 300, width = 6, height = 4.5)
```

## (Appendix 8) ROC curves from best model

```{r}
ggplot(rocdata_months_Landsat_step2$roc, aes(x = FPR, y = TPR, color = Class)) +
  geom_line(linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray") +
  labs(x = "False Positive Rate (1-Specificity",
       y = "True Positive Rate (Sensitivity)",
       color = "Class (AUC)") +
  scale_color_manual(
    values = c(
      "Q (AUC = 0.97)" = "#0072B2",
      "R (AUC = 0.993)" = "#D55E00",
      "S (AUC = 0.984)" = "#009E73",
      "T (AUC = 0.999)" = "#CC79A7"
    ),
    labels = c(
      "Q (AUC = 0.97)" = "Q – Wetlands (AUC = 0.970)",
      "R (AUC = 0.993)" = "R – Grasslands (AUC = 0.993)",
      "S (AUC = 0.984)" = "S – Shrublands (AUC = 0.984)",
      "T (AUC = 0.999)" = "T – Forests (AUC = 0.999)"
    )
  ) +
  theme_minimal() + ggtitle(NULL) +
  theme(legend.position = "right")
ggsave(here("output", "figures", "manuscript", "appendix8.jpeg"), 
       dpi = 300, width = 7, height = 4)
```

## (Appendix 9) Variable importance from best model

```{r}
enframe(
  varimp_months_Landsat_step2$varimp$values,
  name  = "variable",
  value = "importance"
) %>%
  mutate(
    importance100 = 100 * importance / max(importance)
  ) %>%
  arrange(desc(importance100)) %>%
  mutate(variable_long = case_when(
    variable == "NDMI_max" ~ "Maximum NDMI",
    variable == "NDWI_max" ~ "Maximum NDWI",
    variable == "NDVI_max" ~ "Maximum NDVI",
    variable == "SAVI_max" ~ "Maximum SAVI",
    variable == "NDMI_min" ~ "Minimum NDMI",
    variable == "NDWI_min" ~ "Minimum NDWI",
    variable == "NDVI_min" ~ "Minimum NDVI",
    variable == "SAVI_min" ~ "Minimum SAVI",
    variable == "NDVI_avg_value_03" ~ "NDVI value at greening",
    variable == "NDVI_pos_value" ~ "NDVI value at peak",
    variable == "NDVI_avg_value_10" ~ "NDVI value at browning",
    variable == "SAVI_avg_value_03" ~ "SAVI value at greening",
    variable == "SAVI_pos_value" ~ "SAVI value at peak",
    variable == "SAVI_avg_value_10" ~ "SAVI value at browning",
    variable == "NDVI_auc_mar_oct" ~ "NDVI AUC greening - browning",
    variable == "SAVI_auc_mar_oct" ~ "SAVI AUC greening - browning",
    variable == "NDVI_diff_pos_march_value" ~ 
      "NDVI difference in value peak - greening",
    variable == "SAVI_diff_pos_march_value" ~
      "SAVI difference in value peak - greening",
    variable == "NDVI_diff_pos_march_doy" ~
      "NDVI difference in DOY peak - greening",
    variable == "SAVI_diff_pos_march_doy" ~
      "SAVI difference in DOY peak - greening",
    variable == "NDVI_diff_pos_oct_value" ~
      "NDVI difference in value peak - browning",
    variable == "SAVI_diff_pos_oct_value" ~
      "SAVI difference in value peak - browning",
    variable == "NDVI_diff_pos_oct_doy" ~
      "NDVI difference in DOY peak - browning",
    variable == "SAVI_diff_pos_oct_doy" ~
      "SAVI difference in DOY peak - browning",
    TRUE ~ variable
  )) %>%
ggplot(aes(x = reorder(variable_long, importance100), y = importance100)) +
  geom_col(fill = "#2c7fb8") +
  coord_flip() +
  labs(x = "Predictor", y = "Importance (scaled)") +
  theme_minimal()
ggsave(here("output", "figures", "manuscript", "appendix9.jpeg"), 
       dpi = 300, width = 6, height = 4)
```

## (Appendix 10): Maps + graph validated obs

```{r}
plot_obs <- data_validated_clean %>%
  count(EUNIS1, valid_Landsat, name = "n") %>% 
  ggplot(aes(x = EUNIS1, y = n, fill = valid_Landsat)) +
  geom_col(position = position_stack()) +
  xlab(NULL) + ylab("Number of plot observations") +
  scale_y_continuous(labels = label_number(scale = 1e-3, suffix = "k")) +
  scale_fill_manual(values = c("FALSE" = "#D55E00", "TRUE" = "#0072B2",
                               "NA" = "#828282"),
                    labels = c("FALSE" = "Non-validated",
                               "TRUE" = "Validated",
                               "NA"        = "NA"),
                     name = "Validated") +
  scale_x_discrete(
    labels = function(x) {
      x <- ifelse(x == "Q", "Q - Wetlands", x)
      x <- ifelse(x == "R", "R - Grasslands", x)
      x <- ifelse(x == "S", "S - Shrublands", x)
      x <- ifelse(x == "T", "T - Forests", x)
      x
      }
    ) +
  theme_minimal() + theme(legend.position = c(0.9, 0.8)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.title = element_blank(),
        legend.text  = element_text(size = 12),
        legend.key.size = unit(12, "pt"))
plot_obs
```

```{r}
# 1) Create a regular grid over your bbox
cell_km <- 50
grid_3035 <- st_make_grid(
  europe_bbox_3035,
  cellsize = cell_km * 1000,
  what = "polygons",
  square = TRUE
) |> 
  (\(g) st_sf(grid_id = seq_len(length(g)), geometry = g))()

# 2) Keep only grid cells that overlap your area of interest (optional)
grid_3035 <- st_intersection(grid_3035, st_as_sf(world_3035_crop))

points_3035 <- st_as_sf(
  data_validated_clean,
  coords = c("Lon_updated", "Lat_updated"),
  crs = 4326
) |>
  st_transform(3035)

points_3035_crop <- st_crop(points_3035, europe_bbox_3035)

# 3) Spatial join: assign each point to a grid cell
# (st_join on polygons brings grid_id onto points)
pts_with_grid <- st_join(
  points_3035_crop %>% # points, filter only validated / invalidated
    filter(valid_Landsat == "TRUE" | valid_Landsat == "FALSE"),
  grid_3035["grid_id"],           # only keep grid_id column from grid
  left = FALSE                    # drop points outside the grid
)

# 4) Aggregate by grid_id and EUNIS category
agg <- pts_with_grid |>
  st_drop_geometry() |>
  mutate(
    is_valid = case_when(
      valid_Landsat== "TRUE" ~ 1L,
      TRUE ~ 0L
    )
  ) |>
  group_by(grid_id, EUNIS1) |>
  summarise(
    n_total = n(),
    n_valid = sum(is_valid, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    prop_valid = if_else(n_total > 0, n_valid / n_total, NA_real_)
  )

# 5) Join back to grid polygons
grid_agg <- grid_3035 |>
  inner_join(agg, by = "grid_id")

map_obs_plots <- ggplot() +
  geom_sf(data = world_3035_crop, fill = "grey95", color = "grey80", linewidth = 0.2) +
  geom_sf(data = grid_agg, aes(fill = prop_valid), color = NA) +   # no borders for speed
  facet_wrap(
    ~ EUNIS1,
    nrow = 3,
    ncol = 2,
    labeller = labeller(EUNIS1 = c("Q" = "Q - Wetlands", "R" = "R - Grasslands",
                                   "S" = "S - Shrublands", "T" = "T - Forests"))) +
  scale_fill_gradientn(
    colours = c("#D55E00", "#F0E442", "#0072B2"),    # orange → yellow → blue
    values = rescale(c(0, 0.5, 1)),
    limits = c(0, 1),
    oob = squish,
    na.value = "grey90",
    name = "Proportion validated"
  ) +
  theme_minimal() +
  theme(
    legend.position = "top",
    legend.justification = c("center", "top"),
    # Tighten the legend itself
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
    # Pull the legend box closer to the panels (reduce bottom gap)
    legend.box.margin = margin(t = 0, r = 0, b = -6, l = 0),
    # Keep a little top plot margin so nothing clips
    plot.margin = margin(t = 6, r = 0, b = 5, l = 0),
    panel.spacing = unit(6, "pt"),
    legend.key.size = unit(0.3, "lines"),
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 7)
  ) +
  coord_sf(expand = FALSE) +
  guides(
  fill = guide_colorbar(
    barwidth = unit(4, "cm"),   # increase width
    barheight = unit(0.3, "cm"),  # keep it thin
    title.vjust = 0.8
  )
)
map_obs_plots
```

```{r}
(map_obs_plots / 
(
  (plot_spacer() | plot_obs | plot_spacer()) +
    plot_layout(widths = c(1, 3, 1))    # ← side padding : middle column is the plot
)
)  +
  plot_layout(heights = c(4, 2)) +
  theme(plot.margin = margin(1, 1, 1, 1))
ggsave(here("output", "figures", "manuscript", "appendix10.jpeg"), 
       dpi = 300, width = 5, height = 8)
```

# Session info

```{r}
sessionInfo()
```
