---
title: "MOTIVATE_State_BioDiv"
author: "Michael Glaser"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# SETUP

```{r}

### general
library(plyr)
library(tidyverse)
library(readxl)
library(forcats)
library(stringr)
library(data.table)
library(reshape2)
library(purrr)
library(grid)
library(gridExtra)
library(cowplot)

### spatial
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(terra)
library(tidyterra)
library(exactextractr)
library(countrycode)

### occupancy mods
library(spOccupancy)

### taxonomy
library(WorldFlora)


### - ### helper functions ### - ###
  
  ### generalized rounding function
    mround <- function(x, base) {
      ifelse(is.na(x), NA, base * floor(x / base))
    }

  ### get_legend function  
  get_legend2 <- function(plot, legend = NULL) {
  if (is.ggplot(plot)) {
    gt <- ggplotGrob(plot)
  } else {
    if (is.grob(plot)) {
      gt <- plot
    } else {
      stop("Plot object is neither a ggplot nor a grob.")
    }
  }
  pattern <- "guide-box"
  if (!is.null(legend)) {
    pattern <- paste0(pattern, "-", legend)
  }
  indices <- grep(pattern, gt$layout$name)
  not_empty <- !vapply(
    gt$grobs[indices], 
    inherits, what = "zeroGrob", 
    FUN.VALUE = logical(1)
  )
  indices <- indices[not_empty]
  if (length(indices) > 0) {
    return(gt$grobs[[indices[1]]])
  }
  return(NULL)
}  
    
### - ### PATHS and folders ### - ###
    
     ### working directory
      setwd("D:/33_MOTIVATE/05_WP2/01_RWD_DataCleaning")
    
     ### date for versioning
      sys.date <- gsub("-", "", Sys.Date())
      
     ### path to data
     path.data <- ("../00_DATA/")
     
     ### output path
      path.out <- paste0("./00_OUTPUT/output_", sys.date)
      dir.create(path.out)

###-### settings
options(scipen=9999)

```

## plot settings

```{r}

### plot sizes
wide <- 15.8
half <- 7.9

high <- 20

wide.pp <- 20
height.pp <- 15

### plot theme
my.theme <- theme_classic()+
  theme(plot.title=element_text(face="bold", size=12),
        axis.title=element_text(face="bold", size=10),
        axis.text=element_text(size=8),
        legend.title=element_text(face="bold", size=10),
        legend.position="bottom")

### color palettes

col.hab <- c(
  "forests"                       = "#33a02c",  
  "grasslands"                    = "#8c510a",  
  "heathlands, scrub and tundra"  = "#dfc27d",  
  "inland sparsely vegetated"     = "#b2abd2",  
  "marine"                        = "#0571b0",
  "coastal"                       = "#80cdc1",  
  "wetlands"                      = "#74add1",  
  "anthropogenic"                 = "#ff7f00",  
  "class not possible"            = "#d9d9d9",  
  "class failed"                  = "#969696",  
  "multiple class"                = "#525252"   
)


```

## projection settings

```{r}
proj.crs <- 3035
```

# STUDY AREA

## remove islands

```{r}

### SPAIN
  esp <- ne_states(country="Spain")
  esp.excl <- c("Baleares", "Las Palmas", "Santa Cruz de Tenerife", "Ceuta", "Melilla")
  esp <- esp[!esp$name %in% esp.excl,]
  
  esp <- esp %>% summarise(name=first(admin),
                          geometry=st_union(geometry))

### CRETE & GREEK ISLANDS
  gr <- ne_states(country="Greece")
  gr.excl <- c("Kriti", "Notio Aigaio", "Voreio Aigaio", "Ionioi Nisoi" )
  gr <- gr[!gr$name %in% gr.excl,]

  gr.isl <- st_cast(gr, "POLYGON")
  gr.isl$num <- 1:nrow(gr.isl)
  
  
  gr.isl <- gr.isl[!(gr.isl$name=="Anatoliki Makedonia kai Thraki" & !gr.isl$num==4),]
  gr.isl <- gr.isl[!(gr.isl$name=="Thessalia" & !gr.isl$num==8),]
  gr.isl <- gr.isl[!(gr.isl$name=="Stereá Elláda" & !gr.isl$num %in% c(13, 15)),]
  gr.isl <- gr.isl[!(gr.isl$name=="Attiki" & !gr.isl$num %in% c(16)),]
  gr.isl <- gr.isl[!(gr.isl$name=="Peloponnisos" & !gr.isl$num %in% c(27)),]
  
  gr.isl$num <- NULL

  gr <- gr.isl %>% summarise(name=first(admin),
                          geometry=st_union(geometry))

### ITALY
  it <- ne_states(country="Italy")
  it.excl.region <- c("Sicily", "Sardegna")
  it <- it[!it$region %in% it.excl.region,]
  
  it.isl <- it[it$name %in% c("Grosseto", "Livorno", "Latina", "Napoli"),]
  
  it.isl <- st_cast(it.isl, "POLYGON")
  it.isl$num <- LETTERS[1:nrow(it.isl)]
  
  it.isl <- it.isl[it.isl$num %in% c("A", "C", "E", "G"),]
  
  it.isl$num <- NULL
  
  it[it$name=="Grosseto",] <- it.isl[it.isl$name=="Grosseto",]
  it[it$name=="Livorno",] <- it.isl[it.isl$name=="Livorno",]
  it[it$name=="Latina",] <- it.isl[it.isl$name=="Latina",]
  it[it$name=="Napoli",] <- it.isl[it.isl$name=="Napoli",]
  
  it <- it %>% summarise(name=first(admin),
                          geometry=st_union(geometry))
  
  

### FRANCE 
  fr <- ne_states(country="France")
  fr.excl <- c("Guyane française",  "Guadeloupe", "Martinique" ,  "Mayotte", "Haute-Corse",  "Corse-du-Sud", "La Réunion" )
  fr <- fr[!fr$name %in% fr.excl,]
  
  fr <- fr %>% summarise(name=first(admin),
                          geometry=st_union(geometry))

### PORTUGAL
  por <- ne_states(country="Portugal")
  por.excl <- c("Madeira", "Azores")
  por <- por[!por$name %in% por.excl,]
  
  por <- por %>% summarise(name=first(admin),
                          geometry=st_union(geometry))

### NORWAY
  nor <- ne_states(country="Norway")
  nor.excl <- c("Svalbard", "Bouvet Island" )
  nor <- nor[!nor$name %in% nor.excl,]
  
  
  
  
  nor.bbox <- st_bbox(nor)
  nor.bbox["xmin"] <- 0
  nor.bbox["ymin"] <- 57
  
  nor.bbox.sf <- st_as_sfc(nor.bbox)
  nor <- st_crop(nor, nor.bbox.sf)
  
  
  nor <- nor %>% summarise(name=first(admin),
                         geometry=st_union(geometry))
  
### UK
  uk <- ne_states(country="United Kingdom")
  uk.excl <- c("Isle of Wight", "Shetland Islands", "Isles of Scilly", "Orkney", "Eilean Siar")
  uk <- uk[!uk$name %in% uk.excl,]
  uk <- uk %>% summarise(name=first(admin),
                          geometry=st_union(geometry))

### NL
  nl <- ne_states(country="Netherlands")
  nl.excl <- c("St. Eustatius", "Saba")
  nl <- nl[!nl$name %in% nl.excl,]

  nl.isl <- nl[nl$name %in% c("Friesland", "Noord-Holland", "Groningen"),]
  
  
  nl.isl <- st_cast(nl.isl, "POLYGON")
  
  nl.isl$num <- LETTERS[1:nrow(nl.isl)]


  nl.isl <- nl.isl[nl.isl$num %in% c("A", "C", "E"),]
  
  nl.isl$num <- NULL
  
  nl[nl$name=="Friesland",] <- nl.isl[nl.isl$name=="Friesland",]
  nl[nl$name=="Noord-Holland",] <- nl.isl[nl.isl$name=="Noord-Holland",]
  nl[nl$name=="Groningen",] <- nl.isl[nl.isl$name=="Groningen",]
  
  nl <- nl %>% summarise(name=first(admin),
                         geometry=st_union(geometry))
  
### EST
  
  est <- ne_states(country="Estonia")
  

  
  est.excl <- c("Hiiu", "Saare")
  est <- est[!est$name %in% est.excl,]   
  
  est.isl <- est[est$name %in% c("Lääne", "Pärnu", "Harju"),]
  

  est.isl <- st_cast(est.isl, "POLYGON")
  est.isl$num <- LETTERS[1:nrow(est.isl)]


  est.isl <- est.isl[est.isl$num %in% c("A", "C", "E"),]
  
  est.isl$num <- NULL
  
  est[est$name=="Lääne",] <- est.isl[est.isl$name=="Lääne",]
  est[est$name=="Pärnu",] <- est.isl[est.isl$name=="Pärnu",]
  est[est$name=="Harju",] <- est.isl[est.isl$name=="Harju",]
  
  est <- est %>% summarise(name=first(admin),
                         geometry=st_union(geometry))
  
### Finland
  fin <- ne_states(country="Finland")

  fin.isl <- fin[fin$name %in% c("Northern Ostrobothnia", "Ostrobothnia", "Finland Proper"),]
  
  fin.isl <- st_cast(fin.isl, "POLYGON")
  fin.isl$num <- 1:nrow(fin.isl)
  
  fin.isl <- fin.isl[!(fin.isl$name=="Ostrobothnia" & !fin.isl$num==31),]
  fin.isl <- fin.isl[!(fin.isl$name=="Northern Ostrobothnia" & !fin.isl$num==1),]
  fin.isl <- fin.isl[!(fin.isl$name=="Finland Proper" & !fin.isl$num==3),]
  
  fin.isl$num <- NULL
  
  fin[fin$name=="Ostrobothnia",] <- fin.isl[fin.isl$name=="Ostrobothnia",]
  fin[fin$name=="Northern Ostrobothnia",] <- fin.isl[fin.isl$name=="Northern Ostrobothnia",]
  fin[fin$name=="Finland Proper",] <- fin.isl[fin.isl$name=="Finland Proper",]
  
  fin <- fin %>% summarise(name=first(admin),
                         geometry=st_union(geometry))
  
### Sweden
  
  sw <- ne_states(country="Sweden")
  
  sw <- sw[!sw$name %in% "Gotland",]
  
  sw.isl <- sw[sw$name=="Kalmar",]

  sw.isl <- st_cast(sw.isl, "POLYGON")
  sw.isl$num <- LETTERS[1:nrow(sw.isl)]
  
  sw.isl <- sw.isl[sw.isl$num %in% c("A"),]
  sw.isl$num <- NULL
  
  sw[sw$name=="Kalmar",] <- sw.isl[sw.isl$name=="Kalmar",]
  
  sw <- sw %>% summarise(name=first(admin),
                         geometry=st_union(geometry))
  
### Denmark
  
  den <- ne_states(country="Denmark")

  den <- st_cast(den, "POLYGON")
  den$num <- LETTERS[1:nrow(den)]
  den <- den[!den$num %in% c("I", "L", "P"),]
  
  den <- den %>% summarise(name=first(admin),
                         geometry=st_union(geometry))
    
### ISLAND LIST
#island.list <- list(esp, gr, it, fr, uk, por, nl, est, den, sw, fin, nor) ### FULL LIST

island.list <- list(gr, it, fr, uk, nl, est, den) ### removing data-poor regions
  
star.fill <- do.call(rbind, island.list)


```

## retrieve study area

```{r}

### load all European countries
  star <- ne_countries(continent="Europe", scale="large")
  star <- star[!star$name %in% c("Russia", "France", "Greece", "Spain", "Italy", "United Kingdom", "Guernsey", "Jersey", "Faeroe Is.", "Iceland", "Isle of Man", "Åland", "Malta", "Portugal", "Norway", "Netherlands", "Finland", "Estonia", "Sweden", "Denmark"),]
  
### remove extraneous columns
  star <- star[,c("name", "geometry")]

### bind with filler
  star <- rbind(star, star.fill)
  
### cut to lat/long filters
  star.bbox <- st_bbox(star)
  star.bbox["ymax"] <- 60
  star.bbox["xmax"] <- 25
  
  star <- st_crop(star, star.bbox)

### transofrm to project crs
  star.tf <- st_transform(star, proj.crs)

### add area column  
  star.tf$area.km2 <- as.numeric(st_area(star.tf)) /1e6  
  
### unify for plotting (buffer to clean up left over geometries)
  star.union <- st_as_sf(st_union(star.tf))
  star.union <- st_as_sf(st_buffer(star.union, dist=1))
  

### unclutter environment    
rm(list=c("island.list", "rus", "uk", "fr", "star.fill", "gr", "it", "it.isl", "est", "est.isl", "fin", "fin.isl", "esp", "por", "nor", "nor.bbox", "nor.bbox.sf", "nl", "nl.isl", "gr.isl", "den", "sw.isl", "sw"))
rm(list=ls(pattern=".excl"))



```

# BIOGEOGRAPHIC REGION

```{r}

biogeo <- terra::vect(paste0(path.data,"02_BIOGEO/BiogeoRegions2016.shp"))

biogeo.tf <- st_as_sf(st_transform(st_as_sf(biogeo), proj.crs))

biogeo.tf <- st_crop(biogeo.tf, st_bbox(star.tf))

biogeo.tf$area.km2 <- as.numeric(st_area(biogeo.tf)) /1e6

biogeo.tf$name.pretty <- gsub(" Bio-geographical Region", "", biogeo.tf$name)

```

# HEADER

```{r}

### read header
head <- as.data.table(fread(paste0(path.data,"01_EVA/200_Motivate20250415_notJUICE_header_with_precise_coordinates.csv"), 
         header=T,
         sep = "\t"))

### adjust column names
colnames(head) <- gsub(" ", ".", colnames(head))

     ### extract year, day, month
          head$date <- head$Date.of.recording ### storage for original column
          head[, c("day", "month", "year") := tstrsplit(Date.of.recording, "\\.")] ### data.table version is faster
          head$year <- as.numeric(head$year) ### convert to numeric
  
     ### create timesteps going backwards from the latest year of records
          yr.int.length <- 15
          
          max.year <- max(head$year, na.rm=T)
          min.year <- min(head$year, na.rm=T)
          
          yr.breaks <- rev(seq(max.year + 1, min.year - 1, by = -yr.int.length))
          yr.labels <- yr.breaks[-length(yr.breaks)]
          
          head[, yr.interval := as.numeric(as.character(cut(year,
                     breaks = yr.breaks,
                     labels = yr.labels,
                     right = F,
                     include.lowest = T)))]
          
          table(head$yr.interval, useNA="ifany")
          
  ### manually correct country column
    head$country <- revalue(head$Country, replace=c("Slovak Republic"="Slovakia",
                                                    "DE"="Germany",
                                                    "Ukraina"="Ukraine",
                                                    "ES"="Spain",
                                                    "Moldavia"="Moldova",
                                                    "Andorra, Principality of"="Andorra"))
  
  ### create an area column
    head$area <- head$'Relevé.area.(m²)' 
     
    ##head$area[head$area>625] <- NA ### what are unrealistic area sizes?
    
  ### convert project types to lower
    head$RS_PROJTYP <- tolower(head$RS_PROJTYP)

```

### manually correct erroneous entries

received from Ilona Knollova 26-11-2025 pertains to the manipulation column in ReSurvey and the project type column

DE_00039 is likely typo from original e-mail
Czechia_nvd is a "Dataset" not a "RS_CODE"
"RU_0003c" is not in RS_CODE

```{r}

all.codes <- unique(head$RS_CODE)


### vector of all corrections just to double-check
all.cors <- c("AT_0004b",
              "CZ_0019_024", "CZ_0019_034", "CZ_0019_054",
              "DE_0039", "DE_0037_29", "DE_0037_88",
              "FRCH_0001",
              "HU_0002",
              "IT_0001a", "IT_0001b", "IT_0001c", "IT_0001d", "IT_0001e",
              "PL_0009",
              "RU_0003c"
              )

###-### changes to Manipulate (Y/N)

  ### typo (M instead of N)
  head$`Manipulate.(y/n)`[head$Dataset=="Czechia_nvd"] <- "N"
  
  ### not manipulated (according to IK)
  manip.no <- c("AT_0004b", 
                "CZ_0016_054",
                "DE_0039", "DE_0037_29",
                "FRCH_0001",
                "IT_0001_a", "IT_0001_b", "IT_0001c", "IT_0001d", "IT_0001e")

  head$`Manipulate.(y/n)`[head$RS_CODE %in% manip.no] <- "N"

###-### changes to RS_PROJTYP

  ### change study type to Site comparison
  study.site <- c("HU_0002")

  head$RS_PROJTYP[head$RS_CODE %in% study.site] <- "site comparison"


```

# EUNIS

## habitat from data

```{r}

### make a new column to retain original assessments
     head$eunis <- gsub("!", "", head$Expert.system) ### "!" only signified different classification path, not habitat
     head$eunis[head$eunis==""] <- "X" ### classification not possible bc prerequisites missing (usually cover missing)
     head$eunis[head$eunis=="~"] <- "Y" ### Esy fails (ambiguous assignemtn, not enough indicator species, not enough cover of i.s. etc.)

### separate rows
     head.esy <- head[, .(eunis = unlist(strsplit(eunis, ","))), by = PlotObservationID]  # Separate rows by EUNIS

     ### some edits
     head.esy$eunis <- gsub("MA", "M", head.esy$eunis) ### remove the unnecessary MA
     head.esy$eunis <- gsub("Qa", "Q", head.esy$eunis) ### remove non-capital letters that hamper classification
     head.esy$eunis <- gsub("Qb", "Q", head.esy$eunis) ### unofficial category, merge to Q
     head.esy$eunis <- gsub("^S[a-zA-Z]", "S", head.esy$eunis) ### unofficial category, merge to S
     head.esy$eunis <- gsub("^M[a-zA-Z]", "M", head.esy$eunis) ### marine vegetation not target for MOTIVATE
     head.esy$eunis <- gsub("^P([abcdefgh])$", "P3\\1", head.esy$eunis) ### 

### extract eunis level
     
     ### level = number of characters
     head.esy$level <- nchar(head.esy$eunis)
     
     ### level 1
     head.esy$lvl1 <- substr(head.esy$eunis, 1,1) ### level 1 is the first character

     ### level 2
     head.esy$lvl2 <- NA ### empty storage column
     head.esy$lvl2[head.esy$level>1] <- substr(head.esy$eunis[head.esy$level>1], 1,2)
     
     ### level 3
     head.esy$lvl3 <- NA
     head.esy$lvl3[head.esy$level>2] <- substr(head.esy$eunis[head.esy$level>2], 1,3)
     
     ### add missing levels for unclassified etc.
     head.esy$lvl2[head.esy$lvl1=="X"] <- "XX"
     head.esy$lvl2[head.esy$lvl1=="Y"] <- "YY"
     
     head.esy$lvl3[head.esy$lvl1=="X"] <- "XXX"
     head.esy$lvl3[head.esy$lvl1=="Y"] <- "YYY"
     
     ### create a combined column (RUNTIME IS 15-30 mins, get coffee!!!!)
     head.esy <- head.esy[, .(
     eunis.countlvl1 = uniqueN(lvl1),  # Count unique lvl1
     eunis.lvl1 = paste(sort(unique(lvl1)), collapse = ", "), # Concatenate unique names
     eunis.countlvl2 = uniqueN(lvl2),  # Count unique lvl1
     eunis.lvl2 = paste(sort(unique(lvl2)), collapse = ", "),
     eunis.countlvl3 = uniqueN(lvl3),  # Count unique lvl1
     eunis.lvl3 = paste(sort(unique(lvl3)), collapse = ", ")
     ), by = PlotObservationID]
     
     ### add symbols for missing assignments
     head.esy$eunis.lvl1[head.esy$eunis.countlvl1>1] <- "Z"      ### multiple lvl1
     head.esy$eunis.lvl2[head.esy$eunis.countlvl2>1] <- "ZZ"     ### multiple lvl2
     head.esy$eunis.lvl3[head.esy$eunis.countlvl3>1] <- "ZZZ"    ### multiple lvl3
     
     head.esy$eunis.lvl2[head.esy$eunis.lvl2==""] <- "YY"        ### Esy failed lvl 2
     head.esy$eunis.lvl3[head.esy$eunis.lvl3==""] <- "YYY"       ### Esy failed lvl 3
     
     ### add method column
      head.esy$method <- "esy"

```

## Alicias Crosswalk

DK_Naturdata_Res

```{r}
     ### CROSSWALK BY ALICIA
     
          ### read Alicias data and check
          dk.eunis <- fread(paste0(path.data,"04_EUNIS/db_Michael.csv"))
          all(dk.eunis$PlotObservationID %in% head$PlotObservationID)
          
          dk.eunis$method <- "acw"
          
          
          
          ### align with my encoding     
          dk.eunis$eunis.edit <- gsub("MA", "M", dk.eunis$EUNIS)
          dk.eunis$eunis.edit <- gsub("^P([abcdefghi])$", "P", dk.eunis$eunis.edit)
          
          unique(dk.eunis$eunis.edit)
          
          ### define levels
          dk.eunis$level <- nchar(dk.eunis$eunis.edit)
          
          ### extract level 1
          dk.eunis$lvl1 <- substr(dk.eunis$eunis.edit, 1,1)
          
          ### extract level 2
          dk.eunis$lvl2 <- NA
          dk.eunis$lvl2[dk.eunis$level>1] <- substr(dk.eunis$eunis.edit[dk.eunis$level>1], 1,2)
          dk.eunis$lvl2[is.na(dk.eunis$lvl2)] <- "XX"
          
          ### extract level 3
          dk.eunis$lvl3 <- NA
          dk.eunis$lvl3[dk.eunis$level>2] <-  substr(dk.eunis$eunis.edit[dk.eunis$level>2], 1,3)
          dk.eunis$lvl3[is.na(dk.eunis$lvl3)] <- "XXX"
               
```

## Pl@ntBert abundance

```{r}

plantbert <- fread(paste0(path.data,"04_EUNIS/plantbert_habitat_predictions.csv"))
colnames(plantbert) <- gsub(" ", ".", colnames(plantbert))

plantbert$top.lvl2 <- substr(plantbert$`Most.likely.EUNIS.level-2.habitat`, 1,2)
plantbert$top.lvl2 <- gsub("MA", "M2", plantbert$top.lvl2)

plantbert$top.lvl2.perc <- as.numeric(sub(".*\\(([^%]+)%\\).*", "\\1", plantbert$`Most.likely.EUNIS.level-2.habitat`))

plantbert$method <- "pb_abund"

### cutoff table
  cutoff <- c(75, 80,  85, 90, 95, 97, 99)
  
  plantbert.cutoff <- plantbert %>% 
    dplyr::summarise(cutoff=cutoff,
                     n.remaining=sapply(cutoff, function (c) sum(top.lvl2.perc>=c)))

### raw plot
  plot.pb.hist <- ggplot(plantbert)+my.theme+
    geom_histogram(aes(x=top.lvl2.perc))+
    scale_x_continuous(breaks=seq(0,100,10))+
    scale_y_continuous(expand=expansion(mult=c(0,0.15)))+
    labs(x="certainty (%)", y="plots")

### cutoff
  plot.pb.cutoff <- ggplot(plantbert.cutoff)+my.theme+
    geom_bar(aes(x=cutoff, y=n.remaining), stat="identity", width=0.75)+
    scale_x_continuous(breaks=cutoff)+
    scale_y_continuous(expand=expansion(mult=c(0,0.15)))+
    geom_label(aes(x=cutoff, y=n.remaining*1.1, label=paste0(round(n.remaining/1000),"k")),
               size=3)+
    labs(x="cutoff value (%)", y="remaining plots")

### combined plot
  grid.pb <- plot_grid(plot.pb.hist, plot.pb.cutoff, nrow=2)
  
ggsave(filename=paste0(path.out,"/plantbert_abund.png"),
       plot=grid.pb,
       width=wide, unit="cm",
       dpi=900,
       bg="white")

### choose cutoff and filter accordingly
pb.cut.abund <- 75

plantbert <- plantbert %>% 
  filter(top.lvl2.perc>=pb.cut.abund) %>% 
  mutate(method="pb_abund")

nrow(plantbert)

```

## Pl@ntBert pres-abs

```{r}
plantbert.pa <- fread(paste0(path.data,"04_EUNIS/plantbert_habitat_predictions_random.csv"))
colnames(plantbert.pa) <- gsub(" ", ".", colnames(plantbert.pa))

plantbert.pa$top.lvl2 <- substr(plantbert.pa$`Most.likely.EUNIS.level-2.habitat`, 1,2)
plantbert.pa$top.lvl2 <- gsub("MA", "M2", plantbert.pa$top.lvl2)

plantbert.pa$top.lvl2.perc <- as.numeric(sub(".*\\(([^%]+)%\\).*", "\\1", plantbert.pa$`Most.likely.EUNIS.level-2.habitat`))

### cutoff table
  cutoff <- c(75, 80,  85, 90, 95, 97, 99)
  
  plantbert.pa.cutoff <- plantbert.pa %>% 
    dplyr::summarise(cutoff=cutoff,
                     n.remaining=sapply(cutoff, function (c) sum(top.lvl2.perc>=c)))

### raw plot
  plot.pb.pa.hist <- ggplot(plantbert.pa)+my.theme+
    geom_histogram(aes(x=top.lvl2.perc))+
    scale_x_continuous(breaks=seq(0,100,10))+
    scale_y_continuous(expand=expansion(mult=c(0,0.15)))+
    labs(x="certainty (%)", y="plots")

### cutoff
  plot.pb.pa.cutoff <- ggplot(plantbert.pa.cutoff)+my.theme+
    geom_bar(aes(x=cutoff, y=n.remaining), stat="identity", width=0.75)+
    scale_x_continuous(breaks=cutoff)+
    scale_y_continuous(expand=expansion(mult=c(0,0.15)))+
    geom_label(aes(x=cutoff, y=n.remaining*1.1, label=paste0(round(n.remaining/1000),"k")),
               size=3)+
    labs(x="cutoff value (%)", y="remaining plots")

### combined plot
  grid.pb <- plot_grid(plot.pb.hist, plot.pb.cutoff, nrow=2)
  
ggsave(filename=paste0(path.out,"/plantbert_presabs.png"),
       plot=grid.pb,
       width=wide, unit="cm",
       dpi=900,
       bg="white")

### choose cutoff and filter accordingly
pb.cut.presabs <- pb.cut.abund*0.92/0.78

plantbert.pa <- plantbert.pa %>% 
  filter(top.lvl2.perc>=pb.cut.presabs) %>% 
  mutate(method="pb_presabs")

nrow(plantbert.pa)

```

## merge into head.esy

```{r}

### update AWC
markers <- c(eunis.lvl1 = "X", eunis.lvl2 = "XX", eunis.lvl3 = "XXX")

head.esy <- head.esy %>%
  left_join(
    dk.eunis %>%
      select(PlotObservationID, lvl1, lvl2, lvl3, method) %>%
      rename(
        eunis.lvl1.new = lvl1,
        eunis.lvl2.new = lvl2,
        eunis.lvl3.new = lvl3,
        method.new     = method
      ),
    by = "PlotObservationID"
  ) %>%
  mutate(
    rep1 = eunis.lvl1 %in% c("X", "Y", "Z")       & !is.na(eunis.lvl1.new),
    rep2 = eunis.lvl2 %in% c("XX", "YY", "ZZ")    & !is.na(eunis.lvl2.new),
    rep3 = eunis.lvl3 %in% c("XXX", "YYY", "ZZZ") & !is.na(eunis.lvl3.new),

    eunis.lvl1 = ifelse(rep1, eunis.lvl1.new, eunis.lvl1),
    eunis.lvl2 = ifelse(rep2, eunis.lvl2.new, eunis.lvl2),
    eunis.lvl3 = ifelse(rep3, eunis.lvl3.new, eunis.lvl3),

    method = ifelse((rep1 | rep2 | rep3) & !is.na(method.new), method.new, method)
  ) %>%
  select(-ends_with(".new"), -rep1, -rep2, -rep3)

### update Pl@ntBert abundance
head.esy <- head.esy %>%
  left_join(
    plantbert %>%
      select(PlotObservationID, top.lvl2, top.lvl2.perc, method) %>%
      rename(
        eunis.lvl2.new = top.lvl2,
        method.new     = method
      ),
    by = "PlotObservationID"
  ) %>%
  mutate(
    rep2       = eunis.lvl2 %in% c("XX", "YY", "ZZ") & !is.na(eunis.lvl2.new),
    eunis.lvl2 = ifelse(rep2, eunis.lvl2.new, eunis.lvl2),
    method     = ifelse(rep2 & !is.na(method.new), method.new, method)
  ) %>%
  select(-ends_with(".new"), -rep2)

### update Pl@ntBert pres-abs
head.esy <- head.esy %>%
  left_join(
    plantbert.pa %>%
      select(PlotObservationID, top.lvl2, top.lvl2.perc, method) %>%
      rename(
        eunis.lvl2.new = top.lvl2,
        method.new     = method
      ),
    by = "PlotObservationID"
  ) %>%
  mutate(
    rep2       = eunis.lvl2 %in% c("XX", "YY", "ZZ") & !is.na(eunis.lvl2.new),
    eunis.lvl2 = ifelse(rep2, eunis.lvl2.new, eunis.lvl2),
    method     = ifelse(rep2 & !is.na(method.new), method.new, method)
  ) %>%
  select(-ends_with(".new"), -rep2)

table(head.esy$method, useNA="always")

```

## join with legend

```{r}

     ### CONTINUE with head.esy as usual 

     ### Create eunis.lvl1 for augmented datasets (hierarchy respected)
     head.esy$eunis.lvl1 <- substr(head.esy$eunis.lvl2, 1, 1)

     ### add names
     eunis.legend <- read_excel(paste0(path.data,"04_EUNIS/EUNIS-Habitats-2023-06-15-update_MG.xlsx"))
     eunis.legend$`EUNIS 2023 code` <- trimws(eunis.legend$`EUNIS 2023 code`)
     eunis.legend$`EUNIS 2023 code` <- gsub("MA", "M", eunis.legend$`EUNIS 2023 code`)
     
     ### join with legend
     head.esy <- left_join(head.esy, eunis.legend, by=c("eunis.lvl1"="EUNIS 2023 code"))
     head.esy <- left_join(head.esy, eunis.legend, by=c("eunis.lvl2"="EUNIS 2023 code")) 
     
     head.esy$eunis.lvl1.name <- head.esy$`EUNIS-2023 habitat name.x`
     head.esy$eunis.lvl2.name <- head.esy$`EUNIS-2023 habitat name.y`
     head.esy$eunis.method <- head.esy$method


### join with head (names can be done later)
head <- left_join(head, head.esy[,c("PlotObservationID", 
                                    "eunis.lvl1", 
                                    "eunis.lvl2", 
                                    "eunis.lvl1.name", 
                                    "eunis.lvl2.name",
                                    "eunis.method")], by="PlotObservationID")


```

## group eunis classes NEW

```{r}

head.eunis.grp <- head %>% 
     group_by(eunis.lvl2, eunis.lvl2.name) %>% 
     dplyr::summarise(n.group=n())

fwrite(head.eunis.grp, file=paste0(path.out, "/eunis_grouped.csv"), sep=",", dec=".", row.names=F)

```

## group eunis classes

```{r}

##head.eunis.tmp <- head %>% 
     ##group_by(eunis.lvl2, eunis.lvl2.name) %>% 
     ##dplyr::summarise(n.group=n())

##fwrite(head.eunis.tmp, file=paste0(path.out, "/eunis_grouped.csv"), sep=",", dec=".", row.names=F)

eunis.grp <- read_excel(paste0(path.data,"04_EUNIS/EUNIS_level2_groups_20250812.xlsx"))

vec.eunis <- eunis.grp$eunis.lvl2.grp
names(vec.eunis) <- eunis.grp$eunis.lvl2

head$eunis.lvl2 <- revalue(head$eunis.lvl2, replace=vec.eunis)

eunis.grp.legend <- unique(eunis.grp[,c("eunis.lvl2.grp", "eunis.lvl2.grp.name")])

head <- left_join(head, eunis.grp.legend, by=c("eunis.lvl2"="eunis.lvl2.grp"))

rm(eunis.grp, vec.eunis, eunis.grp.legend)

```

# SPATIAL

## head to sf

```{r}

sum(head$Longitude>180, na.rm=T)
sum(head$Latitude>90, na.rm=T)

### filter out coordinates
head <- head[!head$Longitude>180,]
head <- head[!head$Latitude>90,]

### save coordinates
head$lon.wgs <- head$Longitude
head$lat.wgs <- head$Latitude

### transform to sf
head.sf <- st_as_sf(head, coords=c("Longitude", "Latitude"), crs=4326)

### transform to project crs
head.sf <- st_transform(head.sf, crs=proj.crs)

head.coords <- st_coordinates(head.sf)

### save transformed longitude as a separate column
head.sf$lon.tf <- head.coords[,1]
head.sf$lat.tf <- head.coords[,2]

### extract correct country to match with names and for per area density
head.sf <- st_join(head.sf, star.tf["name"])

### extract biogeo region
head.sf <- st_join(head.sf, biogeo.tf["short_name"])
head.sf$biogeo.region <- head.sf$short_name
head.sf$short_name <- NULL

### subset to study area
head.sf$star <- st_intersects(head.sf, star.union, sparse=F)

rm(head.coords)
rm(head)

```

# FILTER

```{r}

### remove old filter when rerunning, otherwise counts are off
head.sf$filter.meta <- NULL

###-### COORDINATES

  ### filtering
  head.sf$filter.lon <- 0
  head.sf$filter.lon[is.na(head.sf$Longitude)] <- 1
  head.sf$filter.lon[head.sf$lon.wgs>180] <- 1
  head.sf$filter.lon[head.sf$lon.wgs<(-10)] <- 1
  
  #head$filter.lat <- 0
  #head$filter.lat[is.na(head$Latitude)] <- 1
  
###-### study area
  head.sf$filter.star <- 0
  head.sf$filter.star[!head.sf$star] <- 1
  
###-### YEARS
  
  head.sf$filter.year <- 0
  head.sf$filter.year[is.na(head.sf$year)] <- 1
  
  #head$filter.month <- 0
  #head$filter.month[is.na(head$month)] <- 1

  #head$filter.day <- 0
  #head$filter.day[is.na(head$day)] <- 1  
  
  
###-### DUPLICATES
  head.sf$filter.dup <- 0
  head.sf$filter.dup[head.sf$Dataset=="Denmark Naturdata"] <- 1
  head.sf$filter.dup[!head.sf$RS_DUPL==""] <- 1
  
###-### MANIPULATION
  
  ### filter out manipulated plots
  head.sf$filter.manip <- 0
  head.sf$filter.manip[head.sf$RS_PROJTYP %in% c("permanent (man)", "Permanent (man)") ] <- 1

  #head.sf$filter.foreva <- 0
  #head.sf$filter.foreva[head.sf$`For.EVA.(Y/N)`=="N" & !head.sf$`ReSurvey.plot.(Y/N)`=="Y"] <- 1
  
###-### RESURVEY
  #head.sf$filter.resobs1 <- 0
  #head.sf$filter.resobs1[head.sf$RS_TIME >1] <- 1

###-### EUNIS
  
  ### non-target habitats
  head.sf$filter.eunis <- 0
  head.sf$filter.eunis[head.sf$eunis.lvl1 %in% c("P")] <- 1 ### remove aquatic habitats
  head.sf$filter.eunis[head.sf$eunis.lvl2 %in% c("XX", "YY", "ZZ")] <- 1 ### remove unclear habiats

###-### LOCATION UNCERTAINTY
  head.sf$filter.location <- 0
  head.sf$filter.location[head.sf$`Location.uncertainty.(m)`>=50000] <- 1
  
###-### YEAR
  
  ### redo for filter breaks (i.e., 20 yr intervals now)
  cutoff.year <- 1950
  head.sf$filter.studyperiod <- 0
  head.sf$filter.studyperiod[head.sf$year<cutoff.year] <- 1
  
###-### AREA
  #head.sf$filter.area <- 0
  #head.sf$filter.area[is.na(head.sf$area)] <- 1

###-### ALL FILTERS
  head.sf <- head.sf %>%
    mutate(filter.meta = if_any(starts_with("filter."), ~ . == 1)+0)

```

## filtering table

```{r}

head.filter <- head.sf %>%
  st_drop_geometry() %>% 
  select(starts_with("filter.")) %>% 
  colSums() %>% 
  as.data.frame()%>%
  rownames_to_column(var = "filter.name") %>%
  mutate(filter_name = gsub("filter\\.", "", filter.name)) %>%
  rename(plots = ".")

head.filter$filter.details <- revalue(head.filter$filter.name, replace=c("filter.lon"="coordinates missing",
                                                                         "filter.resobs1"="ReSurvey observation >1",
                                                                         "filter.star" = "outside study area",
                                                                         "filter.hab"="surface waters",
                                                                         "filter.year"="year missing",
                                                                         "filter.studyperiod"="outside study period",
                                                                         "filter.manip"="manipulated",
                                                                         "filter.foreva"="manipulated (EVA)",
                                                                         "filter.location"="location uncertain\n>50km",
                                                                         "filter.meta"="all filters applied",
                                                                         "filter.eunis"="non-target habitat",
                                                                         "filter.area"="area missing",
                                                                         "filter.dup" ="potential duplicate"))

head.filter$filter.details <- factor(head.filter$filter.details, 
                              levels=head.filter$filter.details[order(head.filter$plots)])

head.filter$final <- "no"
head.filter$final[head.filter$filter.details=="all filters applied"] <- "yes"

### PLOT
plot.head.filter <- ggplot(head.filter) + my.theme +
  geom_bar(aes(y=filter.details, x=plots, fill=final), stat="identity") +
  scale_x_continuous(expand=expansion(mult=c(0,0.1)))+
  labs(title=paste0(nrow(head.sf)-head.filter$plots[head.filter$filter_name=="meta"], " plots remain"),
       x="number of plots",
       y="")+ 
  geom_label(aes(y=filter.details, x=10000, label=plots, hjust=0, fontface="bold")) +
  theme(legend.position="none",
        axis.text.y=element_text(size=10, hjust=0, face="bold"))

ggsave(filename=paste0(path.out,"/head_filterplot.png"),
       plot=plot.head.filter,
       width=10, height=15, unit="cm",
       dpi=900,
       bg="white")

```

## filter head

```{r}

head.sf <- head.sf[head.sf$filter.meta==0,]

### create factors here
head.sf$yr.interval.pretty <- paste0(head.sf$yr.interval, "\n-\n", head.sf$yr.interval+yr.int.length-1)
head.sf$yr.interval.pretty.1line <- gsub("\n-\n", "-", head.sf$yr.interval.pretty)

table(head.sf$eunis.method)

```


# PLOTS

```{r}

plot <- fread(paste0(path.data,"01_EVA/200_Motivate20250415_taxa_concepts_unified.csv"), 
         header=T,
         sep = "\t")

colnames(plot) <- gsub(" ", ".", colnames(plot))

### to be implemented AFTER this server run for better compatibility and file structure ;-)
plot$sPlot.concept.unified <- gsub(" ", "_", plot$sPlot.concept.unified)

### small editing column to create a "presence" column (important later!!!)
plot$occ <- 1

```

## filtering plots

```{r}

### enforce referential integrity between plot and (filtered) header as well as vice versa
plot <- plot[plot$PlotObservationID %in% head.sf$PlotObservationID,]
head.sf <- head.sf[head.sf$PlotObservationID %in% unique(plot$PlotObservationID),]

### remove non-vascular plants
plot <- plot[plot$Taxon.group.ID==1,]

### transform layers into 1
plot <- plot %>% 
  group_by(PlotObservationID, sPlot.concept.unified) %>% 
  dplyr::summarise(n=sum(occ)) %>% 
  mutate(occ=1)

### transform to data table for faster handling (above operations run slower with data.table)
plot <- as.data.table(plot)

```

## species richness

```{r}

sprich <- plot %>% group_by(PlotObservationID) %>% dplyr::summarise(nspec=length(unique(sPlot.concept.unified)))

head.sf <- left_join(head.sf, sprich, by="PlotObservationID")

rm(sprich)

```

# FIGURES I

## raster study area

```{r}
### make a metric bbox
bbox <- st_bbox(star.tf)

### set resolution (in km)
res <- 100

### create raster
my.rast <- rast(
  xmin = bbox["xmin"],
  xmax = bbox["xmax"],
  ymin = bbox["ymin"],
  ymax = bbox["ymax"],
  resolution = c(res*1000, res*1000), # Resolution in meters
  crs = "EPSG:3035" # Match the CRS of the sf object
)

values(my.rast) <- 0

my.rast <- mask(my.rast, star.tf)


```

## find raster cell for every plot

```{r}

head.sf$cell <- cells(my.rast, vect(head.sf))[,2]

head.sf <- head.sf[!is.na(head.sf$cell),] ### remove 46 points that are on the border of the study area

```

# SAMPLING

## density

### overall

```{r}

cell.cts <- data.frame(table(head.sf$cell))
colnames(cell.cts) <- c("cell", "count")
cell.cts$cell <- as.numeric(as.character(cell.cts$cell))

dens.rast <- my.rast

dens.rast[cell.cts$cell] <- cell.cts$count

dens.rast <- mask(dens.rast, star.tf)

dens.df <- as.data.frame(dens.rast, xy=T)

colnames(dens.df) <- c("x", "y", "plots")

range(dens.df$plots)

br.plots <- c(0, 1, 10, 50, 100, 500, 1000, 5000, Inf)
lab.plots <- c("0", "1-9", "10-49", "50-99", "100-499", "500-999", "1000-4999", " ≥5000")

dens.df$plots.cut <- cut(dens.df$plots,
                         breaks=br.plots,
                         labels=lab.plots,
                         include.lowest=T,
                         right=F)

table(dens.df$plots.cut)

pal.plots <- colorRampPalette(c("#a1d99b", "#00441b"))
col.plots <- c("#fde0dd", pal.plots(length(lab.plots)-1))
names(col.plots) <- lab.plots

plot.dens <- ggplot()+my.theme+
  geom_raster(dens.df, mapping=aes(x=x,y=y, fill=plots.cut))+
  scale_fill_manual(name="plots per \n50x50 km cell", values=col.plots)+
  guides(fill=guide_legend(byrow=T))+
  geom_sf(star.tf, mapping=aes(), fill="transparent", color="black")+
  labs(x="",
       y="")+
  theme(legend.position="bottom")

ggsave(filename=paste0(path.out,"/dens_total.png"),
       plot=plot.dens,
       width=wide, unit="cm",
       dpi=900,
       bg="white")

```

### per decade

```{r}

dens.blank <- dens.rast
dens.blank[] <- 0

dens.dec.list <- list()
dens.rast.list <- list()

for (i in 1:length(unique(head.sf$yr.interval.pretty))){

  d <- sort(unique(head.sf$yr.interval.pretty))[i]
  
  ### progress bar
  print(paste0("Procesing ", d))
  
  ### subset head.sf
  head.sf.dec <- head.sf[head.sf$yr.interval.pretty==d,]
  
  ### extract plot counts, mask & convert to df
  dens.dec <- rasterize(vect(head.sf.dec), dens.blank, fun="count", background=0)
  dens.dec <- mask(dens.dec, star.tf)
  
  dens.rast.list[[i]] <- dens.dec
  
  
  dens.dec.df <- as.data.frame(dens.dec, na.rm=T, xy=T)
  

  ### cut df for easier plotting
  dens.dec.df$plots.cut <- cut(dens.dec.df$count,
                               breaks=br.plots,
                               labels=lab.plots,
                               include.lowest=T)
  
  n.dec <- sum(dens.dec.df$count)
  
  ### add a column for the decade
  dens.dec.df$yr.interval.pretty <- d
  
  ### add a binary column
  dens.dec.df$plots.bin <- dens.dec.df$count>0
  
  
  ### save to list
  dens.dec.list[[i]] <- dens.dec.df
  names(dens.dec.list)[i] <- d
  
}

dens.dec.df.all <- bind_rows(dens.dec.list)

plot.dens.dec.all <- ggplot(dens.dec.df.all)+my.theme+
  geom_raster(dens.dec.df.all, mapping=aes(x=x, y=y, fill=plots.cut))+
  scale_fill_manual(name="plots per \n50x50 km cell", values=col.plots, na.value="white")+
  geom_sf(star.tf, mapping=aes(), fill="transparent", color="black")+
  labs(x="",
       y="")+
  theme(legend.position="none",
        legend.direction="horizontal",
        legend.byrow = T,
        axis.text=element_blank(),
        axis.ticks=element_blank())+
  facet_wrap(~yr.interval.pretty, nrow=2)

ggsave(filename=paste0(path.out,"/dens_dec.png"),
       plot=plot.dens.dec.all,
       width=wide, unit="cm",
       dpi=900,
       bg="white")

```

### by country

```{r}

### create a frequency table per decade
  country.dec <- data.frame(table(head.sf$name, head.sf$yr.interval.pretty))
  colnames(country.dec) <- c("country", "yr.interval.pretty", "plots")
  
  country.dec$country <- as.character(country.dec$country)
  #country.dec$yr.interval.pretty <- as.character(country.dec$yr.interval)
  
  ### create an area table and join with frequency table
    star.area <- st_drop_geometry(star.tf[,c("name", "area.km2")])
  
    country.dec <- left_join(country.dec, star.area, by=c("country"="name"))
  
  ### calculate plots per area (multiply with 1000 for better breaks)
    country.dec$plots.km2x1000 <- country.dec$plots/country.dec$area.km2*1000


### make a helper table
  ctry.helper <- country.dec %>% group_by(country) %>% dplyr::summarise(plots.tot=sum(plots))
  ctry.helper <- left_join(ctry.helper, star.area, by=c("country"="name"))
  
  ctry.helper$plots.km2x1000 <- ctry.helper$plots.tot/ctry.helper$area.km2*1000
  
  ctry.helper$country.fac <- factor(ctry.helper$country, levels=ctry.helper$country[order(ctry.helper$plots.km2x1000)], ordered=T)
  
  ctry.helper$plots.km2x1000.rd <- round(ctry.helper$plots.km2x1000)
  
  ctry.helper$ctry.lab <- paste0(ctry.helper$country, "=", ctry.helper$plots.km2x1000.rd)
  
### cut for easier plotting with assigned color vector
br.ctryct <- c(-Inf, 0, 1, 2, 5, 10, 20, 50, 100, Inf) 
lab.ctryct <- c("0", ">0-1", ">1-2", "2-4", "5-9", "10-19", "20-49", "50-99", "100+")

country.dec$plots.km2x1000.cut <- cut(country.dec$plots.km2x1000, breaks=br.ctryct, labels=lab.ctryct, include.lowest=T)

col.ctry <- c("grey75", "#8e0152", "#c51b7d", "#de77ae", "#f1b6da", "#b8e186", "#7fbc41", "#4d9221",  "#276419")
names(col.ctry) <- lab.ctryct

country.dec <- left_join(country.dec, ctry.helper, by="country")

country.dec$ctry.lab.fac <- factor(country.dec$ctry.lab, levels=ctry.helper$ctry.lab[order(ctry.helper$plots.km2x1000)], ordered=T)

### final plot
plot.ctry.dec <- ggplot(country.dec, aes(x=yr.interval.pretty, y=ctry.lab.fac, fill=plots.km2x1000.cut))+my.theme+
  geom_tile(width=0.9, height=0.9)+
  scale_fill_manual(name= "plots/decade\n(per 1000 km²)", values=col.ctry)+
  labs(x="",
       y="country (plots/1000km²)")+
  theme(legend.position="right",
        axis.text.y=element_text(hjust=0))

ggsave(filename=paste0(path.out,"/country_dec.png"),
       plot=plot.ctry.dec,
       width=wide, unit="cm",
       dpi=900,
       bg="white")

```

### by biogeo region

```{r}

### create a frequency table per decade
  biogeo.dec <- data.frame(table(head.sf$biogeo.region, head.sf$yr.interval.pretty))
  colnames(biogeo.dec) <- c("biogeo.region", "yr.interval.pretty", "plots")
  
  biogeo.dec$biogeo.region <- as.character(biogeo.dec$biogeo.region)
  #biogeo.dec$yr.interval <- as.character(biogeo.dec$yr.interval)

  ### create an area table and join with frequency table
    biogeo.area <- st_drop_geometry(biogeo.tf[,c("short_name", "area.km2")])
  
    biogeo.dec <- left_join(biogeo.dec, biogeo.area, by=c("biogeo.region"="short_name"))
    
  ### calculate plots per area (multiply with 1000 for better breaks)
    biogeo.dec$plots.km2x1000 <- biogeo.dec$plots/biogeo.dec$area.km2*1000


### make a helper table
  biogeo.helper <- biogeo.dec %>% group_by(biogeo.region) %>% dplyr::summarise(plots.tot=sum(plots))
  biogeo.helper <- left_join(biogeo.helper, biogeo.area, by=c("biogeo.region"="short_name"))
  
  biogeo.helper$plots.km2x1000 <- biogeo.helper$plots.tot/biogeo.helper$area.km2*1000
  
  colnames(biogeo.helper)
  
  
  biogeo.helper$biogeo.region.fac <- factor(biogeo.helper$biogeo.region, levels=biogeo.helper$biogeo.region[order(biogeo.helper$plots.km2x1000)], ordered=T)
  
  biogeo.helper$plots.km2x1000.rd <- round(biogeo.helper$plots.km2x1000)
  
  biogeo.helper$biogeo.region.lab <- paste0(biogeo.helper$biogeo.region, "=", biogeo.helper$plots.km2x1000.rd)
  
### cut for easier plotting with assigned color vector
br.ctryct <- c(-Inf, 0, 1, 2, 5, 10, 20, 50, 100, Inf) 
lab.ctryct <- c("0", ">0-1", ">1-2", ">2-4", ">5-9", ">10-19", ">20-49", ">50-99", ">100")

biogeo.dec$plots.km2x1000.cut <- cut(biogeo.dec$plots.km2x1000, breaks=br.ctryct, labels=lab.ctryct, include.lowest=T)

col.ctry <- c("grey75", "#8e0152", "#c51b7d", "#de77ae", "#f1b6da", "#b8e186", "#7fbc41", "#4d9221",  "#276419")
names(col.ctry) <- lab.ctryct

biogeo.dec <- left_join(biogeo.dec, biogeo.helper, by="biogeo.region")

biogeo.dec$biogeo.lab.fac <- factor(biogeo.dec$biogeo.region.lab, levels=biogeo.helper$biogeo.region.lab[order(biogeo.helper$plots.km2x1000)], ordered=T)

### final plot
plot.biogeo.region.dec <- ggplot(biogeo.dec, aes(x=yr.interval.pretty, y=biogeo.lab.fac, fill=plots.km2x1000.cut))+my.theme+
  geom_tile(width=0.9, height=0.9)+
  scale_fill_manual(name= "plots\n(per 1000 km²)", values=col.ctry)+
  labs(x="",
       y="biogeographic region (plots/1000 km²)")+
  theme(legend.position="right",
        axis.text.y=element_text(hjust=0))

ggsave(filename=paste0(path.out,"/biogeoregion_dec.png"),
       plot=plot.biogeo.region.dec,
       width=wide, unit="cm",
       dpi=900,
       bg="white")


```

## habitat

### habitats lvl1

```{r}

### helper table
hab.helper <- head.sf %>% 
  st_drop_geometry() %>% 
  group_by(yr.interval.pretty.1line) %>% 
  dplyr::summarise(n.tot=n())

hab.helper$y=90

hab.helper$label.1k <- paste0(round(hab.helper$n.tot/1000), "k")

### plot table
habitat <- head.sf %>% 
  st_drop_geometry() %>% 
  group_by(yr.interval.pretty.1line, eunis.lvl1.name) %>% 
  dplyr::summarise(n=n()) %>% 
  left_join(hab.helper, by="yr.interval.pretty.1line") %>% 
  mutate(hab.rel= (n/n.tot)*100,
         eunis.lvl1.name=tolower(eunis.lvl1.name),
         eunis.lvl1.name=recode(eunis.lvl1.name, 
                                 "forests and other wooded land"="forests",
                                 "inland habitats with no or little soil and mostly with sparse vegetation"="inland sparsely vegetated",
                                 "grasslands and lands dominated by forbs, mosses or lichens"="grasslands",
                                 "vegetated man-made habitats" = "anthropogenic",
                                 "coastal habitats" = "coastal",
                                 "marine habitats" = "marine",
                                .default = eunis.lvl1.name
                                           ),
         eunis.lvl1.name=factor(eunis.lvl1.name, levels=c(  "forests",
                                                            "grasslands",
                                                            "heathlands, scrub and tundra",
                                                            "inland sparsely vegetated",
                                                            "anthropogenic",
                                                            "marine",
                                                            "coastal",
                                                            "wetlands",
                                                            "class not possible",
                                                            "class failed",  
                                                            "class ambiguous"
                                                          ), ordered=T ))


### PLOT
plot.habdec.lvl1 <- ggplot()+my.theme+
  geom_bar(habitat, mapping=aes(x=yr.interval.pretty.1line, y=hab.rel, fill=eunis.lvl1.name), stat="identity", width=0.66)+
  scale_fill_manual(name="", values=col.hab)+
  guides(fill=guide_legend(nrow=3))+
  scale_y_continuous(expand=expansion(mult=c(0,0)), labels=seq(0,100,10), breaks=seq(0,100,10))+
  labs(x="",
       y="EUNIS level 1 (%)")+
  theme(
  legend.key.size = unit(0.4, "cm"),
  legend.text = element_text(size = 8),
  legend.margin = margin(-20, 0, 0, 0),
  legend.spacing = unit(0, "cm"),
  plot.margin=margin(2,2,0,0),
  axis.title.x = element_text(margin = margin(t = 0, r = 0, b = 0, l = 0)),
  axis.text.x = element_text(margin = margin(t = 0)) )+
  geom_label(data=hab.helper, aes(x=yr.interval.pretty.1line, y=y, label=label.1k), size=4, fontface="bold", color="white", label.size=0, fill=NA)

```

### habitats lvl 2

```{r}

eunis.legend$name <- eunis.legend$`EUNIS-2023 habitat name`

habitat2 <- head.sf %>%
     st_drop_geometry() %>% 
     group_by(yr.interval.pretty.1line, eunis.lvl2) %>% 
     dplyr::summarise(n=n()) %>% 
     left_join(hab.helper, by="yr.interval.pretty.1line") %>% 
     mutate(hab.rel= (n/n.tot)*100,
         eunis.lvl1=substr(eunis.lvl2, 1,1)) %>% 
     left_join(eunis.legend, by=c("eunis.lvl1"="EUNIS 2023 code")) %>% 
     mutate(hab.rel= (n/n.tot)*100,
            name=tolower(name),
            name=recode(name, 
                        "forests and other wooded land"="forests",
                        "inland habitats with no or little soil and mostly with sparse vegetation"="inland sparsely vegetated",
                        "grasslands and lands dominated by forbs, mosses or lichens"="grasslands",
                        "vegetated man-made habitats" = "anthropogenic",
                        "coastal habitats" = "coastal",
                        "marine habitats" = "marine",
                                .default = name
                                           ),
         name=factor(name, levels=c("forests",
                                    "grasslands",
                                    "heathlands, scrub and tundra",
                                    "inland sparsely vegetated",
                                    "anthropogenic",
                                    "marine",
                                    "coastal",
                                    "wetlands",
                                    "class not possible",
                                    "class failed",  
                                    "class ambiguous"), ordered=T ))
    



### set levels for grid
ncol <- ceiling(sqrt(length(unique(habitat2$yr.interval.pretty.1line))))


### plot level 2
plot.habdec.lvl2 <- ggplot(habitat2)+my.theme+
     geom_bar(aes(x=eunis.lvl2, y=hab.rel, fill=name), stat="identity", width=0.66)+
     scale_fill_manual(name="", values=col.hab)+
     scale_y_continuous(expand=expansion(mult=c(0,0)))+
     labs(x="",
          y="EUNIS level 2 (%)")+
     theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5, size=6, margin=margin(0,0,0,0)),
           legend.position="none",
           plot.margin=margin(2,2,0,0),
           axis.title.x = element_text(margin = margin(t = 0, r = 0, b = 0, l = 0)))+
     facet_wrap(.~yr.interval.pretty.1line, scales="free", ncol=2)

```

### habitat grid

```{r}

plot.habdec.grid1x2 <- plot_grid(plot.habdec.lvl1, plot.habdec.lvl2,
                                 nrow=2, align="hv", axis="lt", labels=c("A", "B"),
                                 rel_heights=c(0.6,1))

ggsave(filename=paste0(path.out,"/habitat_dec_grid1x2.png"),
       plot=plot.habdec.grid1x2,
       width=15.8, height=21, unit="cm",
       dpi=900,
       bg="white")

```

## raster centroids

exclude cells that are in ocean exclude cells that have no data (make a filter column)

```{r}

centroids <- as.data.frame(xyFromCell(my.rast, 1:ncell(my.rast)), na.rm=T)
centroids$cell <- 1:ncell(my.rast)

### look at only the cells you have any data on
centroids$cell.data <- !centroids$cell %in% head.sf$cell

### remove the cells in the ocean
centroids$ocean <- centroids$cell %in% which(is.na(values(my.rast)))

```

## biogeo region per cell

```{r}

### extract the coverage of every area per raster cell
biogeo.list <- exact_extract(my.rast, biogeo.tf, include_cell=T, include_xy=T)

### add a naming column for every region 
for (i in seq_along(biogeo.list)) {
  if (nrow(biogeo.list[[i]]) > 0) {
    biogeo.list[[i]]$biogeo.region <- biogeo.tf$short_name[i]
  }
}

### bind rows together
biogeo.df <- rbindlist(biogeo.list, fill=T)

### choose individual area contributions
biogeo.area <- biogeo.df %>% 
     pivot_wider(id_cols=cell, names_from=biogeo.region, values_from=coverage_fraction, values_fill = 0) %>% 
     rowwise() %>%
     mutate(across(-cell, ~ .x / sum(c_across(-cell)))) %>%
     mutate(row_total = sum(c_across(-cell))) %>%
     ungroup()
     
     ### rejoin with centroids
     centroids <- left_join(centroids, biogeo.area, by="cell")

### choose maximum
biogeo.max <- biogeo.df %>% 
     group_by(cell) %>% 
     slice_max(coverage_fraction, with_ties=F) %>% 
     mutate(biogeo.region.bymax=biogeo.region) %>% 
     select(-biogeo.region, -x, -y)

     ### rejoin with centroids
     centroids <- left_join(centroids, biogeo.max, by="cell")

colnames(centroids)

```

## sampling effort per site per timestep

```{r}

sample.site <- head.sf %>% 
  st_drop_geometry() %>% 
  group_by(cell, yr.interval.pretty) %>% 
  dplyr::summarise(n.sample=n(), .groups="drop") %>% 
  complete(cell, yr.interval.pretty, fill=list(n.sample=0))

```

# TAXA

## make taxa table

```{r}

### cut n into scale
  br.n <- c(-Inf,50, 100, 200, 500, 1000, 2000, 5000, 10000, Inf)
  lab.n <- c("<50",
            "50\n-\n99", 
            "100\n-\n199", 
            "200\n-\n499", 
            "500\n-\n999", 
            "1000\n-\n1999", 
            "2000\n-\n4999", 
            "5000\n-\n9999", 
            "10000+")

### create taxonomy table from filtered plots
  taxa <- plot %>% 
    left_join(st_drop_geometry(head.sf), by="PlotObservationID")  %>% 
    group_by(sPlot.concept.unified) %>% 
    dplyr::summarise(n=n(),
                     n.rel=n/nrow(plot)*100,
                     n.cut=cut(n, breaks=br.n, labels=lab.n, right=F),
                     year.first=min(year),
                     year.last=max(year),
                     year.span=year.last-year.first,
                     yr.interval.first=min(yr.interval), 
                     yr.interval.last=max(yr.interval),
                     yr.interval.span=length(unique(yr.interval)))%>% 
    mutate(genus.only = !str_detect(sPlot.concept.unified, "_"))

### helber object for total taxa  
  n.taxa.tot <- nrow(taxa)
  
```

## temporal distribution

```{r}

### modelling certainty
  pal.mod <- colorRampPalette(c("#ffffe5", "#00441b"))
  col.mod <- pal.mod(length(lab.n))
  names(col.mod) <- lab.n

### make a helper table
  taxa.helper <- taxa %>%
    group_by(n.cut) %>% 
    dplyr::summarise(n=n(),
                     n.rel=n/n.taxa.tot*100) %>% 
    mutate(r.sum=rev(cumsum(rev(n))))
  
### PLOTS
  
  ### taxa for modelling
    plot.taxa <- ggplot(taxa.helper)+my.theme+
      geom_bar(aes(x=n.cut, fill=n.cut, y=n.rel), color="black", stat="identity", width=0.66)+
      scale_fill_manual("",values=col.mod)+
      scale_y_continuous(expand=expansion(mult=c(0,0)), breaks=seq(0,100,10), limits=c(0,60))+
      scale_x_discrete(expand=expansion(add=c(0.5,0.5)))+
      labs(x="total occurrences",
           y="taxa (%)")+
      theme(legend.position="bottom",
            legend.text.position="bottom",
            legend.text=element_text(size=4))+
      guides(fill=guide_legend(nrow=1))+
      geom_label(mapping=aes(x=n.cut, y=20, label=n), angle=90, vjust=0.5, size=3)+
      geom_segment(aes(x=length(col.mod), xend=0.125, y=40, yend=40),
                   arrow=arrow(length=unit(0.33, "cm"), type="closed"),
                   color="black")+
      geom_label(mapping=aes(x=n.cut, y=40, label=r.sum), color="black", size=3)+
      geom_vline(xintercept=2.5, color="black")+
      annotate("text", x=3.5, y=45, label="cumulative sums",
               color="black")
    
  ### records for clipping
    plot.decfirst <- ggplot(taxa)+my.theme+
      geom_bar(aes(x=yr.interval.first, fill=n.cut),  color="black",width=7.5)+
      scale_fill_manual(values=col.mod)+
      scale_y_continuous(expand=expansion(mult=c(0,0)))+
      scale_x_continuous(expand=expansion(mult=c(0.01,0.01)), breaks=unique(taxa$yr.interval.first))+
      labs(x="first timestep",
           y="number of taxa")+
      theme(legend.position="none")
    
    plot.declast <- ggplot(taxa)+my.theme+
      geom_bar(aes(x=yr.interval.last, fill=n.cut), color="black", width=7.5)+
      scale_fill_manual(values=col.mod)+
      scale_y_continuous(expand=expansion(mult=c(0,0)))+
      scale_x_continuous(expand=expansion(mult=c(0.01,0.01)), breaks=unique(taxa$yr.interval.first))+
      labs(x="last timestep",
           y="number of taxa")+
      theme(legend.position="none")
  
    plot.decspan <- ggplot(taxa)+my.theme+
      geom_bar(aes(x=yr.interval.span, fill=n.cut), color="black", width=0.66)+
      scale_fill_manual(values=col.mod)+
      scale_x_continuous(expand=expansion(mult=c(0.01,0.01)), breaks=seq(0,length(unique(taxa$yr.interval.span)),1))+
      scale_y_continuous(expand=expansion(mult=c(0,0.1)))+
      labs(x="number of timesteps",
           y="number of taxa")+
      theme(legend.position="none")
  
    
  lom.decade <- as.matrix(rbind(c(1,1,1,1),
                                c(1,1,1,1),
                                c(1,1,1,1),
                                c(2,2,3,3),
                                c(2,2,3,3),
                                c(4,4,4,4),
                                c(4,4,4,4)
                                ))
  
  grid.taxa <- grid.arrange(plot.taxa+theme(legend.position="none"),
                              plot.decfirst, plot.declast,
                              plot.decspan,
                              layout_matrix=lom.decade)  

### export
ggsave(filename=paste0(path.out,"/taxa.png"),
       plot=grid.taxa,
       width=wide, 
       height=high,
       unit="cm",
       dpi=900,
       bg="white")


```

## biogeo origin

mail to Irena Axmanova on August 11th 2025 probably November
currently using temporary dataset provided by SK
status = status for all of Europe (?), might also include areas outside of study area...
many species missing, assessments for Austria are weird...

```{r}

origin <- read_excel(paste0(path.data,"05_TAXA/02_biogeo/MOTIVATE_species_for_native_non-native_classification-IA.xlsx"))

colnames(origin) <- tolower(gsub(" ", ".", colnames(origin), fixed=T))

table(origin$origineurope[origin$taxonomy=="AU-Austria"])

### group by species and unify taxa
origin.grp <- origin %>% 
  group_by(splot.concept.unified) %>% 
  dplyr::summarise(status=paste(unique(origineurope),collapse=", "),
                   n.status=length(unique(origineurope)),
                   n.tax=length(unique(taxonomy))
                   ) %>% 
  mutate(status=revalue(status, replace=c("neo"="neophyte",
                                          "arch"="archaeophyte")),
         splot.join=gsub(" ", "_", splot.concept.unified, fixed=T))

taxa <- left_join(taxa, origin.grp, by=c("sPlot.concept.unified"="splot.join"))

table(origin.grp$status)


write.table(taxa, file="./taxtbl.csv", row.names=F, sep=";", na="")

```

## Red List

```{r}

redlist <- read.csv(paste0(path.data,"05_TAXA/03_redlist/redlist_MOTIVATE_subset.csv"), sep="\t")

table(redlist$Threat, redlist$Country)
table(redlist$Redlist_cat, redlist$Country)

table(redlist$redlist_source)

table(redlist$V1)
length(unique(redlist$V1))
nrow(redlist)-sum(is.na(redlist$V1))

nrow(redlist)
length(unique(redlist$sPlot.concept.unified))

unique(redlist$Redlist_cat)

redlist.test <- redlist %>% 
  filter(redlist_source %in% c("rl_NationalRL")) %>% 
  group_by(sPlot.concept.unified, Country) %>% 
  dplyr::summarise(threat.tot=paste(unique(Threat), collapse=", "),
                   n.threat=length(unique(Threat)))

sum(redlist.test$n.threat>1)

### national red lists
redlist$threat.fac <- factor(redlist$Threat, levels=c("not_evaluated_dd", "not_threatened","threatened","extinct"), ordered=T)

redlist.nat <- redlist %>% 
  filter(redlist_source %in% c("rl_NationalRL")) %>% 
  select(sPlot.concept.unified, Country, threat.fac) %>% 
  pivot_wider(names_from=Country, 
              values_from=threat.fac, 
              values_fill=NA, 
              values_fn=~ if (all(is.na(.x))) NA_real_ else max(.x, na.rm = TRUE))

unique(redlist$Threat)



```

## Traits

## EIVT

```{r}


```


## EIVE

```{r}

eive <- read_excel(paste0(path.data,"./05_TAXA/04_eive/vegetation_classification_and_survey-004-007-g008.xlsx"), sheet="mainTable")


sum(taxa$sPlot.concept.unified %in% gsub(" ", "_",eive$TaxonConcept, fixed=T))/n.taxa.tot*100


```

# RUN ALL

```{r}


```

# EXPORT

write head, plots and centroids to the Data Input folder

```{r}

length(unique(head.sf$PlotObservationID))
length(unique(plot$PlotObservationID))

plot <- plot[plot$PlotObservationID %in% head.sf$PlotObservationID,]
head.sf <- head.sf[head.sf$PlotObservationID %in% unique(plot$PlotObservationID),]

### head
head.out <- st_drop_geometry(head.sf)

sum(is.na(head.out$nspec))

fwrite(head.out, file=paste0("../02_FORMATSERVER/01_INPUT/head.csv"), sep=";", dec=",", row.names=F)

### plot
fwrite(plot, file=paste0("../02_FORMATSERVER/01_INPUT/plot.csv"), sep=";", dec=",", row.names=F)

### export to next step
fwrite(centroids, file="../02_FORMATSERVER/01_INPUT/centroids.csv", sep=";", dec=",", row.names=F)

### taxa
fwrite(taxa, file="../04_RWD_Model_Output/01_INPUT/taxa.csv", sep=";", dec=",", row.names=F)

### write raster to input of model output
writeRaster(my.rast, filename=paste0("../04_RWD_Model_Output/01_INPUT/myrast_", res, ".tif"), overwrite=T)

```

# ARCHIVE

## eunis lvl1

```{r}

### read in EUNIS legend (edited by MG)
     eunis<- as.data.table(read_excel("04_EUNIS/EUNIS-ESy-2022-10-16_legendfrom ESYfile.xlsx", sheet = "MG_edit"))

### create a duplicate eunis column for checking
     head$EUNIS <- head$Expert.system
    
### add a column for blanks and unidentified/mixed habitats
     head$EUNIS[head$EUNIS %in% c("")] <- "X" ### replace "" with X for unclassified
     head$EUNIS[head$EUNIS %in% c("~")] <- "Y" ### replace "~" with Y for unclassifiable
     
### extract EUNIS assessment
     head.esy <- head[, .(EUNIS = unlist(tstrsplit(EUNIS, ","))), by = PlotObservationID]  # Separate rows by EUNIS
  
  ### level 1
     head.esy$lvl1 <- substr(head.esy$EUNIS, 1,1)
     
     ### head.esy$lvl1 <- gsub("[^A-Za-z]", "", head.esy$EUNIS) ### to get more precise levels
     ### head.esy$lvl2 <- substr(head.esy$EUNIS, 1,3) ### to extract level 2
  
  ### combine with EUNIS legend
  
     head.esy <- merge(head.esy, eunis, by.x = "lvl1", by.y = "EUNIS.code", all.x = TRUE) # Left join with eunis
     
     head.esy <- head.esy[, .(
     eunis.countlvl1 = uniqueN(lvl1),  # Count unique lvl1
     eunis.lvl1 = paste(sort(unique(lvl1)), collapse = ", ") # Concatenate unique names
     ), by = PlotObservationID]

# Add final column based on condition
     head.esy[, eunis.final := ifelse(eunis.countlvl1 > 1, "Z", eunis.lvl1)]

     head.esy <- left_join(head.esy, eunis, by=c("eunis.final"="EUNIS.code"))

### rejoin with head
     head <- left_join(head, head.esy, by="PlotObservationID")
  
     head$eunis.name <- head$EUNIS.name

### remove unnecessary objects
     rm(list=c("eunis", "head.esy"))
  
```

## eunis lvl 2

```{r}

### read in EUNIS legend (edited by MG)
     eunis<- as.data.table(read_excel("04_EUNIS/EUNIS-ESy-2022-10-16_legendfrom ESYfile.xlsx", sheet = "List1"))

     eunis$EUNIS.code <- gsub("!", "", eunis$`EUNIS.code`, fixed=T) ### remove the !!! for clear codes
     eunis$EUNIS.name <- gsub("[0-9]", "", eunis$EUNIS.name) ### remove the odd numbering for coniferous forests

     eunis <- eunis[,c("EUNIS.code", "EUNIS.name")] ### reduce dt to necessary columns for easier handling
     
     eunis <- unique(eunis) ### subset to unique results
     
     
     
     
### edit header
     head$EUNIS <- head$Expert.system ### make new column for cleaner edits
     
     ### add a column for blanks and unidentified/mixed habitats
     head$EUNIS[head$EUNIS %in% c("")] <- "XXX" ### replace "" with X for unclassified
     head$EUNIS[head$EUNIS %in% c("~")] <- "YYY" ### replace "~" with Y for unclassifiable
     
### separate header rows
     head.esy <- head[, .(EUNIS = unlist(tstrsplit(EUNIS, ","))), by = PlotObservationID]  # Separate rows by EUNIS
     
     head.esy$EUNIS <- gsub("!", "", head.esy$EUNIS)
     
     head.esy$lvl2 <- substr(head.esy$EUNIS, 1,3)

     lvl2 <- sort(unique(head.esy$lvl2))
     

### join with descriptions
     head.esy <- left_join(head.esy, eunis, by=c("lvl2"="EUNIS.code"))
     
     head.esy$EUNIS.name[head.esy$lvl2 %in% c("XXX")] <- "unclassified"
     head.esy$EUNIS.name[head.esy$lvl2 %in% c("YYY")] <- "unclassifiable"
     
     
     
```

### WFO match

```{r}
WFO.download()
WFO.remember()

taxa.match <- WFO.match(spec.data=taxa$sPlot.concept.unified, WFO.file="./classification.csv", no.dates=T,
                        counter=250, Fuzzy=0)


length(unique(taxa.match$family))

tail(unique(taxa.match$spec.name.ORIG[is.na(taxa.match$family)]), 141)

fam.excl <- c("Conocephalaceae", "Ptychomitriaceae", "Orthotrichaceae", "Encalyptaceae", "Ricciaceae")

fam.excl %in% taxa.match$family ### check for typos

taxa.match$spec.name.ORIG[taxa.match$family %in% fam.excl] ### simply incorrect assinments

taxa.excl <- c("Lycopdium x dubium", "Lycopodium zeilleri")
```

### investigation of info columns

```{r}

table(head$`ReSurvey.plot.(Y/N)`, useNA="ifany")

table(head$`For.EVA.(Y/N)`, useNA="ifany")

table(head$`Manipulate.(y/n)`, useNA="ifany")

424166+317-(55645+394483)

length(unique(head$Type.of.manipulation))
View(data.frame(table(head$Type.of.manipulation, useNA="ifany")))

table(head$`For.EVA.(Y/N)`, head$`Manipulate.(y/n)`)

20983+34233+429

table(head$RS_DUPL)

nrow(head)-2382870

table(head$RS_DUPL)

unique(head$RS_DUPL) %in% head$RS_CODE

data.frame(table(head$Location.method))

table(head$RS_PROJTYP, head$`ReSurvey.plot.(Y/N)`)

table(head$RS_TIME)

### make plot ID

resurvey <- head[head$`ReSurvey.plot.(Y/N)`=="Y",]

resurvey$ID.resurvey <- paste(resurvey$ReSurvey.project, resurvey$ReSurvey.site, resurvey$ReSurvey.plot, sep="_")

length(unique(resurvey$ReSurvey.project))

length(unique(resurvey$ReSurvey.site))

tail(unique(resurvey$ReSurvey.plot))

View(unique(resurvey[,c("ReSurvey.plot")]))

sample(resurvey$ReSurvey.plot, 100)

length(unique(resurvey$ID.resurvey))

id.df <- data.frame(table(resurvey$ID.resurvey))

colnames(id.df) <- c("ID", "frequency")

table(id.df$frequency)

resurvey$Country

countcountry <- resurvey %>% group_by(ID.resurvey) %>% dplyr::summarise(count=paste0(length(unique(Country))),
                                                                        country=paste(unique(Country), collapse=","))


### plot-header congruency
all(head$PlotObservationID %in% plot$PlotObservationID)

all(plot$PlotObservationID %in% head$PlotObservationID)

sum(!head$PlotObservationID %in% plot$PlotObservationID)

head$notinplot <- !head$PlotObservationID %in% plot$PlotObservationID


table(head$`ReSurvey.plot.(Y/N)`[head$notinplot])

```

```{r}
    
    
 
###-### FRENCH OVERSEAS TERRITORIES  
  
  ### create bounding box excluding France to exclude French oversees territories
    star.bbox <- st_bbox(star[!star$name =="France",], crs=proj.crs)
    star.bbox["ymin"] <- 20
    star.bbox["ymax"] <- 72
  
  ### and now crop to this bbox
    star <- st_crop(star, star.bbox)
    
    star.bla <- st_intersection(star, st_as_sfc(star.bbox))
    star.bla <- star.bla[st_coordinates(st_centroid(star.bla))[,2]<= 72,]

###-### RUSSIA
    
  ### either the shp provided is broken or the fact that it crosses the date line is an issue, so load separately and exclude provinces
    
    rus <- ne_states(country="Russia")
    
    rus <- rus[!rus$name %in% c("Kamchatka", 
                                "Sakha (Yakutia)",  
                                "Maga Buryatdan", 
                                "Buryat", 
                                "Yevrey",  
                                "Khabarovsk" ,
                                "Primor'ye" ,
                                "Chukchi Autonomous Okrug",
                                "Sakhalin"),] ### removes provinces that span the dividing line of the globe
    
    rus <- st_make_valid(rus)
    
    rus.bbox <- st_bbox(rus)
    rus.bbox["xmax"] <- 45
    rus <- st_crop(rus, rus.bbox)
    
    rus <- rus[,c("name", "geometry")]
    
    rus <- st_make_valid(rus)
    
    rus <- st_union(rus)
    
    rus <- st_as_sf(rus)
    
    rus$name <- "Russia"
    colnames(rus)[1] <- "geometry"  
    st_geometry(rus) <- "geometry"

### FINAL RBIND
    
    star <- rbind(rus, star)

    ggplot()+
      geom_sf(star.bla, mapping=aes(fill=name))
```

```{r}

  
  ###-### OLD OLD OLD OLD OLD
  
  
    
### load Russia 
  rus <- ne_states(country="Russia")
  
  rus <- rus[!rus$name %in% c("Kamchatka", 
                              "Sakha (Yakutia)",  
                              "Maga Buryatdan", 
                              "Buryat", 
                              "Yevrey",  
                              "Khabarovsk" ,
                              "Primor'ye" ,
                              "Chukchi Autonomous Okrug",
                              "Sakhalin"),] ### removes provinces that span the dividing line of the globe
  
  ### correct validity of shapefiles
  rus <- st_make_valid(rus)
  
  ### combine into one & rconvert to sf, then match columns
  rus <- rus %>% summarise(name=first(admin),
                        geometry=st_union(geometry))

### rbind back to star
  #star <- rbind(rus, star)
  
  
  
    
### clip back the bounding box to your study area

  star.bbox <- st_as_sfc(st_bbox(star, crs=proj.crs))
   # star.bbox["ymin"] <- 27.5
   # star.bbox["ymax"] <- 60
   # star.bbox["xmin"] <- -10
   # star.bbox["xmax"] <- 27.5
  
  star.wgs <- st_crop(star, star.bbox)
  
  star.tf <- st_transform(star, proj.crs)
  
  ### unify for subsetting
  star.union <- st_union(star.tf)
  star.union <- st_as_sf(st_buffer(star.union, dist=0.1))
  
### plot to check    
  ggplot()+
    geom_sf(star.union, mapping=aes())+my.theme+
    geom_sf(star.tf, mapping=aes(fill=name))+
    scale_x_continuous(expand=expansion(mult=0.05,0))+
    scale_y_continuous(expand=expansion(mult=0,0))
    
    
rm(list=c("island.list", "rus", "uk", "fr", "star.fill", "gr", "it", "esp"))

### add area column
```

### by country

```{r}


### create a frequency table per decade
  country.dec <- data.frame(table(head.sf$name, head.sf$decade))
  colnames(country.dec) <- c("country", "decade", "plots")
  
  country.dec$country <- as.character(country.dec$country)
  country.dec$decade <- as.character(country.dec$decade)
  
  
  ### create an area table and join with frequency table
    star.area <- st_drop_geometry(star.tf[,c("name", "area.km2")])
  
    country.dec <- left_join(country.dec, star.area, by=c("country"="name"))
  
  ### calculate plots per area (multiply with 1000 for better breaks)
    country.dec$plots.km2x1000 <- country.dec$plots/country.dec$area.km2*1000


### make a helper table
  ctry.helper <- country.dec %>% group_by(country) %>% dplyr::summarise(plots.tot=sum(plots))
  ctry.helper <- left_join(ctry.helper, star.area, by=c("country"="name"))
  
  ctry.helper$plots.km2x1000 <- ctry.helper$plots.tot/ctry.helper$area.km2*1000
  
  ctry.helper$country.fac <- factor(ctry.helper$country, levels=ctry.helper$country[order(ctry.helper$plots.km2x1000)], ordered=T)
  
  ctry.helper$plots.1k <- round(ctry.helper$plots.tot/1000)
  ctry.helper$plots.km2x1000.rd <- round(ctry.helper$plots.km2x1000)
  ctry.helper$ctry.lab <- paste0(ctry.helper$country, "=", ctry.helper$plots.km2x1000, "k")


### cut for easier plotting with assigned color vector
br.ctryct <- c(-Inf, 0, 10, 100, 1000, 10000, 100000, Inf)
lab.ctryct <- c("0", "1-10", "11-99", "100-999", "1000-9999", "10000-99999", ">100000")

country.dec$plots.km2x1000

country.dec$plots.cut <- cut(country.dec$plots, breaks=br.ctryct, labels=lab.ctryct, include.lowest=F)

col.ctry <- c("#c51b7d", "#de77ae", "#f1b6da", "#b8e186", "#7fbc41", "#4d9221",  "#276419")
names(col.ctry) <- lab.ctryct

country.dec <- left_join(country.dec, ctry.helper, by="country")

country.dec$ctry.lab.fac <- factor(country.dec$ctry.lab, levels=ctry.helper$ctry.lab[order(ctry.helper$plots.km2x1000)], ordered=T)

### final plot
plot.ctry.dec <- ggplot(country.dec, aes(x=decade, y=ctry.lab.fac, fill=plots.cut))+my.theme+
  geom_tile(width=0.9, height=0.9)+
  scale_fill_manual(name= "plots/decade", values=col.ctry)+
  labs(x="",
       y="country (n in 1k plots)")+
  theme(legend.position="right")

ggsave(filename=paste0(path.out,"/country_dec.png"),
       plot=plot.ctry.dec,
       width=wide, unit="cm",
       bg="white")

```

### alternative eunis

```{r}

### import eunis legend

eunis<- as.data.table(read_excel("04_EUNIS/EUNIS-ESy-2022-10-16_legendfrom ESYfile_raw.xlsx", 
    sheet = "List1"))

eunis.lvl1 <- eunis[eunis$level==1,]

### create a duplicate eunis column for checking
  head$EUNIS <- head$Expert.System

### add a column for blanks and unidentified/mixed habitats
  head$EUNIS[head$EUNIS %in% c("")] <- "X" ### replace "" with X for unclassified
  head$EUNIS[head$EUNIS %in% c("~")] <- "Y" ### replace "~" with Y for unclassifiable

### separate into unique EUNIS types
  head.esy <- head[, .(EUNIS = unlist(tstrsplit(EUNIS, ","))), by = PlotObservationID]  # Separate rows by EUNIS

### remove !! (they denote variations of hab types)
  head.esy$eunis.noexcl <- gsub("!", "", head.esy$EUNIS, fixed=T)
  
### extract letters
  head.esy$letters <- gsub("([A-Za-z]+).*", "\\1", head.esy$eunis.noexcl)

  head.esy <- left_join(head.esy, eunis.lvl1, by=c("letters"="EUNIS code"))
  
  View(head.esy[head.esy$letters=="RR",])

rm(head.esy, eunis, eunis.lvl1)
  
```

## lat/lon cuts

```{r}

plot.lon <- ggplot(head.sf)+my.theme+
  geom_histogram(aes(x=lon.wgs))+
  geom_vline(xintercept=25, color="darkred")+
  scale_x_continuous(breaks=seq(-10,65, 5), expand=expansion(add=c(0,0)))+
  scale_y_continuous(breaks=seq(0,1000000,50000), expand=expansion(mult=c(0,0.1)))+
  labs(title="longitudinal distribution",
       x="Longitude (WGS84)",
       y="number of plots")

plot.lat <- ggplot(head.sf)+my.theme+
  geom_histogram(aes(y=lat.wgs))+
  geom_hline(yintercept=57.5, color="darkred")+
  scale_y_continuous(breaks=seq(5,85, 2.5), expand=expansion(add=c(0,0)))+
  scale_x_continuous(breaks=seq(0,1000000,50000), expand=expansion(mult=c(0,0.1)))+
  labs(title="latitudinal distribution",
       x="number of plots",
       y="Latitude (WGS84)")

grid.lonlat <- grid.arrange(plot.lon, plot.lat, nrow=1)

ggsave(filename=paste0(path.out,"/head_lonlatfilters.png"),
       plot=grid.lonlat,
       width=wide, height=10, unit="cm",
       bg="white")


```

```{r}

bla <- data.frame(id=1:10,
                  lvl1=c(sample(letters[1:3], 9, replace=T), rep(NA,1)) ,
                  eunis.lvl1=c(rep("x", 3), sample(letters[1:3], 6, replace=T), rep("a",1)))


bla$eunis.lvl1[bla$eunis.lvl1=="x"] <- bla$lvl1[bla$eunis.lvl1=="x"]


```

## combo-table for Irena

```{r}

### join country info with species info
megajoin <- left_join(head, plot, by="PlotObservationID")

specxcountry <- megajoin %>% 
     group_by(sPlot.concept.unified, Country) %>% 
     dplyr::summarise(n=n())

### exclude genera
specxcountry <- specxcountry[grepl("_", specxcountry$sPlot.concept.unified),]

### exclude countries clearly not in study region
ctry.excl <- c("Canada", "Belarus", "Israel", "Tunisia", "Cyprus", "Lebanon", "Morocco", "Malta", "Azerbaidjan", "Algeria","Faroe Islands", "Iran", "Syria", "Kazakhstan", "Greenland", "Libya", "Turkey", "Iceland", "Svalbard and Jan Mayen Is", "Russia")

specxcountry <- specxcountry[!specxcountry$Country %in% ctry.excl,]

### exclude species with less than 100 occurrences
specxcountry <- specxcountry[specxcountry$n>99,]

### export
fwrite(specxcountry, file=paste0(path.out, "/species_x_country.csv"), sep=",", dec=".", row.names=F)

rm(megajoin, specxcountry)

```

### testing cutting year to data

```{r}

header <- st_drop_geometry(head.sf)

list.yrgrp <- list()

for (bin in 5:10){

# Quantile binning
n.yr.grp.br <- bin

# Quantile breaks (keep numeric; no format() here)
yr.grp.br <- quantile(header$year, probs = seq(0, 1, length.out = n.yr.grp.br + 1), na.rm = TRUE)
yr.grp.br <- unique(yr.grp.br)  # guard against duplicate breakpoints with discrete years

# Labels: [start, end)
yr.grp.lab <- paste0(
  format(head(yr.grp.br, -1)+1, scientific = FALSE, trim = TRUE),
  "\n–\n ",
  format(tail(yr.grp.br, -1), scientific = FALSE, trim = TRUE)
)

# Cut as left-closed, right-open
header$yr.grp <- cut(
  header$year,
  breaks = yr.grp.br,
  labels = yr.grp.lab,
  include.lowest = TRUE,
  right = FALSE
)

headsum <- header %>% group_by(yr.grp) %>% dplyr::summarise(n=n())

headsum$groups <- bin

list.yrgrp[[bin-3]] <- headsum

}

df.yrgrp <- bind_rows(list.yrgrp)

library(dplyr)
library(stringr)
library(tidytext)   # for reorder_within()
library(ggplot2)

grp_order <- sort(unique(df.yrgrp$groups))

df.yrgrp2 <- df.yrgrp %>%
  group_by(groups) %>%
  mutate(
    start = as.integer(str_extract(as.character(yr.grp), "^[0-9]+")),
    yr.grp = reorder_within(yr.grp, start, groups)   # order within each facet
  ) %>%
  ungroup() %>% 
      mutate(
    # facet label with global levels in the order you want
    grp.lab = factor(paste0(groups, " groups"),
                     levels = paste0(grp_order, " groups"),
                     ordered = TRUE)
  )

ggplot(df.yrgrp2)+my.theme+
     geom_col(aes(x = yr.grp, y = n), width=0.66)+
     scale_y_continuous(expand=expansion(mult=c(0,0.05)))+
     theme(axis.text.x=element_text(size=6))+
     facet_wrap(~ grp.lab, scales = "free_x")+
     scale_x_reordered()+
     labs(x = NULL, y = "number of plots")
```
