---
title: "Script to validate points in ReSurvey database using RS data (Landsat)"
subtitle: "Fitting Random Forest models with ntree = 1500"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

This R script is used to validate the points in the ReSurvey database using RS indicators from Landsat (indices + phenology).

# Load libraries

```{r}
library(tidyverse)
library(here)
library(parallel)
library(doParallel)
library(moreparty)
library(caret)
library(randomForest)
library(pROC)
library(rlang)
library(stringr)
library(beepr)
library(permimp)
library(purrr)
library(gghalves)
library(patchwork)
library(scales)
```

# Load functions

```{r}
knitr::purl("functions_20260611_reduced.Rmd", output = "functions_20260611_reduced.R")
source("functions_20260611_reduced.R")
```

# Load previously created objects

```{r}
# Define the folder path
folder_path <- here("objects", "RF_1500_20260608")

# List all .RData or .rda files in the folder
rdata_files <- list.files(folder_path, full.names = TRUE)

# Load each file
lapply(rdata_files, load, envir = .GlobalEnv)
```

```{r}
# Load varimp objects
lapply(list.files(here("objects", "RF_1500_varimp_20260608"), full.names = TRUE), load,
       envir = .GlobalEnv)
```

```{r}
# Load ROC objects
lapply(list.files(here("objects", "RF_1500_ROC_20260608"), full.names = TRUE), load,
       envir = .GlobalEnv)
```

# Read data

```{r}
data_validation_Landsat <- read_tsv(here(
  "data", "clean","final_RS_data_bands_Landsat_all_20260618.csv"))
```

No parsing issues!

```{r}
nrow(data_validation_Landsat)
```

# Define lists of predictor variables

## Moments with slope method

```{r}
vars_RF_slope_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI and SAVI
  "NDVI_auc", "SAVI_auc",
  # Values of NDVI and SAVI at sos, pos and eos
  "NDVI_sos_value", "NDVI_pos_value", "NDVI_eos_value",
  "SAVI_sos_value", "SAVI_pos_value", "SAVI_eos_value",
  # Differences pos-sos in value and doy
  "NDVI_diff_pos_sos_value","SAVI_diff_pos_sos_value",
  "NDVI_diff_pos_sos_doy", "SAVI_diff_pos_sos_doy",
  # Differences pos-eos in value and doy
  "NDVI_diff_pos_eos_value", "SAVI_diff_pos_eos_value",
  "NDVI_diff_pos_eos_doy", "SAVI_diff_pos_eos_doy",
  # Growing season duration
  "NDVI_gsd", "SAVI_gsd")
```

## Months method

```{r}
vars_RF_months_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, and SAVI between march and oct
  "NDVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI and SAVI at march, pos and oct
  "NDVI_avg_value_03", "NDVI_pos_value", "NDVI_avg_value_10",
  "SAVI_avg_value_03", "SAVI_pos_value", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "SAVI_diff_pos_march_value",
  "NDVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "SAVI_diff_pos_oct_value",
  "NDVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy")
```

# Filter data to use in RF models

## Functions

```{r}
get_full_datasets <- function(data, variables) {
  data %>%
    mutate(EUNIS1 = as.factor(EUNIS1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Keep only rows with differences > 0 in variables
    dplyr::filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0)) %>%
    # Filter based on max NDWI
    mutate(valid_1_NDWI = case_when(NDWI_max > 0.3 ~ "wrong", 
                                    TRUE ~ NA_character_),
           valid_1_count = rowSums(across(c(valid_1_NDWI), ~ . == "wrong"), 
                                   na.rm = TRUE),
           valid_1 = if_else(valid_1_count > 0,
                             "At least 1 rule broken", "No rules broken so far")
           ) %>%
    # Keep only rows with no rules broken
    dplyr::filter(valid_1 == "No rules broken so far") %>%
    # Remove unneeded cols
    select(-starts_with("valid"))
  }
```

```{r}
get_training_subsets <- function(data, variables) {
  data %>%
    mutate(EUNIS1 = as.factor(EUNIS1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Keep only rows with differences > 0 in variables
    dplyr::filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0)) %>%
    # Filter based on max NDWI
    mutate(valid_1_NDWI = case_when(NDWI_max > 0.3 ~ "wrong", 
                                    TRUE ~ NA_character_),
           valid_1_count = rowSums(across(c(valid_1_NDWI), ~ . == "wrong"), 
                                   na.rm = TRUE),
           valid_1 = if_else(valid_1_count > 0,
                             "At least 1 rule broken", "No rules broken so far")
           ) %>%
    # Keep only rows with no rules broken
    dplyr::filter(valid_1 == "No rules broken so far") %>%
    # Remove unneeded cols
    select(-starts_with("valid")) %>%
    # Select only GPS points
    dplyr::filter(`Location method` == "Location with GPS" | 
                    `Location method` == "Location with differential GPS") %>%
    # Filter based on NDVI percentiles
    filter_by_group_percentiles("NDVI_pos_value")
  }
```

## Get full datasets and training subsets

```{r}
full_dataset_slope <- data_validation_Landsat %>%
  get_full_datasets(vars_RF_slope_Landsat)
full_dataset_months <- data_validation_Landsat %>%
  get_full_datasets(vars_RF_months_Landsat)  
training_subset_slope <- data_validation_Landsat %>%
  get_training_subsets(vars_RF_slope_Landsat)
training_subset_months <- data_validation_Landsat %>%
  get_training_subsets(vars_RF_months_Landsat)
```

## N obs

```{r}
nrow(full_dataset_slope)
nrow(full_dataset_months)
nrow(training_subset_slope)
nrow(training_subset_months)
```

# (Appendix 5) Plots distribution max NDVI

```{r}
plot_perc_Landsat <- full_dataset_months %>%
    # Select only GPS points
    dplyr::filter(`Location method` == "Location with GPS" | 
                    `Location method` == "Location with differential GPS") %>%
  distr_plot("NDVI_pos_value", "Maximum NDVI", grey = T) +
  scale_x_discrete(limits = c("T", "S", "R", "Q"),
                   labels = c("T" = "T - Forests", "S" = "S - Shrublands",
                              "R" = "R - Grasslands", "Q" = "Q - Wetlands"))
plot_perc_Landsat
ggsave(here("output", "figures", "manuscript", "appendix5.jpeg"), 
       dpi = 300, width = 5, height = 4)
```

# Fit RF models (level 1)

## Without refinement

### Split into training and test data sets

Split data:

```{r}
# Slope
set.seed(123)
split_data_slope_Landsat <- split_train_test(training_subset_slope)

train_data_slope_Landsat <- split_data_slope_Landsat$train
test_data_slope_Landsat <- split_data_slope_Landsat$test

# Check overlap between training and testing datasets
sum(train_data_slope_Landsat$PlotObservationID %in% test_data_slope_Landsat$PlotObservationID)

# Months
set.seed(123)
split_data_months_Landsat <- split_train_test(training_subset_months)

train_data_months_Landsat <- split_data_months_Landsat$train
test_data_months_Landsat <- split_data_months_Landsat$test

# Check overlap between training and testing datasets
sum(train_data_months_Landsat$PlotObservationID %in% test_data_months_Landsat$PlotObservationID)
```

### Fit models

```{r eval=FALSE, include=FALSE}
rf_slope_Landsat <- run_rf_parallel(vars_RF_slope_Landsat, train_data_slope_Landsat, 
                           "EUNIS1", ntree = 1500)
rf_months_Landsat <- run_rf_parallel(vars_RF_months_Landsat, train_data_months_Landsat,
                            "EUNIS1", ntree = 1500)
```

```{r}
models_no_refin <- c("rf_slope_Landsat", "rf_months_Landsat")
```

Print the time:

```{r}
for (model_name in models_no_refin) {
  obj <- get(model_name)
  cat("Time for", model_name, ":\n")
  print(obj$time)
}
```

Save:

```{r eval=FALSE, include=FALSE}
for (model_name in models_no_refin) {
  file_path <- paste0("objects/RF_1500_20260608/", model_name, "_model.Rdata")
  save(list = model_name, file = file_path)
}
```

### Predictions

```{r eval=FALSE, include=FALSE}
for (model_name in models_no_refin) {
  # Derive test data name by replacing "rf_" with "test_data_"
  test_data_name <- sub("^rf_", "test_data_", model_name)
  
  # Create prediction variable name
  prediction_name <- paste0(model_name, "_prediction")
  
  # Get model and test data
  model <- get(model_name)$model
  test_data <- get(test_data_name)
  
  # Run prediction
  prediction <- predict(model, newdata = test_data, 
                        type = "response")
  
  # Assign prediction to a variable
  assign(prediction_name, prediction)
}
```

```{r}
predictions_no_refin <- c("rf_slope_Landsat_prediction", 
                          "rf_months_Landsat_prediction")
```

Save:

```{r eval=FALSE, include=FALSE}
for (prediction_name in predictions_no_refin) {
  file_path <- paste0("objects/RF_1500_20260608/", prediction_name, ".Rdata")
  save(list = prediction_name, file = file_path)
}
```

### Confusion matrices

```{r}
for (prediction_name in predictions_no_refin) {
  # Derive test data name by replacing "_prediction" with "" 
  # and "rf_" with "test_data_"
  test_data_name <- paste0("test_data_", sub("^rf_", "", sub("_prediction$", "",
                                                             prediction_name)))
  
  # Get prediction and test data
  prediction <- get(prediction_name)
  test_data <- get(test_data_name)
  
  # Compute confusion matrix
  cat("Confusion Matrix for", sub("_prediction$", "", prediction_name), ":\n")
  print(confusionMatrix(prediction, test_data$EUNIS1))
  cat("\n")
}
```

#### Plots

```{r}
for (prediction_name in predictions_no_refin) {
  # Derive name of test data
  test_data_name <- paste0("test_data_", sub("^rf_", "", sub("_prediction$", "",
                                                             prediction_name)))
  
  # Obtain objects
  prediction <- get(prediction_name)
  test_data <- get(test_data_name)

  # Personalized title
  title <- paste(#"Confusion Matrix for", 
    sub("_prediction$", "",
                                              prediction_name))

  # Plot
  print(plot_confusion_matrix(prediction, test_data$EUNIS1, title))
}
```

## Refinement step 1

Based on PAs "Morralla protocol".

### Compute predictions for all data and combine with original data

```{r eval=FALSE, include=FALSE}
# Slope
training_subset_slope_with_probs <- bind_cols(
  training_subset_slope %>% select(PlotObservationID, EUNIS1),
  as.data.frame(predict(rf_slope_Landsat$model, 
                        # Predicting in all data (training + testing)
                                          newdata = training_subset_slope,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_slope_Landsat$model, 
                         # Predicting in all data (training + testing)
                         newdata = training_subset_slope,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNIS1, prob_Q = EUNIS1.Q,
         prob_R = EUNIS1.R, prob_S = EUNIS1.S, prob_T = EUNIS1.T)

# Months
training_subset_months_with_probs <- bind_cols(
  training_subset_months %>% select(PlotObservationID, EUNIS1),
  as.data.frame(predict(rf_months_Landsat$model, 
                        # Predicting in all data (training + testing)
                                          newdata = training_subset_months,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_months_Landsat$model, 
                         # Predicting in all data (training + testing)
                         newdata = training_subset_months,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNIS1, prob_Q = EUNIS1.Q,
         prob_R = EUNIS1.R, prob_S = EUNIS1.S, prob_T = EUNIS1.T)
```

### Detect probability columns

```{r eval=FALSE, include=FALSE}
prob_cols_slope_Landsat <- names(training_subset_slope_with_probs)[grepl("^prob_", names(training_subset_slope_with_probs))]
prob_cols_months_Landsat <- names(training_subset_months_with_probs)[grepl("^prob_", names(training_subset_months_with_probs))]
```

### Calculate auxiliary metrics

```{r eval=FALSE, include=FALSE}
df_aux_slope_Landsat <- calculate_df_aux(training_subset_slope_with_probs,
                                     prob_cols_slope_Landsat)
df_aux_months_Landsat <- calculate_df_aux(training_subset_months_with_probs,
                                     prob_cols_months_Landsat)
```

### Apply filtering function

Applying "Morralla" protocol once, removing negatives until 20% of the class and, if there is room, uncertain positives until 10% of the rest.

```{r eval=FALSE, include=FALSE}
df_final_20_10_slope_Landsat <- df_aux_slope_Landsat %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20_pos10)
df_final_20_10_months_Landsat <- df_aux_months_Landsat %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20_pos10)
```

```{r eval=FALSE, include=FALSE}
filtered_data_20_10_slope_Landsat_probs <- filtered_data_slope_Landsat %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_slope_Landsat$PlotObservationID)
filtered_data_20_10_months_Landsat_probs <- filtered_data_months_Landsat %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_months_Landsat$PlotObservationID)
```

### Save

```{r}
filtered_data_with_probs <- c("filtered_data_20_10_slope_Landsat_probs", 
                              "filtered_data_20_10_months_Landsat_probs")
```

Save:

```{r eval=FALSE, include=FALSE}
for (filtered_data_name in filtered_data_with_probs) {
  file_path <- paste0("objects/RF_1500_20260608/", filtered_data_name, ".Rdata")
  save(list = filtered_data_name, file = file_path)
}
```

### Compare classes before and after

```{r}
compare_classes(training_subset_slope, filtered_data_20_10_slope_Landsat_probs)
compare_classes(training_subset_months, filtered_data_20_10_months_Landsat_probs)
```

### Split into training and testing

```{r}
# Landsat
# Slope
set.seed(123)
split_data_20_10_slope_Landsat <- split_train_test(filtered_data_20_10_slope_Landsat_probs)

train_data_20_10_slope_Landsat <- split_data_20_10_slope_Landsat$train
test_data_20_10_slope_Landsat <- split_data_20_10_slope_Landsat$test

# Check overlap between training and testing datasets
sum(train_data_20_10_slope_Landsat$PlotObservationID %in% test_data_20_10_slope_Landsat$PlotObservationID)

# Months
set.seed(123)
split_data_20_10_months_Landsat <- split_train_test(filtered_data_20_10_months_Landsat_probs)

train_data_20_10_months_Landsat <- split_data_20_10_months_Landsat$train
test_data_20_10_months_Landsat <- split_data_20_10_months_Landsat$test

# Check overlap between training and testing datasets
sum(train_data_20_10_months_Landsat$PlotObservationID %in% test_data_20_10_months_Landsat$PlotObservationID)
```

### Fit models

```{r eval=FALSE, include=FALSE}
rf_slope_Landsat_refin_20_10 <- run_rf_parallel(vars_RF_slope_Landsat,
                                       train_data_20_10_slope_Landsat, 
                                       "EUNIS1", ntree = 1500)
rf_months_Landsat_refin_20_10 <- run_rf_parallel(vars_RF_months_Landsat,
                                        train_data_20_10_months_Landsat, 
                                        "EUNIS1", ntree = 1500)
```

```{r}
models_refin_20_10 <- c("rf_slope_Landsat_refin_20_10", 
                        "rf_months_Landsat_refin_20_10")
```

Save:

```{r eval=FALSE, include=FALSE}
for (model_name in models_refin_20_10) {
  file_path <- paste0("objects/RF_1500_20260608/", model_name, "_model.Rdata")
  save(list = model_name, file = file_path)
}
```

### Predictions

```{r eval=FALSE, include=FALSE}
for (model_name in models_refin_20_10) {
  # Derive test data name
  test_data_name <- paste0(gsub("_refin_20_10$", "", 
                                sub("^rf_", "test_data_20_10_", model_name)))

  # Create prediction variable name
  prediction_name <- paste0(model_name, "_prediction")
  
  # Get model and test data
  model <- get(model_name)$model
  test_data <- get(test_data_name)
  
  # Run prediction
  prediction <- predict(model, newdata = test_data, 
                        type = "response")
  
  # Assign prediction to a variable
  assign(prediction_name, prediction)
}
```

```{r}
predictions_refin_20_10 <- c("rf_slope_Landsat_refin_20_10_prediction",
                        "rf_months_Landsat_refin_20_10_prediction")
```

Save:

```{r eval=FALSE, include=FALSE}
for (prediction_name in predictions_refin_20_10) {
  file_path <- paste0("objects/RF_1500_20260608/", prediction_name, ".Rdata")
  save(list = prediction_name, file = file_path)
  }
```

### Confusion matrices

```{r}
for (prediction_name in predictions_refin_20_10) {
  # Derive test data name
  test_data_name <- paste0(gsub("_refin_20_10$", "",
                              sub("^rf_", "test_data_20_10_", 
                                  sub("_prediction$", "", prediction_name))))
  
  # Get prediction and test data
  prediction <- get(prediction_name)
  test_data <- get(test_data_name)
  
  # Compute confusion matrix
  cat("Confusion Matrix for", sub("_prediction$", "", prediction_name), ":\n")
  print(confusionMatrix(prediction, test_data$EUNIS1))
  cat("\n")
}
```

#### Plots

```{r}
for (prediction_name in predictions_refin_20_10) {
  # Derive test data name
  test_data_name <- paste0(gsub("_refin_20_10$", "",
                              sub("^rf_", "test_data_20_10_", 
                                  sub("_prediction$", "", prediction_name))))
  
  # Obtain objects
  prediction <- get(prediction_name)
  test_data <- get(test_data_name)

  # Personalized title
  title <- paste("Confusion Matrix for", sub("_prediction$", "",
                                              prediction_name))

  # Plot
  print(plot_confusion_matrix(prediction, test_data$EUNIS1, title))
}
```

## Refinement step 2

Then refitting the model and applying "Morralla" protocol again, removing only negatives until 20% of the class.

#### Compute predictions for all data and combine with original data

```{r eval=FALSE, include=FALSE}
# Landsat, slope
filtered_data_slope_Landsat_with_probs_2x <- bind_cols(
  filtered_data_20_10_slope_Landsat_probs %>% select(PlotObservationID, EUNIS1),
  as.data.frame(predict(rf_slope_Landsat_refin_20_10$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_20_10_slope_Landsat_probs,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_slope_Landsat_refin_20_10$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_20_10_slope_Landsat_probs,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNIS1, prob_Q = EUNIS1.Q,
         prob_R = EUNIS1.R, prob_S = EUNIS1.S, prob_T = EUNIS1.T)

# Landsat, months
filtered_data_months_Landsat_with_probs_2x <- bind_cols(
  filtered_data_20_10_months_Landsat_probs %>% select(PlotObservationID, EUNIS1),
  as.data.frame(predict(rf_months_Landsat_refin_20_10$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_20_10_months_Landsat_probs,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(rf_months_Landsat_refin_20_10$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_20_10_months_Landsat_probs,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNIS1, prob_Q = EUNIS1.Q,
         prob_R = EUNIS1.R, prob_S = EUNIS1.S, prob_T = EUNIS1.T)
```

#### Calculate auxiliary metrics

```{r eval=FALSE, include=FALSE}
df_aux_slope_Landsat_2x <- calculate_df_aux(filtered_data_slope_Landsat_with_probs_2x,
                                     prob_cols_slope_Landsat)
df_aux_months_Landsat_2x <- calculate_df_aux(filtered_data_months_Landsat_with_probs_2x,
                                     prob_cols_months_Landsat)
```

#### Apply filtering function

```{r eval=FALSE, include=FALSE}
df_final_20_10_slope_Landsat_2x <- df_aux_slope_Landsat_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)
df_final_20_10_months_Landsat_2x <- df_aux_months_Landsat_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)
```

```{r eval=FALSE, include=FALSE}
filtered_data_20_10_slope_Landsat_probs_2x <- filtered_data_slope_Landsat %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_slope_Landsat_2x$PlotObservationID)
filtered_data_20_10_months_Landsat_probs_2x <- filtered_data_months_Landsat %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_months_Landsat_2x$PlotObservationID)
```

### Save

```{r}
filtered_data_with_probs_2x <- c("filtered_data_20_10_slope_Landsat_probs_2x", 
                              "filtered_data_20_10_months_Landsat_probs_2x")
```

Save:

```{r eval=FALSE, include=FALSE}
for (filtered_data_name in filtered_data_with_probs_2x) {
  file_path <- paste0("objects/RF_1500_20260608/", filtered_data_name, ".Rdata")
  save(list = filtered_data_name, file = file_path)
}
```

#### Compare classes before and after

```{r}
compare_classes(training_subset_slope,
                filtered_data_20_10_slope_Landsat_probs,
                filtered_data_20_10_slope_Landsat_probs_2x)
compare_classes(training_subset_months, 
                filtered_data_20_10_months_Landsat_probs,
                filtered_data_20_10_months_Landsat_probs_2x)
```

##### (Appendix 6)

```{r}
(compare_classes(training_subset_slope, 
                  filtered_data_20_10_slope_Landsat_probs,
                  filtered_data_20_10_slope_Landsat_probs_2x,
                  return = "plot") +
  compare_classes(training_subset_months, 
                  filtered_data_20_10_months_Landsat_probs,
                  filtered_data_20_10_months_Landsat_probs_2x,
                  return = "plot")
 ) +
  plot_annotation(tag_levels = "A", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  scale_fill_manual(
    breaks = c("before", "step1", "step2"),
    labels = c("No refinement", "Refinement step 1", "Refinement step 2"),
    values = c(
      "before" = "#edf8b1", 
      "step1"  = "#7fcdbb",
      "step2"  = "#2c7fb8"
    )
  ) &
  scale_y_continuous(limits = c(0, 45000),
                  labels = label_number(scale = 1e-3, suffix = "k")) &
  theme(legend.position = "bottom", legend.title = element_blank()) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
ggsave(here("output", "figures", "manuscript", "appendix6.jpeg"), 
       dpi = 300, width = 6, height = 4.5)
```

#### Split into training and testing

```{r}
# Landsat
# Slope
set.seed(123)
split_data_20_10_2x_slope_Landsat <- split_train_test(filtered_data_20_10_slope_Landsat_probs_2x)

train_data_20_10_2x_slope_Landsat <- split_data_20_10_2x_slope_Landsat$train
test_data_20_10_2x_slope_Landsat <- split_data_20_10_2x_slope_Landsat$test

# Check overlap between training and testing datasets
sum(train_data_20_10_2x_slope_Landsat$PlotObservationID %in% test_data_20_10_2x_slope_Landsat$PlotObservationID)

# Months
set.seed(123)
split_data_20_10_2x_months_Landsat <- split_train_test(filtered_data_20_10_months_Landsat_probs_2x)

train_data_20_10_2x_months_Landsat <- split_data_20_10_2x_months_Landsat$train
test_data_20_10_2x_months_Landsat <- split_data_20_10_2x_months_Landsat$test

# Check overlap between training and testing datasets
sum(train_data_20_10_2x_months_Landsat$PlotObservationID %in% test_data_20_10_2x_months_Landsat$PlotObservationID)
```

#### Fit models

```{r eval=FALSE, include=FALSE}
rf_slope_Landsat_refin_20_10_2x <- run_rf_parallel(vars_RF_slope_Landsat,
                                       train_data_20_10_2x_slope_Landsat, 
                                       "EUNIS1", ntree = 1500)
rf_months_Landsat_refin_20_10_2x <- run_rf_parallel(vars_RF_months_Landsat,
                                        train_data_20_10_2x_months_Landsat, 
                                        "EUNIS1", ntree = 1500)
```

```{r}
models_refin_20_10_2x <- c("rf_slope_Landsat_refin_20_10_2x", 
                           "rf_months_Landsat_refin_20_10_2x")
```

Save:

```{r eval=FALSE, include=FALSE}
for (model_name in models_refin_20_10_2x) {
  file_path <- paste0("objects/RF_1500_20260608/", model_name, "_model.Rdata")
  save(list = model_name, file = file_path)
}
```

#### Predictions

```{r eval=FALSE, include=FALSE}
for (model_name in models_refin_20_10_2x) {
  # Derive test data name
  test_data_name <- paste0(gsub("_refin_20_10_2x$", "", 
                                sub("^rf_", "test_data_20_10_2x_", model_name)))

  # Create prediction variable name
  prediction_name <- paste0(model_name, "_prediction")
  
  # Get model and test data
  model <- get(model_name)$model
  test_data <- get(test_data_name)
  
  # Run prediction
  prediction <- predict(model, newdata = test_data, 
                        type = "response")
  
  # Assign prediction to a variable
  assign(prediction_name, prediction)
}
```

```{r}
predictions_refin_20_10_2x <- c("rf_slope_Landsat_refin_20_10_2x_prediction",
                                "rf_months_Landsat_refin_20_10_2x_prediction")
```

Save:

```{r eval=FALSE, include=FALSE}
for (prediction_name in predictions_refin_20_10_2x) {
  file_path <- paste0("objects/RF_1500_20260608/", prediction_name, ".Rdata")
  save(list = prediction_name, file = file_path)
  }
```

#### Confusion matrices

```{r}
for (prediction_name in predictions_refin_20_10_2x) {
  # Derive test data name
  test_data_name <- paste0(gsub("_refin_20_10_2x$", "",
                              sub("^rf_", "test_data_20_10_2x_", 
                                  sub("_prediction$", "", prediction_name))))
  
  # Get prediction and test data
  prediction <- get(prediction_name)
  test_data <- get(test_data_name)
  
  # Compute confusion matrix
  cat("Confusion Matrix for", sub("_prediction$", "", prediction_name), ":\n")
  print(confusionMatrix(prediction, test_data$EUNIS1))
  cat("\n")
}
```

##### Plots

```{r}
for (prediction_name in predictions_refin_20_10_2x) {
  # Derive test data name
  test_data_name <- paste0(gsub("_refin_20_10_2x$", "",
                              sub("^rf_", "test_data_20_10_2x_", 
                                  sub("_prediction$", "", prediction_name))))
  
  # Obtain objects
  prediction <- get(prediction_name)
  test_data <- get(test_data_name)

  # Personalized title
  title <- paste("Confusion Matrix for", sub("_prediction$", "",
                                              prediction_name))

  # Plot
  print(plot_confusion_matrix(prediction, test_data$EUNIS1, title))
}
```

#### (Appendix 9) Variable importance

```{r eval=FALSE, include=FALSE}
varimp_months_Landsat_step2 <-
  compute_varimp(rf_months_Landsat_refin_20_10_2x$model, nperm = 100)
```

```{r eval=FALSE, include=FALSE}
save(varimp_months_Landsat_step2,
     file = "objects/RF_1500_20260608/varimp_months_Landsat_step2.Rdata")
```

```{r}
enframe(
  varimp_months_Landsat_step2$varimp$values,
  name  = "variable",
  value = "importance"
) %>%
  mutate(
    importance100 = 100 * importance / max(importance)
  ) %>%
  arrange(desc(importance100)) %>%
  mutate(variable_long = case_when(
    variable == "NDMI_max" ~ "Maximum NDMI",
    variable == "NDWI_max" ~ "Maximum NDWI",
    variable == "NDVI_max" ~ "Maximum NDVI",
    variable == "SAVI_max" ~ "Maximum SAVI",
    variable == "NDMI_min" ~ "Minimum NDMI",
    variable == "NDWI_min" ~ "Minimum NDWI",
    variable == "NDVI_min" ~ "Minimum NDVI",
    variable == "SAVI_min" ~ "Minimum SAVI",
    variable == "NDVI_avg_value_03" ~ "NDVI value at greening",
    variable == "NDVI_pos_value" ~ "NDVI value at peak",
    variable == "NDVI_avg_value_10" ~ "NDVI value at browning",
    variable == "SAVI_avg_value_03" ~ "SAVI value at greening",
    variable == "SAVI_pos_value" ~ "SAVI value at peak",
    variable == "SAVI_avg_value_10" ~ "SAVI value at browning",
    variable == "NDVI_auc_mar_oct" ~ "NDVI AUC greening - browning",
    variable == "SAVI_auc_mar_oct" ~ "SAVI AUC greening - browning",
    variable == "NDVI_diff_pos_march_value" ~ 
      "NDVI difference in value peak - greening",
    variable == "SAVI_diff_pos_march_value" ~
      "SAVI difference in value peak - greening",
    variable == "NDVI_diff_pos_march_doy" ~
      "NDVI difference in DOY peak - greening",
    variable == "SAVI_diff_pos_march_doy" ~
      "SAVI difference in DOY peak - greening",
    variable == "NDVI_diff_pos_oct_value" ~
      "NDVI difference in value peak - browning",
    variable == "SAVI_diff_pos_oct_value" ~
      "SAVI difference in value peak - browning",
    variable == "NDVI_diff_pos_oct_doy" ~
      "NDVI difference in DOY peak - browning",
    variable == "SAVI_diff_pos_oct_doy" ~
      "SAVI difference in DOY peak - browning",
    TRUE ~ variable
  )) %>%
ggplot(aes(x = reorder(variable_long, importance100), y = importance100)) +
  geom_col(fill = "#2c7fb8") +
  coord_flip() +
  labs(x = "Predictor", y = "Importance (scaled)") +
  theme_minimal()
ggsave(here("output", "figures", "manuscript", "appendix9.jpeg"), 
       dpi = 300, width = 6, height = 4)
```

### (Appendix 8) ROC curves

```{r eval=FALSE, include=FALSE}
rocdata_months_Landsat_step2 <-
  compute_roc_level1(rf_months_Landsat_refin_20_10_2x$model,
                     test_data_20_10_2x_months_Landsat)
```

```{r eval=FALSE, include=FALSE}
save(rocdata_months_Landsat_step2,
     file = "objects/RF_1500_20260608/rocdata_months_Landsat_step2.Rdata")
```

```{r}
ggplot(rocdata_months_Landsat_step2$roc, aes(x = FPR, y = TPR, color = Class)) +
  geom_line(linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray") +
  labs(title = paste("Multiclass ROC Curve -", model_name),
       x = "False Positive Rate (1-Specificity",
       y = "True Positive Rate (Sensitivity)",
       color = "Class (AUC)") +
  scale_color_manual(
    values = c(
      "Q (AUC = 0.915)" = "#0072B2",
      "R (AUC = 0.978)" = "#D55E00",
      "S (AUC = 0.967)" = "#009E73",
      "T (AUC = 0.997)" = "#CC79A7"
    ),
    labels = c(
      "Q (AUC = 0.915)" = "Q – Wetlands (AUC = 0.915)",
      "R (AUC = 0.978)" = "R – Grasslands (AUC = 0.978)",
      "S (AUC = 0.967)" = "S – Shrublands (AUC = 0.967)",
      "T (AUC = 0.997)" = "T – Forests (AUC = 0.997)"
    )
  ) +
  theme_minimal() + ggtitle(NULL) +
  theme(legend.position = "right")
ggsave(here("output", "figures", "manuscript", "appendix8.jpeg"), 
       dpi = 300, width = 7, height = 4)
```

# Session info

```{r}
sessionInfo()
```
