---
title: "Script to validate points in ReSurvey database using RS data (Landsat)"
subtitle: "K-fold cross validation of all models"
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
```

This R script is used to validate the points in the ReSurvey database using RS indicators from Landsat (indices + phenology).
Here, we perform a K-fold cross validation of all the models in 41_validation_models_onlyLandsat.Rmd. 

# Load libraries

```{r}
library(tidyverse)
library(parallel)
library(doParallel)
library(foreach)
library(partykit)
library(caret)
library(here)
library(readr)
library(progressr)
library(party)
library(permimp)
library(tibble)
library(moreparty)
library(patchwork)
library(scales)
```

# Load previously created objects

```{r}
# Load cross-validation objects
lapply(list.files(here("objects", "RF_1500_cv_20260609"), full.names = TRUE), load,
       envir = .GlobalEnv)
```

# Read data

```{r}
data_validation_Landsat <- read_tsv(here(
  "data", "clean","final_RS_data_bands_Landsat_all_20260618.csv"))
```

No parsing issues!

```{r}
nrow(data_validation_Landsat)
```

# Define lists of predictor vars

## Moments with slope method

```{r}
# Excluding EVI because many values outside of range -1, 1
# Not using canopy height!
vars_RF_slope_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI and SAVI
  "NDVI_auc", "SAVI_auc",
  # Values of NDVI and SAVI at sos, pos and eos
  "NDVI_sos_value", "NDVI_pos_value", "NDVI_eos_value",
  "SAVI_sos_value", "SAVI_pos_value", "SAVI_eos_value",
  # Differences pos-sos in value and doy
  "NDVI_diff_pos_sos_value","SAVI_diff_pos_sos_value",
  "NDVI_diff_pos_sos_doy", "SAVI_diff_pos_sos_doy",
  # Differences pos-eos in value and doy
  "NDVI_diff_pos_eos_value", "SAVI_diff_pos_eos_value",
  "NDVI_diff_pos_eos_doy", "SAVI_diff_pos_eos_doy",
  # Growing season duration
  "NDVI_gsd", "SAVI_gsd")
```

## Months method

```{r}
# Excluding EVI because many values outside of range -1, 1
# Not using canopy height!
vars_RF_months_Landsat <- c(
  # Min values of all indices
  "NDVI_min", "NDMI_min", "NDWI_min", "SAVI_min",
  # Max values of NDMI and NDWI
  "NDMI_max", "NDWI_max",
  # AUC of NDVI, and SAVI between march and oct
  "NDVI_auc_mar_oct", "SAVI_auc_mar_oct",
  # Values of NDVI and SAVI at march, pos and oct
  "NDVI_avg_value_03", "NDVI_pos_value", "NDVI_avg_value_10",
  "SAVI_avg_value_03", "SAVI_pos_value", "SAVI_avg_value_10",
  # Differences pos-march in value and doy
  "NDVI_diff_pos_march_value", "SAVI_diff_pos_march_value",
  "NDVI_diff_pos_march_doy", "SAVI_diff_pos_march_doy",
  # Differences pos-oct in value and doy
  "NDVI_diff_pos_oct_value", "SAVI_diff_pos_oct_value",
  "NDVI_diff_pos_oct_doy", "SAVI_diff_pos_oct_doy")
```

# Filter data to use in RF models

## Functions

### Function to get full datasets

```{r}
get_full_datasets <- function(data, variables) {
  data %>%
    mutate(EUNIS1 = as.factor(EUNIS1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Keep only rows with differences > 0 in variables
    dplyr::filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0)) %>%
    # Filter based on max NDWI
    mutate(valid_1_NDWI = case_when(NDWI_max > 0.3 ~ "wrong", 
                                    TRUE ~ NA_character_),
           valid_1_count = rowSums(across(c(valid_1_NDWI), ~ . == "wrong"), 
                                   na.rm = TRUE),
           valid_1 = if_else(valid_1_count > 0,
                             "At least 1 rule broken", "No rules broken so far")
           ) %>%
    # Keep only rows with no rules broken
    dplyr::filter(valid_1 == "No rules broken so far") %>%
    # Remove unneeded cols
    select(-starts_with("valid"))
  }
```

### Function to get training subsets

```{r}
get_training_subsets <- function(data, variables) {
  data %>%
    mutate(EUNIS1 = as.factor(EUNIS1)) %>%
    # Remove rows with missing values in predictors
    dplyr::filter(if_all(all_of(variables), ~ !is.na(.))) %>%
    # Remove all rows with wrong values of indices (not between -1 and 1)
    dplyr::filter(NDVI_max <= 1 & NDVI_min >= -1) %>%
    dplyr::filter(SAVI_max <= 1 & SAVI_min >= -1) %>%
    dplyr::filter(NDMI_max <= 1 & NDMI_min >= -1) %>%
    dplyr::filter(NDWI_max <= 1 & NDWI_min >= -1) %>%
    # Keep only rows with differences > 0 in variables
    dplyr::filter(if_all(any_of(variables[grepl("diff", variables)]), 
                  ~ .x > 0)) %>%
    # Filter based on max NDWI
    mutate(valid_1_NDWI = case_when(NDWI_max > 0.3 ~ "wrong", 
                                    TRUE ~ NA_character_),
           valid_1_count = rowSums(across(c(valid_1_NDWI), ~ . == "wrong"), 
                                   na.rm = TRUE),
           valid_1 = if_else(valid_1_count > 0,
                             "At least 1 rule broken", "No rules broken so far")
           ) %>%
    # Keep only rows with no rules broken
    dplyr::filter(valid_1 == "No rules broken so far") %>%
    # Remove unneeded cols
    select(-starts_with("valid")) %>%
    # Select only GPS points
    dplyr::filter(`Location method` == "Location with GPS" | 
                    `Location method` == "Location with differential GPS") %>%
    # Filter based on NDVI percentiles
    filter_by_group_percentiles("NDVI_pos_value")
  }
```

### Function for NDVI-based filtering

```{r}
# data: your tibble/data.frame
# y_var: name of the numeric column as a string, e.g., "NDVI_pos_value"
filter_by_group_percentiles <- function(data, y_var) {
  cats <- c("T","Q","R","S")

  # Compute thresholds per group for the selected column
  qs <- data %>%
    dplyr::group_by(EUNIS1) %>%
    dplyr::summarise(
      p10 = stats::quantile(.data[[y_var]], probs = 0.10, na.rm = TRUE),
      p90 = stats::quantile(.data[[y_var]], probs = 0.90, na.rm = TRUE),
      .groups = "drop"
    )

  # Join thresholds and filter rows according to the rule
  d_filtered <- data %>%
    dplyr::left_join(qs, by = "EUNIS1") %>%
    dplyr::filter(
      dplyr::case_when(
        EUNIS1 == "T" ~ .data[[y_var]] >= p10,        # keep >= 10% for T
        EUNIS1 %in% c("Q","R","S") ~ .data[[y_var]] <= p90,  # keep <= 90% for Q/R/S
        TRUE ~ TRUE
      )
    ) %>%
    dplyr::select(-p10, -p90)  # drop helper columns

  return(d_filtered)
}
```

## Get full datasets and training subsets

```{r}
full_dataset_slope <- data_validation_Landsat %>%
  get_full_datasets(vars_RF_slope_Landsat)
full_dataset_months <- data_validation_Landsat %>%
  get_full_datasets(vars_RF_months_Landsat)  
training_subset_slope <- data_validation_Landsat %>%
  get_training_subsets(vars_RF_slope_Landsat)
training_subset_months <- data_validation_Landsat %>%
  get_training_subsets(vars_RF_months_Landsat)
```

## N obs

```{r}
nrow(full_dataset_slope)
nrow(full_dataset_months)
nrow(training_subset_slope)
nrow(training_subset_months)
```

# Random forest model with k-fold cross-validation 

Using 4 folds.

## Functions

### Function to run CV

```{r}
run_fastcforest_cv <- function(
    data, vars_RF, response_var,
    ntree = 500, k = 4
) {
  library(party)
  library(moreparty)
  library(doParallel)
  library(foreach)
  library(parallel)
  library(caret)

  # Create folds
  set.seed(123)
  folds <- createFolds(data[[response_var]], k = k)

  # Parallel cluster for CV
  ncores <- max(1, detectCores() - 1)
  cl <- makeCluster(ncores)
  registerDoParallel(cl)

  # Export needed objects
  clusterExport(cl, c("folds", "data", "vars_RF", "response_var"), envir = environment())

  # Load required packages inside workers
  clusterEvalQ(cl, {
    library(party)
    library(moreparty)
  })

  results <- list()

  for (i in seq_along(folds)) {

    cat("\n========== Starting fold", i, "of", k, "==========\n")
    t_start <- Sys.time()

    fold_result <- foreach(j = i, .combine = rbind,
                           .packages = c("party", "moreparty")) %dopar% {

      test_idx  <- folds[[j]]
      train_idx <- setdiff(seq_len(nrow(data)), test_idx)

      train_data <- data[train_idx, ]
      test_data  <- data[test_idx, ]

      # Build formula
      fmla <- reformulate(vars_RF, response = response_var)

      # ✅ Fast parallel conditional forest from moreparty
      model <- moreparty::fastcforest(
        formula  = fmla,
        data     = train_data,
        controls = party::cforest_unbiased(
          mtry  = round(sqrt(length(vars_RF))),
          ntree = ntree
        ),
        parallel = TRUE
      )

      pred <- predict(model, newdata = test_data, type = "response")

      data.frame(
        obs  = test_data[[response_var]],
        pred = pred,
        fold = j
      )
    }

    t_end <- Sys.time()
    cat(
      "✅ Finished fold", i, "in",
      round(difftime(t_end, t_start, units = "mins"), 2),
      "minutes\n"
    )

    results[[i]] <- fold_result
  }

  stopCluster(cl)

  do.call(rbind, results)
}
```

### Function to compute results from CV

```{r}
compute_rf_cv_metrics <- function(cv_results) {
  library(caret)
  library(pROC)
  library(dplyr)
  
  # Ensure pred is factor with same levels as obs
  cv_results$obs  <- factor(cv_results$obs)
  cv_results$pred <- factor(cv_results$pred, levels = levels(cv_results$obs))
  
  classes <- levels(cv_results$obs)
  folds   <- sort(unique(cv_results$fold))
  
  ## ---------------------------
  ## 1. PER-FOLD METRICS
  ## ---------------------------
  fold_metrics <- lapply(folds, function(f) {
    df <- cv_results %>% filter(fold == f)
    
    cm <- confusionMatrix(df$pred, df$obs)
    
    list(
      fold = f,
      accuracy = cm$overall["Accuracy"],
      kappa    = cm$overall["Kappa"]
    )
  })
  
  fold_metrics_df <- bind_rows(lapply(fold_metrics, as.data.frame))
  
  ## ---------------------------
  ## 2. OVERALL CONFUSION MATRIX
  ## ---------------------------
  overall_cm <- confusionMatrix(cv_results$pred, cv_results$obs)
  
  ## ---------------------------
  ## 3. CLASS-WISE METRICS
  ## ---------------------------
  class_metrics <- data.frame(
    class = classes,
    users_accuracy     = overall_cm$byClass[, "Pos Pred Value"],
    producers_accuracy = overall_cm$byClass[, "Sensitivity"],
    balanced_accuracy  = overall_cm$byClass[, "Balanced Accuracy"]
  )
  
  ## ---------------------------
  ## 4. FINAL OUTPUT STRUCTURE
  ## ---------------------------
  list(
    overall_confusion_matrix = overall_cm,
    
    accuracy_mean_sd = c(
      mean = mean(fold_metrics_df$accuracy),
      sd   = sd(fold_metrics_df$accuracy)
    ),
    
    kappa_mean_sd = c(
      mean = mean(fold_metrics_df$kappa),
      sd   = sd(fold_metrics_df$kappa)
    ),
    
    per_fold_metrics = fold_metrics_df,
    
    per_class_metrics = class_metrics
  )
}
```

## Function for fitting RF models on full dataset

```{r}
run_rf_parallel <- function(vars_RF, train_data, response_var, ntree = 500) {
  # Build formula locally (no NSE)
  fmla <- reformulate(vars_RF, response = response_var)

  # Start cluster
  n_cores <- max(1, parallel::detectCores() - 1)
  cl <- parallel::makeCluster(n_cores)
  doParallel::registerDoParallel(cl)

  # Make sure packages are available on workers
  parallel::clusterEvalQ(cl, {
    library(party)
  })

  # Export the *data* to workers under a simple name
  local_data <- train_data
  parallel::clusterExport(cl, varlist = c("local_data"), envir = environment())

  # Fit in parallel
  execution_time <- system.time({
    rf_model <- fastcforest(
      formula  = fmla,
      data     = local_data,   # use the exported name
      controls = party::cforest_control(
        mtry  = max(1, round(sqrt(length(vars_RF)))),
        ntree = ntree,
        # Decision trees were developed using subsampling without replacement,
        # which is appropriate to use when predictors vary in their scale of
        # measurement (Strobl et al., 2007).
        replace = FALSE
      ),
      parallel = TRUE
    )
  })

  parallel::stopCluster(cl)

  list(model = rf_model, time = execution_time)
}
```

## Without refinement

### Landsat, slope

#### CV

```{r eval=FALSE, include=FALSE}
cv_slope_Landsat <- run_fastcforest_cv(
  data = training_subset_slope,
  vars_RF = vars_RF_slope_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_slope_Landsat,
     file = "objects/RF_1500_cv_20260609/cv_slope_Landsat.Rdata") 
```

##### Results

```{r}
cv_slope_Landsat_results <-
  compute_rf_cv_metrics(cv_slope_Landsat)
cv_slope_Landsat_results
```

#### Fit model on full dataset

```{r eval=FALSE, include=FALSE}
RF_slope_Landsat <- run_rf_parallel(vars_RF_slope_Landsat, 
                                    training_subset_slope, 
                                    "EUNIS1", ntree = 1500)
```

```{r eval=FALSE, include=FALSE}
save(RF_slope_Landsat,
     file = "objects/RF_1500_cv_20260609/RF_slope_Landsat.Rdata") 
```

### Landsat, months

#### CV

```{r eval=FALSE, include=FALSE}
cv_months_Landsat <- run_fastcforest_cv(
  data = training_subset_months,
  vars_RF = vars_RF_months_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_months_Landsat,
     file = "objects/RF_1500_cv_20260609/cv_months_Landsat.Rdata") 
```

##### Results

```{r}
cv_months_Landsat_results <-
  compute_rf_cv_metrics(cv_months_Landsat)
cv_months_Landsat_results
```

#### Fit model on full dataset

```{r eval=FALSE, include=FALSE}
RF_months_Landsat <- run_rf_parallel(vars_RF_months_Landsat, 
                                    training_subset_slope, 
                                    "EUNIS1", ntree = 1500)
```

```{r eval=FALSE, include=FALSE}
save(RF_months_Landsat,
     file = "objects/RF_1500_cv_20260609/RF_months_Landsat.Rdata") 
```

## Refinement step 1

### Compute predictions for all data and combine with original data

```{r eval=FALSE, include=FALSE}
# Slope
training_subset_slope_with_probs <- bind_cols(
  training_subset_slope %>% select(PlotObservationID, EUNIS1),
  as.data.frame(predict(RF_slope_Landsat$model, 
                        # Predicting in all data
                                          newdata = training_subset_slope,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(RF_slope_Landsat$model, 
                         # Predicting in all
                         newdata = training_subset_slope,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNIS1, prob_Q = EUNIS1.Q,
         prob_R = EUNIS1.R, prob_S = EUNIS1.S, prob_T = EUNIS1.T)

# Months
training_subset_months_with_probs <- bind_cols(
  training_subset_months %>% select(PlotObservationID, EUNIS1),
  as.data.frame(predict(RF_months_Landsat$model, 
                        # Predicting in all data
                                          newdata = training_subset_months,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(RF_months_Landsat$model, 
                         # Predicting in all data (training + testing)
                         newdata = training_subset_months,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNIS1, prob_Q = EUNIS1.Q,
         prob_R = EUNIS1.R, prob_S = EUNIS1.S, prob_T = EUNIS1.T)
```

### Detect probability columns

```{r eval=FALSE, include=FALSE}
prob_cols_slope_Landsat <- names(training_subset_slope_with_probs)[grepl("^prob_", names(training_subset_slope_with_probs))]
prob_cols_months_Landsat <- names(training_subset_months_with_probs)[grepl("^prob_", names(training_subset_months_with_probs))]
```

### Calculate auxiliary metrics

```{r}
calculate_df_aux <- function(data, prob_cols) {
  data %>%
    rowwise() %>%
    mutate(
      # Probability of reference (original) class
      p_ref = get(paste0("prob_", original_EUNIS)),
      # Maximum probability between all classes
      max_prob = max(c_across(all_of(prob_cols)), na.rm = TRUE),
      # Second maximum probability
      second_max = sort(c_across(all_of(prob_cols)), decreasing = TRUE)[2],
      # Difference between original class and second best
      delta = p_ref - second_max,
      # TRUE if original class is not the most probable
      neg_flag = p_ref < max_prob,
      # TRUE if difference is small (uncertainty)
      small_flag = delta < 0.1,
      # How much worse is the original class with respect to the best class
      gap_neg = max_prob - p_ref
    ) %>%
    ungroup()
}
```

```{r eval=FALSE, include=FALSE}
df_aux_slope_Landsat <- calculate_df_aux(training_subset_slope_with_probs,
                                     prob_cols_slope_Landsat)
df_aux_months_Landsat <- calculate_df_aux(training_subset_months_with_probs,
                                     prob_cols_months_Landsat)
```

### Apply filtering function

Removing negatives until 20% of the class and, if there is room, uncertain positives until 10% of the rest.

```{r}
filter_probs_neg20_pos10 <- function(gdf) {
  # Removes negatives until 20% of the class
  cap_total <- ceiling(0.20 * nrow(gdf))  # max 20% of the class

  neg <- gdf %>% dplyr::filter(neg_flag) %>% arrange(desc(gap_neg))
  drop_neg <- head(neg, cap_total)
  cap_left <- cap_total - nrow(drop_neg)

  drop_pos <- tibble()
  if (cap_left > 0) {
    remaining <- nrow(gdf) - nrow(drop_neg)
    # If there is room, removes positives with doubts (small_flag = TRUE)
    # until 10% of the rest
    pos_cap <- ceiling(0.10 * remaining)

    drop_pos <- gdf %>%
      dplyr::filter(!neg_flag & small_flag) %>%
      arrange(delta) %>%
      slice_head(n = min(cap_left, pos_cap))
  }

  drop_ids <- c(drop_neg$PlotObservationID, drop_pos$PlotObservationID)
  gdf %>% dplyr::filter(!PlotObservationID %in% drop_ids) %>%
    select(-max_prob, -second_max, -delta, -neg_flag, -small_flag, -gap_neg)
}
```

```{r eval=FALSE, include=FALSE}
df_final_20_10_slope_Landsat <- df_aux_slope_Landsat %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20_pos10)
df_final_20_10_months_Landsat <- df_aux_months_Landsat %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20_pos10)
```

```{r eval=FALSE, include=FALSE}
filtered_data_20_10_slope_Landsat_probs <- training_subset_slope %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_slope_Landsat$PlotObservationID)
filtered_data_20_10_months_Landsat_probs <- training_subset_months %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_months_Landsat$PlotObservationID)
```

### Save

```{r}
filtered_data_with_probs <- c("filtered_data_20_10_slope_Landsat_probs", 
                              "filtered_data_20_10_months_Landsat_probs")
```

Save:

```{r eval=FALSE, include=FALSE}
for (filtered_data_name in filtered_data_with_probs) {
  file_path <- paste0("objects/RF_1500_cv_20260609/", filtered_data_name, ".Rdata")
  save(list = filtered_data_name, file = file_path)
}
```

### Fit models

#### Landsat, slope

##### CV

```{r eval=FALSE, include=FALSE}
cv_slope_Landsat_step1 <- run_fastcforest_cv(
  data = filtered_data_20_10_slope_Landsat_probs,
  vars_RF = vars_RF_slope_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_slope_Landsat_step1, 
     file = "objects/RF_1500_cv_20260609/cv_slope_Landsat_step1.Rdata") 
```

##### Results

```{r}
cv_slope_Landsat_step1_results <-
  compute_rf_cv_metrics(cv_slope_Landsat_step1)
cv_slope_Landsat_step1_results
```

##### Fit model on full dataset

```{r eval=FALSE, include=FALSE}
RF_slope_Landsat_step1 <- run_rf_parallel(vars_RF_slope_Landsat, 
                                          filtered_data_20_10_slope_Landsat_probs, 
                                          "EUNIS1", ntree = 1500)
```

```{r eval=FALSE, include=FALSE}
save(RF_slope_Landsat_step1,
     file = "objects/RF_1500_cv_20260609/RF_slope_Landsat_step1.Rdata") 
```

#### Landsat, months

##### CV

```{r eval=FALSE, include=FALSE}
cv_months_Landsat_step1 <- run_fastcforest_cv(
  data = filtered_data_20_10_months_Landsat_probs,
  vars_RF = vars_RF_months_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_months_Landsat_step1, 
     file = "objects/RF_1500_cv_20260609/cv_months_Landsat_step1.Rdata") 
```

##### Results

```{r}
cv_months_Landsat_step1_results <-
  compute_rf_cv_metrics(cv_months_Landsat_step1)
cv_months_Landsat_step1_results
```

##### Fit model on full dataset

```{r eval=FALSE, include=FALSE}
RF_months_Landsat_step1 <- run_rf_parallel(vars_RF_months_Landsat, 
                                          filtered_data_20_10_months_Landsat_probs, 
                                          "EUNIS1", ntree = 1500)
```

```{r eval=FALSE, include=FALSE}
save(RF_months_Landsat_step1,
     file = "objects/RF_1500_cv_20260609/RF_months_Landsat_step1.Rdata") 
```

## Refinement step 2

#### Compute predictions for all data and combine with original data

```{r eval=FALSE, include=FALSE}
# Landsat, slope
filtered_data_slope_Landsat_with_probs_2x <- bind_cols(
  filtered_data_20_10_slope_Landsat_probs %>% select(PlotObservationID, EUNIS1),
  as.data.frame(predict(RF_slope_Landsat_step1$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_20_10_slope_Landsat_probs,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(RF_slope_Landsat_step1$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_20_10_slope_Landsat_probs,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNIS1, prob_Q = EUNIS1.Q,
         prob_R = EUNIS1.R, prob_S = EUNIS1.S, prob_T = EUNIS1.T)

# Landsat, months
filtered_data_months_Landsat_with_probs_2x <- bind_cols(
  filtered_data_20_10_months_Landsat_probs %>% select(PlotObservationID, EUNIS1),
  as.data.frame(predict(RF_months_Landsat_step1$model, 
                        # Predicting in all data (training + testing)
                                          newdata = filtered_data_20_10_months_Landsat_probs,
                      type = "response")) %>%
    rename(predicted_EUNIS = 1),
  do.call(rbind, predict(RF_months_Landsat_step1$model, 
                         # Predicting in all data (training + testing)
                         newdata = filtered_data_20_10_months_Landsat_probs,
                      type = "prob")) %>% as.data.frame()
  ) %>%
  rename(original_EUNIS = EUNIS1, prob_Q = EUNIS1.Q,
         prob_R = EUNIS1.R, prob_S = EUNIS1.S, prob_T = EUNIS1.T)
```

### Detect probability columns

```{r eval=FALSE, include=FALSE}
prob_cols_slope_Landsat <- names(filtered_data_slope_Landsat_with_probs_2x)[grepl("^prob_", names(filtered_data_slope_Landsat_with_probs_2x))]
prob_cols_months_Landsat <- names(filtered_data_months_Landsat_with_probs_2x)[grepl("^prob_", names(filtered_data_months_Landsat_with_probs_2x))]
```

#### Calculate auxiliary metrics

```{r eval=FALSE, include=FALSE}
df_aux_slope_Landsat_2x <- calculate_df_aux(filtered_data_slope_Landsat_with_probs_2x,
                                     prob_cols_slope_Landsat)
df_aux_months_Landsat_2x <- calculate_df_aux(filtered_data_months_Landsat_with_probs_2x,
                                     prob_cols_months_Landsat)
```

#### Apply filtering function

```{r}
filter_probs_neg20 <- function(gdf) {
  # Removes negatives until 20% of the class
  cap_total <- ceiling(0.20 * nrow(gdf))  # max 20% of the class

  neg <- gdf %>% filter(neg_flag) %>% arrange(desc(gap_neg))
  drop_neg <- head(neg, cap_total)
  cap_left <- cap_total - nrow(drop_neg)

  drop_ids <- c(drop_neg$PlotObservationID)
  gdf %>% filter(!PlotObservationID %in% drop_ids) %>%
    select(-max_prob, -second_max, -delta, -neg_flag, -small_flag, -gap_neg)
}
```

```{r eval=FALSE, include=FALSE}
df_final_20_10_slope_Landsat_2x <- df_aux_slope_Landsat_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)
df_final_20_10_months_Landsat_2x <- df_aux_months_Landsat_2x %>%
  group_split(original_EUNIS) %>%
  map_dfr(filter_probs_neg20)
```

```{r eval=FALSE, include=FALSE}
filtered_data_20_10_slope_Landsat_probs_2x <- filtered_data_20_10_slope_Landsat_probs %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_slope_Landsat_2x$PlotObservationID)
filtered_data_20_10_months_Landsat_probs_2x <- filtered_data_20_10_months_Landsat_probs %>%
  dplyr::filter(PlotObservationID %in% df_final_20_10_months_Landsat_2x$PlotObservationID)
```

### Save

```{r}
filtered_data_with_probs_2x <- c("filtered_data_20_10_slope_Landsat_probs_2x", 
                              "filtered_data_20_10_months_Landsat_probs_2x")
```

Save:

```{r eval=FALSE, include=FALSE}
for (filtered_data_name in filtered_data_with_probs_2x) {
  file_path <- paste0("objects/RF_1500_cv_20260609/", filtered_data_name, ".Rdata")
  save(list = filtered_data_name, file = file_path)
}
```

#### (Appendix 6) Compare classes before and after

```{r}
compare_classes <- function(
  data_before, data_step1, data_step2 = NULL,
  class_col = "EUNIS1",
  x_lab = NULL,
  title = NULL,
  return = c("plot", "both", "counts")
) {
  return <- match.arg(return)

  # Helper: safe counting that works with empty data frames
  count_safe <- function(df, out_name) {
    if (is.null(df) || nrow(df) == 0) {
      tibble(!!class_col := character(), !!out_name := integer())
    } else {
      if (!(class_col %in% names(df))) {
        abort(paste0("Column '", class_col, "' is missing in one of the inputs."))
      }
      df %>% count(.data[[class_col]], name = out_name)
    }
  }

  # Build counts table (before + step1; optionally step2)
  counts <- full_join(
    count_safe(data_before, "before"),
    count_safe(data_step1,  "step1"),
    by = class_col
  )

  has_step2 <- !is.null(data_step2)
  if (has_step2) {
    counts <- full_join(
      counts,
      count_safe(data_step2, "step2"),
      by = class_col
    )
  } else {
    # Ensure the column exists for consistent downstream handling (as NA)
    counts <- counts %>% mutate(step2 = NA_integer_)
  }

  # Replace missing counts with zero for plotting
  counts <- counts %>%
    mutate(
      before = replace_na(before, 0L),
      step1  = replace_na(step1,  0L),
      step2  = replace_na(step2,  0L)
    )

  # Prepare long format depending on availability of step2
  if (has_step2) {
    counts_long <- counts %>%
      pivot_longer(
        cols = c(before, step1, step2),
        names_to = "state",
        values_to = "n"
      ) %>%
      mutate(state = factor(state, levels = c("before", "step1", "step2")))
    legend_title <- "State"
  } else {
    counts_long <- counts %>%
      select(any_of(c(class_col, "before", "step1"))) %>%
      pivot_longer(
        cols = c(before, step1),
        names_to = "state",
        values_to = "n"
      ) %>%
      mutate(state = factor(state, levels = c("before", "step1")))
    legend_title <- "State"
  }

    # Plot
  p <- ggplot(counts_long, aes(x = .data[[class_col]], y = n, fill = state)) +
    geom_col(position = position_dodge(width = 0.85),
             width = 0.8, color = "black", linewidth = 0.25) +
    labs(
      title = title,
      x = x_lab,
      y = "Number of plot observations",
      fill = legend_title
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 0, hjust = 1),
      panel.grid.minor = element_blank()
    ) +
    scale_x_discrete(
    breaks = c("Q", "R", "S", "T"),
    
    labels = function(x) {
      x <- ifelse(x == "Q", "Q - Wetlands", x)
      x <- ifelse(x == "R", "R - Grasslands", x)
      x <- ifelse(x == "S", "S - Shrublands", x)
      x <- ifelse(x == "T", "T - Forests", x)
      x
      }
    ) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  # Optionally attach counts to the plot as an attribute for quick retrieval
  attr(p, "counts") <- counts

  # Return mode
  if (return == "plot") {
    return(p)
  } else if (return == "counts") {
    return(counts)
  } else {
    # both
    return(list(plot = p, counts = counts))
  }
}
```

```{r}
(compare_classes(training_subset_slope, 
                  filtered_data_20_10_slope_Landsat_probs,
                  filtered_data_20_10_slope_Landsat_probs_2x,
                  return = "plot") +
  compare_classes(training_subset_months, 
                  filtered_data_20_10_months_Landsat_probs,
                  filtered_data_20_10_months_Landsat_probs_2x,
                  return = "plot")
 ) +
  plot_annotation(tag_levels = "A", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  scale_fill_manual(
    breaks = c("before", "step1", "step2"),
    labels = c("No refinement", "Refinement step 1", "Refinement step 2"),
    values = c(
      "before" = "#edf8b1", 
      "step1"  = "#7fcdbb",
      "step2"  = "#2c7fb8"
    )
  ) &
  scale_y_continuous(limits = c(0, 45000),
                  labels = label_number(scale = 1e-3, suffix = "k")) &
  theme(legend.position = "bottom", legend.title = element_blank()) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
ggsave(here("output", "figures", "manuscript", "appendix6.jpeg"), 
       dpi = 300, width = 6, height = 4.5)
```

### Fit models

#### Landsat, slope

##### CV

```{r eval=FALSE, include=FALSE}
cv_slope_Landsat_step2 <- run_fastcforest_cv(
  data = filtered_data_20_10_slope_Landsat_probs_2x,
  vars_RF = vars_RF_slope_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_slope_Landsat_step2, 
     file = "objects/RF_1500_cv_20260609/cv_slope_Landsat_step2.Rdata") 
```

##### Results

```{r}
cv_slope_Landsat_step2_results <-
  compute_rf_cv_metrics(cv_slope_Landsat_step2)
cv_slope_Landsat_step2_results
```

##### Fit model on full dataset

```{r eval=FALSE, include=FALSE}
RF_slope_Landsat_step2 <- run_rf_parallel(vars_RF_slope_Landsat, 
                                          filtered_data_20_10_slope_Landsat_probs_2x, 
                                          "EUNIS1", ntree = 1500)
```

```{r eval=FALSE, include=FALSE}
save(RF_slope_Landsat_step2,
     file = "objects/RF_1500_cv_20260609/RF_slope_Landsat_step2.Rdata") 
```

#### Landsat, months

##### CV

```{r eval=FALSE, include=FALSE}
cv_months_Landsat_step2 <- run_fastcforest_cv(
  data = filtered_data_20_10_months_Landsat_probs_2x,
  vars_RF = vars_RF_months_Landsat,
  response_var = "EUNIS1",
  ntree = 1500,
  k = 4
)
```

```{r eval=FALSE, include=FALSE}
save(cv_months_Landsat_step2, file = "objects/RF_1500_cv_20260609/cv_months_Landsat_step2.Rdata") 
```

##### (Table 2) Results

```{r}
cv_months_Landsat_step2_results <-
  compute_rf_cv_metrics(cv_months_Landsat_step2)
cv_months_Landsat_step2_results
```

##### Fit model on full dataset

```{r eval=FALSE, include=FALSE}
RF_months_Landsat_step2 <- run_rf_parallel(vars_RF_months_Landsat, 
                                          filtered_data_20_10_months_Landsat_probs_2x, 
                                          "EUNIS1", ntree = 1500)
```

```{r eval=FALSE, include=FALSE}
save(RF_months_Landsat_step2,
     file = "objects/RF_1500_cv_20260609/RF_months_Landsat_step2.Rdata") 
```

# Best model

## (Appendix 9) Variable importance from best model

```{r}
compute_varimp <- function(model, nperm = 100) {

  # Measure execution time
  execution_time <- system.time({
    varimp_result <- permimp(model, conditional = FALSE, progressBar = TRUE)
  })

  return(list(varimp = varimp_result, time = execution_time))
}
```

```{r eval=FALSE, include=FALSE}
varimp_months_Landsat_step2 <-
  compute_varimp(RF_months_Landsat_step2$model, nperm = 100)
```

```{r eval=FALSE, include=FALSE}
save(varimp_months_Landsat_step2,
     file = "objects/RF_1500_cv_20260609/varimp_months_Landsat_step2.Rdata")
```

```{r}
enframe(
  varimp_months_Landsat_step2$varimp$values,
  name  = "variable",
  value = "importance"
) %>%
  mutate(
    importance100 = 100 * importance / max(importance)
  ) %>%
  arrange(desc(importance100)) %>%
  mutate(variable_long = case_when(
    variable == "NDMI_max" ~ "Maximum NDMI",
    variable == "NDWI_max" ~ "Maximum NDWI",
    variable == "NDVI_max" ~ "Maximum NDVI",
    variable == "SAVI_max" ~ "Maximum SAVI",
    variable == "NDMI_min" ~ "Minimum NDMI",
    variable == "NDWI_min" ~ "Minimum NDWI",
    variable == "NDVI_min" ~ "Minimum NDVI",
    variable == "SAVI_min" ~ "Minimum SAVI",
    variable == "NDVI_avg_value_03" ~ "NDVI value at greening",
    variable == "NDVI_pos_value" ~ "NDVI value at peak",
    variable == "NDVI_avg_value_10" ~ "NDVI value at browning",
    variable == "SAVI_avg_value_03" ~ "SAVI value at greening",
    variable == "SAVI_pos_value" ~ "SAVI value at peak",
    variable == "SAVI_avg_value_10" ~ "SAVI value at browning",
    variable == "NDVI_auc_mar_oct" ~ "NDVI AUC greening - browning",
    variable == "SAVI_auc_mar_oct" ~ "SAVI AUC greening - browning",
    variable == "NDVI_diff_pos_march_value" ~ 
      "NDVI difference in value peak - greening",
    variable == "SAVI_diff_pos_march_value" ~
      "SAVI difference in value peak - greening",
    variable == "NDVI_diff_pos_march_doy" ~
      "NDVI difference in DOY peak - greening",
    variable == "SAVI_diff_pos_march_doy" ~
      "SAVI difference in DOY peak - greening",
    variable == "NDVI_diff_pos_oct_value" ~
      "NDVI difference in value peak - browning",
    variable == "SAVI_diff_pos_oct_value" ~
      "SAVI difference in value peak - browning",
    variable == "NDVI_diff_pos_oct_doy" ~
      "NDVI difference in DOY peak - browning",
    variable == "SAVI_diff_pos_oct_doy" ~
      "SAVI difference in DOY peak - browning",
    TRUE ~ variable
  )) %>%
ggplot(aes(x = reorder(variable_long, importance100), y = importance100)) +
  geom_col(fill = "#2c7fb8") +
  coord_flip() +
  labs(x = "Predictor", y = "Importance (scaled)") +
  theme_minimal()
ggsave(here("output", "figures", "manuscript", "appendix9.jpeg"), 
       dpi = 300, width = 6, height = 4)
```

## (Appendix 8) ROC curves from best model

```{r}
compute_roc_level1 <- function(model, test_data) {
  # Measure execution time
  execution_time <- system.time({
    # Step 1: Predict probabilities
    probabilities <- predict(model, newdata = test_data, type = "prob")
    
    # Step 2: Convert list of matrices to a proper data frame
    prob_matrix <- t(sapply(probabilities, as.vector))
    prob_df <- as.data.frame(prob_matrix)
    colnames(prob_df) <- c("Q", "R", "S", "T")
    
    # Step 3: Prepare actual class labels
    actual <- factor(test_data$EUNIS1, levels = c("Q", "R", "S", "T"))
    
    # Step 4: Binarize actual labels
    actual_bin <- model.matrix(~ actual - 1)
    colnames(actual_bin) <- gsub("actual", "", colnames(actual_bin))
    
    # Step 5: Compute ROC data for each class
    roc_data <- lapply(levels(actual), function(class) {
      roc_obj <- roc(actual_bin[, class], prob_df[[class]])
      auc_val <- round(auc(roc_obj), 3)
      data.frame(
        FPR = 1 - roc_obj$specificities,
        TPR = roc_obj$sensitivities,
        Class = paste0(class, " (AUC = ", auc_val, ")")
      )
    }) %>% bind_rows()
  })
  
  # Return both ROC data and execution time
  return(list(roc = roc_data, time = execution_time))
}
```

```{r eval=FALSE, include=FALSE}
rocdata_months_Landsat_step2 <-
  compute_roc_level1(RF_months_Landsat_step2$model,
                     filtered_data_20_10_months_Landsat_probs_2x)
```

```{r eval=FALSE, include=FALSE}
save(rocdata_months_Landsat_step2,
     file = "objects/RF_1500_cv_20260609/rocdata_months_Landsat_step2.Rdata")
```

```{r}
ggplot(rocdata_months_Landsat_step2$roc, aes(x = FPR, y = TPR, color = Class)) +
  geom_line(linewidth = 1.2) +
  geom_abline(linetype = "dashed", color = "gray") +
  labs(title = paste("Multiclass ROC Curve -", model_name),
       x = "False Positive Rate (1-Specificity",
       y = "True Positive Rate (Sensitivity)",
       color = "Class (AUC)") +
  scale_color_manual(
    values = c(
      "Q (AUC = 0.915)" = "#0072B2",
      "R (AUC = 0.978)" = "#D55E00",
      "S (AUC = 0.967)" = "#009E73",
      "T (AUC = 0.997)" = "#CC79A7"
    ),
    labels = c(
      "Q (AUC = 0.915)" = "Q – Wetlands (AUC = 0.915)",
      "R (AUC = 0.978)" = "R – Grasslands (AUC = 0.978)",
      "S (AUC = 0.967)" = "S – Shrublands (AUC = 0.967)",
      "T (AUC = 0.997)" = "T – Forests (AUC = 0.997)"
    )
  ) +
  theme_minimal() + ggtitle(NULL) +
  theme(legend.position = "right")
ggsave(here("output", "figures", "manuscript", "appendix8.jpeg"), 
       dpi = 300, width = 7, height = 4)
```

# Session info

```{r}
sessionInfo()
```
